---
title: "Phase 0 Prototype: P(true Sharpe > 0) column"
subtitle: "Revision 2 (owner feedback 2026-09-29, draft PR #927): short cell labels + hover/focus pop-up. Throwaway prototype for owner approval — see .claude/rules/dashboard-output-first.md. NOT the production build."
format:
  html:
    theme:
      light: cosmo
      dark: darkly
    embed-resources: true
    toc: true
    include-in-header:
      - text: |
          <!-- accessibility rule Part 2 Clause 0/6: any color-scheme
               declaration suppresses Chrome's Auto Dark Mode heuristic,
               regardless of value -- see docs/_includes/color-scheme-meta.html,
               copied inline here rather than referenced by relative path
               since this prototype does not render inside the docs/
               Quarto project. -->
          <meta name="color-scheme" content="dark" />
          <style>:root, html, body { color-scheme: dark; }</style>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(dplyr)
library(DT)
library(htmltools)

# READ-ONLY: points at the MAIN CHECKOUT's built store, never a worktree
# store (this worktree has no _targets store of its own). This prototype
# does not build or write anything -- it only reads the `leaderboard`
# target as it exists on origin/main today.
tar_config_set(store = "/Users/johngavin/docs_gh/proj/finance/data/historical/docs/_targets")

lb <- tar_read(leaderboard)
full <- lb |>
  dplyr::filter(period == "Full Period") |>
  dplyr::arrange(dplyr::desc(sharpe))

# ── HD_PROB_SHARPE_POSITIVE_THRESHOLD, extracted from its OWN source line in
# R/plan_leaderboard.R rather than hand-typed here or fully sourced (see
# Revision 1's comment for why a full source() is disproportionate for a
# throwaway prototype).
plan_lb_path <- if (file.exists("../../R/plan_leaderboard.R")) {
  "../../R/plan_leaderboard.R"
} else {
  file.path(here::here(), "R/plan_leaderboard.R")
}
threshold_line <- grep(
  "^HD_PROB_SHARPE_POSITIVE_THRESHOLD\\s*<-",
  readLines(plan_lb_path),
  value = TRUE
)
stopifnot(length(threshold_line) == 1L)
HD_PROB_SHARPE_POSITIVE_THRESHOLD <- eval(parse(text = threshold_line))

pct     <- function(x) sprintf("%.1f%%", x * 100)
fmt_yr  <- function(x) if (is.na(x)) "?" else sprintf("%.0f", x)
fmt_yr1 <- function(x) if (is.na(x)) "?" else sprintf("%.1f", x)
fmt_keff <- function(x) if (is.na(x)) "n/a" else sprintf("%.1f", x)

# ── "years available" (full$years) is genuinely NA for 15 of 18 Full-Period
# strategies as of this dispatch (verified via a direct tar_read() query,
# not assumed) -- some per-strategy metrics builders in R/plan_leaderboard.R
# never populate the `years` column, unlike `months`/`detection_min_n_years`
# which are populated for every positive-Sharpe row. Per fail-loud-not-null,
# an unavailable value is stated plainly rather than silently reverse-
# engineered from `months`/`obs_ann_factor` (which would require sourcing
# STRATEGY_OBS_ANN_FACTOR from R/plan_leaderboard.R -- out of scope for a
# throwaway prototype confined to explorations/**). Flagged as a follow-up
# note in this revision's PR body, not fixed here.
avail_phrase <- function(avail) {
  if (is.na(avail)) {
    "years available not recorded for this strategy in the leaderboard target (data gap -- see #726)"
  } else {
    sprintf("<b>%s yr</b> available", fmt_yr1(avail))
  }
}

# ── Drilldown link targets (owner feedback, PR #927 revision 2) ────────────
# Relative paths from explorations/prob_sharpe_prototype/ to docs/ -- this
# prototype is opened locally from a repo checkout, not deployed, so a
# relative link (not a github.io URL) is the correct, verifiable target.
# Anchor confirmed against docs/leaderboard.qmd's `# Biases and Caveats`
# heading (line ~2098) -- Quarto's default heading-to-anchor slug.
CAVEATS_HREF <- "../../docs/leaderboard.html#biases-and-caveats"
ISSUE_BASE   <- "https://github.com/JohnGavin/historical/issues/"

# Strategy -> its own dashboard page, ONLY where one genuinely exists.
# Verified by grepping docs/*.qmd for each strategy's short_name (see
# dispatch notes) rather than guessing from filenames -- e.g. "Risk State"
# and "Macro Defense Rotation" are NOT the same strategy despite both being
# VIX-related, so Risk State deliberately has NO entry here rather than a
# wrong link. Anchors confirmed against docs/stock-backtest.qmd headings
# (`# Stock MAX`, `# Stock DRIF`, `# XGBoost DRIF`,
# `# Factor-Level Deep Dive {#factor-level-deep-dive}`).
STRATEGY_DASHBOARD_HREF <- list(
  "Avoid Worst"   = "../../docs/avoid-worst-days.html",
  "Factor DRIF"   = "../../docs/stock-backtest.html#factor-level-deep-dive",
  "Factor MAX"    = "../../docs/stock-backtest.html#factor-level-deep-dive",
  "Stock MAX"     = "../../docs/stock-backtest.html#stock-max",
  "Stock DRIF"    = "../../docs/stock-backtest.html#stock-drif",
  "XGB DRIF"      = "../../docs/stock-backtest.html#xgboost-drif",
  "Mom Pre-Peak"  = "../../docs/momentum-prepeak.html",
  "Mom Post-Peak" = "../../docs/momentum-prepeak.html",
  "Mom 12-2"      = "../../docs/momentum-prepeak.html"
)

issue_link <- function(n) {
  sprintf("<a href='%s%s' target='_blank' rel='noopener'>#%s</a>", ISSUE_BASE, n, n)
}

pop_links <- function(strategy, issue_nums) {
  parts <- sprintf(
    "<a href='%s' target='_blank' rel='noopener'>Detection &amp; Rigour caveats &#8599;</a>",
    CAVEATS_HREF
  )
  sh <- STRATEGY_DASHBOARD_HREF[[strategy]]
  if (!is.null(sh)) {
    parts <- c(parts, sprintf(
      "<a href='%s' target='_blank' rel='noopener'>%s dashboard &#8599;</a>", sh, strategy
    ))
  }
  parts <- c(parts, vapply(issue_nums, issue_link, character(1)))
  paste(parts, collapse = " &middot; ")
}

# ── pop_wrap(): the SHORT badge + hover/focus pop-up shell ─────────────────
# A CSS-driven hover/focus popover (accessibility rule's "Do not rely on
# the native HTML title attribute" clause) -- NOT `title=`. `tabindex='0'`
# makes the wrapper itself keyboard-focusable so `:focus`/`:focus-within`
# (see the <style> block below) reveal the pop-up without a mouse. `family`
# is the existing badge-family class (det-badge/psr-badge, unchanged from
# revision 1) and `state` is the existing state class (detectable/mtfail/
# underpowered/napp/notcomputed for Detection; sufficient/mtfail/
# insufficient/napp/notcomputed for P(SR>0)) -- BOTH still carried on the
# element so the colour-mapping CSS selectors from revision 1 keep working
# unchanged, just re-targeted at `.pop-short` instead of the whole badge.
pop_wrap <- function(family, state, short, body_html, aria) {
  sprintf(
    "<span class='pop-wrap %s %s' tabindex='0' aria-label='%s'><span class='pop-short'>%s</span><span class='pop-body'>%s</span></span>",
    family, state, aria, short, body_html
  )
}

DET_MEANING <- "<span class='pop-meaning'>Detection asks: is the sample long enough to reliably detect this Sharpe if it is real? (prospective power test, one-sided, &alpha;=0.05)</span>"
PSR_MEANING <- "<span class='pop-meaning'>P(SR&gt;0) asks: given what we observed, how confident are we the TRUE Sharpe is positive? (retrospective confidence, Bailey &amp; Lopez de Prado)</span>"

# ── Detection badge (short label + pop-up), mirrors historicaldata::
# hd_detection_power()'s 5-state verdict shape from revision 1's
# detection_badge(), restructured for a SHORT cell (icon + years
# available/years needed, <=~11 chars) with the full wording moved into a
# hover/focus pop-up. See .claude/rules/detection-power-required.md.
#
# Design choice, kept consistent across every state: the SECOND number in
# the short label is always the "years needed" figure that is OPERATIVE
# for that row's final verdict -- the single-test figure for every state
# except `mtfail`, where the MT-CORRECTED figure is shown instead (that is
# the number that actually fails). This keeps the available/needed
# comparison in the cell always readable as "pass" (available >= needed
# shown) or "fail" (available < needed shown), regardless of which test
# produced the verdict.
detection_badge2 <- function(strategy, sr, avail, under, dmin, under_mt, dmin_mt, keff) {
  links <- pop_links(strategy, c(726, 903))

  if (is.na(sr) || sr <= 0) {
    body <- paste0(
      "<strong>&ndash; Not applicable</strong><br>",
      "Sharpe &le; 0: the one-sided detection test has no positive effect to test for. Not applicable &mdash; not the same as unproven.<br>",
      DET_MEANING, "<br>",
      "<span class='pop-links'>", links, "</span>"
    )
    return(list(rank = 4L, html = pop_wrap(
      "det-badge", "napp", "&ndash; n/a", body,
      "Detection: not applicable, Sharpe is not positive"
    )))
  }

  if (is.na(under)) {
    body <- paste0(
      "<strong>&#10007; NOT COMPUTED</strong><br>",
      "Positive Sharpe with NO detection verdict &mdash; this should never render live (QA gate S20, #726 item 3, is meant to block the build until it is fixed).<br>",
      DET_MEANING, "<br>",
      "<span class='pop-links'>", links, "</span>"
    )
    return(list(rank = 5L, html = pop_wrap(
      "det-badge", "notcomputed", "&#10007; ERR", body,
      "Detection: defect, verdict was not computed"
    )))
  }

  if (isTRUE(under)) {
    mt_line <- if (!is.na(under_mt)) {
      "Multiple-testing correction is academic here &mdash; the single test already fails."
    } else {
      "MT-corrected verdict not computed (k_eff unavailable for this strategy)."
    }
    body <- paste0(
      "<strong>&#9888; Underpowered</strong><br>",
      sprintf("Needs <b>%s yr</b> at &alpha;=0.05, one-sided; %s.<br>", fmt_yr1(dmin), avail_phrase(avail)),
      mt_line, "<br>",
      DET_MEANING, "<br>",
      "<span class='pop-links'>", links, "</span>"
    )
    short <- sprintf("&#9888; %s/%sy", fmt_yr(avail), fmt_yr(dmin))
    return(list(rank = 3L, html = pop_wrap(
      "det-badge", "underpowered", short, body,
      sprintf("Detection: underpowered, needs %s years, %s years available", fmt_yr(dmin), fmt_yr(avail))
    )))
  }

  if (is.na(under_mt)) {
    body <- paste0(
      "<strong>&#10003; Detectable</strong><br>",
      sprintf("Needs <b>%s yr</b> at &alpha;=0.05, one-sided; %s.<br>", fmt_yr1(dmin), avail_phrase(avail)),
      "MT-corrected verdict not computed (k_eff unavailable for this strategy).<br>",
      DET_MEANING, "<br>",
      "<span class='pop-links'>", links, "</span>"
    )
    short <- sprintf("&#10003; %s/%sy", fmt_yr(avail), fmt_yr(dmin))
    return(list(rank = 1L, html = pop_wrap(
      "det-badge", "detectable", short, body,
      sprintf("Detection: detectable, needs %s years, %s years available", fmt_yr(dmin), fmt_yr(avail))
    )))
  }

  if (isTRUE(under_mt)) {
    body <- paste0(
      "<strong>&#9670; Fails MT-correction</strong><br>",
      sprintf(
        "Passes the single test (needs <b>%s yr</b>; %s) but FAILS once corrected for testing multiple strategies (needs <b>%s yr</b> corrected, k_eff=%s).<br>",
        fmt_yr1(dmin), avail_phrase(avail), fmt_yr1(dmin_mt), fmt_keff(keff)
      ),
      DET_MEANING, "<br>",
      "<span class='pop-links'>", links, "</span>"
    )
    short <- sprintf("&#9670; %s/%sy", fmt_yr(avail), fmt_yr(dmin_mt))
    return(list(rank = 2L, html = pop_wrap(
      "det-badge", "mtfail", short, body,
      sprintf("Detection: passes single test, fails multiple-testing correction, needs %s years corrected", fmt_yr(dmin_mt))
    )))
  }

  body <- paste0(
    "<strong>&#10003; Detectable</strong><br>",
    sprintf(
      "Needs <b>%s yr</b> at &alpha;=0.05, one-sided (<b>%s yr</b> corrected for multiple testing, k_eff=%s); %s.<br>",
      fmt_yr1(dmin), fmt_yr1(dmin_mt), fmt_keff(keff), avail_phrase(avail)
    ),
    DET_MEANING, "<br>",
    "<span class='pop-links'>", links, "</span>"
  )
  short <- sprintf("&#10003; %s/%sy", fmt_yr(avail), fmt_yr(dmin))
  list(rank = 1L, html = pop_wrap(
    "det-badge", "detectable", short, body,
    "Detection: detectable, passes single test and multiple-testing correction"
  ))
}

# ── P(SR>0) badge (short label + pop-up) -- mirrors detection_badge2()'s
# exact 5-state shape and severity-rank convention, applied to
# historicaldata::hd_prob_sharpe_positive() instead of hd_detection_power().
# Same "operative number" design choice: the short label shows the
# single-test percentage for every state except `mtfail`, where it shows
# the MT-CORRECTED percentage (the number that fails).
psr_badge2 <- function(strategy, sr, psr, psr_insuff, psr_mt, psr_mt_insuff, threshold, keff) {
  links <- pop_links(strategy, 851)

  if (is.na(sr) || sr <= 0) {
    body <- paste0(
      "<strong>&ndash; Not applicable</strong><br>",
      "Sharpe &le; 0: P(true Sharpe &gt; 0) is not a meaningful question when the point estimate itself is non-positive. Not applicable &mdash; not the same as unproven.<br>",
      PSR_MEANING, "<br>",
      "<span class='pop-links'>", links, "</span>"
    )
    return(list(rank = 4L, html = pop_wrap(
      "psr-badge", "napp", "&ndash; n/a", body,
      "P(SR>0): not applicable, Sharpe is not positive"
    )))
  }

  if (is.na(psr)) {
    body <- paste0(
      "<strong>&#10007; NOT COMPUTED</strong><br>",
      "Positive Sharpe with NO P(SR&gt;0) verdict &mdash; this should never render live (QA gate S40, #851, is meant to block the build until it is fixed).<br>",
      PSR_MEANING, "<br>",
      "<span class='pop-links'>", links, "</span>"
    )
    return(list(rank = 5L, html = pop_wrap(
      "psr-badge", "notcomputed", "&#10007; ERR", body,
      "P(SR>0): defect, verdict was not computed"
    )))
  }

  if (isTRUE(psr_insuff)) {
    mt_line <- if (!is.na(psr_mt_insuff)) {
      "Multiple-testing correction is academic here &mdash; the single test already fails."
    } else {
      "MT-corrected verdict not computed (k_eff unavailable for this strategy)."
    }
    body <- paste0(
      "<strong>&#9888; Insufficient</strong><br>",
      sprintf("P(true Sharpe &gt; 0) = <b>%s</b>, below the <b>%s</b> bar.<br>", pct(psr), pct(threshold)),
      mt_line, "<br>",
      PSR_MEANING, "<br>",
      "<span class='pop-links'>", links, "</span>"
    )
    short <- sprintf("&#9888; %s", pct(psr))
    return(list(rank = 3L, html = pop_wrap(
      "psr-badge", "insufficient", short, body,
      sprintf("P(SR>0): insufficient, %s below the %s threshold", pct(psr), pct(threshold))
    )))
  }

  if (is.na(psr_mt)) {
    body <- paste0(
      "<strong>&#10003; Sufficient</strong><br>",
      sprintf("P(true Sharpe &gt; 0) = <b>%s</b> (single test), at or above the <b>%s</b> bar.<br>", pct(psr), pct(threshold)),
      "MT-corrected verdict not computed (k_eff unavailable for this strategy).<br>",
      PSR_MEANING, "<br>",
      "<span class='pop-links'>", links, "</span>"
    )
    short <- sprintf("&#10003; %s", pct(psr))
    return(list(rank = 1L, html = pop_wrap(
      "psr-badge", "sufficient", short, body,
      sprintf("P(SR>0): sufficient, %s single test", pct(psr))
    )))
  }

  if (isTRUE(psr_mt_insuff)) {
    body <- paste0(
      "<strong>&#9670; Fails MT-correction</strong><br>",
      sprintf(
        "Passes the single-test bar (<b>%s</b>) but FAILS once corrected for multiple testing (<b>%s</b>, below the <b>%s</b> bar, k_eff=%s).<br>",
        pct(psr), pct(psr_mt), pct(threshold), fmt_keff(keff)
      ),
      PSR_MEANING, "<br>",
      "<span class='pop-links'>", links, "</span>"
    )
    short <- sprintf("&#9670; %s", pct(psr_mt))
    return(list(rank = 2L, html = pop_wrap(
      "psr-badge", "mtfail", short, body,
      sprintf("P(SR>0): passes single test, fails multiple-testing correction at %s", pct(psr_mt))
    )))
  }

  body <- paste0(
    "<strong>&#10003; Sufficient</strong><br>",
    sprintf(
      "P(true Sharpe &gt; 0): <b>%s</b> single-test, <b>%s</b> corrected for multiple testing (k_eff=%s) &mdash; both at or above the <b>%s</b> bar.<br>",
      pct(psr), pct(psr_mt), fmt_keff(keff), pct(threshold)
    ),
    PSR_MEANING, "<br>",
    "<span class='pop-links'>", links, "</span>"
  )
  short <- sprintf("&#10003; %s", pct(psr))
  list(rank = 1L, html = pop_wrap(
    "psr-badge", "sufficient", short, body,
    sprintf("P(SR>0): sufficient, %s single test, %s corrected", pct(psr), pct(psr_mt))
  ))
}

det <- mapply(
  detection_badge2,
  strategy = full$strategy,
  sr       = full$sharpe,
  avail    = full$years,
  under    = full$detection_underpowered,
  dmin     = full$detection_min_n_years,
  under_mt = full$detection_underpowered_mt,
  dmin_mt  = full$detection_min_n_years_mt,
  keff     = full$k_eff_leaderboard,
  SIMPLIFY = FALSE
)
psr <- mapply(
  psr_badge2,
  strategy      = full$strategy,
  sr            = full$sharpe,
  psr           = full$prob_sharpe_positive,
  psr_insuff    = full$prob_sharpe_positive_insufficient,
  psr_mt        = full$prob_sharpe_positive_mt,
  psr_mt_insuff = full$prob_sharpe_positive_insufficient_mt,
  keff          = full$k_eff_leaderboard,
  MoreArgs      = list(threshold = HD_PROB_SHARPE_POSITIVE_THRESHOLD),
  SIMPLIFY = FALSE
)

metrics_table <- full |>
  mutate(
    .sharpe_rank = ifelse(is.na(sharpe), -Inf, sharpe),
    Sharpe       = ifelse(is.na(sharpe), "—", as.character(round(sharpe, 2))),
    Detection    = vapply(det, function(x) x$html, character(1)),
    .det_rank    = vapply(det, function(x) x$rank, integer(1)),
    `P(SR>0)`    = vapply(psr, function(x) x$html, character(1)),
    .psr_rank    = vapply(psr, function(x) x$rank, integer(1))
  ) |>
  select(Strategy = strategy, Sharpe, .sharpe_rank, Detection, .det_rank,
         `P(SR>0)`, .psr_rank)

sharpe_col      <- which(names(metrics_table) == "Sharpe") - 1L
sharpe_rank_col <- which(names(metrics_table) == ".sharpe_rank") - 1L
det_col         <- which(names(metrics_table) == "Detection") - 1L
det_rank_col    <- which(names(metrics_table) == ".det_rank") - 1L
psr_col         <- which(names(metrics_table) == "P(SR>0)") - 1L
psr_rank_col    <- which(names(metrics_table) == ".psr_rank") - 1L

# ── Dynamically computed disagreement diagnostic (unchanged from revision 1,
# never hardcoded) ──────────────────────────────────────────────────────────
pos <- full |> dplyr::filter(!is.na(sharpe), sharpe > 0)
det_overall <- !is.na(pos$detection_underpowered) & !pos$detection_underpowered &
  (is.na(pos$detection_underpowered_mt) | !pos$detection_underpowered_mt)
psr_overall <- !is.na(pos$prob_sharpe_positive_insufficient) & !pos$prob_sharpe_positive_insufficient &
  (is.na(pos$prob_sharpe_positive_insufficient_mt) | !pos$prob_sharpe_positive_insufficient_mt)
overall_disagree <- pos$strategy[det_overall != psr_overall]

det_single_ok <- !is.na(pos$detection_underpowered) & !pos$detection_underpowered
psr_single_ok <- !is.na(pos$prob_sharpe_positive_insufficient) & !pos$prob_sharpe_positive_insufficient
single_disagree <- pos$strategy[det_single_ok != psr_single_ok]

keff_nonNA <- full$k_eff_leaderboard[!is.na(full$k_eff_leaderboard)]
kraw_nonNA <- full$k_raw_leaderboard[!is.na(full$k_raw_leaderboard)]
keff <- if (length(keff_nonNA) > 0) round(keff_nonNA[1], 1) else NA_real_
kraw <- if (length(kraw_nonNA) > 0) kraw_nonNA[1] else NA_integer_
```

## Revision 2 (owner feedback, 2026-09-29): short labels + hover pop-up

**Problem with revision 1:** the P(SR>0) and Detection columns' cell text
(e.g. `Sufficient, fails MT-correction (98.7% MT)`,
`Detectable, fails MT-correction (needs 21yr corrected)`) was cut off at
the right edge of the table.

**Fix:** every cell now shows a **short badge** (icon + one compact number,
≤ ~11 characters) instead of a full sentence. The full verdict wording,
numbers, meaning, and drilldown links move into a **CSS hover/focus
pop-up** (not the native `title=` attribute — per the accessibility rule,
`title` is unreliable on hover-timing and not keyboard-accessible).
`tabindex='0'` on each badge makes it focusable, so `Tab` + the pop-up
reveals the same content a mouse hover does.

### Short-label scheme

Deliberately NOT a markdown table here — this Quarto/pandoc build (1.8.26)
throws 18 non-fatal `jog.lua: Don't know how to traverse TableBody` errors
on ANY literal `<table>` (markdown pipe table or raw HTML `<table>`) in the
prose, confirmed via a minimal isolated reproduction outside this file
during this dispatch. `DT::datatable()`'s widget output is unaffected (its
`<table>` is constructed by JS at runtime, not present as a literal tag at
pandoc-processing time), so only PROSE tables are avoided here; the actual
Rankings table below is unaffected.

- **Not applicable** (Sharpe ≤ 0) — icon `–`, grey italic — Detection:
  `– n/a` · P(SR>0): `– n/a`
- **Not computed** (defect) — icon `✗`, red — Detection: `✗ ERR` ·
  P(SR>0): `✗ ERR`
- **Fails the single test** — icon `⚠`, amber — Detection:
  `⚠ avail/needy` · P(SR>0): `⚠ XX.X%`
- **Passes** (MT not computed) — icon `✓`, green — Detection:
  `✓ avail/needy` · P(SR>0): `✓ XX.X%`
- **Passes single, fails MT-correction** — icon `◆`, purple — Detection:
  `◆ avail/need_MTy` · P(SR>0): `◆ XX.X%_MT`
- **Passes single and MT-correction** — icon `✓`, green — Detection:
  `✓ avail/needy` · P(SR>0): `✓ XX.X%`

Design choice: the **second number** in the Detection short label (and the
**single number** shown for P(SR>0)) is always the figure **operative** for
that row's final verdict — the single-test figure everywhere except the
`◆` (fails-MT) state, where the MT-corrected figure is shown instead. This
keeps every cell readable the same way: "available ≥ needed" always reads
as a pass, "available < needed" always reads as a fail, regardless of
which test produced the verdict.

### Pop-up contents (every state)

1. The full verdict, bolded (e.g. "Detectable", "Fails MT-correction").
2. The numbers behind it — for Detection: years needed (single-test and,
   where computed, MT-corrected) and years available; for P(SR>0): the
   single-test and MT-corrected percentages and the 95% threshold (read
   from `HD_PROB_SHARPE_POSITIVE_THRESHOLD` in `R/plan_leaderboard.R`, not
   hand-typed) — plus `k_eff` where a correction was applied.
3. One italicised line on what the metric means.
4. Drilldown links: the [Biases and
   Caveats](../../docs/leaderboard.html#biases-and-caveats) page's
   Detection & Rigour table, the strategy's **own dashboard page** where
   one genuinely exists (`STRATEGY_DASHBOARD_HREF` above — deliberately
   left blank for strategies with no dedicated page, rather than guessing
   a wrong link), and the explaining issue(s)
   ([#726](https://github.com/JohnGavin/historical/issues/726) /
   [#903](https://github.com/JohnGavin/historical/issues/903) for
   Detection, [#851](https://github.com/JohnGavin/historical/issues/851)
   for P(SR>0)).

## Step 1 — Relationship to the existing Detection column

**P(true Sharpe > 0)** (`historicaldata::hd_prob_sharpe_positive()`, Bailey &
Lopez de Prado's Probabilistic Sharpe Ratio) and **Detection**
(`historicaldata::hd_detection_power()`) answer two different questions about
the same row, and neither implies the other — see
`.claude/rules/detection-power-required.md`. In the live data read for this
prototype, that relationship is exactly what happens for
**`r if (length(single_disagree) > 0) single_disagree[1] else "no strategy currently"`**:
Detection reports the single-test verdict as `Underpowered`, while P(SR>0)'s
own single-test verdict is `Sufficient` at
`r if (length(single_disagree) > 0) { s <- single_disagree[1]; pct(pos$prob_sharpe_positive[pos$strategy == s]) } else "n/a"`.
Once both are corrected for testing `r kraw` strategies (K_eff = `r keff`),
the two diagnostics converge back to agreement on that row — see the
disagreement callout below the table.

**Diagram:** unchanged from revision 1 — no new Mermaid node is proposed.

## Step 2 — Home: exact placement in `docs/leaderboard.qmd`

Unchanged from revision 1 — new column **`P(SR>0)`** inserted immediately
right of the existing **Detection** column, both immediately right of
**Sharpe**. See revision 1's PR text for the exact line-by-line plan; this
revision only changes the CELL CONTENT and adds a `<style>` block for the
pop-up (below), not the placement.

## Step 3 — Comparable outputs this must match (updated for the pop-up)

- **Badge markup**: `.pop-wrap` (new wrapper, replaces the plain
  `.det-badge`/`.psr-badge span` from revision 1) carries the SAME family
  class (`det-badge` / `psr-badge`) and state class (`detectable` /
  `mtfail` / `underpowered` / `napp` / `notcomputed`; `sufficient` /
  `mtfail` / `insufficient` / `napp` / `notcomputed`) as before, so the
  existing colour-mapping convention needs no redesign — only its
  selectors now target `.pop-short` instead of the whole span.
- **Hover/focus pop-up, not `title=`**: per the accessibility rule's "Do
  not rely on the native HTML `title` attribute" clause — a CSS
  `:hover`/`:focus`/`:focus-within` popover instead, keyboard-accessible
  via `tabindex='0'`.
- **Number formatting**: unchanged — percentages to one decimal place,
  years to one decimal place in the pop-up (whole years in the short
  label, for compactness).
- **Threshold sourced, not hand-typed**: unchanged.
- **Sort-rank convention**: unchanged — hidden `.psr_rank`/`.det_rank`
  columns, `orderData`-mapped.
- **New: dashboard-table-styling compliance** — the `<style>` block below
  now also shrinks the table to its content width (`dashboard-table-styling`
  rule), which revision 1's screenshot did not need to worry about because
  its overflow problem was the CELL TEXT, not the table shell.

## Step 4 — Prototype (below)

Full Period, live `leaderboard` target read from the main checkout's build
(`docs/_targets`) at render time — the SAME data the production Rankings
table shows today.

```{r}
#| label: prototype-table
DT::datatable(
  metrics_table,
  escape = FALSE,
  rownames = FALSE,
  options = list(
    pageLength = 20,
    dom = "t",
    columnDefs = list(
      list(visible = FALSE, targets = c(sharpe_rank_col, det_rank_col, psr_rank_col)),
      list(orderData = sharpe_rank_col, targets = sharpe_col),
      list(orderData = det_rank_col, targets = det_col),
      list(orderData = psr_rank_col, targets = psr_col)
    ),
    order = list(list(sharpe_col, "desc"))
  )
)
```

```{=html}
<style>
/* dashboard-table-styling: shrink to content, don't stretch to 100% */
table.dataTable { width: auto !important; max-width: 100% !important; table-layout: auto !important; }
.dataTables_wrapper { width: auto !important; display: inline-block; }
table.dataTable th, table.dataTable td { white-space: nowrap; vertical-align: middle; padding: 0.35rem 0.8rem !important; }

/* ── Short badge + hover/focus pop-up (accessibility rule: no native title=) ── */
.pop-wrap {
  position: relative;
  display: inline-block;
  cursor: help;
}
.pop-short {
  font-weight: 600;
  padding: 1px 6px;
  border-radius: 3px;
  white-space: nowrap;
  border-bottom: 1px dotted currentColor;
}
.pop-body {
  position: absolute;
  left: 0;
  top: 125%;
  z-index: 60;
  min-width: 230px;
  max-width: min(340px, calc(100vw - 32px));
  background: #ffffff;
  color: #1a1a1a;
  border: 1px solid #ccc;
  border-radius: 6px;
  box-shadow: 0 4px 14px rgba(0, 0, 0, 0.18);
  padding: 8px 10px;
  font-size: 0.82rem;
  font-weight: 400;
  line-height: 1.45;
  white-space: normal;
  text-align: left;
  opacity: 0;
  visibility: hidden;
  pointer-events: none;
  transition: opacity 0.12s ease;
}
.pop-wrap:hover .pop-body,
.pop-wrap:focus .pop-body,
.pop-wrap:focus-within .pop-body {
  opacity: 1;
  visibility: visible;
  pointer-events: auto;
}
.pop-body .pop-meaning { display: block; margin-top: 4px; font-style: italic; opacity: 0.85; }
.pop-body .pop-links   { display: block; margin-top: 6px; }
.pop-body a { color: #0d6efd; text-decoration: underline; }

/* Colour mapping -- identical to revision 1, re-targeted at .pop-short */
.det-badge.detectable .pop-short,   .psr-badge.sufficient .pop-short   { color: #1a7f37; }
.det-badge.mtfail .pop-short,       .psr-badge.mtfail .pop-short       { color: #6f42c1; }
.det-badge.underpowered .pop-short, .psr-badge.insufficient .pop-short { color: #9a5b00; }
.det-badge.napp .pop-short,         .psr-badge.napp .pop-short         { color: #6c757d; font-style: italic; font-weight: 400; }
.det-badge.notcomputed .pop-short,  .psr-badge.notcomputed .pop-short  { color: #fff; background: #c0392b; }

[data-bs-theme="dark"] .pop-body { background: #1c1f24; color: #f5f5f5; border-color: #3a3f47; }
[data-bs-theme="dark"] .pop-body a { color: #7ab8ff; }
[data-bs-theme="dark"] .det-badge.detectable .pop-short,   [data-bs-theme="dark"] .psr-badge.sufficient .pop-short   { color: #4ade80; }
[data-bs-theme="dark"] .det-badge.mtfail .pop-short,       [data-bs-theme="dark"] .psr-badge.mtfail .pop-short       { color: #c084fc; }
[data-bs-theme="dark"] .det-badge.underpowered .pop-short, [data-bs-theme="dark"] .psr-badge.insufficient .pop-short { color: #fbbf24; }
[data-bs-theme="dark"] .det-badge.napp .pop-short,         [data-bs-theme="dark"] .psr-badge.napp .pop-short         { color: #9ca3af; }

@media (prefers-color-scheme: dark) {
  .pop-body { background: #1c1f24; color: #f5f5f5; border-color: #3a3f47; }
  .pop-body a { color: #7ab8ff; }
  .det-badge.detectable .pop-short,   .psr-badge.sufficient .pop-short   { color: #4ade80; }
  .det-badge.mtfail .pop-short,       .psr-badge.mtfail .pop-short       { color: #c084fc; }
  .det-badge.underpowered .pop-short, .psr-badge.insufficient .pop-short { color: #fbbf24; }
  .det-badge.napp .pop-short,         .psr-badge.napp .pop-short         { color: #9ca3af; }
}
</style>
```

::: {.callout-note title="Disagreement between Detection and P(SR>0), this build"}
Of `r nrow(pos)` strategies with a positive Full-Period Sharpe,
`r length(single_disagree)` disagree between the two diagnostics' own
**single-test** verdicts alone
(`r if (length(single_disagree) > 0) paste(single_disagree, collapse = ", ") else "none"`),
and `r length(overall_disagree)` disagree once each diagnostic's **overall**
verdict (single-test AND MT-corrected both required to pass) is compared
(`r if (length(overall_disagree) > 0) paste(overall_disagree, collapse = ", ") else "none"`).
Disagreement is expected and informative, not a bug — see Step 1.
:::

## Step 5 — Consolidation review

Unchanged from revision 1: no existing column becomes redundant, but this
addition motivates a flagged (not executed) follow-up to fold the
Biases-and-Caveats page's plain-column detection/rigour table into a
per-strategy drilldown popover — which this revision's pop-up mechanism is
now itself a working example of.

## Build info

`r paste0("Rendered ", format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z"), " | leaderboard rows (Full Period): ", nrow(full), " | commit: ", tryCatch(system("git rev-parse --short HEAD", intern = TRUE), error = function(e) "unknown"))`
