---
title: "Phase 0 Prototype: P(true Sharpe > 0) column"
subtitle: "Throwaway prototype for owner approval — see .claude/rules/dashboard-output-first.md. NOT the production build."
format:
  html:
    theme:
      light: cosmo
      dark: darkly
    embed-resources: true
    toc: true
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(dplyr)
library(DT)
library(htmltools)

# READ-ONLY: points at the MAIN CHECKOUT's built store, never a worktree
# store (this worktree has no _targets store of its own). This prototype
# does not build or write anything -- it only reads the `leaderboard`
# target as it exists on origin/main today.
#
# tar_config_set() writes a `_targets.yaml` config file alongside this
# .qmd (tar_read()/tar_load() read the store path back from that file, not
# from an in-session option -- redirecting it to tempfile() left tar_read()
# unable to find the store at all). That file holds a machine-specific
# absolute path, which `portable-build-artifacts` forbids committing -- it
# is covered by this repo's existing top-level `.gitignore` entry for the
# bare pattern `_targets.yaml` (already used for docs/_targets.yaml for the
# identical reason), so it is never staged.
tar_config_set(store = "/Users/johngavin/docs_gh/proj/finance/data/historical/docs/_targets")

lb <- tar_read(leaderboard)
full <- lb |>
  dplyr::filter(period == "Full Period") |>
  dplyr::arrange(dplyr::desc(sharpe))

# ── HD_PROB_SHARPE_POSITIVE_THRESHOLD, extracted from its OWN source line in
# R/plan_leaderboard.R rather than hand-typed here (per this dispatch's
# instructions) or fully sourced (sourcing the whole 1700+ line plan file
# would require loading the historicaldata package plus every upstream
# R/plan_strategy_names.R dependency just to read one scalar constant --
# disproportionate for a throwaway prototype that touches nothing else in
# that file). A regex extraction of the literal assignment line is not
# hand-typing: the value is read from the committed source at render time,
# so it tracks any future change to the threshold with zero edits here.
plan_lb_path <- if (file.exists("../../R/plan_leaderboard.R")) {
  "../../R/plan_leaderboard.R"
} else {
  file.path(here::here(), "R/plan_leaderboard.R")
}
threshold_line <- grep(
  "^HD_PROB_SHARPE_POSITIVE_THRESHOLD\\s*<-",
  readLines(plan_lb_path),
  value = TRUE
)
stopifnot(length(threshold_line) == 1L)
HD_PROB_SHARPE_POSITIVE_THRESHOLD <- eval(parse(text = threshold_line))

pct <- function(x) sprintf("%.1f%%", x * 100)

# ── Detection badge -- copied VERBATIM from docs/leaderboard.qmd's
# `detection_badge()` (5-state verdict, symbol AND text on every state per
# the accessibility rule) so this prototype's Detection column renders
# pixel-identically to production. Do NOT edit independently of that copy;
# if this prototype is approved, the real change updates leaderboard.qmd
# only -- this file stays a frozen snapshot for the approval record.
detection_badge <- function(sr, under, dmin, under_mt, dmin_mt) {
  if (is.na(sr) || sr <= 0) {
    return(list(
      html = "<span class='det-badge napp' title='Sharpe &le;0: the one-sided detection test has no positive effect to test for. Not applicable -- not the same as unproven.'>&ndash; n/a</span>",
      rank = 4L
    ))
  }
  if (is.na(under)) {
    return(list(
      html = "<span class='det-badge notcomputed' title='Positive Sharpe with NO detection verdict. This should never render live -- QA gate S20 (#726 item 3) is meant to block the build until it is fixed.'>&#10007; NOT COMPUTED</span>",
      rank = 5L
    ))
  }
  if (isTRUE(under)) {
    mt_note <- if (!is.na(under_mt)) " Multiple-testing correction is academic here -- the single test already fails." else ""
    return(list(
      html = sprintf(
        "<span class='det-badge underpowered' title='Needs %.0f years of data at alpha=0.05, one-sided.%s See #726.'>&#9888; Underpowered (needs %.0fyr)</span>",
        dmin, mt_note, dmin
      ),
      rank = 3L
    ))
  }
  if (is.na(under_mt)) {
    return(list(
      html = sprintf(
        "<span class='det-badge detectable' title='Needs %.1f years at alpha=0.05, one-sided. Multiple-testing-corrected verdict not computed (K_eff unavailable for this strategy). See #726.'>&#10003; Detectable (needs %.1fyr)</span>",
        dmin, dmin
      ),
      rank = 1L
    ))
  }
  if (isTRUE(under_mt)) {
    return(list(
      html = sprintf(
        "<span class='det-badge mtfail' title='Passes the single test (needs %.1fyr) but FAILS once corrected for testing multiple strategies (needs %.0fyr corrected). See #726.'>&#9686; Detectable, fails MT-correction (needs %.0fyr corrected)</span>",
        dmin, dmin_mt, dmin_mt
      ),
      rank = 2L
    ))
  }
  list(
    html = sprintf(
      "<span class='det-badge detectable' title='Needs %.1f years at alpha=0.05, one-sided (%.1fyr corrected for multiple testing). See #726.'>&#10003; Detectable (needs %.1fyr; %.1fyr corrected)</span>",
      dmin, dmin_mt, dmin, dmin_mt
    ),
    rank = 1L
  )
}

# ── NEW: P(SR>0) badge -- mirrors detection_badge()'s exact 5-state shape
# and severity-rank convention (1 = best ... 5 = defect), applied to the
# DIFFERENT statistical object historicaldata::hd_prob_sharpe_positive()
# computes (P(true Sharpe > 0), Bailey & Lopez de Prado's Probabilistic
# Sharpe Ratio) instead of hd_detection_power()'s T_min/power. See
# .claude/rules/detection-power-required.md and the roxygen atop
# packages/historicaldata/R/hd_prob_sharpe_positive.R for why the two are
# complementary, not redundant, questions -- and R/plan_qa_gates.R's S40
# gate for the coverage/monotonicity properties this badge assumes hold.
psr_badge <- function(sr, psr, psr_insuff, psr_mt, psr_mt_insuff, threshold) {
  if (is.na(sr) || sr <= 0) {
    return(list(
      html = "<span class='psr-badge napp' title='Sharpe &le;0: P(true Sharpe &gt; 0) is not a meaningful question when the point estimate itself is non-positive. Not applicable -- not the same as unproven.'>&ndash; n/a</span>",
      rank = 4L
    ))
  }
  if (is.na(psr)) {
    return(list(
      html = "<span class='psr-badge notcomputed' title='Positive Sharpe with NO P(SR&gt;0) verdict. This should never render live -- QA gate S40 (#851) is meant to block the build until it is fixed.'>&#10007; NOT COMPUTED</span>",
      rank = 5L
    ))
  }
  if (isTRUE(psr_insuff)) {
    mt_note <- if (!is.na(psr_mt_insuff)) " Multiple-testing correction is academic here -- the single test already fails." else ""
    return(list(
      html = sprintf(
        "<span class='psr-badge insufficient' title='P(true Sharpe &gt; 0) = %s, below the %s bar (the SAME 95%% accept/reject threshold the StratProof article that motivated #851 uses).%s See #851.'>&#9888; Insufficient (%s)</span>",
        pct(psr), pct(threshold), mt_note, pct(psr)
      ),
      rank = 3L
    ))
  }
  if (is.na(psr_mt)) {
    return(list(
      html = sprintf(
        "<span class='psr-badge sufficient' title='P(true Sharpe &gt; 0) = %s (single test), at or above the %s bar. Multiple-testing-corrected verdict not computed (K_eff unavailable for this strategy). See #851.'>&#10003; Sufficient (%s)</span>",
        pct(psr), pct(threshold), pct(psr)
      ),
      rank = 1L
    ))
  }
  if (isTRUE(psr_mt_insuff)) {
    return(list(
      html = sprintf(
        "<span class='psr-badge mtfail' title='Passes the single-test bar (%s) but FAILS once corrected for testing multiple strategies (%s, below the %s bar). See #851.'>&#9686; Sufficient, fails MT-correction (%s)</span>",
        pct(psr), pct(psr_mt), pct(threshold), pct(psr_mt)
      ),
      rank = 2L
    ))
  }
  list(
    html = sprintf(
      "<span class='psr-badge sufficient' title='P(true Sharpe &gt; 0): %s single-test, %s corrected for multiple testing -- both at or above the %s bar. See #851.'>&#10003; Sufficient (%s; %s MT)</span>",
      pct(psr), pct(psr_mt), pct(threshold), pct(psr), pct(psr_mt)
    ),
    rank = 1L
  )
}

det <- mapply(detection_badge, full$sharpe, full$detection_underpowered,
              full$detection_min_n_years, full$detection_underpowered_mt,
              full$detection_min_n_years_mt, SIMPLIFY = FALSE)
psr <- mapply(psr_badge, full$sharpe, full$prob_sharpe_positive,
              full$prob_sharpe_positive_insufficient, full$prob_sharpe_positive_mt,
              full$prob_sharpe_positive_insufficient_mt,
              MoreArgs = list(threshold = HD_PROB_SHARPE_POSITIVE_THRESHOLD),
              SIMPLIFY = FALSE)

metrics_table <- full |>
  mutate(
    .sharpe_rank = ifelse(is.na(sharpe), -Inf, sharpe),
    Sharpe       = ifelse(is.na(sharpe), "—", as.character(round(sharpe, 2))),
    Detection    = vapply(det, function(x) x$html, character(1)),
    .det_rank    = vapply(det, function(x) x$rank, integer(1)),
    # ---- THE PROPOSED NEW COLUMN, inserted immediately right of Detection ----
    `P(SR>0)`    = vapply(psr, function(x) x$html, character(1)),
    .psr_rank    = vapply(psr, function(x) x$rank, integer(1))
  ) |>
  select(Strategy = strategy, Sharpe, .sharpe_rank, Detection, .det_rank,
         `P(SR>0)`, .psr_rank)

sharpe_col   <- which(names(metrics_table) == "Sharpe") - 1L
sharpe_rank_col <- which(names(metrics_table) == ".sharpe_rank") - 1L
det_col      <- which(names(metrics_table) == "Detection") - 1L
det_rank_col <- which(names(metrics_table) == ".det_rank") - 1L
psr_col      <- which(names(metrics_table) == "P(SR>0)") - 1L
psr_rank_col <- which(names(metrics_table) == ".psr_rank") - 1L

# ── Dynamically computed disagreement diagnostic (never hardcoded) --
# compares the OVERALL verdict from each diagnostic (both single-test AND
# MT-corrected must clear their own bar) for every positive-Sharpe strategy,
# and separately reports rows where the two diagnostics' SINGLE-TEST
# verdicts alone disagree (the more visually striking case for a reader
# scanning two adjacent badges).
pos <- full |> dplyr::filter(!is.na(sharpe), sharpe > 0)
det_overall <- !is.na(pos$detection_underpowered) & !pos$detection_underpowered &
  (is.na(pos$detection_underpowered_mt) | !pos$detection_underpowered_mt)
psr_overall <- !is.na(pos$prob_sharpe_positive_insufficient) & !pos$prob_sharpe_positive_insufficient &
  (is.na(pos$prob_sharpe_positive_insufficient_mt) | !pos$prob_sharpe_positive_insufficient_mt)
overall_disagree <- pos$strategy[det_overall != psr_overall]

det_single_ok <- !is.na(pos$detection_underpowered) & !pos$detection_underpowered
psr_single_ok <- !is.na(pos$prob_sharpe_positive_insufficient) & !pos$prob_sharpe_positive_insufficient
single_disagree <- pos$strategy[det_single_ok != psr_single_ok]

# K_eff/K_raw for the prose below -- the SAME values production's
# `detection_summary_text` reports (docs/leaderboard.qmd), NOT nrow(pos)
# (that is merely the count of positive-Sharpe rows on this table, a
# different number from how many strategies the multiple-testing
# correction is actually run across).
keff_nonNA <- full$k_eff_leaderboard[!is.na(full$k_eff_leaderboard)]
kraw_nonNA <- full$k_raw_leaderboard[!is.na(full$k_raw_leaderboard)]
keff <- if (length(keff_nonNA) > 0) round(keff_nonNA[1], 1) else NA_real_
kraw <- if (length(kraw_nonNA) > 0) kraw_nonNA[1] else NA_integer_
```

## Step 1 — Relationship to the existing Detection column

**P(true Sharpe > 0)** (`historicaldata::hd_prob_sharpe_positive()`, Bailey &
Lopez de Prado's Probabilistic Sharpe Ratio) and **Detection**
(`historicaldata::hd_detection_power()`) answer two different questions about
the same row, and neither implies the other:

- **Detection** (prospective, frequentist): *is this strategy's sample long
  enough to reliably detect its effect as real, if it is real?* A strategy
  can be flagged `Underpowered` here even when its observed Sharpe is
  strongly positive, simply because the sample is short relative to Cohen's
  standard 80%-power convention.
- **P(SR>0)** (retrospective): *given the sample we actually have — its
  Sharpe, length, skewness and kurtosis — how confident are we that the
  TRUE population Sharpe is positive?* This is the accept/reject question
  the StratProof article that motivated
  [#851](https://github.com/JohnGavin/historical/issues/851) asks directly
  ("need 95%+ to call this signal, not noise").

They are complementary, not redundant: a strategy can be power-sufficient
with a weak point estimate (low P(SR>0)), or power-insufficient with a
strong point estimate that nonetheless clears the 95% confidence bar. In the
live data read for this prototype (below), that second case is exactly what
happens for **`r if (length(single_disagree) > 0) single_disagree[1] else "no strategy currently"`**:
Detection reports the single-test verdict as `Underpowered`, while P(SR>0)'s
own single-test verdict is `Sufficient` at
`r if (length(single_disagree) > 0) { s <- single_disagree[1]; pct(pos$prob_sharpe_positive[pos$strategy == s]) } else "n/a"`.
Once both are corrected for testing `r kraw` strategies (K_eff = `r keff`),
the two diagnostics converge back to agreement on that row — see the
disagreement callout below the table for the full, dynamically computed
list.

**Diagram:** no new Mermaid node is proposed. The one relationship diagram
this leaderboard page's own domain has —
[`falsification.qmd`'s "Strategy & Dashboard Relationship Map"](../../docs/falsification.qmd) —
is scoped to the MVO/portfolio-construction subsystem (covariance estimator,
returns shrinkage, weight engines) and carries no node for either
`hd_detection_power()` or `hd_prob_sharpe_positive()` today; both are
per-row statistical diagnostics computed inside the `leaderboard` target
itself, not separate pipeline stages with their own data-flow edges to
draw. Extending that diagram to cover every diagnostic column on the
Rankings table would broaden its scope well past what this one-column
addition needs. No diagram change is proposed.

## Step 2 — Home: exact placement in `docs/leaderboard.qmd`

New column **`P(SR>0)`** inserted immediately to the right of the existing
**Detection** column (both immediately right of **Sharpe**, per
`detection-power-required` requirement 2 — "beside its Sharpe, not in a
distant column"), in the `metrics_table` `mutate()`/`select()` block of the
`#| label: detection-rigour-summary` → main table chunk (the unlabelled
chunk starting `lb <- safe_tar_read("leaderboard")`, around
`docs/leaderboard.qmd:411-556` on `main` as of this dispatch). Concretely:

- A `psr_badge()` function defined alongside `detection_badge()`/
  `rigour_badge()` (same file, same section, ~line 247 in `main`).
- `psr <- mapply(psr_badge, full$sharpe, full$prob_sharpe_positive, ...)`
  added alongside the existing `det <- mapply(detection_badge, ...)` call
  (~line 383).
- ``` `P(SR>0)` = vapply(psr, function(x) x$html, character(1)) ``` and
  `.psr_rank = vapply(psr, function(x) x$rank, integer(1))` added to the
  `mutate()` block, immediately after `Detection`/`.det_rank`.
- `select()` updated to `select(Strategy = strategy, Credible,
  .credible_rank, Sharpe, .sharpe_rank, Detection, .det_rank,` **`` `P(SR>0)`,
  .psr_rank, ``** `CAGR, ...)`.
- `columnDefs` extended with one more `visible = FALSE` rank column and one
  more `orderData` mapping, following the exact pattern already used for
  `.det_rank`/`.rig_rank`/`.sharpe_rank` (~`docs/leaderboard.qmd:849-857`).
- The "Statistical Detection Power and Rigour" callout
  (`detection_summary_text`, ~line 118) and its `detection_note` caption
  text (~line 740) each get one added sentence reporting P(SR>0) coverage,
  mirroring the existing K_eff/deflated-Sharpe sentence shape.
- The "Statistical Detection Power and Rigour" tab's plain-column table on
  the Biases and Caveats page (`caveats-detection-rigour-table`, ~line 2184)
  gets two added columns (`P(SR>0)`, `P(SR>0) MT-corrected`), matching its
  existing `Years Needed` / `Years Needed (MT-corrected)` pair.

## Step 3 — Comparable outputs this must match

- **Badge markup and CSS class family**: reuses the exact `.det-badge`
  shape (`font-weight: 600; padding: 1px 6px; border-radius: 3px;
  white-space: nowrap;`) under a new `.psr-badge` class, with the SAME
  4-state colour mapping (`sufficient` → the `detectable` green, `mtfail`
  → the same purple, `insufficient` → the same amber as `underpowered`,
  `napp` → the same muted italic grey, `notcomputed` → the same red
  background/white text), including the light-mode / `[data-bs-theme="dark"]`
  / `@media (prefers-color-scheme: dark)` triad already present for every
  other badge family in `docs/leaderboard.qmd` (see the `<style>` block
  below this prototype's own table for the exact rules, copied 1:1 in
  spirit from `docs/leaderboard.qmd:877-920`).
- **Hover-popover convention**: every state's `title=` attribute states the
  numeric verdict, the threshold it's compared against, and links back to
  [#851](https://github.com/JohnGavin/historical/issues/851) — matching
  every existing badge's "hover for detail" pattern (`accessibility` rule,
  `dashboard-output-first` step 3).
- **Number formatting**: percentages to one decimal place
  (`sprintf("%.1f%%", x*100)`), matching `CAGR`/`Net CAGR`/`Vol`/`CVaR 95%`'s
  existing one-decimal convention on this same table.
- **Threshold sourced, not hand-typed**: `HD_PROB_SHARPE_POSITIVE_THRESHOLD`
  is read from `R/plan_leaderboard.R`'s own constant (0.95, documented there
  as the SAME 95% bar the StratProof article uses) — never a literal in the
  display layer.
- **Caption / disclosure-note conventions**: a dynamically-computed
  coverage sentence (`n` positive-Sharpe strategies, `n` below the 95% bar,
  singular/plural handled the way every other `*_note`/`*_summary_text`
  block on this page already does) — see the "Statistical Detection Power
  and Rigour" callout text mirrored in Step 2 above.
- **Sort-rank convention**: a hidden `.psr_rank` integer column
  (1 = sufficient … 5 = defect), `orderData`-mapped exactly like
  `.det_rank`/`.rig_rank`/`.sharpe_rank`, so DT's default alphabetical
  string sort never scrambles the verdict order (same rationale as the
  existing hidden-rank columns, `docs/leaderboard.qmd:807-816`).

## Step 4 — Prototype (below)

Full Period, live `leaderboard` target read from the main checkout's build
(`docs/_targets`) at render time — the SAME data the production Rankings
table shows today, not a stub or a hand-typed sample.

```{r}
#| label: prototype-table
DT::datatable(
  metrics_table,
  escape = FALSE,
  rownames = FALSE,
  options = list(
    pageLength = 20,
    dom = "t",
    columnDefs = list(
      list(visible = FALSE, targets = c(sharpe_rank_col, det_rank_col, psr_rank_col)),
      list(orderData = sharpe_rank_col, targets = sharpe_col),
      list(orderData = det_rank_col, targets = det_col),
      list(orderData = psr_rank_col, targets = psr_col)
    ),
    order = list(list(sharpe_col, "desc"))
  )
)
```

```{=html}
<style>
.det-badge, .psr-badge { font-weight: 600; padding: 1px 6px; border-radius: 3px; white-space: nowrap; }
.det-badge.detectable   { color: #1a7f37; }
.det-badge.mtfail       { color: #6f42c1; }
.det-badge.underpowered { color: #9a5b00; }
.det-badge.napp         { color: #6c757d; font-style: italic; font-weight: 400; }
.det-badge.notcomputed  { color: #fff; background: #c0392b; }
/* .psr-badge deliberately mirrors .det-badge's colour mapping 1:1 (Step 3) */
.psr-badge.sufficient   { color: #1a7f37; }
.psr-badge.mtfail       { color: #6f42c1; }
.psr-badge.insufficient { color: #9a5b00; }
.psr-badge.napp         { color: #6c757d; font-style: italic; font-weight: 400; }
.psr-badge.notcomputed  { color: #fff; background: #c0392b; }

[data-bs-theme="dark"] .det-badge.detectable   { color: #4ade80; }
[data-bs-theme="dark"] .det-badge.mtfail       { color: #c084fc; }
[data-bs-theme="dark"] .det-badge.underpowered { color: #fbbf24; }
[data-bs-theme="dark"] .det-badge.napp         { color: #9ca3af; }
[data-bs-theme="dark"] .psr-badge.sufficient   { color: #4ade80; }
[data-bs-theme="dark"] .psr-badge.mtfail       { color: #c084fc; }
[data-bs-theme="dark"] .psr-badge.insufficient { color: #fbbf24; }
[data-bs-theme="dark"] .psr-badge.napp         { color: #9ca3af; }
@media (prefers-color-scheme: dark) {
  .det-badge.detectable   { color: #4ade80; }
  .det-badge.mtfail       { color: #c084fc; }
  .det-badge.underpowered { color: #fbbf24; }
  .psr-badge.sufficient   { color: #4ade80; }
  .psr-badge.mtfail       { color: #c084fc; }
  .psr-badge.insufficient { color: #fbbf24; }
}
</style>
```

::: {.callout-note title="Disagreement between Detection and P(SR>0), this build"}
Of `r nrow(pos)` strategies with a positive Full-Period Sharpe,
`r length(single_disagree)` disagree between the two diagnostics' own
**single-test** verdicts alone
(`r if (length(single_disagree) > 0) paste(single_disagree, collapse = ", ") else "none"`),
and `r length(overall_disagree)` disagree once each diagnostic's **overall**
verdict (single-test AND MT-corrected both required to pass) is compared
(`r if (length(overall_disagree) > 0) paste(overall_disagree, collapse = ", ") else "none"`).
Disagreement is expected and informative, not a bug: it is exactly the
"complementary, not redundant" relationship described in Step 1 — a reader
scanning both badges on the same row sees where a strategy is
under-sampled (Detection) versus where its point estimate itself is weak
(P(SR>0)), which a single verdict column could not show.
:::

## Step 5 — Consolidation review

**Proposed simplification:** none of the four existing columns become
redundant — Detection and P(SR>0) test different objects (prospective power
vs. retrospective confidence in sign), and the disagreement callout above
demonstrates real cases where they diverge, so collapsing them into one
badge would destroy information a reader currently cannot get any other
way. However, this addition **does** motivate one flagged future
consolidation, not made here: the "Statistical Detection Power and Rigour"
tab's plain-column table on the Biases and Caveats page
(`caveats-detection-rigour-table`) restates every Detection/Rigour number
already visible as a hover-badge on the Rankings table. Adding two more
plain columns there (P(SR>0), P(SR>0) MT-corrected) grows that
already-restating table further — a stronger Step 5 answer, once this
column ships and both hover badges exist, is to fold that plain-column
table's content into a single per-strategy detail popover/drilldown
reachable from the Rankings table itself, rather than maintaining two
parallel renderings (badge tooltip + plain table) of the same six numbers.
Filed as a note for a follow-up issue rather than done in this dispatch,
per "no change" not being the right answer but also not being this
dispatch's scope to execute.

## Build info

`r paste0("Rendered ", format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z"), " | leaderboard rows (Full Period): ", nrow(full), " | commit: ", tryCatch(system("git rev-parse --short HEAD", intern = TRUE), error = function(e) "unknown"))`
