---
title: "PROTOTYPE — Leverage Allocator (Phase 0, #626)"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    embed-resources: false
    include-in-header:
      - text: |
          <link rel="stylesheet" href="../vignette-shared.css">
          <script src="../vignette-shared.js"></script>
execute:
  echo: false
  warning: false
  message: false
---

::: {.callout-important title="Phase 0 prototype -- NOT for merge"}
This page is a throwaway design artifact for [dashboard-output-first](https://github.com/JohnGavin/historical/blob/main/.claude/rules/dashboard-output-first.md)
Phase 0, steps 1 and 4, covering [issue #626](https://github.com/JohnGavin/historical/issues/626)
(the leverage allocator). It reads the live targets store **read-only** --
no new pipeline target is created, `tar_make()` is never called, and it is
not wired into `_quarto.yml`. It replicates the allocator arithmetic
already shipped as `compute_vol_per_unit_gross()` /
`compute_budget_neutral_sigma()` in
[`R/plan_leverage.R`](https://github.com/JohnGavin/historical/blob/de4a4e99fb1335c19f79032ee86b40e8c29d85ac/R/plan_leverage.R)
(landed today via [#826](https://github.com/JohnGavin/historical/pull/826)) against the
real `leaderboard` target, so every number below is current, not a hand-typed
stub. Production implementation waits for explicit approval (Class C gate
per `human-in-the-loop-decision-points`) and, once approved, belongs in
`leaderboard.qmd` (exposure columns) and `falsification.qmd` (leverage
diagnostics) -- see "Step 2 -- home" below, which is the placement #626's own
issue body already specifies.
:::

```{r}
#| label: setup
#| include: false
library(targets)
library(dplyr)
library(tidyr)
library(DT)

# This is a worktree checkout (agent-identity-and-task-scopes rule) -- the
# live targets store is gitignored and exists only in the main checkout, not
# here. Try the normal in-project relative locations first (so this chunk
# still works if the file is ever moved into the main checkout's docs/), then
# fall back to the absolute main-checkout path this dispatch was told to read
# from (READ ONLY -- see the dispatch prompt; never tar_make() against it).
store_path <- if (dir.exists("../_targets")) "../_targets" else
  if (dir.exists(file.path(here::here(), "docs/_targets"))) file.path(here::here(), "docs/_targets") else
  "/Users/johngavin/docs_gh/proj/finance/data/historical/docs/_targets"
tar_config_set(store = store_path)

utils_path <- if (file.exists("../vignette_utils.R")) "../vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)

# compute_vol_per_unit_gross() / compute_budget_neutral_sigma() /
# compute_regime_stress_ratio() -- the real, already-merged functions this
# prototype's numbers come from. Sourcing the plain functions (not calling
# plan_leverage(), which only returns a tar_target() list) is read-only and
# creates no pipeline target.
leverage_fns_path <- if (file.exists("../../R/plan_leverage.R")) "../../R/plan_leverage.R" else
  if (file.exists("../R/plan_leverage.R")) "../R/plan_leverage.R" else
  file.path(here::here(), "R/plan_leverage.R")
source(leverage_fns_path)

git_sha <- tryCatch(system("git -C .. rev-parse HEAD 2>/dev/null", intern = TRUE),
                     error = function(e) "main")
git_sha <- if (length(git_sha) == 0 || !nzchar(git_sha)) "main" else git_sha
gh_base <- paste0("https://github.com/JohnGavin/historical/blob/", git_sha)

lb <- safe_tar_read("leaderboard")
```

# Step 1 -- Relationship Diagram

## Row

Extends the existing causal-DAG surface in `falsification.qmd` (the
"Strategy & Dashboard Relationship Map" tab, itself a #481/#514 seed diagram,
rendered via `docs/causal-diagrams.js`'s `dag-portcons-mount`) rather than
starting a new standalone diagram. This is the missing branch: how a
per-strategy volatility target and a gross-exposure convention become an
*allocated* leverage figure, and where that figure meets the still-open
backstop decision.

```{mermaid}
%%| label: fig-leverage-allocator-dag
%%| fig-cap: "Budget-neutral sigma_target and per-strategy gross_convention/vol feed compute_vol_per_unit_gross() and compute_budget_neutral_sigma() (both live in R/plan_leverage.R, #635/#719); the resulting implied G_i meets an UNRESOLVED gross backstop before reaching either dashboard."
graph LR
  LB["leaderboard target<br/>vol, strategy, period"]
  SGC["strategy_gross_convention<br/>gross_convention, is_cap (#626)"]
  VPUG["compute_vol_per_unit_gross()<br/>vol / gross_convention"]
  SIGMA["compute_budget_neutral_sigma()<br/>leverage_sigma_target (live target, #635)"]
  GI["Implied G_i = sigma_target / vol_per_unit_gross<br/>per-strategy allocator output"]
  STRESS["compute_regime_stress_ratio()<br/>leverage_regime_stress_ratio (#635 DoD 7)"]
  BACKSTOP{{"Gross backstop cap<br/>UNRESOLVED -- #626"}}
  GCAPPED["G_i capped at backstop"]
  LBDASH["leaderboard.qmd<br/>exposure columns"]
  FALSDASH["falsification.qmd<br/>leverage diagnostics"]

  LB --> VPUG
  SGC --> VPUG
  VPUG --> SIGMA
  VPUG --> GI
  SIGMA --> GI
  VPUG --> STRESS
  GI --> BACKSTOP
  STRESS -. informs .-> BACKSTOP
  BACKSTOP --> GCAPPED
  GCAPPED -.->|"this prototype"| LBDASH
  GI -.->|"this prototype"| FALSDASH
  STRESS -.->|"this prototype"| FALSDASH

  click LB "https://github.com/JohnGavin/historical/blob/de4a4e99fb1335c19f79032ee86b40e8c29d85ac/R/plan_leaderboard.R#L120" _blank
  click SGC "https://github.com/JohnGavin/historical/blob/de4a4e99fb1335c19f79032ee86b40e8c29d85ac/R/plan_exposure.R#L34" _blank
  click VPUG "https://github.com/JohnGavin/historical/blob/de4a4e99fb1335c19f79032ee86b40e8c29d85ac/R/plan_leverage.R#L65" _blank
  click SIGMA "https://github.com/JohnGavin/historical/blob/de4a4e99fb1335c19f79032ee86b40e8c29d85ac/R/plan_leverage.R#L105" _blank
  click STRESS "https://github.com/JohnGavin/historical/blob/de4a4e99fb1335c19f79032ee86b40e8c29d85ac/R/plan_leverage.R#L144" _blank
  click BACKSTOP "https://github.com/JohnGavin/historical/issues/626" _blank
  click LBDASH "https://github.com/JohnGavin/historical/blob/de4a4e99fb1335c19f79032ee86b40e8c29d85ac/docs/leaderboard.qmd#L373" _blank
  click FALSDASH "https://github.com/JohnGavin/historical/blob/de4a4e99fb1335c19f79032ee86b40e8c29d85ac/docs/falsification.qmd#L901" _blank
```

Every node link is pinned to commit `de4a4e9` (the #826 merge that landed
`R/plan_leverage.R`) with a `#L<n>` line anchor per
[`mermaid-click-anchors`](https://github.com/JohnGavin/llm/blob/main/.claude/rules/mermaid-click-anchors.md).
**Known limitation of this prototype (same as the #735 prototype before
it):** Quarto's native `{mermaid}` fenced block runs mermaid.js in an iframe
whose click-through and hover-tooltip behaviour is less reliable than the
project's own `causal-diagrams.js` mount system --
`docs/MERMAID_LESSONS.md` documents exactly why `falsification.qmd` moved
off native fenced blocks. Production wiring should extend the
`dag-portcons-mount` diagram already in `causal-diagrams.js` (adding these 6
nodes and their `nodeTooltips`/`edgeMetadata` entries) rather than shipping
a standalone native block -- the diagram above is the design for that
extension, not the artifact itself.

# Step 4 -- Table Prototype

## Row {height="25%"}

```{r}
#| label: allocator-summary
#| results: asis
if (!is.null(lb)) {
  vpug <- compute_vol_per_unit_gross(lb)
  sigma_tbl <- compute_budget_neutral_sigma(vpug)
  sigma_headline <- sigma_tbl$sigma_target_budget_neutral[sigma_tbl$is_headline]

  full_vpug <- vpug |>
    filter(period == "Full Period") |>
    mutate(G_implied = sigma_headline / vol_per_unit_gross) |>
    arrange(desc(G_implied))

  # Candidate backstop levels for comparison -- p90 of the observed G_implied
  # distribution (the #635 methodology) vs the highest level that binds no
  # strategy today (the "changes nothing on day one" property #635 wanted).
  p90 <- as.numeric(quantile(full_vpug$G_implied, 0.9))
  max_g <- max(full_vpug$G_implied)
  n_over_p90 <- sum(full_vpug$G_implied > p90)
  n_cap_rows <- sum(full_vpug$is_cap)
  n_cap_over2 <- sum(full_vpug$is_cap & full_vpug$G_implied > 2)

  cat(paste0(
    '<div class="callout-warning" style="border-left:4px solid #b35900;padding:0.6em 1em;margin-bottom:0.5em;">',
    "<strong>Live allocator arithmetic, this build (commit <code>", substr(git_sha, 1, 7), "</code>):</strong> ",
    "budget-neutral sigma_target = <strong>", sprintf("%.4f", sigma_headline), "</strong> (", nrow(full_vpug),
    " strategies with a measurable gross_convention). Implied G_i ranges from ",
    sprintf("%.2f", min(full_vpug$G_implied)), "x to <strong>", sprintf("%.2f", max_g),
    "x</strong>. A p90-based backstop candidate sits at <strong>", sprintf("%.2f", p90),
    "x</strong>, which binds ", n_over_p90, " strateg", if (n_over_p90 == 1) "y" else "ies",
    " today (by construction, ~10%). This is materially LOWER than the 3.5x candidate #626's ",
    "2026-08-05 comment derived by hand: that derivation predates the #717/#722 CMR vol-annualisation ",
    "fixes, which moved CMR's implied G_i from a stale 4.56x down to a live ",
    sprintf("%.2f", full_vpug$G_implied[full_vpug$strategy == "CMR"]),
    "x -- the exact kind of drift `R/plan_leverage.R`'s header comment (four prior hand-derivations, three ",
    "invalidated by data bugs) exists to stop. Today the only strategy near any plausible backstop is ",
    "<strong>Managed Futures</strong> (", sprintf("%.2f", max_g),
    "x), and it is an <code>is_cap</code> row -- exactly the \"measurement artefact\" #626's own comment ",
    "thread already flagged as inflating its apparent G_i (a 3.0x ceiling it rarely reaches, so its true ",
    "vol-per-unit-realised-gross is understated). See the callout below the table.",
    '</div>'
  ))
} else {
  cat("*`leaderboard` target not available from the read-only store.*")
}
```

## Row

```{r}
#| label: allocator-table
#| results: asis
if (!is.null(lb) && exists("full_vpug") && nrow(full_vpug) > 0) {

  proto <- full_vpug |>
    mutate(
      `Vol (ann.)`          = paste0(round(vol * 100, 1), "%"),
      `Gross now`           = paste0(format(gross_convention, nsmall = 1), "x", ifelse(is_cap, "*", "")),
      `Vol / unit gross`    = round(vol_per_unit_gross, 4),
      `Implied G_i`         = round(G_implied, 2),
      `G_i @ p90 backstop`  = round(pmin(G_implied, p90), 2),
      `Backstop binds?`     = ifelse(G_implied > p90,
                                      paste0("YES (-", round((G_implied - pmin(G_implied, p90)) / G_implied * 100, 0), "%)"),
                                      "no"),
      .g_rank = -G_implied
    ) |>
    select(Strategy = strategy, `Vol (ann.)`, `Gross now`, `Vol / unit gross`,
           `Implied G_i`, .g_rank, `G_i @ p90 backstop`, `Backstop binds?`)

  g_col    <- which(names(proto) == "Implied G_i") - 1L
  rank_col <- which(names(proto) == ".g_rank") - 1L

  cap <- paste0(
    "PROTOTYPE allocator table -- Full Period, ", nrow(proto), " strategies with a measurable ",
    "gross_convention (Value (HML) and PSO Optimal excluded by construction, matching every #635 ",
    "hand-derivation). Implied G_i = sigma_target (", sprintf("%.4f", sigma_headline),
    ") / vol_per_unit_gross, i.e. the vol-normalised leverage this strategy would receive under the ",
    "layered allocator design #626's 2026-08-03 comment adopted (vol-normalised allocator sets relative ",
    "sizing; a gross-exposure cap binds the total). \"Gross now\" is the SAME column already live in ",
    "`leaderboard.qmd` (`strategy_gross_convention`, trailing `*` = construction ceiling, not always ",
    "realised). A trailing `*` on Implied G_i is not shown because this column is itself a proposal, not ",
    "yet a measured or capped figure. Source: R/plan_leverage.R (compute_vol_per_unit_gross, ",
    "compute_budget_neutral_sigma), R/plan_exposure.R (strategy_gross_convention)."
  )

  DT::datatable(
    proto,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold;",
      class = "dt-caption", cap
    ),
    escape = FALSE,
    rownames = FALSE,
    filter = "top",
    options = list(
      pageLength = 20,
      scrollX = TRUE,
      autoWidth = FALSE,
      dom = "frtip",
      search = list(regex = TRUE, caseInsensitive = TRUE),
      columnDefs = list(
        list(visible = FALSE, targets = rank_col),
        list(orderData = rank_col, targets = g_col)
      ),
      initComplete = DT::JS(
        "function(settings, json) {",
        "  var isDark = document.documentElement.getAttribute('data-bs-theme') === 'dark'",
        "    || document.body.classList.contains('dark-mode');",
        "  if (isDark) {",
        "    $(this.api().table().container()).css({'color': '#ddd', 'background-color': '#1a1a1a'});",
        "    $(this.api().table().header()).css({'color': '#ddd', 'background-color': '#222'});",
        "    $('input', this.api().table().container()).css({'color': '#ddd', 'background-color': '#333', 'border-color': '#555'});",
        "  }",
        "}"
      )
    )
  )
}
```

## Row

```{r}
#| label: regime-stress-table
#| results: asis
if (!is.null(lb)) {
  stress <- tryCatch(compute_regime_stress_ratio(vpug), error = function(e) NULL)
  if (!is.null(stress) && nrow(stress) > 0) {
    n_covered <- nrow(stress)
    n_total   <- length(unique(full_vpug$strategy))
    hd_dt(
      as.data.frame(stress) |> mutate(across(c(Training, Testing, stress_ratio), \(x) round(x, 4))),
      paste0(
        "Training-vs-Testing stress ratio of vol_per_unit_gross (compute_regime_stress_ratio(), ",
        "R/plan_leverage.R -- the live replacement for #635 DoD item 7's hand-run regime check). Ratio > 1 ",
        "means a strategy's risk-per-unit-of-gross rose in the 2020-22 Testing partition relative to the calm ",
        "Training partition -- informs how conservative the still-open backstop decision should be, per this ",
        "prototype's Step 1 diagram. Coverage: ", n_covered, " of ", n_total,
        " strategies with a measurable gross_convention have BOTH a Training and a Testing row -- the ",
        "remaining strategies (CMR, Mom Pre-Peak, Mom Post-Peak, Mom 12-2, Managed Futures) are the #648 ",
        "coverage gap already documented in R/plan_leverage.R's own header comment, and are silently absent ",
        "here, not a bug."
      )
    )
  } else {
    cat("*Regime stress ratio not available.*")
  }
}
```

```{=html}
<style>
.dt-caption { margin-bottom: 4px; }
</style>
```

# Design Notes

## Row

### Step 2 -- Locate the home

**No new dashboard.** #626's own issue body already states the placement:
"exposure columns extend `leaderboard.qmd`; leverage diagnostics extend
`falsification.qmd`." This prototype's two artifacts map directly onto that
split:

- The **allocator table** (Implied G_i, G_i @ backstop, Backstop binds?)
  belongs in `docs/leaderboard.qmd`, immediately beside the existing
  [`Gross` column](https://github.com/JohnGavin/historical/blob/de4a4e99fb1335c19f79032ee86b40e8c29d85ac/docs/leaderboard.qmd#L373)
  (same table, same row per strategy) -- not a new tab, an extension of the
  existing exposure disclosure.
- The **regime stress-ratio table** and the **relationship diagram** belong
  in `docs/falsification.qmd`'s existing "Strategy & Dashboard Relationship
  Map" tab
  ([`#L901`](https://github.com/JohnGavin/historical/blob/de4a4e99fb1335c19f79032ee86b40e8c29d85ac/docs/falsification.qmd#L901))
  and its `dag-portcons-mount` diagram, alongside the covariance-estimator
  subsystem already documented there.

### Step 3 -- Comparable-output audit

This output sits beside, and must stay consistent with:

- **`Gross` column** (`docs/leaderboard.qmd`
  [`#L373-377`](https://github.com/JohnGavin/historical/blob/de4a4e99fb1335c19f79032ee86b40e8c29d85ac/docs/leaderboard.qmd#L373)) --
  same `strategy_gross_convention` source, same trailing `*` convention for
  `is_cap` ceiling rows. `Implied G_i` is deliberately placed as a SEPARATE
  proposed column, not a replacement, because `Gross` reports what a
  strategy's own code implements today and `Implied G_i` reports what the
  vol-normalised allocator would assign -- conflating them would erase the
  gap the whole allocator design exists to close.
- **`gross_note` dynamic disclosure** (same file,
  [`#L512-523`](https://github.com/JohnGavin/historical/blob/de4a4e99fb1335c19f79032ee86b40e8c29d85ac/docs/leaderboard.qmd#L512)) --
  this prototype's summary callout follows the identical pattern: every
  number computed from the live store, never hardcoded, with the same `*`
  convention explained inline.
- **Strategy naming**
  ([`strategy-name-consistency`](https://github.com/JohnGavin/historical/blob/main/.claude/rules/strategy-name-consistency.md)):
  `strategy` values are used unchanged from `leaderboard`/`strategy_gross_convention` --
  no new naming introduced.
- **Captions**
  ([`visualization-standards`](https://github.com/JohnGavin/historical/blob/main/.claude/rules/) family):
  both table captions are assembled from live values (sigma_target, counts,
  coverage), matching the project's established caption style, with source
  links to `R/plan_leverage.R` and `R/plan_exposure.R`.
- **`falsification.qmd`'s covariance-regularisation block**
  ([`#L410-468`](https://github.com/JohnGavin/historical/blob/de4a4e99fb1335c19f79032ee86b40e8c29d85ac/docs/falsification.qmd)) --
  named in `docs/DASHBOARDS.md`'s Template Recipe as the model for a
  walk-forward-estimator-style diagnostic; the regime stress-ratio table
  above follows its "comprehensive table + explicit coverage-gap disclosure"
  shape rather than hiding the #648 gap.

### Step 5 -- Consolidation review (mandatory)

**Proposal: fold `Implied G_i` into the existing `Gross` column's cell as a
secondary figure, rather than adding it as a wholly separate column.**

`docs/leaderboard.qmd` already carries 21 columns in its main ranking table
(`Strategy` through `Breach Risk`). Adding `Implied G_i`, `G_i @ backstop`,
and `Backstop binds?` as three more columns repeats the "column 60 of 61"
sprawl pattern #726 already diagnosed and #735's prototype already proposed
fixing for `Rigour`. Once the backstop level is actually decided (this
prototype's live numbers argue for something much closer to the p90 ≈
`r if (exists("p90")) sprintf("%.1f", p90) else "?"`x than the stale 3.5x),
the concrete proposal is:

- **One new column, `Gross (target)`**, showing `Implied G_i` alongside the
  existing `Gross` (now relabelled `Gross (now)`), with the backstop-bound
  state as a hover tooltip on that cell (same badge pattern as
  `Detection`/`Rigour`) rather than three separate columns.
- The regime stress-ratio table does **not** need its own dashboard tab --
  it folds into the same `falsification.qmd` tab that already documents the
  covariance-estimator subsystem, as one more diagnostic accordion, per the
  Template Recipe.

**Net effect on `docs/DASHBOARDS.md` inventory: dashboard count unchanged
(11).** Within `leaderboard.qmd`, net column count grows by one (not three)
relative to a naive per-metric-per-column implementation. This is a smaller
consolidation than a dashboard merge, but "no change" would understate what
this prototype's own comparable-output audit found: three raw columns
compress to one badge column, the same simplification #735 already
identified for `Rigour`/`Detection` and shipped in #741.

### Data sources

- `tar_read("leaderboard")` -- [`R/plan_leaderboard.R#L120`](https://github.com/JohnGavin/historical/blob/de4a4e99fb1335c19f79032ee86b40e8c29d85ac/R/plan_leaderboard.R#L120)
- `tar_read("strategy_gross_convention")` -- [`R/plan_exposure.R#L34`](https://github.com/JohnGavin/historical/blob/de4a4e99fb1335c19f79032ee86b40e8c29d85ac/R/plan_exposure.R#L34)
- `compute_vol_per_unit_gross()` / `compute_budget_neutral_sigma()` / `compute_regime_stress_ratio()` -- [`R/plan_leverage.R`](https://github.com/JohnGavin/historical/blob/de4a4e99fb1335c19f79032ee86b40e8c29d85ac/R/plan_leverage.R)
- [#626](https://github.com/JohnGavin/historical/issues/626) (this issue), [#635](https://github.com/JohnGavin/historical/issues/635) (sigma_target measurement), [#719](https://github.com/JohnGavin/historical/issues/719) (periodicity + plausibility gates, prerequisite), [#648](https://github.com/JohnGavin/historical/issues/648) (regime-coverage gap), [`dashboard-output-first`](https://github.com/JohnGavin/historical/blob/main/.claude/rules/dashboard-output-first.md)
