---
title: "Beat the Market by Avoiding Its Worst Days"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    embed-resources: false
    include-in-header:
      - text: |
          <link rel="stylesheet" href="vignette-shared.css">
          <script src="vignette-shared.js"></script>
          <style>
          body { font-size: 1.3em; }
          .navbar .navbar-title, h1.title { font-size: 1.3em !important; }
          .nav-link { font-size: 0.95em !important; }
          .cell-output-display img { min-height: auto; max-height: 500px; object-fit: contain; }
          .navbar, .quarto-dashboard-header, .navbar-dark,
          .bslib-page-dashboard > .navbar {
            background-color: black !important;
            color: white !important;
          }
          .navbar .navbar-brand, .navbar .nav-link, .navbar-text,
          .navbar .navbar-nav .nav-link {
            color: white !important;
          }
          .navbar .nav-link.active, .navbar .nav-link:hover {
            color: white !important;
            background-color: black !important;
            border: 1px solid white !important;
            border-radius: 4px;
          }
          .navbar .nav-link { color: #aaa !important; }
          .navbar-nav { flex-wrap: wrap !important; }
          .navbar .nav-link { font-size: 0.9em; padding: 0.3em 0.6em !important; }
          .bslib-page-fill, .html-fill-container, .html-fill-item,
          .bslib-card, .card, .tab-pane, .tab-content, .card-body,
          .bslib-grid, .bslib-grid-item {
            height: auto !important; min-height: 0 !important;
            max-height: none !important; overflow: visible !important;
            flex: none !important;
          }
          .cell-output-display { display: block !important; position: relative !important;
            overflow: visible !important; margin-bottom: 1em; }
          .cell-output-display img { display: block; width: 100%; height: auto; min-height: 400px; }
          .fig-caption { display: block; position: relative; clear: both;
            margin-top: 0.5em; margin-bottom: 1em; font-size: 0.85em; color: #e0e0e0; }
          .datatables, .dataTables_wrapper { display: block !important;
            position: relative !important; overflow-x: auto !important; margin-bottom: 1em; }
          table.dataTable { white-space: nowrap; }
          table caption { caption-side: top !important; }
          details { margin-bottom: 0.5em; }
          details summary { cursor: pointer; color: #6c9bd2; font-size: 0.9em; }
          .fullscreen-overlay {
            position: fixed !important; top: 0 !important; left: 0 !important;
            width: 100vw !important; height: 100vh !important;
            z-index: 9999 !important; background: black !important;
            padding: 10px !important; cursor: zoom-out !important;
          }
          .fullscreen-overlay img { width: 100% !important; height: 100% !important;
            object-fit: contain !important; }
          .cell-output-display:has(img) { cursor: zoom-in; }
          </style>
          <script>
          if (!localStorage.getItem('quarto-color-scheme')) {
            document.documentElement.setAttribute('data-bs-theme', 'dark');
          }
          document.addEventListener('click', function(e) {
            // Exit fullscreen: click anywhere on overlay or its img
            var overlay = e.target.closest('.fullscreen-overlay');
            if (overlay) { overlay.remove(); return; }
            var img = e.target.closest('.cell-output-display img');
            if (!img) return;
            var div = document.createElement('div');
            div.className = 'fullscreen-overlay';
            var clone = document.createElement('img');
            clone.src = img.src;
            div.appendChild(clone);
            document.body.appendChild(div);
          });
          // Scroll to top when clicking a tab
          document.addEventListener('shown.bs.tab', function(e) {
            var pane = document.querySelector(e.target.getAttribute('data-bs-target') || e.target.getAttribute('href'));
            if (pane) pane.scrollIntoView({ behavior: 'smooth', block: 'start' });
          });
          </script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)
library(dplyr)

pkgload::load_all(here::here("packages/historicaldata"), quiet = TRUE)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)
```

::: {.callout-warning}
**THEORETICAL EXERCISE.** This vignette examines the asymmetry of daily returns -- removing the worst days improves performance more than removing the best days hurts it. You cannot know which days to avoid in advance. This is not investment advice.
:::

# Definition

## Row

### {.tabset}

#### What

**Beat the Market by Avoiding Its Worst Days.** A thought experiment demonstrating that returns are asymmetric: losses compound more destructively than gains compound constructively.

If you could magically skip the market's N worst days each year, your returns would improve dramatically. But if you also missed the N best days, you'd lose far less. The asymmetry arises from the mathematics of compounding:

- Lose 10% and you need +11.1% to recover
- Lose 20% and you need +25% to recover
- Lose 50% and you need +100% to recover

This is why downside protection matters more than upside capture.

Source: [Morningstar -- "You Can Beat the Stock Market by Avoiding Its Worst Days. You Won't."](https://www.morningstar.com/funds/you-can-beat-stock-market-by-avoiding-its-worst-days-you-wont)

#### Key Distinction

This is **not a tradeable strategy**. The clustering analysis (Tab 3) shows why: the worst and best days occur near each other in crisis periods. Any scheme to avoid the worst days would almost certainly also miss the best days.

The practical implication: strategies with built-in downside protection (like [Macro Defense Rotation](macro-defense-rotation.html)) have a structural advantage because they reduce exposure before crashes, not after.

#### Worked Example

Consider \$100 invested in the S&P 500 over one year with 252 trading days:

1. **All days**: \$100 grows to \$112 (+12% annual return)
2. **Remove 10 worst days** (just 4% of days): \$100 grows to \$165 (+65%)
3. **Remove 10 best days**: \$100 grows to \$75 (-25%)
4. **Asymmetry**: Removing worst days adds \$53; removing best days costs \$37. The downside matters more.

These are illustrative numbers -- the actual analysis with real data follows.

# S&P 500 Analysis

## Row

### {.tabset}

#### Equity Curves

```{r}
#| fig-height: 6
safe_tar_read("aw_cumret_plot")
```

::: {.fig-caption}
```{r}
#| echo: false
#| results: asis
cap <- safe_tar_read("aw_cumret_caption")
if (!is.null(cap)) cat(cap) else cat("*Caption not available -- run tar_make() to generate.*")
```
:::

#### Asymmetry Table

```{r}
asym <- safe_tar_read("aw_asymmetry_table")
if (!is.null(asym)) {
  DT::datatable(asym,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold; color: #e0e0e0;",
      "Asymmetry of removing N worst vs N best days. Ratio > 1 means avoiding losses matters more than capturing gains."
    ),
    rownames = FALSE,
    options = list(pageLength = 15, scrollX = TRUE, dom = "t")
  )
}
```

#### Metrics by Period

```{r}
met <- safe_tar_read("aw_metrics")
if (!is.null(met)) {
  DT::datatable(met,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold; color: #e0e0e0;",
      "SPY metrics by scenario and partition. Training: pre-2020, Testing: 2020+."
    ),
    rownames = FALSE,
    options = list(pageLength = 15, scrollX = TRUE, dom = "t")
  )
}
```

# Temporal Clustering

## Row

### {.tabset}

#### Timeline

```{r}
#| fig-height: 6
safe_tar_read("aw_clustering_plot")
```

::: {.fig-caption}
```{r}
#| echo: false
#| results: asis
cap <- safe_tar_read("aw_clustering_caption")
if (!is.null(cap)) cat(cap) else cat("*Caption not available.*")
```
:::

#### Worst Days Detail

```{r}
cl <- safe_tar_read("aw_clustering")
if (!is.null(cl)) {
  worst <- cl$worst |> select(date, ret_pct, nearest_best, days_to_nearest_best)
  DT::datatable(worst,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold; color: #e0e0e0;",
      "SPY's 10 worst days and distance to the nearest best day. Short distances confirm clustering."
    ),
    rownames = FALSE,
    options = list(pageLength = 15, scrollX = TRUE, dom = "t")
  )
}
```

#### Best Days Detail

```{r}
cl <- safe_tar_read("aw_clustering")
if (!is.null(cl)) {
  best <- cl$best |> select(date, ret_pct, nearest_worst, days_to_nearest_worst)
  DT::datatable(best,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold; color: #e0e0e0;",
      "SPY's 10 best days and distance to the nearest worst day."
    ),
    rownames = FALSE,
    options = list(pageLength = 15, scrollX = TRUE, dom = "t")
  )
}
```

#### Implications

The clustering of worst and best days in the same periods (2008-2009, March 2020) means any practical scheme to avoid worst days would also miss best days.

This is why the "remove both" scenario in the asymmetry table is informative: it shows the net effect of missing both extremes, which is closer to what a real risk-reduction strategy achieves.

The practical takeaway: instead of trying to time individual days, use systematic risk reduction (trend following, momentum, or defensive rotation) that reduces exposure during high-volatility regimes.

# Rolling Analysis

## Row

### {.tabset}

#### Rolling 1-Year Returns

```{r}
#| fig-height: 6
safe_tar_read("aw_rolling_plot")
```

::: {.fig-caption}
Rolling 1-year returns for SPY under three scenarios: all days, remove 10 worst, remove 10 best. The gap between "all days" and "remove worst" varies over time but is consistently positive.
:::

#### Net Benefit

```{r}
#| fig-height: 6
safe_tar_read("aw_net_benefit_plot")
```

::: {.fig-caption}
Rolling 1-year net benefit = gain from avoiding 10 worst days minus loss from also missing 10 best days (since they cluster together). Green = gain from avoiding worst, red = loss from missing best, blue = net. Positive net means the asymmetry favours downside protection even accounting for missed upside.
:::

# Cross-Asset

## Row

### {.tabset}

#### Index Comparison

```{r}
mi <- safe_tar_read("aw_multi_index")
if (!is.null(mi)) {
  DT::datatable(mi,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold; color: #e0e0e0;",
      "Asymmetry analysis across US index ETFs. Ratio > 1 = avoiding worst days matters more."
    ),
    rownames = FALSE,
    options = list(pageLength = 15, scrollX = TRUE, dom = "t")
  )
}
```

#### Long History (Mkt-RF)

The asymmetry table on the **S&P 500 Analysis** tab includes Mkt-RF rows (Fama-French market factor, 1926+). The same pattern holds over nearly a century: the asymmetry ratio exceeds 1 at every N, confirming this is not a recent or data-period-specific phenomenon.

Mkt-RF is not identical to the S&P 500 — it is the value-weighted CRSP market return minus the risk-free rate. The longer history (1926 vs 1993 for SPY) provides stronger evidence that return asymmetry is a persistent structural feature of equity markets.
```

# Assumptions

## Row

### {.tabset}

#### Assumptions

| Assumption | Value |
|------------|-------|
| Returns | Daily close-to-close, adjusted for splits and dividends |
| Transaction costs | None modelled (theoretical exercise) |
| Execution | Perfect foresight (impossible in practice) |
| Data source | SPY via HuggingFace equity_daily; Mkt-RF via Fama-French |
| Cash return | Not modelled (days "removed" earn 0%) |

#### Caveats

- **You cannot know in advance** which days will be worst or best
- **Worst and best days cluster** in the same crisis periods (see Temporal Clustering tab)
- **Survivorship bias**: SPY exists today because the US market survived; other markets did not
- **Mkt-RF is not S&P 500**: it is the value-weighted CRSP market return minus risk-free rate
- **This is not a strategy**: it is a demonstration of return asymmetry

# References

## Row

### {.tabset}

#### References

- [Morningstar: "You Can Beat the Stock Market by Avoiding Its Worst Days. You Won't."](https://www.morningstar.com/funds/you-can-beat-stock-market-by-avoiding-its-worst-days-you-wont)
- [Macro Defense Rotation](macro-defense-rotation.html) -- a practical downside-protection strategy
- [Strategy Leaderboard](leaderboard.html) -- comparison of all strategies
- Data: [`historicaldata` R package](https://johngavin.github.io/historical/examples.html)
