---
title: "BDBB Queueing Diagnostics — SOL/USD (Phase 1)"
description: |
  M/G/infinity queueing model (Varma 2026) applied to SOL/USD Kraken hourly data.
  Replicates resilience half-life estimation and tail-risk predictivity test.
date: "`r Sys.Date()`"
format:
  html:
    theme:
      dark: darkly
      light: cosmo
    include-in-header:
      - _includes/color-scheme-meta.html
      - text: |
          <link rel="stylesheet" href="vignette-shared.css">
          <script src="vignette-shared.js"></script>
---

```{r setup, include=FALSE}
library(targets)
library(dplyr)
library(ggplot2)
store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)
```

## Overview

This vignette replicates Varma (2026)'s M/G/∞ queueing model on SOL/USD
using Kraken hourly OHLCVT bars.  The model characterises the market's
resilience — its tendency to absorb order-flow shocks and revert to
equilibrium price — as a queueing process.  Three outputs are tracked:

1. **R-metric**: variance per unit of signed order flow (volume × direction).
   Higher R implies prices move more per unit of flow — thinner liquidity.
2. **θ (theta)**: mean-reversion rate estimated from ACF decay of signed
   log-returns.  Positive θ → mean reversion; negative → momentum.
3. **Half-life**: `log(2) / θ` in hours — the time for half the price
   impact of a shock to dissipate.

```{r data-check}
dv <- tar_read(bdbb_sol_dv)
knitr::kable(dv, caption = "SOL/USD data coverage")
```

## Rolling Diagnostics (R, θ, Half-Life)

```{r fit-summary}
fit     <- tar_read(bdbb_sol_fit)
metrics <- tar_read(bdbb_sol_metrics)
knitr::kable(metrics, caption = tar_read(bdbb_sol_caption))
```

## Regime Distribution

```{r regime-plot}
fit |>
  dplyr::filter(!is.na(regime)) |>
  ggplot2::ggplot(ggplot2::aes(x = regime, fill = regime)) +
  ggplot2::geom_bar() +
  ggplot2::labs(
    title   = "BDBB Regime Distribution — SOL/USD",
    x       = "Regime",
    y       = "Window count",
    caption = paste0(
      "Mean-reversion vs momentum vs ambiguous across 30-day rolling windows. ",
      "Varma (2026) BTC benchmark: ~39% mean-reversion windows."
    )
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(legend.position = "bottom")
```

## Tail-Risk Predictivity (R vs Amihud vs Kyle)

```{r tail-predict}
tp <- tar_read(bdbb_sol_tail_predict)
pull_spread <- function(pred) {
  round(dplyr::filter(tp, predictor == pred)$spread_pp, 2)
}

knitr::kable(
  tp,
  digits  = 2,
  caption = paste0(
    "Tail-risk predictivity by predictor. ",
    "Spread = P(top-decile hourly move | high tercile) - P(... | low tercile). ",
    "SOL/USD spreads computed here: R = ", pull_spread("R"), " pp, ",
    "Amihud = ", pull_spread("amihud_mean"), " pp, ",
    "Kyle = ", pull_spread("kyle_mean"), " pp. ",
    "Varma (2026) BTC benchmarks, shown for reference only: R = 8.6 pp, ",
    "Amihud = 4.7 pp, Kyle = 1.1 pp. Kyle's lambda is estimated as the OLS ",
    "price-impact slope of log-returns on signed order flow within each ",
    "rolling window (cov/var; see ",
    "[`bdbb_fit()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/bdbb.R#L46)",
    "), not a per-bar ratio. Prior to [#624](https://github.com/JohnGavin/historical/issues/624), ",
    "the `kyle_mean` column was numerically identical to `amihud_mean`, so any ",
    "earlier apparent agreement with the Varma Kyle benchmark was spurious and ",
    "the corrected figures above should not be assumed to reproduce it."
  )
)
```

## Related vignettes

- **[Falsification](falsification.html)** — statistical testing framework applied to all systematic strategies; the same hypothesis-testing approach used to evaluate the BDBB diagnostics
- **[Negative Results](falsification.html#failed-strategies)** — documents approaches that failed empirical scrutiny; the BDBB mean-reversion findings sit alongside these investigations
- **[Strategy Leaderboard](leaderboard.html)** — performance comparison across all strategies; provides context for the SOL/USD queueing model relative to other systematic approaches

## Build Info

{{< include _includes/build-info-footer.qmd >}}
