---
title: "DRIF: Daily Return Information Factor"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    embed-resources: false
    include-in-header:
      - text: |
          <link rel="stylesheet" href="vignette-shared.css">
          <script src="vignette-shared.js"></script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)
library(dplyr)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)
```

# Strategy

## Row

::: {.callout-note}
## Daily Return Information Factor (DRIF)

Based on [Alpha Architect research](https://alphaarchitect.com/daily-stock-returns/): an elastic net regression on the **full distribution of past 21 daily returns** — both chronological order and rank order — predicts next-month returns. This captures more information than single-signal approaches like MAX or momentum.

**Factor-level implementation:** Each month, use an expanding-window elastic net (`glmnet::cv.glmnet`, alpha = 0.5, 5-fold CV) to predict next month's factor return from 42 features (21 chronological + 21 rank daily returns). Go long the top-2 predicted factors. Monthly rebalance.

**Key insight from the paper:** Chronological order (when returns occurred) is more predictive than rank order (how extreme they were). Timing > magnitude. The DRIF signal subsumes many existing anomalies including short-term reversal, volatility effects, and the MAX signal.
:::

## Row

### {.tabset}

#### Performance

```{r}
#| fig-height: 6
safe_tar_read("drif_cumret_plot")
```

::: {.fig-caption}
**DRIF vs Market.** Y: growth of $1 (log scale). Blue = DRIF (top-2 factors by elastic net prediction, equal weight). Red = Market factor (Mkt-RF). Dashed line = OOS start (2020). Uses PCA-OLS fallback (glmnet not in current Nix shell). Source: Fama-French FF5 + Momentum via `hd_factors()`.
:::

#### Metrics

```{r}
metrics <- safe_tar_read("drif_metrics")
if (!is.null(metrics)) {
  display <- metrics |>
    dplyr::mutate(across(c(cagr, vol, max_dd, hit_rate, bench_cagr, bench_vol),
                  ~ paste0(round(. * 100, 1), "%"))) |>
    dplyr::mutate(across(c(sharpe, bench_sharpe), ~ round(., 2)))
  hd_dt_wide(display, "DRIF strategy metrics: in-sample (1968-2019), out-of-sample (2020+), and full period. Uses 60-month minimum expanding window.")
}
```

#### Factor Selection

```{r}
sf <- safe_tar_read("drif_selection_freq")
if (!is.null(sf)) {
  sf <- sf |> dplyr::mutate(pct = paste0(round(pct * 100, 1), "%"))
  hd_dt(sf, "How often each factor is selected by the DRIF signal. Momentum leads but less dominant than in Factor MAX (44% vs 68%).")
}
```

#### DRIF vs Factor MAX

```{r}
#| fig-height: 6
safe_tar_read("drif_vs_max_plot")
```

::: {.fig-caption}
**DRIF vs Factor MAX.** Both strategies applied to the same FF5+Momentum factor universe. DRIF uses 42 features (21 chronological + 21 rank daily returns) via elastic net; Factor MAX uses only the single highest daily return. DRIF achieves higher returns by capturing the full daily return distribution.
:::

# Methodology

## Row

### {.tabset}

#### How It Works

::: {.callout-tip}
## The DRIF Signal

For each factor, each month:

1. **Collect** the past 21 trading days of daily returns
2. **Create two feature sets:**
   - **Chronological** (c1...c21): returns in time order (c1 = oldest, c21 = most recent)
   - **Rank** (r1...r21): returns sorted by magnitude (r1 = worst, r21 = best)
3. **Predict** next month's return using elastic net (or PCA-OLS fallback) trained on expanding window
4. **Select** top-2 factors by predicted return, equal weight
5. **Rebalance** monthly

**Why two dimensions?** Chronological features capture price pressure, liquidity effects, and momentum-reversal dynamics. Rank features capture behavioral biases around extreme returns. Together they explain more variance than either alone.

**Expanding window:** Training starts after 60 months of data. Each subsequent month adds one month of training data — no lookahead bias.
:::

#### Parameters

```{r}
params <- safe_tar_read("drif_params")
if (!is.null(params)) {
  param_df <- dplyr::tibble(
    Parameter = c("Factors", "Lookback Days", "Top N", "Elastic Net Alpha",
                  "Min Train Months", "Start Date", "IS End", "OOS Start"),
    Value = c(paste(params$factors, collapse = ", "),
              as.character(params$lookback_days),
              as.character(params$top_n),
              as.character(params$alpha),
              as.character(params$min_train_months),
              as.character(params$start_date),
              as.character(params$is_end),
              as.character(params$oos_start))
  )
  hd_dt(param_df, "Strategy parameters. Change drif_params target and tar_make() to update all.")
}
```

#### Chronological vs Rank

The Alpha Architect paper finds chronological features are ~1.7x more predictive than rank features:

| Dimension | Monthly Spread | Intuition |
|-----------|---------------|-----------|
| **Chronological** | ~1.5% | Recent price pressure, liquidity, order flow |
| **Rank** | ~0.9% | Behavioral biases around extreme returns |
| **Combined (DRIF)** | ~1.6% | Full daily return distribution |

Our factor-level implementation applies the same logic: the sequence of daily factor returns contains more information than just the extremes (MAX signal) or the average (momentum).

#### References

1. **Research:** [A Unified Framework for Anomalies based on Daily Returns](https://alphaarchitect.com/daily-stock-returns/) — Alpha Architect (2026)
2. **Comparison:** [Factor MAX](factor-max.html) — simpler signal using only max daily return
3. **Data:** FF5 + Momentum via [Ken French Data Library](https://mba.tuck.dartmouth.edu/pages/faculty/ken.french/data_library.html)
4. **Methodology:** Expanding window, PCA-OLS (glmnet unavailable in current Nix shell)

#### Build Info

```{r}
#| label: build-info
#| echo: false
#| results: asis
build_info()
```

# Details

## Row

### {.tabset}

#### Universe

The strategy rotates among 5 Fama-French factors (daily data from 1963):

| Factor | Description | Source |
|--------|-------------|--------|
| **HML** | High Minus Low (value) | FF5 |
| **SMB** | Small Minus Big (size) | FF5 |
| **RMW** | Robust Minus Weak (profitability) | FF5 |
| **CMA** | Conservative Minus Aggressive (investment) | FF5 |
| **Mom** | Momentum (winners minus losers) | Ken French |

Benchmark: **Mkt-RF** (market excess return). 42 features per factor-month (21 chronological + 21 rank).

#### Limitations

::: {.callout-warning}
## Factor-Level vs Stock-Level

This is a **factor-level** implementation — rotating among 5-6 academic factors. The original DRIF paper applies the signal **cross-sectionally to individual stocks** (3000+ in CRSP).

Key differences:
- **Cross-section size:** 5 factors vs 3000+ stocks — much less statistical power for decile sorts
- **Feature independence:** Factor daily returns are more correlated than stock returns
- **No shorting:** We go long top-2 factors; the paper shorts bottom decile stocks
- **PCA-OLS:** Using 5-PC OLS as fallback (glmnet would give elastic net regularisation)

A proper stock-level implementation requires ~500+ US equity tickers (see #6 for international equities expansion).
:::
