---
title: "European Overlay: CISS vs VIX Regime"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    embed-resources: false
    css: falsification-theme.css
    include-in-header:
      - text: |
          <link rel="stylesheet" href="vignette-shared.css">
          <script src="vignette-shared.js"></script>
          <script>
          if (!localStorage.getItem('quarto-color-scheme')) {
            document.documentElement.setAttribute('data-bs-theme', 'dark');
          }
          document.addEventListener('shown.bs.tab', function(e) {
            var tabBar = e.target.closest('.nav-tabs');
            if (tabBar) tabBar.scrollIntoView({ behavior: 'smooth', block: 'start' });
          });
          </script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)
library(dplyr)

pkgload::load_all(here::here("packages/historicaldata"), quiet = TRUE)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)
```

::: {.callout-note .small}
European overlay strategies. Tests whether US VIX-based regime signals generalise to European equities, and compares with the ECB's own [CISS](https://www.ecb.europa.eu/pub/pdf/scpwps/ecbwp1426.pdf){target="_blank"} (<abbr title="Composite Indicator of Systemic Stress — ECB's flagship daily systemic risk indicator combining bond, equity, FX, money market, and financial intermediary stress">Composite Indicator of Systemic Stress</abbr>) equity sub-index as a European-native vol proxy. CISS equity correlates r≈0.75 with VIX over 6,000+ daily observations (2000-2026). See also: [ECB CISS page](https://www.ecb.europa.eu/stats/macroprudential_indicators/systemic_risk/html/index.en.html){target="_blank"} | [Wikipedia](https://en.wikipedia.org/wiki/Composite_Indicator_of_Systemic_Stress){target="_blank"}.
:::

# What is CISS?

## Row

### {.tabset}

#### Definition

The **Composite Indicator of Systemic Stress** (CISS) is the ECB's flagship risk indicator, published daily since 2000. It measures systemic stress across 5 financial market segments:

| Sub-market | What it captures | Key variable |
|-----------|-----------------|-------------|
| **Equity** | Stock market volatility + returns | Realised vol, max drawdown, index returns |
| **Bond** | Government bond spreads + volatility | Yield spread, realised vol |
| **FX** | Currency volatility | EUR exchange rate volatility |
| **Money market** | Interbank stress | EURIBOR-OIS spread, bid-ask |
| **Financial intermediary** | Bank/insurance sector | Bank stock vol, CDS spreads |

The composite index combines all 5 using **time-varying cross-correlations** — when markets stress simultaneously (contagion), the index spikes more than the sum of parts.

**Range:** 0 (calm) to 1 (systemic crisis). Peak: 0.84 during the GFC (Oct 2008).

**Why CISS, not VSTOXX?** VSTOXX (European implied vol, equivalent to VIX) is a commercial Qontigo product with no free API. CISS equity sub-index correlates r=0.75 with VIX and is freely available from the ECB.

**Source:** [ECB SDMX API](https://data.ecb.europa.eu/){target="_blank"} — series `CISS/D.U2.Z0Z.4F.EC.SS_CIN.IDX`. R function: [`hd_ecb()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/ecb.R){target="_blank"}. Methodology: [Holló, Kremer & Lo Duca (2012), ECB WP #1426](https://www.ecb.europa.eu/pub/pdf/scpwps/ecbwp1426.pdf){target="_blank"}.

#### CISS vs VIX

```{r}
#| label: ciss-vix-correlation
ecb_raw <- safe_tar_read("ecb_raw")
data_available <- FALSE

if (!is.null(ecb_raw)) {
  ciss_eq <- ecb_raw |> filter(series_name == "ciss_equity") |> select(date, ciss = value)
  ciss_comp <- ecb_raw |> filter(series_name == "ciss_composite") |> select(date, ciss_composite = value)

  # VIX from aw_vix_daily target
  vix_raw <- safe_tar_read("aw_vix_daily")
  if (!is.null(vix_raw)) {
    vix <- vix_raw |>
      mutate(date = as.Date(date)) |>
      select(date, vix) |>
      distinct(date, .keep_all = TRUE)

    merged <- inner_join(vix, ciss_eq, by = "date") |>
      inner_join(ciss_comp, by = "date")

    if (nrow(merged) > 0) {
      data_available <- TRUE
      cor_tbl <- tibble::tibble(
        Pair = c("CISS equity vs VIX", "CISS composite vs VIX", "CISS equity vs CISS composite"),
        `Spearman r` = round(c(
          cor(merged$ciss, merged$vix, method = "spearman"),
          cor(merged$ciss_composite, merged$vix, method = "spearman"),
          cor(merged$ciss, merged$ciss_composite, method = "spearman")
        ), 3),
        N = nrow(merged),
        From = as.character(min(merged$date)),
        To = as.character(max(merged$date))
      )

      DT::datatable(cor_tbl, rownames = FALSE, filter = "none",
        options = list(dom = "t"),
        caption = htmltools::tags$caption(
          style = "caption-side: top; text-align: left;",
          htmltools::HTML(paste0(
            "Spearman rank correlations between CISS sub-indices and VIX. ",
            "N = ", format(nrow(merged), big.mark = ","), " overlapping daily observations. ",
            "CISS equity is the strongest VIX proxy (r = 0.75). ",
            "Source: <a href='https://github.com/JohnGavin/historical/blob/main/R/plan_ecb.R' target='_blank'>plan_ecb.R</a> + FRED VIXCLS."
          ))
        )
      ) |> DT::formatRound("Spearman r", 3)
    }
  }
}

if (!data_available) {
  htmltools::p(class = "text-muted", "Data not available. Run tar_make() to build ecb_raw and aw_vix_daily targets.")
}
```

#### Sub-Market Breakdown

```{r}
#| label: ciss-sub-markets
if (!is.null(ecb_raw)) {
  sub_markets <- c("ciss_equity", "ciss_bond", "ciss_fx", "ciss_money", "ciss_financial", "ciss_correlation", "ciss_sovereign")
  sub_labels <- c("Equity", "Bond", "FX", "Money market", "Financial intermediary", "Cross-correlation", "Sovereign")

  sub_summary <- ecb_raw |>
    filter(series_name %in% sub_markets) |>
    group_by(series_name) |>
    summarise(
      N = n(),
      From = as.character(min(date)),
      To = as.character(max(date)),
      Min = round(min(value, na.rm = TRUE), 4),
      Max = round(max(value, na.rm = TRUE), 4),
      Mean = round(mean(value, na.rm = TRUE), 4),
      .groups = "drop"
    ) |>
    mutate(`Sub-market` = sub_labels[match(series_name, sub_markets)]) |>
    select(`Sub-market`, N, From, To, Min, Max, Mean)

  DT::datatable(sub_summary, rownames = FALSE, filter = "none",
    options = list(dom = "t"),
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left;",
      htmltools::HTML(paste0(
        "CISS sub-market components. All daily, 2000-2026. ",
        "Cross-correlation is negative in calm markets (independence) ",
        "and flips positive in crises (contagion). ",
        "Source: <a href='https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/ecb.R' target='_blank'>hd_ecb_registry()</a>."
      ))
    )
  ) |> DT::formatRound(c("Min", "Max", "Mean"), 4)
} else {
  htmltools::p(class = "text-muted", "Data not available. Run tar_make() to build ecb_raw target.")
}
```

# CISS Overlay Results

## Row

### {.tabset}

#### OOS Performance

```{r}
#| label: ciss-results-table
ciss_res <- safe_tar_read("eur_ciss_results")
if (!is.null(ciss_res)) {
  display <- ciss_res |>
    select(ETF = asset, Strategy = strategy,
           `CAGR (%)` = cagr, `Vol (%)` = vol,
           `Max DD (%)` = max_dd, `Sharpe` = hac_sharpe)

  DT::datatable(display, rownames = FALSE, filter = "none",
    options = list(dom = "t", pageLength = 15),
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left;",
      htmltools::HTML(paste0(
        "CISS equity overlay vs Buy & Hold for 5 European ETFs (OOS, 2020-present). ",
        "CISS regime: benign (100% equity), cautious (50%), hostile (10%). ",
        "Thresholds: 33rd/67th percentile of CISS equity stress history. ",
        "4 of 5 ETFs improve on Sharpe ratio. Max drawdown halved across all ETFs. ",
        "Source: <a href='https://github.com/JohnGavin/historical/blob/main/R/plan_european_overlay.R' target='_blank'>plan_european_overlay.R</a>."
      ))
    )
  ) |> DT::formatRound(c("CAGR (%)", "Vol (%)", "Max DD (%)", "Sharpe"), 2)
}
```

#### VIX vs CISS Comparison

```{r}
#| label: vix-vs-ciss
eur_vix <- safe_tar_read("eur_results")
if (!is.null(eur_vix) && !is.null(ciss_res)) {
  # VIX overlay OOS
  vix_oos <- eur_vix |>
    filter(period == "OOS", strategy == "RSC Overlay") |>
    select(asset, vix_cagr = cagr, vix_vol = vol, vix_dd = max_dd, vix_sharpe = hac_sharpe)

  # CISS overlay OOS
  ciss_oos <- ciss_res |>
    filter(strategy == "CISS Overlay") |>
    select(asset, ciss_cagr = cagr, ciss_vol = vol, ciss_dd = max_dd, ciss_sharpe = hac_sharpe)

  # Buy & hold
  bh_oos <- eur_vix |>
    filter(period == "OOS", strategy == "Buy & Hold") |>
    select(asset, bh_cagr = cagr, bh_vol = vol, bh_dd = max_dd, bh_sharpe = hac_sharpe)

  comparison <- bh_oos |>
    left_join(vix_oos, by = "asset") |>
    left_join(ciss_oos, by = "asset") |>
    select(ETF = asset,
           `BH Sharpe` = bh_sharpe, `VIX Sharpe` = vix_sharpe, `CISS Sharpe` = ciss_sharpe,
           `BH DD (%)` = bh_dd, `VIX DD (%)` = vix_dd, `CISS DD (%)` = ciss_dd)

  DT::datatable(comparison, rownames = FALSE, filter = "none",
    options = list(dom = "t"),
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left;",
      htmltools::HTML(paste0(
        "Three-way comparison: Buy & Hold vs US VIX overlay vs ECB CISS overlay. ",
        "CISS overlay uses European-native stress signal; VIX uses US vol. ",
        "Both reduce drawdowns dramatically. CISS outperforms VIX overlay on Sharpe for 3 of 5 ETFs. ",
        "Source: <a href='https://github.com/JohnGavin/historical/blob/main/R/plan_european_overlay.R' target='_blank'>plan_european_overlay.R</a>."
      ))
    )
  ) |> DT::formatRound(c("BH Sharpe", "VIX Sharpe", "CISS Sharpe", "BH DD (%)", "VIX DD (%)", "CISS DD (%)"), 2)
}
```

#### Attribution

The CISS overlay reduces risk through regime-dependent exposure scaling:

| Regime | CISS equity level | Equity exposure | When it fires |
|--------|------------------|----------------|---------------|
| **Benign** | Below 33rd percentile | 100% | Normal markets — fully invested |
| **Cautious** | 33rd-67th percentile | 50% | Elevated stress — half equity, half cash |
| **Hostile** | Above 67th percentile | 10% | Crisis — near-cash, avoid drawdown |

**Return attribution:**

- **Return drag in calm markets:** CISS overlay slightly underperforms buy & hold because cautious regime reduces exposure even when markets are fine (STOXX 600: 5.0% vs 10.6% CAGR)
- **Drawdown protection:** Max DD halved across all ETFs (-36% to -43% buy & hold → -8% to -14% CISS overlay)
- **Sharpe improvement:** Lower vol more than compensates for lower CAGR in 4 of 5 ETFs (Euro Stoxx 50: Sharpe 1.03 vs 0.56)
- **STOXX 600 exception:** Broad diversification already provides some crisis protection, so the overlay's cost (reduced upside) exceeds its benefit (reduced drawdown) on a Sharpe basis

**Why CISS works better than VIX for European equities:**

1. CISS is contemporaneous — it measures European stress directly, not a lagged US signal
2. CISS captures bond, FX, and money market stress that VIX misses (pure equity vol)
3. CISS cross-correlation component detects contagion — markets moving together in crisis

#### Country CISS

```{r}
#| label: country-ciss
if (!is.null(ecb_raw)) {
  countries <- c("ciss_de", "ciss_fr", "ciss_it", "ciss_gb", "ciss_us")
  c_labels <- c("Germany", "France", "Italy", "UK", "US (ECB-computed)")

  country_summary <- ecb_raw |>
    filter(series_name %in% countries) |>
    group_by(series_name) |>
    summarise(
      N = n(),
      Mean = round(mean(value, na.rm = TRUE), 4),
      Max = round(max(value, na.rm = TRUE), 4),
      `Max date` = as.character(date[which.max(value)]),
      .groups = "drop"
    ) |>
    mutate(Country = c_labels[match(series_name, countries)]) |>
    select(Country, N, Mean, Max, `Max date`) |>
    arrange(desc(Max))

  DT::datatable(country_summary, rownames = FALSE, filter = "none",
    options = list(dom = "t"),
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left;",
      htmltools::HTML(paste0(
        "Country-level CISS: systemic stress per country (daily, 2000-2026). ",
        "Italy and France peaked during the eurozone debt crisis (2011-2012). All countries peaked during GFC (2008). ",
        "The ECB also computes a <strong>US CISS</strong> for cross-Atlantic comparison — a potential overlay signal for US equities too. ",
        "Source: <a href='https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/ecb.R' target='_blank'>hd_ecb_registry()</a>. ",
        "Reference: <a href='https://www.ecb.europa.eu/stats/macroprudential_indicators/systemic_risk/html/index.en.html' target='_blank'>ECB CISS dashboard</a>."
      ))
    )
  ) |> DT::formatRound(c("Mean", "Max"), 4)
}
```

#### Build Info

`r build_info()`
