---
title: "Evidence & Context"
subtitle: "Strategy Verdicts and Long-Run Historical Evidence"
format:
  html:
    embed-resources: false
    theme:
      dark: darkly
      light: cosmo
    toc: true
    toc-depth: 2
    code-fold: true
    include-in-header:
      - text: |
          <style>
          body { font-size: 1.3em; line-height: 1.6; }
          caption, figcaption, .figure-caption { color: #fff !important; }
          .cell-output-display { margin: 1.5em 0; }
          figure { margin: 2em 0; }
          figcaption { font-size: 0.9em; opacity: 0.8; margin-top: 0.5em; }
          table { margin: 1.5em auto; }
          .negative-card {
            background: #1a1a2e; border: 1px solid #d73027; border-radius: 8px;
            padding: 16px; margin: 16px 0;
          }
          .negative-card h3 { color: #d73027; margin-top: 0; }
          .lesson { background: #111; border-left: 4px solid #e6a817; padding: 12px 16px; margin: 12px 0; }
          #ui-controls {
            position: fixed; top: 10px; right: 10px; z-index: 9999;
            display: flex; gap: 5px; background: var(--bs-body-bg);
            padding: 5px; border-radius: 4px; border: 1px solid var(--bs-border-color);
          }
          #ui-controls button {
            padding: 5px 10px; border: 1px solid var(--bs-border-color);
            background: var(--bs-body-bg); color: var(--bs-body-color);
            cursor: pointer; border-radius: 3px;
          }
          #ui-controls button:hover { opacity: 0.8; }
          </style>
          <script>
          if (!localStorage.getItem('quarto-color-scheme'))
            document.documentElement.setAttribute('data-bs-theme', 'dark');
          document.addEventListener('DOMContentLoaded', function() {
            var toolbar = document.createElement('div');
            toolbar.id = 'ui-controls';
            toolbar.innerHTML = '<button id="font-down" title="Decrease font">A-</button>' +
              '<button id="font-up" title="Increase font">A+</button>' +
              '<button id="theme-toggle" title="Toggle dark/light">◐</button>';
            document.body.appendChild(toolbar);
            var baseSize = parseFloat(localStorage.getItem('font-size-px') || '20.8');
            document.documentElement.style.fontSize = baseSize + 'px';
            document.getElementById('font-up').addEventListener('click', function() {
              baseSize = Math.min(baseSize + 1, 28);
              document.documentElement.style.fontSize = baseSize + 'px';
              localStorage.setItem('font-size-px', baseSize);
            });
            document.getElementById('font-down').addEventListener('click', function() {
              baseSize = Math.max(baseSize - 1, 12);
              document.documentElement.style.fontSize = baseSize + 'px';
              localStorage.setItem('font-size-px', baseSize);
            });
            document.getElementById('theme-toggle').addEventListener('click', function() {
              var current = document.documentElement.getAttribute('data-bs-theme');
              var next = current === 'dark' ? 'light' : 'dark';
              document.documentElement.setAttribute('data-bs-theme', next);
              localStorage.setItem('quarto-color-scheme', next);
            });
          });
          </script>
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| label: setup
#| include: false
library(targets)
library(dplyr)
library(here)
library(ggplot2)
library(tidyr)
library(DT)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)

jst_equity_premium <- safe_tar_read("jst_equity_premium")
jst_pervasiveness  <- safe_tar_read("jst_pervasiveness")
jst_crises         <- safe_tar_read("jst_crises")
jst_summary        <- safe_tar_read("jst_summary")
```

::: {.panel-tabset}

## Negative Results

### Why Document Negative Results?

Negative results are as valuable as positive ones. A strategy that fails the falsification framework teaches us:

- Which signals are **disguised beta** (market exposure dressed up as alpha)
- Which intuitions about markets are **wrong** (or at least not tradeable)
- Where the **boundary** lies between genuine alpha and noise

Publication bias in finance means failed strategies are rarely documented. This page corrects that.

### Failed Strategies

```{r neg-scorecard}
#| label: neg-scorecard
summary <- safe_tar_read("fals_summary")
nms <- safe_tar_read("fals_vig_names")
if (!is.null(summary) && !is.null(nms)) {
  neg <- summary |>
    filter(ff_alpha_tstat < 2.0) |>
    mutate(
      Strategy = nms$long_name[match(strategy, nms$code_name)],
      `HAC t` = round(hac_tstat, 2),
      `Alpha t` = round(ff_alpha_tstat, 2),
      `R² (%)` = round(ff_r_squared * 100, 1),
      `Alpha (%)` = round(ff_alpha_annual * 100, 0),
      Diagnosis = case_when(
        ff_r_squared > 0.50 ~ "Pure market beta",
        ff_r_squared > 0.15 ~ "Significant factor exposure",
        TRUE ~ "Noise — neither alpha nor beta"
      )
    ) |>
    select(Strategy, `HAC t`, `Alpha t`, `R² (%)`, `Alpha (%)`, Diagnosis) |>
    arrange(`Alpha t`)

  hd_dt(neg, paste0(
    nrow(neg), " of ", nrow(summary),
    " strategies failed the falsification framework (Alpha t < 2.0). ",
    "These strategies' returns are explained by known factor exposure or noise."
  ))
}
```

### Avoid Worst Days (VIX Protection) {#avoid-worst}

::: {.negative-card}
#### Verdict: Pure market beta (R² = 41%)

**What it does:** Exit to cash when yesterday's <abbr title="CBOE Volatility Index — implied volatility of S&P 500 options">VIX</abbr> > 30 or yesterday's absolute return > 3%. Re-enter when VIX < 25 and 5-day cooloff expires. All signals use t-1 (previous day) values.

**Why it fails:**

- The strategy is essentially a crude vol-timing filter on SPY
- When VIX is high, you're out of the market — but so is the VIX signal embedded in the market itself
- The 41% R² means the strategy's returns are 41% explained by Mkt-RF (market factor alone)
- Alpha = -1% per year: the strategy **destroys** value relative to just holding the market
- The t+1 execution constraint is critical: the pre-fix version (using same-day VIX) showed 18.4% CAGR; honest t+1 execution drops it to 5.3%

**Key lesson:** VIX is not a leading indicator — it's a concurrent indicator. By the time VIX > 30, the worst days have already happened.
:::

::: {.lesson}
**What we learned:** Any strategy that triggers on elevated VIX is timing exits AFTER the crash, not before it. The worst days and best days cluster together (median 5 calendar days apart). You cannot avoid one without missing the other. See the [temporal clustering analysis](avoid-worst-days.html).
:::

### Risk State Classification (VIX Overlay) {#risk-state}

::: {.negative-card}
#### Verdict: Pure market beta (R² = 88%)

**What it does:** Three-signal design: VVIX (vol-of-vol) as early warning, VIX term structure 5-day change, VIX term structure level. Market-specific percentile thresholds from training data only. Overlay on SPY: reduce exposure when signals are elevated.

**Why it fails:**

- 88% R² — this is almost entirely market beta
- Alpha = -1% per year (negative): the overlay HURTS performance
- Despite using VVIX (which IS an early warning — it rises before VIX spikes), the signal doesn't translate into tradeable alpha
- The 5-day change in term structure adds noise, not signal
- Percentile thresholds from training data don't generalise: crisis regimes are structurally different from calm regimes

**Key lesson:** Adding more VIX-derived signals doesn't help. VVIX, term structure, and VIX level are all highly correlated. Three correlated signals ≈ one signal with more noise.
:::

::: {.lesson}
**What we learned:** Implied volatility signals (VIX, VVIX, term structure) are priced in. The market already incorporates this information. A strategy based on these signals is, by construction, replicating what the market already does — hence the 88% R² with the market factor.
:::

### LTR Cross-Sectional Momentum (Borderline) {#ltr}

::: {.negative-card}
#### Verdict: Borderline (Alpha t = 1.99, R² = 7%)

**What it does:** XGBoost LambdaMART ranks ~500 US stocks monthly using 21 momentum/volatility features. Long top decile, short bottom decile.

**Why it's borderline:**

- Alpha t = 1.99 — just below the 2.0 threshold
- R² = 7% — low factor exposure, which is good
- Alpha = 166% per year — suspiciously high
- The demo uses ~51 stocks (not the full US universe) which inflates returns
- Walk-forward from 2005 with annual retraining — but limited out-of-sample validation

**Key lesson:** The signal may be genuine but the implementation is too small-scale to confirm. Needs: full US universe (3000+ stocks), transaction costs for monthly L/S rebalancing, and multi-market replication.
:::

::: {.lesson}
**What we learned:** Cross-sectional momentum is well-documented academically. Our implementation is a proof-of-concept, not a production signal. The 166% annual alpha is an artifact of the small universe, not evidence of a genuine edge.
:::

### Pattern: What Predicts Failure?

```{r neg-pattern}
#| label: neg-pattern
if (!is.null(summary) && !is.null(nms)) {
  pattern <- summary |>
    mutate(
      Strategy = nms$long_name[match(strategy, nms$code_name)],
      `Asset Class` = nms$asset_class[match(strategy, nms$code_name)],
      `Frequency` = nms$frequency[match(strategy, nms$code_name)],
      `R² (%)` = round(ff_r_squared * 100, 1),
      `Alpha t` = round(ff_alpha_tstat, 2),
      Verdict = ifelse(ff_alpha_tstat > 2.0, "Alpha", "No alpha")
    ) |>
    select(Strategy, `Asset Class`, Frequency, `R² (%)`, `Alpha t`, Verdict) |>
    arrange(desc(`Alpha t`))

  hd_dt(pattern, paste0(
    "All 5 strategies sorted by Alpha t. ",
    "The two overlay strategies (daily, VIX-based) both fail. ",
    "The two factor strategies (monthly, momentum-based) both pass. ",
    "Pattern: daily VIX-based overlays = beta; monthly factor rotation = alpha."
  ))
}
```

### Implications for Portfolio Construction

1. **Do not include** Avoid Worst Days or Risk State in the multi-strategy portfolio (#52)
2. **VIX-based overlays** add cost and complexity without alpha — use passive SPY instead
3. **Factor rotation** (<abbr title="Dynamic Rotation Into Factors — selects best-performing FF factor each month">DRIF</abbr>, Factor MAX) is where the genuine alpha lives
4. **Cross-sectional momentum** (LTR) needs scaling before inclusion — promising but unconfirmed at production scale

### References

- Full falsification methodology: [Strategy Falsification](falsification.html)
- Avoid Worst Days analysis: [avoid-worst-days.html](avoid-worst-days.html)
- Strategy leaderboard: [leaderboard.html](leaderboard.html)
- Harvey, Liu & Zhu (2016). "...and the Cross-Section of Expected Returns"

## Historical Evidence

### Overview

The [Jordà-Schularick-Taylor Macrohistory Database](https://www.macrohistory.net/database/){target="_blank" rel="noopener noreferrer"} provides 150+ years of macroeconomic and financial data across `r jst_summary$n_countries` advanced economies, spanning `r jst_summary$year_range[1]` to `r jst_summary$year_range[2]`. This tab analyzes the equity premium (equity total return minus short-term government bill rate) to test the cross-geography pervasiveness criterion from the Swedroe evidence-based investing framework.

**Key findings:**

- Equity premium is positive in `r jst_summary$pervasive_count` of `r jst_summary$pervasive_total` countries over the full sample (mean > 0, t-stat > 2, majority of years positive)
- USA equity premium from JST correlates `r jst_summary$correlation_with_ff` with the Fama-French market factor (1926-2020 overlap period)
- Financial crises occurred in `r nrow(jst_crises)` countries, with total crisis-years ranging from `r min(jst_crises$n_crises)` to `r max(jst_crises$n_crises)` per country

**Source:** Òscar Jordà, Moritz Schularick, and Alan M. Taylor (2017). "Macrofinancial History and the New Business Cycle Facts." In *NBER Macroeconomics Annual 2016*, volume 31, edited by Martin Eichenbaum and Jonathan A. Parker. Chicago: University of Chicago Press. [DOI: 10.1086/690241](https://doi.org/10.1086/690241){target="_blank" rel="noopener noreferrer"}

---

### Cross-Geography Pervasiveness

The table below shows full-sample equity premium statistics by country, sorted by mean premium. A factor is considered *pervasive* if it has a positive mean premium, t-statistic > 2, and is positive in more than 50% of years.

```{r jst-pervasiveness}
#| label: jst-pervasiveness
#| tbl-cap: "Equity premium pervasiveness test: full-sample statistics by country. Countries are sorted by mean premium. The 'Pervasive' column indicates whether the premium meets all three criteria: positive mean, t-stat > 2, and positive in >50% of years. Data source: [JST Macrohistory Database](https://www.macrohistory.net/database/)."

jst_pervasiveness |>
  select(
    Country = iso,
    Period = from_year,
    `To Year` = to_year,
    `N Years` = n_years,
    `Mean (%)` = mean_premium,
    `SD (%)` = sd_premium,
    `t-stat` = t_stat,
    `% Positive` = pct_positive,
    Pervasive = pervasive
  ) |>
  mutate(across(where(is.numeric), ~round(., 2))) |>
  datatable(
    options = list(
      pageLength = 20,
      dom = 't',
      columnDefs = list(
        list(className = 'dt-center', targets = 1:8)
      )
    ),
    rownames = FALSE
  ) |>
  formatStyle(
    'Mean (%)',
    background = styleColorBar(range(jst_pervasiveness$mean_premium, na.rm = TRUE), '#69d4a0'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  ) |>
  formatStyle(
    'Pervasive',
    backgroundColor = styleEqual(c(TRUE, FALSE), c('#69d4a0', '#555555'))
  )
```

---

### Equity Premium Heatmap

The heatmap below shows mean equity premium by decade and country. Darker green indicates higher premium; red indicates negative premium. Missing cells represent periods with insufficient data.

```{r jst-heatmap}
#| label: jst-heatmap
#| fig-cap: !expr 'paste0("Equity premium heatmap: mean excess return (equity total return minus bill rate) by decade and country, ", jst_summary$year_range[1], "-", jst_summary$year_range[2], ". Green shades indicate positive premium, red indicates negative. Data source: [JST Macrohistory Database](https://www.macrohistory.net/database/). Generated by [R/plan_jst.R](https://github.com/JohnGavin/historical/blob/main/R/plan_jst.R).")'
#| fig-width: 12
#| fig-height: 8

ggplot(jst_equity_premium, aes(x = decade, y = iso, fill = mean_premium)) +
  geom_tile(color = "white", linewidth = 0.5) +
  scale_fill_gradient2(
    low = "#f08080",
    mid = "#1a1a2e",
    high = "#69d4a0",
    midpoint = 0,
    na.value = "grey50",
    name = "Mean Premium (%)"
  ) +
  scale_x_continuous(breaks = seq(1870, 2020, 10)) +
  labs(
    title = "Equity Premium by Decade and Country",
    subtitle = paste0(jst_summary$year_range[1], "-", jst_summary$year_range[2]),
    x = "Decade",
    y = "Country"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 12, color = "grey60"),
    legend.position = "right"
  )
```

---

### Financial Crisis Timeline

Financial crises (identified by the JST `crisisJST` indicator) are shown below as segments spanning crisis years for each country.

```{r jst-crises-timeline}
#| label: jst-crises-timeline
#| fig-cap: !expr 'paste0("Financial crisis timeline: crisis years by country, ", jst_summary$year_range[1], "-", jst_summary$year_range[2], ". Each segment represents a crisis year as identified by the JST crisisJST indicator. Data source: [JST Macrohistory Database](https://www.macrohistory.net/database/).")'
#| fig-width: 12
#| fig-height: 6

crises_long <- jst_crises |>
  separate_rows(crisis_years, sep = ", ") |>
  mutate(year = as.integer(crisis_years)) |>
  select(iso, year) |>
  arrange(iso, year)

ggplot(crises_long, aes(x = year, y = iso)) +
  geom_point(color = "#f08080", size = 2) +
  geom_segment(
    aes(xend = year + 1, yend = iso),
    color = "#f08080",
    linewidth = 1,
    alpha = 0.6
  ) +
  scale_x_continuous(breaks = seq(1870, 2020, 10)) +
  labs(
    title = "Financial Crisis Timeline",
    subtitle = paste0(jst_summary$year_range[1], "-", jst_summary$year_range[2]),
    x = "Year",
    y = "Country"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 12, color = "grey60")
  )
```

---

### Crisis Frequency Table

The table below shows the number of crisis years per country, sorted by total crisis-years.

```{r jst-crises-table}
#| label: jst-crises-table
#| tbl-cap: "Financial crisis frequency by country: total crisis-years identified by the JST crisisJST indicator. Countries are sorted by total crisis count. Data source: [JST Macrohistory Database](https://www.macrohistory.net/database/)."

jst_crises |>
  select(
    Country = iso,
    `Total Crisis Years` = n_crises,
    `Crisis Years` = crisis_years
  ) |>
  datatable(
    options = list(
      pageLength = 20,
      dom = 't',
      columnDefs = list(
        list(className = 'dt-center', targets = 1)
      )
    ),
    rownames = FALSE
  ) |>
  formatStyle(
    'Total Crisis Years',
    background = styleColorBar(range(jst_crises$n_crises), '#f08080'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )
```

---

### Data Notes

- **Equity premium** = equity total return (`eq_tr`) - short-term government bill rate (`bill_rate`)
- **Pervasiveness criteria**: mean premium > 0, t-stat > 2, and positive in >50% of years
- **Crisis indicator**: `crisisJST` binary flag (1 = crisis year)
- **Country coverage**: `r jst_summary$n_countries` countries: `r paste(sort(unique(jst_pervasiveness$iso)), collapse = ", ")`
- **Data period**: `r jst_summary$year_range[1]` to `r jst_summary$year_range[2]` (`r format(jst_summary$n_obs, big.mark = ",")` country-year observations)

All target computations are defined in [`R/plan_jst.R`](https://github.com/JohnGavin/historical/blob/main/R/plan_jst.R).

:::

```{r build-info}
#| label: build-info
#| results: asis
build_info()
```
