---
title: "historicaldata: Usage Examples"
format:
  html:
    theme: darkly
    embed-resources: true
    toc: false
    include-in-header:
      - text: |
          <style>
          /* Bigger base font */
          body { font-size: 1.15em; }
          /* Click-to-zoom: fill entire screen */
          .fullscreen-overlay {
            position: fixed !important;
            top: 0 !important; left: 0 !important;
            width: 100vw !important; height: 100vh !important;
            z-index: 9999 !important;
            background: rgba(0,0,0,0.95) !important;
            padding: 10px !important;
            overflow: auto !important;
            cursor: zoom-out !important;
          }
          .fullscreen-overlay img {
            width: 100% !important; height: 100% !important;
            object-fit: contain !important;
          }
          .cell-output-display:has(img) { cursor: zoom-in; }
          /* Captions */
          .fig-caption {
            font-size: 0.85em; color: #aaa; margin-top: 0.4em;
            text-align: left; line-height: 1.5;
          }
          </style>
          <script>
          document.addEventListener('click', function(e) {
            var overlay = e.target.closest('.fullscreen-overlay');
            if (overlay) { overlay.classList.remove('fullscreen-overlay'); return; }
            var cell = e.target.closest('.cell-output-display');
            if (!cell || !cell.querySelector('img')) return;
            cell.classList.add('fullscreen-overlay');
          });
          </script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)
```

# Installation

::: {.panel-tabset}

## R (pak)

```r
pak::pak("JohnGavin/historical", subdir = "packages/historicaldata")
library(historicaldata)
```

## Nix

```bash
nix develop github:JohnGavin/historical
# All R + Python packages available in the Nix shell
Rscript -e 'pkgload::load_all("packages/historicaldata")'
```

## T language

```bash
git clone https://github.com/JohnGavin/historical.git && cd historical
nix develop              # enter Nix shell with T + R + Python
t run src/pipeline.t     # run the full data pipeline
```

:::

# Setup

Run this once at the start of your session. All subsequent code assumes these libraries and theme are loaded.

```{r}
#| results: asis
show_code("vig_setup")
```

# Equity

::: {.panel-tabset}

## AAPL Moving Averages

AAPL daily close with 50-day and 200-day simple moving averages. The [golden cross](https://www.investopedia.com/terms/g/goldencross.asp) (50d crosses above 200d) and death cross (50d crosses below) are classic trend-following signals.

```{r}
#| results: asis
show_code("vig_eq_aapl")
```

```{r}
safe_tar_read("vig_eq_aapl")
```

::: {.fig-caption}
**AAPL close with 50d and 200d simple moving averages (SMA).** Y-axis: USD. Grey = close, blue dashed = 50d SMA, red dashed = 200d SMA. MA = simple moving average (trailing, equal-weighted). Source: Yahoo Finance via `hd_ohlcv()`. See Volatility tab for risk-adjusted view. Click plot to zoom fullscreen.
:::

## FAANG Returns

FAANG = Meta (META), Apple (AAPL), Amazon (AMZN), Netflix (NFLX), Alphabet (GOOGL). Cumulative returns rebased to 0% on 2024-01-01 show relative performance divergence.

```{r}
#| results: asis
show_code("vig_eq_faang")
```

```{r}
safe_tar_read("vig_eq_faang")
```

::: {.fig-caption}
**FAANG cumulative returns from 2024-01-01 (0% baseline).** Y-axis: cumulative return (%). Colours: ticker. FAANG = Meta, Apple, Amazon, Netflix, Alphabet. Split-adjusted close. Source: `hd_ohlcv()`. See AAPL MA tab for single-stock detail.
:::

## Realised Volatility

21-day rolling annualised volatility (SD of log returns x sqrt(252)). Higher values = greater price uncertainty. SPY (S&P 500 ETF) is the benchmark.

```{r}
#| results: asis
show_code("vig_eq_vol")
```

```{r}
safe_tar_read("vig_eq_vol")
```

::: {.fig-caption}
**21-day rolling realised volatility (annualised).** Y-axis: vol (%). Colours: ticker. Vol = SD(log returns) x sqrt(252). SPY = S&P 500 ETF. Source: `hd_ohlcv()`. See FAANG tab for returns. [Realised volatility](https://www.investopedia.com/terms/r/realizedvolatility.asp).
:::

## Coverage

Click column headers to sort. Type in filter boxes to search.

```{r}
#| results: asis
show_code("vig_eq_coverage")
```

```{r}
df <- safe_tar_read("vig_eq_coverage")
if (!is.null(df)) {
  top <- df[1, ]
  hd_dt(df, paste0(
    "Equity: ", nrow(df), " tickers, daily OHLCV. ",
    "Longest: ", top$Ticker, " (", top$`Trading Days`, " days). ",
    "Source: Yahoo Finance via hd_ohlcv(). See AAPL MA tab."
  ))
}
```

:::

# Crypto

::: {.panel-tabset}

## Major Coins

BTC, ETH, SOL, BNB daily close on log scale since 2022. Log scale: equal vertical distances = equal percentage changes.

```{r}
#| results: asis
show_code("vig_cr_major")
```

```{r}
safe_tar_read("vig_cr_major")
```

::: {.fig-caption}
**BTC, ETH, SOL, BNB daily close on log scale.** Y-axis: USD (log10). Colours: token. Source: Yahoo Finance via `hd_ohlcv()`. See Stablecoins tab for peg analysis.
:::

## Stablecoin Peg

USDC and USDT should trade at $1.00. Deviations = [depeg](https://www.investopedia.com/terms/d/depeg.asp) events. Y-axis zoomed to $0.97–$1.03.

```{r}
#| results: asis
show_code("vig_cr_stable")
```

```{r}
safe_tar_read("vig_cr_stable")
```

::: {.fig-caption}
**USDC and USDT daily close vs $1.00 peg.** Y-axis: USD, zoomed $0.97–$1.03. Dashed = $1.00. Source: `hd_ohlcv()`. See Major Coins tab.
:::

## Correlation

Pairwise Pearson correlation of daily log returns for 6 crypto assets since 2023. Correlation >0.7 = limited diversification.

```{r}
#| results: asis
show_code("vig_cr_corr")
```

```{r}
safe_tar_read("vig_cr_corr")
```

::: {.fig-caption}
**Pairwise Pearson correlation of daily log returns, 6 tokens, 2023+.** Fill: correlation 0–1. High (>0.7) = limited diversification. Source: `hd_ohlcv()`. See Major Coins tab.
:::

## Coverage

```{r}
#| results: asis
show_code("vig_cr_coverage")
```

```{r}
df <- safe_tar_read("vig_cr_coverage")
if (!is.null(df)) {
  top <- df[1, ]
  hd_dt(df, paste0(
    "Crypto: ", nrow(df), " tokens. Longest: ", top$Token,
    " (", top$Days, " days). Source: hd_ohlcv()."
  ))
}
```

:::

# Macro

::: {.panel-tabset}

## Interest Rates

US Treasury yields (2Y, 10Y, 30Y) + Fed Funds rate. DFF = policy rate; Treasury yields = market expectations.

```{r}
#| results: asis
show_code("vig_ma_rates")
```

```{r}
safe_tar_read("vig_ma_rates")
```

::: {.fig-caption}
**US constant-maturity Treasury yields + Fed Funds rate (DFF).** Y-axis: yield (%). DGS = daily constant-maturity. Source: FRED via `hd_macro()`. See Yield Curve tab.
:::

## Yield Curve

10Y–2Y spread: recession indicator. Negative = [inverted](https://www.investopedia.com/terms/i/invertedyieldcurve.asp), historically precedes recessions by 6–18 months. Red shading = inversion.

```{r}
#| results: asis
show_code("vig_ma_yc")
```

```{r}
safe_tar_read("vig_ma_yc")
```

::: {.fig-caption}
**10Y minus 2Y Treasury spread.** Y-axis: spread (pp). Red shading = inverted. Source: FRED T10Y2Y via `hd_macro()`. See Interest Rates tab.
:::

## Credit Spreads

ICE BofA option-adjusted spreads: HY (High Yield) and BBB (lowest investment grade). Wider = more default risk.

```{r}
#| results: asis
show_code("vig_ma_spreads")
```

```{r}
safe_tar_read("vig_ma_spreads")
```

::: {.fig-caption}
**ICE BofA OAS over Treasuries.** Y-axis: spread (pp). HY = junk bonds. BBB = investment grade. OAS = option-adjusted spread. Source: FRED via `hd_macro()`.
:::

## All Series

```{r}
#| results: asis
show_code("vig_ma_coverage")
```

```{r}
df <- safe_tar_read("vig_ma_coverage")
if (!is.null(df)) {
  top <- df |> dplyr::slice_max(Obs, n = 1)
  hd_dt(df, paste0(
    "FRED: ", nrow(df), " series. Most obs: ", top$Series,
    " (", format(top$Obs, big.mark = ","), "). Source: hd_macro()."
  ))
}
```

:::

# Factors

::: {.panel-tabset}

## FF3 Daily

[Fama-French 3-factor model](https://www.investopedia.com/terms/f/famaandfrenchthreefactormodel.asp): Mkt-RF (market), SMB (size), HML (value), RF (risk-free). Daily since 2020.

```{r}
#| results: asis
show_code("vig_fa_ff3")
```

```{r}
#| fig-height: 7
safe_tar_read("vig_fa_ff3")
```

::: {.fig-caption}
**FF3 daily returns (2020+).** Y-axis: return (%). Faceted. Mkt-RF = market excess. SMB = small minus big. HML = high minus low B/M. RF = T-bill. Source: Ken French via `hd_factors()`. See FF5 tab.
:::

## FF5 Cumulative

5-factor model cumulative returns since 2000. Adds RMW (profitability) and CMA (investment). Mkt-RF dominates; HML underperformed post-2010.

```{r}
#| results: asis
show_code("vig_fa_ff5")
```

```{r}
safe_tar_read("vig_fa_ff5")
```

::: {.fig-caption}
**FF5 cumulative factor returns (2000–2026).** Y-axis: cum return (%). RMW = Robust Minus Weak. CMA = Conservative Minus Aggressive. Source: Ken French via `hd_factors()`. See [Momentum](https://www.investopedia.com/terms/m/momentum.asp) tab.
:::

## Momentum

[Momentum](https://www.investopedia.com/terms/m/momentum.asp) (WML = Winners Minus Losers). Catastrophic March 2009 crash, post-2020 recovery. Risk of factor concentration.

```{r}
#| results: asis
show_code("vig_fa_mom")
```

```{r}
safe_tar_read("vig_fa_mom")
```

::: {.fig-caption}
**Momentum (WML) cumulative return (2000–2026).** Y-axis: cum return (%). WML = long winners, short losers. Note 2009 reversal. Source: Ken French via `hd_factors()`. See FF5 tab.
:::

## Summary

```{r}
#| results: asis
show_code("vig_fa_coverage")
```

```{r}
df <- safe_tar_read("vig_fa_coverage")
if (!is.null(df)) {
  hd_dt(df, paste0(
    dplyr::n_distinct(df$Factor), " factors, ",
    dplyr::n_distinct(df$Dataset), " datasets. ",
    "Total: ", format(sum(df$Obs), big.mark = ","), " obs from ", min(df$From),
    ". Source: Ken French via hd_factors()."
  ))
}
```

:::
