---
title: "historicaldata"
format:
  dashboard:
    orientation: rows
    theme: cosmo
    embed-resources: true
    include-in-header:
      - text: |
          <style>
          body { font-size: 1.15em; }
          .fullscreen-overlay {
            position: fixed !important;
            top: 0 !important; left: 0 !important;
            width: 100vw !important; height: 100vh !important;
            z-index: 9999 !important;
            background: rgba(0,0,0,0.95) !important;
            padding: 10px !important;
            overflow: auto !important;
            cursor: zoom-out !important;
          }
          .fullscreen-overlay img {
            width: 100% !important; height: 100% !important;
            object-fit: contain !important;
          }
          .cell-output-display:has(img) { cursor: zoom-in; }
          .fig-caption {
            font-size: 0.85em; color: #888; margin-top: 0.4em;
            text-align: left; line-height: 1.5;
          }
          details { margin-bottom: 0.5em; }
          details summary { cursor: pointer; color: #6c9bd2; font-size: 0.9em; }
          </style>
          <script>
          document.addEventListener('click', function(e) {
            var overlay = e.target.closest('.fullscreen-overlay');
            if (overlay) { overlay.classList.remove('fullscreen-overlay'); return; }
            var cell = e.target.closest('.cell-output-display');
            if (!cell || !cell.querySelector('img')) return;
            cell.classList.add('fullscreen-overlay');
          });
          </script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)
```

# Installation

## Row

### {.tabset}

#### R (pak)

```r
# Install the historicaldata package from GitHub
pak::pak("JohnGavin/historical", subdir = "packages/historicaldata")

# Load all required libraries
library(historicaldata)
library(dplyr)
library(ggplot2)
library(scales)
library(purrr)

# Set plot theme (gray70 background, used by all examples)
theme_set(theme_minimal(base_size = 14) + theme(
  plot.background = element_rect(fill = "gray70", colour = NA),
  panel.background = element_rect(fill = "gray70", colour = NA),
  text = element_text(colour = "grey20"),
  axis.text = element_text(colour = "grey30"),
  panel.grid.minor = element_blank(),
  panel.grid.major = element_line(colour = "grey60"),
  legend.position = "bottom"
))
```

#### Nix

```bash
# Enter the reproducible Nix shell (all R + Python packages pre-installed)
nix develop github:JohnGavin/historical
```

Then in R:

```r
# Load the package from source (all dependencies are in the Nix shell)
pkgload::load_all("packages/historicaldata")
library(dplyr)
library(ggplot2)
library(scales)
library(purrr)

# Set plot theme (same as R (pak) tab)
theme_set(theme_minimal(base_size = 14) + theme(
  plot.background = element_rect(fill = "gray70", colour = NA),
  panel.background = element_rect(fill = "gray70", colour = NA),
  text = element_text(colour = "grey20"),
  axis.text = element_text(colour = "grey30"),
  panel.grid.minor = element_blank(),
  panel.grid.major = element_line(colour = "grey60"),
  legend.position = "bottom"
))
```

#### T language

```bash
# Clone, enter Nix shell, run the full data pipeline
git clone https://github.com/JohnGavin/historical.git && cd historical
nix develop
python scripts/fetch_equity.py     # fetch 51 equity tickers
Rscript scripts/fetch_crypto.R     # fetch 14 crypto tokens
Rscript scripts/fetch_macro.R      # fetch 19 FRED series
Rscript scripts/fetch_factors.R    # fetch Fama-French factors
t run src/pipeline.t               # run the T pipeline
```

Then in R (same as Nix tab above).


# Equity

## Row

### {.tabset}

#### AAPL Moving Averages

AAPL daily close with 50d and 200d simple moving averages. The [golden cross](https://www.investopedia.com/terms/g/goldencross.asp) (50d crosses above 200d) and death cross (50d crosses below) are classic trend-following signals.

```{r}
#| results: asis
show_code("vig_eq_aapl")
```

```{r}
safe_tar_read("vig_eq_aapl")
```

::: {.fig-caption}
**AAPL close + 50d/200d SMA.** Y: USD. Grey = close, blue dashed = 50d, red dashed = 200d. SMA = simple moving average. Source: Yahoo Finance via `hd_ohlcv()`. See Volatility tab. [Golden cross](https://www.investopedia.com/terms/g/goldencross.asp). Click to zoom.
:::

#### FAANG Returns

FAANG = Meta (META), Apple (AAPL), Amazon (AMZN), Netflix (NFLX), Alphabet (GOOGL). Cumulative returns rebased to 0% on 2024-01-01.

```{r}
#| results: asis
show_code("vig_eq_faang")
```

```{r}
safe_tar_read("vig_eq_faang")
```

::: {.fig-caption}
**FAANG cumulative returns from 2024-01-01.** Y: cumulative return (%). Colours: ticker. Split-adjusted close. Source: `hd_ohlcv()`. See AAPL tab for single-stock detail.
:::

#### Realised Volatility

21-day rolling annualised vol (SD of log returns x sqrt(252)). SPY = S&P 500 ETF benchmark.

```{r}
#| results: asis
show_code("vig_eq_vol")
```

```{r}
safe_tar_read("vig_eq_vol")
```

::: {.fig-caption}
**21d realised volatility (annualised).** Y: vol (%). Colours: ticker. Vol = SD(log ret) x sqrt(252). SPY = benchmark. Source: `hd_ohlcv()`. See FAANG tab. [Realised vol](https://www.investopedia.com/terms/r/realizedvolatility.asp).
:::

#### Coverage

```{r}
#| results: asis
show_code("vig_eq_coverage")
```

```{r}
df <- safe_tar_read("vig_eq_coverage")
if (!is.null(df)) {
  top <- df[1, ]
  hd_dt(df, paste0(
    nrow(df), " tickers, daily OHLCV. Longest: ", top$Ticker,
    " (", top$`Trading Days`, " days). Source: hd_ohlcv()."
  ))
}
```


# Crypto

## Row

### {.tabset}

#### Major Coins

BTC, ETH, SOL, BNB daily close on log scale since 2022.

```{r}
#| results: asis
show_code("vig_cr_major")
```

```{r}
safe_tar_read("vig_cr_major")
```

::: {.fig-caption}
**BTC, ETH, SOL, BNB log-scale close.** Y: USD (log10). Equal vertical distance = equal % change. Source: `hd_ohlcv()`. See Stablecoins tab.
:::

#### Stablecoin Peg

USDC/USDT vs $1.00 peg. Y-axis zoomed to $0.97–$1.03. [Depeg](https://www.investopedia.com/terms/d/depeg.asp) = liquidity crisis signal.

```{r}
#| results: asis
show_code("vig_cr_stable")
```

```{r}
safe_tar_read("vig_cr_stable")
```

::: {.fig-caption}
**USDC/USDT vs $1.00 peg.** Y: USD, zoomed $0.97–$1.03. Dashed = peg. Source: `hd_ohlcv()`. See Major Coins tab.
:::

#### Correlation

Pearson correlation of daily log returns, 6 tokens, 2023+. >0.7 = limited diversification.

```{r}
#| results: asis
show_code("vig_cr_corr")
```

```{r}
safe_tar_read("vig_cr_corr")
```

::: {.fig-caption}
**Pairwise log-return correlation, 6 tokens, 2023+.** Fill: correlation 0–1. Source: `hd_ohlcv()`.
:::

#### Coverage

```{r}
#| results: asis
show_code("vig_cr_coverage")
```

```{r}
df <- safe_tar_read("vig_cr_coverage")
if (!is.null(df)) {
  top <- df[1, ]
  hd_dt(df, paste0(
    nrow(df), " tokens. Longest: ", top$Token,
    " (", top$Days, " days). Source: hd_ohlcv()."
  ))
}
```


# Macro

## Row

### {.tabset}

#### Interest Rates

Treasury yields (2Y, 10Y, 30Y) + Fed Funds rate (DFF) since 2020.

```{r}
#| results: asis
show_code("vig_ma_rates")
```

```{r}
safe_tar_read("vig_ma_rates")
```

::: {.fig-caption}
**Treasury yields + Fed Funds rate.** Y: yield (%). DGS = constant-maturity. DFF = Fed Funds. Source: FRED via `hd_macro()`. See Yield Curve tab.
:::

#### Yield Curve

10Y–2Y spread. Negative = [inverted](https://www.investopedia.com/terms/i/invertedyieldcurve.asp), precedes recessions by 6–18 months.

```{r}
#| results: asis
show_code("vig_ma_yc")
```

```{r}
safe_tar_read("vig_ma_yc")
```

::: {.fig-caption}
**10Y–2Y spread with inversion shading.** Y: spread (pp). Red = inverted. Source: FRED T10Y2Y via `hd_macro()`. See Rates tab.
:::

#### Credit Spreads

ICE BofA OAS: HY (junk) + BBB (investment grade). Wider = more default risk.

```{r}
#| results: asis
show_code("vig_ma_spreads")
```

```{r}
safe_tar_read("vig_ma_spreads")
```

::: {.fig-caption}
**ICE BofA OAS over Treasuries.** Y: spread (pp). HY = High Yield. BBB = investment grade. Source: FRED via `hd_macro()`.
:::

#### All Series

```{r}
#| results: asis
show_code("vig_ma_coverage")
```

```{r}
df <- safe_tar_read("vig_ma_coverage")
if (!is.null(df)) {
  top <- df |> dplyr::slice_max(Obs, n = 1)
  hd_dt(df, paste0(
    nrow(df), " series. Most obs: ", top$Series,
    " (", format(top$Obs, big.mark = ","), "). Source: hd_macro()."
  ))
}
```


# Factors

## Row

### {.tabset}

#### FF3 Daily

[Fama-French 3 factors](https://www.investopedia.com/terms/f/famaandfrenchthreefactormodel.asp): Mkt-RF, SMB, HML, RF. Daily since 2020.

```{r}
#| results: asis
show_code("vig_fa_ff3")
```

```{r}
#| fig-height: 7
safe_tar_read("vig_fa_ff3")
```

::: {.fig-caption}
**FF3 daily returns.** Y: return (%). Faceted. Mkt-RF = market excess. SMB = size. HML = value. RF = T-bill. Source: Ken French via `hd_factors()`. See FF5 tab.
:::

#### FF5 Cumulative

5-factor cumulative returns since 2000. RMW = profitability, CMA = investment.

```{r}
#| results: asis
show_code("vig_fa_ff5")
```

```{r}
safe_tar_read("vig_fa_ff5")
```

::: {.fig-caption}
**FF5 cumulative returns (2000–2026).** Y: cum return (%). RMW = Robust Minus Weak. CMA = Conservative Minus Aggressive. Mkt-RF dominates. HML weak post-2010. Source: Ken French via `hd_factors()`.
:::

#### Momentum

[Momentum](https://www.investopedia.com/terms/m/momentum.asp) (WML). March 2009 crash, post-2020 recovery.

```{r}
#| results: asis
show_code("vig_fa_mom")
```

```{r}
safe_tar_read("vig_fa_mom")
```

::: {.fig-caption}
**Momentum (WML) cumulative return (2000–2026).** WML = long winners, short losers. 2009 = reversal crash. Source: Ken French via `hd_factors()`.
:::

#### Summary

```{r}
#| results: asis
show_code("vig_fa_coverage")
```

```{r}
df <- safe_tar_read("vig_fa_coverage")
if (!is.null(df)) {
  hd_dt(df, paste0(
    dplyr::n_distinct(df$Factor), " factors, ",
    dplyr::n_distinct(df$Dataset), " datasets. Total: ",
    format(sum(df$Obs), big.mark = ","), " obs. Source: hd_factors()."
  ))
}
```
