---
title: "historicaldata: Usage Examples"
format:
  html:
    theme: darkly
    embed-resources: true
    toc: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)

# Quarto renders from docs/ so _targets is local, but here::here() points to project root
# Use _targets in current dir if it exists, otherwise try project root
store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)
```

::: {.callout-note appearance="minimal"}
Every code block below is **copy-pasteable R** that reproduces the output. Install: `pak::pak("JohnGavin/historical", subdir = "packages/historicaldata")`. Click "Show code" to reveal.
:::

# Equity

::: {.panel-tabset}


## AAPL Moving Averages

AAPL daily close with 50-day and 200-day simple moving averages. The golden cross (50d crosses above 200d) and death cross (50d crosses below) are classic trend-following signals.

```{r}
#| results: asis
show_code("vig_eq_aapl")
```

```{r}
safe_tar_read("vig_eq_aapl")
```

## FAANG Returns

FAANG = Meta (META), Apple (AAPL), Amazon (AMZN), Netflix (NFLX), Alphabet (GOOGL). Cumulative returns rebased to 0% on 2024-01-01 show relative performance divergence among the five largest US tech companies.

```{r}
#| results: asis
show_code("vig_eq_faang")
```

```{r}
safe_tar_read("vig_eq_faang")
```

## Realised Volatility

21-day rolling annualised volatility (standard deviation of log returns, scaled by sqrt(252)). Higher values indicate greater price uncertainty. SPY (S&P 500 ETF) serves as the benchmark.

```{r}
#| results: asis
show_code("vig_eq_vol")
```

```{r}
safe_tar_read("vig_eq_vol")
```

## Coverage

All 51 equity tickers with trading day count and date range. Click column headers to sort. Type in filter boxes to search.

```{r}
#| results: asis
show_code("vig_eq_coverage")
```

```{r}
df <- safe_tar_read("vig_eq_coverage")
top <- df[1, ]
hd_dt(df, paste0(
  "Equity coverage: ", nrow(df), " tickers, daily OHLCV. ",
  "Columns: ticker symbol, trading days, first/last date. ",
  "Longest history: ", top$Ticker, " (", top$`Trading Days`, " days). ",
  "Source: Yahoo Finance via historicaldata::hd_ohlcv(). ",
  "See AAPL MA tab for single-ticker example."
))
```

:::

# Crypto

::: {.panel-tabset}


## Major Coins

BTC, ETH, SOL, and BNB daily close on log scale since 2022. Log scale reveals percentage moves — equal vertical distances represent equal percentage changes regardless of price level.

```{r}
#| results: asis
show_code("vig_cr_major")
```

```{r}
safe_tar_read("vig_cr_major")
```

## Stablecoin Peg

USDC and USDT should trade at $1.00. Deviations indicate depeg events — potential liquidity crises. Y-axis zoomed to $0.97–$1.03 to make small deviations visible.

```{r}
#| results: asis
show_code("vig_cr_stable")
```

```{r}
safe_tar_read("vig_cr_stable")
```

## Correlation

Pairwise correlation of daily log returns for 6 major crypto assets since 2023. High correlations (>0.7) indicate assets move together — less diversification benefit.

```{r}
#| results: asis
show_code("vig_cr_corr")
```

```{r}
safe_tar_read("vig_cr_corr")
```

## Coverage

```{r}
#| results: asis
show_code("vig_cr_coverage")
```

```{r}
df <- safe_tar_read("vig_cr_coverage")
top <- df[1, ]
hd_dt(df, paste0(
  "Crypto coverage: ", nrow(df), " tokens, daily OHLCV. ",
  "Columns: token, trading days, first/last date. ",
  "Longest: ", top$Token, " (", top$Days, " days). ",
  "Source: Yahoo Finance via historicaldata::hd_ohlcv(). ",
  "See Major Coins tab for price chart."
))
```

:::

# Macro

::: {.panel-tabset}


## Interest Rates

US Treasury yields (2Y, 10Y, 30Y) and the effective Federal Funds rate since 2020. The Fed Funds rate (DFF) is the policy rate; Treasury yields reflect market expectations of future rates and inflation.

```{r}
#| results: asis
show_code("vig_ma_rates")
```

```{r}
safe_tar_read("vig_ma_rates")
```

## Yield Curve

The 10Y–2Y Treasury spread is a recession indicator. When negative (inverted), short-term rates exceed long-term rates — historically preceding recessions by 6–18 months. Red shading marks the inversion period.

```{r}
#| results: asis
show_code("vig_ma_yc")
```

```{r}
safe_tar_read("vig_ma_yc")
```

## Credit Spreads

ICE BofA option-adjusted spreads for High Yield (HY) and BBB-rated corporate bonds. Wider spreads indicate higher perceived default risk. Spikes correspond to market stress events (e.g. COVID March 2020).

```{r}
#| results: asis
show_code("vig_ma_spreads")
```

```{r}
safe_tar_read("vig_ma_spreads")
```

## All Series

```{r}
#| results: asis
show_code("vig_ma_coverage")
```

```{r}
df <- safe_tar_read("vig_ma_coverage")
top <- df |> dplyr::slice_max(Obs, n = 1)
hd_dt(df, paste0(
  "FRED macro: ", nrow(df), " series at mixed frequencies (daily/monthly/quarterly). ",
  "Columns: FRED series ID, observation count, first/last date. ",
  "Most observations: ", top$Series, " (", format(top$Obs, big.mark = ","), "). ",
  "Source: FRED via historicaldata::hd_macro(). ",
  "See Interest Rates tab for yield chart."
))
```

:::

# Factors

::: {.panel-tabset}


## FF3 Daily

The Fama-French 3-factor model decomposes equity returns into: Market excess return (Mkt-RF), Size (SMB = Small Minus Big), and Value (HML = High Minus Low book-to-market). RF = risk-free rate. Daily returns since 2020.

```{r}
#| results: asis
show_code("vig_fa_ff3")
```

```{r}
#| fig-height: 7
safe_tar_read("vig_fa_ff3")
```

## FF5 Cumulative

Cumulative returns for the 5-factor model since 2000. Adds RMW (Robust Minus Weak profitability) and CMA (Conservative Minus Aggressive investment). Shows which factors delivered positive long-run premia: Mkt-RF dominates; HML (value) struggled post-2010.

```{r}
#| results: asis
show_code("vig_fa_ff5")
```

```{r}
safe_tar_read("vig_fa_ff5")
```

## Momentum

Cumulative return of the momentum factor (WML = Winners Minus Losers). Momentum suffered a catastrophic crash in March 2009 (reversal) but recovered over the following decade. Demonstrates the risk of factor concentration.

```{r}
#| results: asis
show_code("vig_fa_mom")
```

```{r}
safe_tar_read("vig_fa_mom")
```

## Summary

```{r}
#| results: asis
show_code("vig_fa_coverage")
```

```{r}
df <- safe_tar_read("vig_fa_coverage")
hd_dt(df, paste0(
  "Fama-French factors: ", dplyr::n_distinct(df$Factor), " factors across ",
  dplyr::n_distinct(df$Dataset), " datasets. ",
  "Columns: dataset (FF3/FF5/Mom), frequency, factor name, observation count, date range. ",
  "Daily data back to ", min(df$From), ". Total: ",
  format(sum(df$Obs), big.mark = ","), " observations. ",
  "Source: Ken French Data Library via historicaldata::hd_factors(). ",
  "See FF3 tab for daily returns."
))
```

:::

