---
title: "historicaldata: Usage Examples"
format:
  dashboard:
    orientation: rows
    theme: darkly
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)

source(here::here("docs/vignette_utils.R"))

# Dark theme for all plots
theme_set(theme_minimal(base_size = 12) + theme(
  plot.background = element_rect(fill = "grey10", colour = NA),
  panel.background = element_rect(fill = "grey10", colour = NA),
  text = element_text(colour = "grey80"),
  axis.text = element_text(colour = "grey70"),
  panel.grid.minor = element_blank(),
  panel.grid.major = element_line(colour = "grey25"),
  legend.position = "bottom",
  plot.caption = element_text(size = 8, colour = "grey60", hjust = 0)
))
```

::: {.callout-note appearance="minimal"}
All data is fetched live from [Hugging Face](https://huggingface.co/datasets/dsfefvx/finance-historical-data) via `historicaldata` package functions: `hd_ohlcv()`, `hd_macro()`, `hd_factors()`. Install with: `pak::pak("JohnGavin/historical", subdir = "packages/historicaldata")`.
:::

# Equity

## Row

### {.tabset}

#### AAPL Moving Averages

AAPL daily close with 50-day and 200-day simple moving averages. The golden cross (50d crosses above 200d) and death cross (50d crosses below) are classic trend-following signals.

```{r}
#| code-fold: true
#| code-summary: "Generating code: hd_ohlcv('AAPL') + moving averages"
# library(historicaldata)
# aapl <- hd_ohlcv("AAPL", from = "2023-01-01") |>
#   arrange(date) |>
#   mutate(
#     cs = cumsum(close),
#     ma_50  = if_else(row_number() >= 50, (cs - lag(cs, 50)) / 50, NA_real_),
#     ma_200 = if_else(row_number() >= 200, (cs - lag(cs, 200)) / 200, NA_real_)
#   )
```

```{r}
safe_tar_read("vig_eq_aapl_plot")
```

#### FAANG Returns

FAANG = Meta (META), Apple (AAPL), Amazon (AMZN), Netflix (NFLX), Alphabet (GOOGL). Cumulative returns rebased to 0% on 2024-01-01. Divergence shows which mega-cap tech stocks outperformed.

```{r}
#| code-fold: true
#| code-summary: "Generating code: hd_ohlcv() for 5 tickers + cumulative returns"
# library(historicaldata)
# faang <- bind_rows(lapply(
#   c("AAPL", "AMZN", "GOOGL", "META", "NFLX"),
#   \(t) hd_ohlcv(t, from = "2024-01-01")
# )) |>
#   group_by(ticker) |>
#   mutate(cum_ret = adjusted / first(adjusted) - 1) |>
#   ungroup()
```

```{r}
safe_tar_read("vig_eq_faang_plot")
```

#### Realised Volatility

21-day rolling annualised volatility (SD of log returns x sqrt(252)). Higher values = greater price uncertainty. SPY (S&P 500 ETF) is the benchmark.

```{r}
#| code-fold: true
#| code-summary: "Generating code: hd_ohlcv() + rolling volatility"
# library(historicaldata)
# vol <- bind_rows(lapply(
#   c("AAPL", "NVDA", "TSLA", "SPY"),
#   \(t) hd_ohlcv(t, from = "2023-06-01")
# )) |>
#   group_by(ticker) |>
#   arrange(date) |>
#   mutate(log_ret = log(adjusted / lag(adjusted)),
#          vol_21d = ...) # rolling SD
```

```{r}
safe_tar_read("vig_eq_vol_plot")
```

#### Coverage

All 51 equity tickers. Click column headers to sort. Filter by typing in the search boxes below headers.

```{r}
hd_dt(safe_tar_read("vig_eq_coverage"),
  "Equity: 51 tickers, daily OHLCV. Source: Yahoo Finance via hd_ohlcv(). Click headers to sort.")
```


# Crypto

## Row

### {.tabset}

#### Major Coins

BTC, ETH, SOL, BNB daily close on log scale since 2022. Log scale reveals percentage moves — equal vertical distances represent equal percentage changes regardless of price level.

```{r}
#| code-fold: true
#| code-summary: "Generating code: hd_ohlcv() for 4 crypto tickers"
# library(historicaldata)
# major <- bind_rows(lapply(
#   c("BTC", "ETH", "SOL", "BNB"),
#   \(t) hd_ohlcv(t, from = "2022-01-01")
# ))
```

```{r}
safe_tar_read("vig_cr_major_plot")
```

#### Stablecoin Peg

USDC and USDT should trade at \$1.00. Deviations indicate depeg events — potential liquidity crises. Y-axis zoomed to \$0.97–\$1.03 to make small deviations visible.

```{r}
#| code-fold: true
#| code-summary: "Generating code: hd_ohlcv('USDC') + hd_ohlcv('USDT')"
# library(historicaldata)
# stable <- bind_rows(
#   hd_ohlcv("USDC", from = "2022-01-01"),
#   hd_ohlcv("USDT", from = "2022-01-01")
# )
```

```{r}
safe_tar_read("vig_cr_stable_plot")
```

#### Correlation

Pairwise correlation of daily log returns for 6 major crypto assets since 2023. High correlations (>0.7) mean assets move together — less diversification benefit.

```{r}
#| code-fold: true
#| code-summary: "Generating code: hd_ohlcv() for 6 tokens + cor()"
# library(historicaldata)
# wide <- bind_rows(lapply(
#   c("BTC","ETH","SOL","BNB","ADA","XRP"),
#   \(t) hd_ohlcv(t, from = "2023-01-01")
# )) |> ... |> pivot_wider() |> cor()
```

```{r}
safe_tar_read("vig_cr_corr_plot")
```

#### Coverage

```{r}
hd_dt(safe_tar_read("vig_cr_coverage"),
  "Crypto: 14 tokens, daily OHLCV. Source: Yahoo Finance via hd_ohlcv(). Click headers to sort.")
```


# Macro

## Row

### {.tabset}

#### Interest Rates

US Treasury yields (2Y, 10Y, 30Y) and the effective Federal Funds rate since 2020. DFF is the policy rate set by the Fed; Treasury yields reflect market expectations of future rates and inflation.

```{r}
#| code-fold: true
#| code-summary: "Generating code: hd_macro() for DGS2, DGS10, DGS30, DFF"
# library(historicaldata)
# rates <- bind_rows(lapply(
#   c("DGS2", "DGS10", "DGS30", "DFF"),
#   \(s) hd_macro(s, from = "2020-01-01")
# ))
```

```{r}
safe_tar_read("vig_ma_rates_plot")
```

#### Yield Curve

The 10Y–2Y Treasury spread is a recession indicator. When negative (inverted), short-term rates exceed long-term rates — this has historically preceded recessions by 6–18 months. Red shading marks the inversion period.

```{r}
#| code-fold: true
#| code-summary: "Generating code: hd_macro('T10Y2Y')"
# library(historicaldata)
# yc <- hd_macro("T10Y2Y", from = "2018-01-01") |>
#   filter(!is.na(value))
```

```{r}
safe_tar_read("vig_ma_yc_plot")
```

#### Credit Spreads

ICE BofA option-adjusted spreads for High Yield (HY) and BBB-rated corporate bonds. Wider spreads indicate higher perceived default risk. Spikes correspond to market stress events (e.g. COVID March 2020).

```{r}
#| code-fold: true
#| code-summary: "Generating code: hd_macro('BAMLH0A0HYM2') + hd_macro('BAMLC0A4CBBB')"
# library(historicaldata)
# spreads <- bind_rows(
#   hd_macro("BAMLH0A0HYM2", from = "2020-01-01"),
#   hd_macro("BAMLC0A4CBBB", from = "2020-01-01")
# )
```

```{r}
safe_tar_read("vig_ma_spreads_plot")
```

#### All Series

```{r}
hd_dt(safe_tar_read("vig_ma_coverage"),
  "FRED macro: 19 series (daily/monthly/quarterly). Source: FRED via hd_macro(). Click headers to sort.")
```


# Factors

## Row

### {.tabset}

#### FF3 Daily

The Fama-French 3-factor model decomposes equity returns into: Market excess return (Mkt-RF), Size (SMB = Small Minus Big), and Value (HML = High Minus Low book-to-market). RF = risk-free rate. Daily returns since 2020.

```{r}
#| code-fold: true
#| code-summary: "Generating code: hd_factors('FF3', 'daily')"
# library(historicaldata)
# ff3 <- hd_factors("FF3", "daily", from = "2020-01-01")
```

```{r}
safe_tar_read("vig_fa_ff3_plot")
```

#### FF5 Cumulative

Cumulative returns for the 5-factor model since 2000. Adds RMW (Robust Minus Weak profitability) and CMA (Conservative Minus Aggressive investment) to FF3. Shows which factors delivered positive long-run premia: Mkt-RF dominates; HML (value) struggled post-2010.

```{r}
#| code-fold: true
#| code-summary: "Generating code: hd_factors('FF5', 'daily') + cumprod()"
# library(historicaldata)
# ff5 <- hd_factors("FF5", "daily", from = "2000-01-01") |>
#   filter(factor_name != "RF") |>
#   group_by(factor_name) |>
#   mutate(cum_ret = cumprod(1 + value/100) - 1)
```

```{r}
safe_tar_read("vig_fa_ff5_plot")
```

#### Momentum

Cumulative return of the momentum factor (WML = Winners Minus Losers). Momentum suffered a catastrophic crash in March 2009 (reversal) but recovered over the following decade. Demonstrates the risk of factor concentration — even well-documented premia can experience multi-year drawdowns.

```{r}
#| code-fold: true
#| code-summary: "Generating code: hd_factors('Mom', 'daily') + cumprod()"
# library(historicaldata)
# mom <- hd_factors("Mom", "daily", from = "2000-01-01") |>
#   mutate(cum_ret = cumprod(1 + value/100) - 1)
```

```{r}
safe_tar_read("vig_fa_mom_plot")
```

#### Summary

```{r}
hd_dt(safe_tar_read("vig_fa_coverage"),
  "Fama-French factors: FF3, FF5, Momentum. Daily + monthly, 1926-2026. Source: Ken French via hd_factors().")
```
