---
title: "historicaldata: Usage Examples"
format:
  html:
    theme: darkly
    embed-resources: true
    toc: false
    include-in-header:
      - text: |
          <style>
          /* Click-to-zoom: plots expand to fullscreen */
          .fullscreen-overlay {
            position: fixed !important;
            top: 0 !important; left: 0 !important;
            width: 100vw !important; height: 100vh !important;
            z-index: 9999 !important;
            background: rgba(0,0,0,0.95) !important;
            padding: 20px !important;
            overflow: auto !important;
            cursor: zoom-out !important;
            display: flex; align-items: center; justify-content: center;
          }
          .fullscreen-overlay img { max-width: 95vw; max-height: 95vh; object-fit: contain; }
          .cell-output-display:has(img) { cursor: zoom-in; }
          /* Captions */
          figcaption, .fig-caption {
            font-size: 0.85em; color: #aaa; margin-top: 0.3em;
            text-align: left; line-height: 1.4;
          }
          </style>
          <script>
          document.addEventListener('click', function(e) {
            var overlay = e.target.closest('.fullscreen-overlay');
            if (overlay) { overlay.classList.remove('fullscreen-overlay'); return; }
            var cell = e.target.closest('.cell-output-display');
            if (!cell || !cell.querySelector('img')) return;
            cell.classList.add('fullscreen-overlay');
          });
          </script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)
```

# Installation {.panel-tabset}

## R (pak)

```r
pak::pak("JohnGavin/historical", subdir = "packages/historicaldata")
library(historicaldata)
```

## Nix

```bash
nix develop github:JohnGavin/historical
# R packages available: dplyr, arrow, duckdb, duckplyr, ggplot2, ...
Rscript -e 'pkgload::load_all("packages/historicaldata")'
```

## T language

```bash
# Clone and enter the T pipeline environment
git clone https://github.com/JohnGavin/historical.git
cd historical
nix develop
t run src/pipeline.t
```

#

---

# Equity {.panel-tabset}

## AAPL Moving Averages

AAPL daily close with 50-day and 200-day simple moving averages. The [golden cross](https://www.investopedia.com/terms/g/goldencross.asp) (50d crosses above 200d) and death cross (50d crosses below) are classic trend-following signals.

```{r}
#| results: asis
show_code("vig_eq_aapl")
```

```{r}
safe_tar_read("vig_eq_aapl")
```

::: {.fig-caption}
**AAPL close with 50d and 200d simple moving averages (SMA).** Y-axis: USD. Grey = close, blue dashed = 50d SMA, red dashed = 200d SMA. MA = simple moving average (trailing, equal-weighted). Source: Yahoo Finance via `historicaldata::hd_ohlcv()`. See Volatility tab for risk-adjusted view. Click plot to zoom.
:::

## FAANG Returns

FAANG = Meta (META), Apple (AAPL), Amazon (AMZN), Netflix (NFLX), Alphabet (GOOGL). Cumulative returns rebased to 0% on 2024-01-01 show relative performance divergence among the five largest US tech companies.

```{r}
#| results: asis
show_code("vig_eq_faang")
```

```{r}
safe_tar_read("vig_eq_faang")
```

::: {.fig-caption}
**FAANG cumulative returns from 2024-01-01 (0% baseline).** Y-axis: cumulative return (%). Colours: ticker symbol. FAANG = Meta (META), Apple (AAPL), Amazon (AMZN), Netflix (NFLX), Alphabet (GOOGL). Split-adjusted close prices. Source: Yahoo Finance via `historicaldata::hd_ohlcv()`. See AAPL MA tab for single-stock detail.
:::

## Realised Volatility

21-day rolling annualised volatility (standard deviation of log returns, scaled by sqrt(252)). Higher values indicate greater price uncertainty. SPY (S&P 500 ETF) serves as the benchmark.

```{r}
#| results: asis
show_code("vig_eq_vol")
```

```{r}
safe_tar_read("vig_eq_vol")
```

::: {.fig-caption}
**21-day rolling realised volatility (annualised).** Y-axis: annualised vol (%). Colours: ticker. Vol = SD(log returns) x sqrt(252). SPY = S&P 500 ETF benchmark. Source: `historicaldata::hd_ohlcv()`. See FAANG tab for return comparison. [Realised volatility](https://www.investopedia.com/terms/r/realizedvolatility.asp).
:::

## Coverage

All 51 equity tickers with trading day count and date range. Click column headers to sort. Type in filter boxes to search.

```{r}
#| results: asis
show_code("vig_eq_coverage")
```

```{r}
df <- safe_tar_read("vig_eq_coverage")
if (!is.null(df)) {
  top <- df[1, ]
  hd_dt(df, paste0(
    "Equity coverage: ", nrow(df), " tickers, daily OHLCV. ",
    "Columns: ticker, trading days, first/last date. ",
    "Longest: ", top$Ticker, " (", top$`Trading Days`, " days). ",
    "Source: Yahoo Finance via hd_ohlcv(). See AAPL MA tab."
  ))
}
```


# Crypto {.panel-tabset}

## Major Coins

BTC, ETH, SOL, and BNB daily close on log scale since 2022. Log scale reveals percentage moves — equal vertical distances represent equal percentage changes regardless of price level.

```{r}
#| results: asis
show_code("vig_cr_major")
```

```{r}
safe_tar_read("vig_cr_major")
```

::: {.fig-caption}
**BTC, ETH, SOL, BNB daily close on log scale.** Y-axis: USD (log10). Colours: token symbol. Log scale: equal vertical distances = equal % changes. Source: Yahoo Finance via `historicaldata::hd_ohlcv()`. See Stablecoins tab for peg analysis.
:::

## Stablecoin Peg

USDC and USDT should trade at $1.00. Deviations indicate [depeg](https://www.investopedia.com/terms/d/depeg.asp) events — potential liquidity crises. Y-axis zoomed to $0.97–$1.03 to make small deviations visible.

```{r}
#| results: asis
show_code("vig_cr_stable")
```

```{r}
safe_tar_read("vig_cr_stable")
```

::: {.fig-caption}
**USDC and USDT daily close vs $1.00 peg.** Y-axis: USD, zoomed to $0.97–$1.03. Dashed line = $1.00 peg. Colours: token. Source: Yahoo Finance via `historicaldata::hd_ohlcv()`. See Major Coins tab for price comparison.
:::

## Correlation

Pairwise Pearson correlation of daily log returns for 6 major crypto assets since 2023. High correlations (>0.7) indicate assets move together — less diversification benefit.

```{r}
#| results: asis
show_code("vig_cr_corr")
```

```{r}
safe_tar_read("vig_cr_corr")
```

::: {.fig-caption}
**Pairwise Pearson correlation of daily log returns, 6 tokens, 2023+.** Fill: correlation coefficient 0–1. High correlation (>0.7) = limited diversification benefit. Source: `historicaldata::hd_ohlcv()`. See Major Coins tab.
:::

## Coverage

```{r}
#| results: asis
show_code("vig_cr_coverage")
```

```{r}
df <- safe_tar_read("vig_cr_coverage")
if (!is.null(df)) {
  top <- df[1, ]
  hd_dt(df, paste0(
    "Crypto coverage: ", nrow(df), " tokens, daily OHLCV. ",
    "Columns: token, days, first/last date. ",
    "Longest: ", top$Token, " (", top$Days, " days). ",
    "Source: Yahoo Finance via hd_ohlcv(). See Major Coins tab."
  ))
}
```


# Macro {.panel-tabset}

## Interest Rates

US Treasury yields (2Y, 10Y, 30Y) and the effective Federal Funds rate since 2020. DFF is the policy rate set by the Fed; Treasury yields reflect market expectations of future rates and inflation.

```{r}
#| results: asis
show_code("vig_ma_rates")
```

```{r}
safe_tar_read("vig_ma_rates")
```

::: {.fig-caption}
**US constant-maturity Treasury yields (2Y, 10Y, 30Y) and effective Fed Funds rate (DFF).** Y-axis: yield (%). Colours: FRED series ID. DGS = daily constant-maturity yield. DFF = effective Federal Funds rate. Source: FRED via `historicaldata::hd_macro()`. See Yield Curve tab for inversion analysis.
:::

## Yield Curve

The 10Y–2Y Treasury spread is a recession indicator. When negative (inverted), short-term rates exceed long-term rates — historically preceding recessions by 6–18 months. Red shading marks the [inversion](https://www.investopedia.com/terms/i/invertedyieldcurve.asp) period.

```{r}
#| results: asis
show_code("vig_ma_yc")
```

```{r}
safe_tar_read("vig_ma_yc")
```

::: {.fig-caption}
**10-year minus 2-year US Treasury yield spread.** Y-axis: spread (percentage points). Red line = spread. Dashed = zero. Red shading = yield curve inversion (short rates > long rates). Inversion historically precedes recessions by 6–18 months. Source: FRED T10Y2Y via `historicaldata::hd_macro()`. See Interest Rates tab.
:::

## Credit Spreads

ICE BofA option-adjusted spreads for High Yield (HY) and BBB-rated corporate bonds. Wider spreads indicate higher perceived default risk. Spikes correspond to market stress events (e.g. COVID March 2020).

```{r}
#| results: asis
show_code("vig_ma_spreads")
```

```{r}
safe_tar_read("vig_ma_spreads")
```

::: {.fig-caption}
**ICE BofA option-adjusted spreads (OAS) over US Treasuries.** Y-axis: spread (percentage points). HY = High Yield (junk bonds). BBB = lowest investment-grade rating. OAS = option-adjusted spread, accounts for embedded options. Wider = more default risk. Source: FRED via `historicaldata::hd_macro()`. See Interest Rates tab.
:::

## All Series

```{r}
#| results: asis
show_code("vig_ma_coverage")
```

```{r}
df <- safe_tar_read("vig_ma_coverage")
if (!is.null(df)) {
  top <- df |> dplyr::slice_max(Obs, n = 1)
  hd_dt(df, paste0(
    "FRED macro: ", nrow(df), " series (daily/monthly/quarterly). ",
    "Columns: FRED series ID, obs count, first/last date. ",
    "Most obs: ", top$Series, " (", format(top$Obs, big.mark = ","), "). ",
    "Source: FRED via hd_macro(). See Interest Rates tab."
  ))
}
```


# Factors {.panel-tabset}

## FF3 Daily

The [Fama-French 3-factor model](https://www.investopedia.com/terms/f/famaandfrenchthreefactormodel.asp) decomposes equity returns into: Market excess return (Mkt-RF), Size (SMB = Small Minus Big), and Value (HML = High Minus Low book-to-market). RF = risk-free rate.

```{r}
#| results: asis
show_code("vig_fa_ff3")
```

```{r}
#| fig-height: 7
safe_tar_read("vig_fa_ff3")
```

::: {.fig-caption}
**Fama-French 3-factor model daily returns (2020+).** Y-axis: daily return (%). Faceted by factor. Mkt-RF = market excess return. SMB = small minus big (size). HML = high minus low B/M (value). RF = risk-free rate (1-month T-bill). Source: Ken French Data Library via `historicaldata::hd_factors()`. See FF5 tab.
:::

## FF5 Cumulative

Cumulative returns for the 5-factor model since 2000. Adds RMW (Robust Minus Weak profitability) and CMA (Conservative Minus Aggressive investment). Shows which factors delivered positive long-run premia.

```{r}
#| results: asis
show_code("vig_fa_ff5")
```

```{r}
safe_tar_read("vig_fa_ff5")
```

::: {.fig-caption}
**FF5 cumulative factor returns (2000–2026).** Y-axis: cumulative return (%). Colours: factor name. RF excluded. RMW = Robust Minus Weak (profitability). CMA = Conservative Minus Aggressive (investment). Mkt-RF dominates. HML (value) underperformed post-2010. Source: Ken French via `historicaldata::hd_factors()`. See [Momentum](https://www.investopedia.com/terms/m/momentum.asp) tab.
:::

## Momentum

Cumulative return of the momentum factor (WML = Winners Minus Losers). [Momentum](https://www.investopedia.com/terms/m/momentum.asp) suffered a catastrophic crash in March 2009 (reversal) but recovered over the following decade. Demonstrates the risk of factor concentration.

```{r}
#| results: asis
show_code("vig_fa_mom")
```

```{r}
safe_tar_read("vig_fa_mom")
```

::: {.fig-caption}
**Momentum (WML = Winners Minus Losers) cumulative return (2000–2026).** Y-axis: cumulative return (%). Green line. WML = long past winners, short past losers (12-1 month formation). Note 2009 crash (momentum reversal) and post-2020 recovery. Source: Ken French via `historicaldata::hd_factors()`. See FF5 tab.
:::

## Summary

```{r}
#| results: asis
show_code("vig_fa_coverage")
```

```{r}
df <- safe_tar_read("vig_fa_coverage")
if (!is.null(df)) {
  hd_dt(df, paste0(
    "Fama-French factors: ", dplyr::n_distinct(df$Factor), " factors, ",
    dplyr::n_distinct(df$Dataset), " datasets. ",
    "Cols: dataset, freq, factor, obs, date range. ",
    "Daily data from ", min(df$From), ". Total: ",
    format(sum(df$Obs), big.mark = ","), " obs. ",
    "Source: Ken French via hd_factors(). See FF3 tab."
  ))
}
```
