---
title: "historicaldata: Usage Examples"
format:
  dashboard:
    orientation: rows
    theme: darkly
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(DBI)
library(duckdb)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(DT)

BASE <- "https://huggingface.co/datasets/dsfefvx/finance-historical-data/resolve/main"

con <- dbConnect(duckdb())
invisible(dbExecute(con, "INSTALL httpfs; LOAD httpfs;"))

# Register remote Parquet as views — all subsequent code uses dplyr, not SQL
invisible(dbExecute(con, sprintf(
  "CREATE VIEW equity AS SELECT * FROM read_parquet('%s/equity_daily.parquet')", BASE)))
invisible(dbExecute(con, sprintf(
  "CREATE VIEW crypto AS SELECT * FROM read_parquet('%s/crypto_daily.parquet')", BASE)))
invisible(dbExecute(con, sprintf(
  "CREATE VIEW macro AS SELECT * FROM read_parquet('%s/macro_daily.parquet')", BASE)))
invisible(dbExecute(con, sprintf(
  "CREATE VIEW factors AS SELECT * FROM read_parquet('%s/factors.parquet')", BASE)))

eq  <- tbl(con, "equity")
cr  <- tbl(con, "crypto")
ma  <- tbl(con, "macro")
fa  <- tbl(con, "factors")

theme_set(theme_minimal(base_size = 12) + theme(
  plot.background = element_rect(fill = "grey10", colour = NA),
  panel.background = element_rect(fill = "grey10", colour = NA),
  text = element_text(colour = "grey80"),
  axis.text = element_text(colour = "grey70"),
  panel.grid.minor = element_blank(),
  panel.grid.major = element_line(colour = "grey25"),
  legend.position = "bottom",
  plot.caption = element_text(size = 8, colour = "grey60", hjust = 0)
))

# DT helper: sortable table with caption, compact columns
hd_dt <- function(df, caption_text, ...) {
  DT::datatable(
    df,
    caption = htmltools::tags$caption(
      style = "caption-side: top; font-size: 1.05em; font-weight: bold; color: #ddd;",
      caption_text
    ),
    rownames = FALSE,
    filter = "top",
    options = list(
      pageLength = 15,
      scrollX = TRUE,
      autoWidth = TRUE,
      dom = "frtip"
    ),
    ...
  )
}
```

# Equity

## Row

### {.tabset}

#### AAPL Moving Averages

AAPL daily close with 50-day and 200-day simple moving averages, computed via DuckDB window functions. The golden cross (50d crosses above 200d) and death cross (50d crosses below 200d) are classic trend-following signals.

```{r}
#| code-fold: true
#| code-summary: "Show code"
# DuckDB window functions for moving averages (dplyr can't express ROWS BETWEEN)
aapl <- dbGetQuery(con, sprintf("
  SELECT date, close, ma_50, ma_200 FROM (
    SELECT date, close,
      AVG(close) OVER (ORDER BY date ROWS BETWEEN 49 PRECEDING AND CURRENT ROW) AS ma_50,
      AVG(close) OVER (ORDER BY date ROWS BETWEEN 199 PRECEDING AND CURRENT ROW) AS ma_200
    FROM equity
    WHERE ticker = 'AAPL' AND date >= '2022-01-01'
  ) WHERE date >= '2023-01-01'
  ORDER BY date")) |> as_tibble()
```

```{r}
#| code-fold: true
ggplot(aapl, aes(date)) +
  geom_line(aes(y = close), colour = "white", linewidth = 0.4) +
  geom_line(aes(y = ma_50), colour = "#3498db", linewidth = 0.5, linetype = "dashed") +
  geom_line(aes(y = ma_200), colour = "#e74c3c", linewidth = 0.5, linetype = "dashed") +
  scale_y_continuous(labels = dollar) +
  labs(
    x = NULL, y = "Close (USD)",
    title = "AAPL daily close with 50d and 200d moving averages",
    caption = "White = close. Blue dashed = 50d MA. Red dashed = 200d MA. Source: Yahoo Finance via historicaldata."
  )
```

#### FAANG Returns

FAANG = Facebook (META), Apple (AAPL), Amazon (AMZN), Netflix (NFLX), Google (GOOGL). Cumulative returns since 2024-01-01, rebased to 0%. Shows relative performance divergence among the five largest US tech companies.

```{r}
#| code-fold: true
faang_tickers <- c("AAPL", "AMZN", "GOOGL", "META", "NFLX")

faang <- eq |>
  filter(ticker %in% faang_tickers, date >= "2024-01-01") |>
  select(ticker, date, adjusted) |>
  collect() |>
  group_by(ticker) |>
  mutate(cum_ret = adjusted / first(adjusted) - 1) |>
  ungroup()
```

```{r}
#| code-fold: true
ggplot(faang, aes(date, cum_ret, colour = ticker)) +
  geom_line(linewidth = 0.5) +
  geom_hline(yintercept = 0, colour = "grey50", linetype = "dashed") +
  scale_y_continuous(labels = percent) +
  labs(
    x = NULL, y = "Cumulative return", colour = NULL,
    title = "FAANG cumulative returns rebased to 2024-01-01",
    caption = "FAANG = META, AAPL, AMZN, NFLX, GOOGL. Adjusted close. Source: Yahoo Finance."
  )
```

#### Realised Volatility

21-day rolling annualised volatility (standard deviation of log returns, scaled by sqrt(252)). Higher values indicate greater price uncertainty. SPY (S&P 500 ETF) serves as the benchmark.

```{r}
#| code-fold: true
vol_raw <- eq |>
  filter(ticker %in% c("AAPL", "NVDA", "TSLA", "SPY"), date >= "2023-06-01") |>
  select(ticker, date, adjusted) |>
  collect() |>
  group_by(ticker) |>
  arrange(date) |>
  mutate(log_ret = log(adjusted / lag(adjusted))) |>
  filter(!is.na(log_ret)) |>
  mutate(
    cum_sq = cumsum(log_ret^2),
    vol_21d = sqrt(pmax((cum_sq - lag(cum_sq, 21, default = 0)) / 21, 0)) * sqrt(252)
  ) |>
  filter(date >= as.Date("2024-01-01"), !is.na(vol_21d)) |>
  ungroup()
```

```{r}
#| code-fold: true
ggplot(vol_raw, aes(date, vol_21d, colour = ticker)) +
  geom_line(linewidth = 0.4) +
  scale_y_continuous(labels = percent) +
  labs(
    x = NULL, y = "21d annualised volatility", colour = NULL,
    title = "Realised volatility: AAPL, NVDA, TSLA vs SPY benchmark",
    caption = "21-day rolling SD of log returns x sqrt(252). Source: Yahoo Finance."
  )
```

#### Coverage

All 51 equity tickers with trading day count and date range. Click column headers to sort.

```{r}
#| code-fold: true
eq_cov <- eq |>
  group_by(ticker) |>
  summarise(
    days = n(),
    first_date = min(date, na.rm = TRUE),
    last_date = max(date, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(days)) |>
  collect()

hd_dt(eq_cov,
  "Equity coverage: 51 tickers, daily OHLCV from Yahoo Finance. Click headers to sort.")
```


# Crypto

## Row

### {.tabset}

#### Major Coins

BTC, ETH, SOL, and BNB daily close on log scale since 2022. Log scale reveals percentage moves: equal vertical distances represent equal percentage changes regardless of price level.

```{r}
#| code-fold: true
major <- cr |>
  filter(ticker %in% c("BTC", "ETH", "SOL", "BNB"), date >= "2022-01-01") |>
  select(ticker, date, close) |>
  collect()
```

```{r}
#| code-fold: true
ggplot(major, aes(date, close, colour = ticker)) +
  geom_line(linewidth = 0.4) +
  scale_y_log10(labels = dollar) +
  labs(
    x = NULL, y = "Close USD (log scale)", colour = NULL,
    title = "BTC, ETH, SOL, BNB daily close prices",
    caption = "Log scale. Source: Yahoo Finance (BTC-USD, ETH-USD, SOL-USD, BNB-USD)."
  )
```

#### Stablecoin Peg

USDC and USDT should trade at $1.00. Deviations indicate depeg events — potential liquidity crises. The chart zooms to $0.97–$1.03 to make small deviations visible.

```{r}
#| code-fold: true
stable <- cr |>
  filter(ticker %in% c("USDC", "USDT"), date >= "2022-01-01") |>
  select(ticker, date, close) |>
  collect()
```

```{r}
#| code-fold: true
ggplot(stable, aes(date, close, colour = ticker)) +
  geom_line(linewidth = 0.4) +
  geom_hline(yintercept = 1.0, linetype = "dashed", colour = "grey50") +
  scale_y_continuous(limits = c(0.97, 1.03)) +
  labs(
    x = NULL, y = "USD price", colour = NULL,
    title = "Stablecoin peg: USDC and USDT deviation from $1.00",
    caption = "Dashed line = $1.00 peg. Y-axis zoomed to $0.97–$1.03. Source: Yahoo Finance."
  )
```

#### Correlation

Pairwise correlation of daily log returns for 6 major crypto assets since 2023. High correlations (>0.7) indicate assets move together — less diversification benefit.

```{r}
#| code-fold: true
wide <- cr |>
  filter(ticker %in% c("BTC", "ETH", "SOL", "BNB", "ADA", "XRP"),
         date >= "2023-01-01") |>
  select(date, ticker, close) |>
  collect() |>
  group_by(ticker) |>
  arrange(date) |>
  mutate(ret = log(close / lag(close))) |>
  filter(!is.na(ret)) |>
  ungroup() |>
  select(date, ticker, ret) |>
  pivot_wider(names_from = ticker, values_from = ret) |>
  filter(if_all(everything(), ~ !is.na(.)))

cor_mat <- cor(wide |> select(-date), use = "complete.obs")

cor_long <- cor_mat |>
  as.data.frame() |>
  mutate(row = rownames(cor_mat)) |>
  pivot_longer(-row, names_to = "col", values_to = "cor")
```

```{r}
#| code-fold: true
ggplot(cor_long, aes(row, col, fill = cor)) +
  geom_tile() +
  geom_text(aes(label = round(cor, 2)), colour = "white", size = 3.5) +
  scale_fill_gradient2(low = "#3498db", mid = "grey20", high = "#e74c3c",
                       midpoint = 0.5, limits = c(0, 1)) +
  labs(
    x = NULL, y = NULL, fill = "Corr",
    title = "Crypto log-return correlation matrix (2023+)",
    caption = "Daily log returns. All pairs >0.5 = limited diversification. Source: Yahoo Finance."
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Coverage

```{r}
#| code-fold: true
cr_cov <- cr |>
  group_by(ticker) |>
  summarise(
    days = n(),
    first_date = min(date, na.rm = TRUE),
    last_date = max(date, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(days)) |>
  collect()

hd_dt(cr_cov,
  "Crypto coverage: 14 tokens, daily OHLCV from Yahoo Finance. Click headers to sort.")
```


# Macro

## Row

### {.tabset}

#### Interest Rates

US Treasury yields (2Y, 10Y, 30Y) and the effective Federal Funds rate since 2020. The Fed Funds rate (DFF) is the policy rate; Treasury yields reflect market expectations of future rates and inflation.

```{r}
#| code-fold: true
rates <- ma |>
  filter(series_id %in% c("DGS2", "DGS10", "DGS30", "DFF"),
         date >= "2020-01-01", !is.na(value)) |>
  select(series_id, date, value) |>
  collect()
```

```{r}
#| code-fold: true
ggplot(rates, aes(date, value, colour = series_id)) +
  geom_line(linewidth = 0.4) +
  labs(
    x = NULL, y = "Yield (%)", colour = NULL,
    title = "US Treasury yields and Fed Funds rate (2020+)",
    caption = "DGS2/10/30 = constant-maturity Treasury yields. DFF = effective Fed Funds rate. Source: FRED."
  )
```

#### Yield Curve

The 10Y–2Y Treasury spread is a recession indicator. When negative (inverted), short rates exceed long rates — historically preceding recessions by 6–18 months. Red shading marks the inversion period.

```{r}
#| code-fold: true
yc <- ma |>
  filter(series_id == "T10Y2Y", date >= "2018-01-01", !is.na(value)) |>
  select(date, value) |>
  collect()

inv_dates <- yc |> filter(value < 0)
inv_start <- min(inv_dates$date)
inv_end <- max(inv_dates$date)
```

```{r}
#| code-fold: true
ggplot(yc, aes(date, value)) +
  geom_line(linewidth = 0.4, colour = "#e74c3c") +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  annotate("rect", xmin = inv_start, xmax = inv_end,
           ymin = -Inf, ymax = 0, fill = "#e74c3c", alpha = 0.15) +
  labs(
    x = NULL, y = "10Y − 2Y spread (%)",
    title = "Yield curve: 10Y–2Y Treasury spread with inversion shading",
    caption = paste0("Red shading = inverted (", inv_start, " to ", inv_end,
                     "). Inversion historically precedes recessions. Source: FRED T10Y2Y.")
  )
```

#### Credit Spreads

ICE BofA option-adjusted spreads for High Yield (HY) and BBB-rated corporate bonds. Wider spreads indicate higher default risk perception. Spikes correspond to market stress events.

```{r}
#| code-fold: true
spreads <- ma |>
  filter(series_id %in% c("BAMLH0A0HYM2", "BAMLC0A4CBBB"),
         date >= "2020-01-01", !is.na(value)) |>
  select(series_id, date, value) |>
  collect() |>
  mutate(series_id = recode(series_id,
    BAMLH0A0HYM2 = "HY Spread", BAMLC0A4CBBB = "BBB Spread"))
```

```{r}
#| code-fold: true
ggplot(spreads, aes(date, value, colour = series_id)) +
  geom_line(linewidth = 0.4) +
  labs(
    x = NULL, y = "OAS (percentage points)", colour = NULL,
    title = "ICE BofA credit spreads: High Yield vs BBB (2020+)",
    caption = "Option-adjusted spread over Treasuries. Higher = more default risk. Source: FRED."
  )
```

#### VIX vs S&P 500

VIX (CBOE Volatility Index) measures expected 30-day equity volatility — often called the 'fear index'. It typically spikes when the S&P 500 falls, reflecting the negative correlation between returns and volatility.

```{r}
#| code-fold: true
vix_sp <- ma |>
  filter(series_id %in% c("VIXCLS", "SP500"),
         date >= "2020-01-01", !is.na(value)) |>
  select(series_id, date, value) |>
  collect()
```

```{r}
#| code-fold: true
ggplot(vix_sp, aes(date, value)) +
  geom_line(linewidth = 0.4, colour = "#3498db") +
  facet_wrap(~series_id, ncol = 1, scales = "free_y") +
  labs(
    x = NULL, y = NULL,
    title = "VIX (fear index) and S&P 500 (2020+)",
    caption = "VIX = expected 30d vol. SP500 = index level. Negative correlation visible in 2020 crash. Source: FRED."
  )
```

#### All Series

```{r}
#| code-fold: true
ma_cov <- ma |>
  group_by(series_id) |>
  summarise(
    obs = n(),
    first_date = min(date, na.rm = TRUE),
    last_date = max(date, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(series_id) |>
  collect()

hd_dt(ma_cov,
  "FRED macro series: 19 series at mixed frequencies (daily/monthly/quarterly). Click headers to sort.")
```


# Factors

## Row

### {.tabset}

#### FF3 Daily

The Fama-French 3-factor model decomposes equity returns into: Market excess return (Mkt-RF), Size (SMB = Small Minus Big), and Value (HML = High Minus Low book-to-market). RF = risk-free rate. Daily returns since 2020.

```{r}
#| code-fold: true
ff3 <- fa |>
  filter(dataset == "FF3", frequency == "daily", date >= "2020-01-01") |>
  select(factor_name, date, value) |>
  collect()
```

```{r}
#| code-fold: true
#| fig-height: 7
ggplot(ff3, aes(date, value, colour = factor_name)) +
  geom_line(alpha = 0.6, linewidth = 0.3) +
  geom_hline(yintercept = 0, colour = "grey50", linewidth = 0.2) +
  facet_wrap(~factor_name, ncol = 1, scales = "free_y") +
  labs(
    x = NULL, y = "Return (%)",
    title = "Fama-French 3 factors: daily returns (2020+)",
    caption = "Mkt-RF = market excess return. SMB = small minus big. HML = high minus low B/M. RF = risk-free. Source: Ken French Data Library."
  ) +
  theme(legend.position = "none")
```

#### FF5 Cumulative

Cumulative returns for the 5-factor model (adds RMW = Robust Minus Weak profitability, CMA = Conservative Minus Aggressive investment) since 2000. Shows which factors have delivered positive long-run premia and which have not.

```{r}
#| code-fold: true
ff5 <- fa |>
  filter(dataset == "FF5", frequency == "daily",
         factor_name != "RF", date >= "2000-01-01") |>
  select(factor_name, date, value) |>
  collect() |>
  group_by(factor_name) |>
  arrange(date) |>
  mutate(cum_ret = cumprod(1 + value / 100) - 1) |>
  ungroup()
```

```{r}
#| code-fold: true
ggplot(ff5, aes(date, cum_ret, colour = factor_name)) +
  geom_line(linewidth = 0.5) +
  geom_hline(yintercept = 0, colour = "grey50", linetype = "dashed") +
  scale_y_continuous(labels = percent) +
  labs(
    x = NULL, y = "Cumulative return", colour = NULL,
    title = "FF5 cumulative factor returns (2000–2026)",
    caption = "Mkt-RF dominates. HML (value) struggled post-2010. RMW/CMA = profitability/investment. Source: Ken French."
  )
```

#### Momentum

Cumulative return of the momentum factor (WML = Winners Minus Losers). Momentum suffered a catastrophic crash in 2009 (March reversal) but recovered over the following decade. Shows the risk of factor concentration.

```{r}
#| code-fold: true
mom <- fa |>
  filter(dataset == "Mom", frequency == "daily", date >= "2000-01-01") |>
  select(date, value) |>
  collect() |>
  arrange(date) |>
  mutate(cum_ret = cumprod(1 + value / 100) - 1)
```

```{r}
#| code-fold: true
ggplot(mom, aes(date, cum_ret)) +
  geom_line(linewidth = 0.5, colour = "#2ecc71") +
  geom_hline(yintercept = 0, colour = "grey50", linetype = "dashed") +
  scale_y_continuous(labels = percent) +
  labs(
    x = NULL, y = "Cumulative return",
    title = "Momentum factor cumulative return (2000–2026)",
    caption = "WML = Winners Minus Losers. Note 2009 crash (momentum reversal) and post-2020 recovery. Source: Ken French."
  )
```

#### Distributions

Distribution of daily returns by factor. Fat tails (kurtosis > 3) indicate more extreme events than a normal distribution would predict. All factors show leptokurtic distributions.

```{r}
#| code-fold: true
ff5_dist <- fa |>
  filter(dataset == "FF5", frequency == "daily",
         factor_name != "RF", date >= "2000-01-01") |>
  select(factor_name, value) |>
  collect()
```

```{r}
#| code-fold: true
ggplot(ff5_dist, aes(value, fill = factor_name)) +
  geom_histogram(bins = 100, alpha = 0.7, show.legend = FALSE) +
  facet_wrap(~factor_name, scales = "free_y") +
  geom_vline(xintercept = 0, colour = "grey50", linetype = "dashed") +
  labs(
    x = "Daily return (%)", y = "Count",
    title = "FF5 daily return distributions (2000+)",
    caption = "All factors show fat tails (leptokurtic). Mkt-RF has widest range. Source: Ken French."
  )
```

#### Summary

```{r}
#| code-fold: true
fa_cov <- fa |>
  group_by(dataset, frequency, factor_name) |>
  summarise(
    obs = n(),
    first_date = min(date, na.rm = TRUE),
    last_date = max(date, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(dataset, frequency, factor_name) |>
  collect()

hd_dt(fa_cov,
  "Fama-French factors: FF3, FF5, Momentum. Daily and monthly, back to 1926. Source: Ken French Data Library.")
```


```{r}
#| label: cleanup
#| include: false
dbDisconnect(con, shutdown = TRUE)
```
