---
title: "historicaldata: Usage Examples"
subtitle: "Live queries against HF-hosted Parquet via DuckDB httpfs"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    code-tools: true
    theme: darkly
    embed-resources: true
execute:
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(DBI)
library(duckdb)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)

BASE <- "https://huggingface.co/datasets/dsfefvx/finance-historical-data/resolve/main"

con <- dbConnect(duckdb())
invisible(dbExecute(con, "INSTALL httpfs; LOAD httpfs;"))

# Helper: query HF parquet
hf <- function(file) paste0(BASE, "/", file)

theme_set(theme_minimal(base_size = 12) + theme(
  plot.background = element_rect(fill = "grey10", colour = NA),
  panel.background = element_rect(fill = "grey10", colour = NA),
  text = element_text(colour = "grey80"),
  axis.text = element_text(colour = "grey70"),
  panel.grid.minor = element_blank(),
  panel.grid.major = element_line(colour = "grey25"),
  legend.position = "bottom"
))
```

# Equity {.tabset}

51 US stocks and ETFs, daily OHLCV from 1980, sourced from Yahoo Finance.

## Single Ticker

Retrieve AAPL daily data and compute 50/200-day moving averages.

```{r}
aapl <- dbGetQuery(con, sprintf("
  SELECT date, close, adjusted, volume,
    AVG(close) OVER (ORDER BY date ROWS BETWEEN 49 PRECEDING AND CURRENT ROW) AS ma_50,
    AVG(close) OVER (ORDER BY date ROWS BETWEEN 199 PRECEDING AND CURRENT ROW) AS ma_200
  FROM read_parquet('%s')
  WHERE ticker = 'AAPL' AND date >= '2022-01-01'
  ORDER BY date", hf("equity_daily.parquet"))) |>
  as_tibble() |>
  filter(date >= as.Date("2023-01-01"))

glimpse(aapl)
```

::: {.callout-note collapse="true" title="Output: AAPL with moving averages"}
```{r}
ggplot(aapl, aes(date)) +
  geom_line(aes(y = close), colour = "white", linewidth = 0.4) +
  geom_line(aes(y = ma_50), colour = "#3498db", linewidth = 0.5, linetype = "dashed") +
  geom_line(aes(y = ma_200), colour = "#e74c3c", linewidth = 0.5, linetype = "dashed") +
  scale_y_continuous(labels = dollar) +
  labs(x = NULL, y = "Close", title = "AAPL: Close + 50d MA (blue) + 200d MA (red)")
```
:::

## Multi-Ticker Comparison

Compare FAANG cumulative returns from 2024.

```{r}
faang <- dbGetQuery(con, sprintf("
  SELECT ticker, date, adjusted
  FROM read_parquet('%s')
  WHERE ticker IN ('AAPL','AMZN','GOOGL','META','NFLX')
    AND date >= '2024-01-01'
  ORDER BY ticker, date", hf("equity_daily.parquet"))) |>
  as_tibble() |>
  group_by(ticker) |>
  mutate(cum_ret = adjusted / first(adjusted) - 1) |>
  ungroup()
```

::: {.callout-note collapse="true" title="Output: FAANG cumulative returns"}
```{r}
ggplot(faang, aes(date, cum_ret, colour = ticker)) +
  geom_line(linewidth = 0.5) +
  geom_hline(yintercept = 0, colour = "grey50", linetype = "dashed") +
  scale_y_continuous(labels = percent) +
  labs(x = NULL, y = "Cumulative return", colour = NULL,
       title = "FAANG cumulative returns (2024+)")
```
:::

## Sector Rotation

Group tickers by sector and compare recent performance.

```{r}
sectors <- tribble(
  ~ticker, ~sector,
  "AAPL",  "Tech",    "MSFT",  "Tech",    "NVDA", "Tech",    "GOOGL", "Tech",
  "JPM",   "Finance", "GS",    "Finance", "BAC",  "Finance",
  "JNJ",   "Health",  "UNH",   "Health",  "PFE",  "Health",
  "XOM",   "Energy",  "CVX",   "Energy",  "COP",  "Energy",
  "WMT",   "Consumer","COST",  "Consumer","HD",   "Consumer"
)

sector_ret <- dbGetQuery(con, sprintf("
  SELECT ticker, date, adjusted
  FROM read_parquet('%s')
  WHERE ticker IN (%s) AND date >= '2025-01-01'
  ORDER BY ticker, date",
  hf("equity_daily.parquet"),
  paste0("'", sectors$ticker, "'", collapse = ","))) |>
  as_tibble() |>
  inner_join(sectors, by = "ticker") |>
  group_by(sector, date) |>
  summarise(avg_adj = mean(adjusted, na.rm = TRUE), .groups = "drop") |>
  group_by(sector) |>
  mutate(cum_ret = avg_adj / first(avg_adj) - 1) |>
  ungroup()
```

::: {.callout-note collapse="true" title="Output: Sector rotation"}
```{r}
ggplot(sector_ret, aes(date, cum_ret, colour = sector)) +
  geom_line(linewidth = 0.6) +
  geom_hline(yintercept = 0, colour = "grey50", linetype = "dashed") +
  scale_y_continuous(labels = percent) +
  labs(x = NULL, y = "Cumulative equal-weight return", colour = NULL,
       title = "Sector performance (2025+)")
```
:::

## Volatility

Rolling 21-day realised volatility for selected tickers.

```{r}
vol_data <- dbGetQuery(con, sprintf("
  SELECT ticker, date, adjusted,
    LN(adjusted / LAG(adjusted) OVER (PARTITION BY ticker ORDER BY date)) AS log_ret
  FROM read_parquet('%s')
  WHERE ticker IN ('AAPL','NVDA','TSLA','SPY') AND date >= '2023-06-01'
  ORDER BY ticker, date", hf("equity_daily.parquet"))) |>
  as_tibble() |>
  filter(!is.na(log_ret), date >= as.Date("2024-01-01")) |>
  group_by(ticker) |>
  mutate(
    cum_var = cumsum(log_ret^2),
    vol_21d = sqrt(pmax((cum_var - lag(cum_var, 21, default = 0)) / 21, 0)) * sqrt(252)
  ) |>
  filter(!is.na(vol_21d)) |>
  ungroup()
```

::: {.callout-note collapse="true" title="Output: Realised volatility"}
```{r}
ggplot(vol_data, aes(date, vol_21d, colour = ticker)) +
  geom_line(linewidth = 0.4) +
  scale_y_continuous(labels = percent) +
  labs(x = NULL, y = "21d annualised vol", colour = NULL,
       title = "Realised volatility: AAPL, NVDA, TSLA vs SPY")
```
:::

## Coverage Table

```{r}
eq_cov <- dbGetQuery(con, sprintf("
  SELECT ticker, COUNT(*) as days, MIN(date) as first_date, MAX(date) as last_date
  FROM read_parquet('%s')
  GROUP BY ticker ORDER BY days DESC", hf("equity_daily.parquet"))) |>
  as_tibble()

knitr::kable(eq_cov, col.names = c("Ticker", "Trading Days", "From", "To"))
```


# Crypto {.tabset}

14 tokens including BTC, ETH, SOL, and stablecoins. Daily data from Yahoo Finance.

## Major Coins

```{r}
major <- dbGetQuery(con, sprintf("
  SELECT ticker, date, close, volume
  FROM read_parquet('%s')
  WHERE ticker IN ('BTC','ETH','SOL','BNB') AND date >= '2022-01-01'
  ORDER BY ticker, date", hf("crypto_daily.parquet"))) |>
  as_tibble()
```

::: {.callout-note collapse="true" title="Output: Major crypto prices"}
```{r}
ggplot(major, aes(date, close, colour = ticker)) +
  geom_line(linewidth = 0.4) +
  scale_y_log10(labels = dollar) +
  labs(x = NULL, y = "Close (log)", colour = NULL,
       title = "BTC, ETH, SOL, BNB daily close")
```
:::

## Stablecoin Depeg Monitor

Check USDC and USDT for deviations from $1.00.

```{r}
stable <- dbGetQuery(con, sprintf("
  SELECT ticker, date, close
  FROM read_parquet('%s')
  WHERE ticker IN ('USDC','USDT') AND date >= '2022-01-01'
  ORDER BY ticker, date", hf("crypto_daily.parquet"))) |>
  as_tibble() |>
  mutate(depeg = abs(close - 1.0))
```

::: {.callout-note collapse="true" title="Output: Stablecoin peg"}
```{r}
ggplot(stable, aes(date, close, colour = ticker)) +
  geom_line(linewidth = 0.4) +
  geom_hline(yintercept = 1.0, linetype = "dashed", colour = "grey50") +
  scale_y_continuous(limits = c(0.97, 1.03)) +
  labs(x = NULL, y = "USD price", colour = NULL, title = "Stablecoin peg: USDC vs USDT")
```
:::

::: {.callout-note collapse="true" title="Output: Absolute depeg"}
```{r}
ggplot(stable, aes(date, depeg, colour = ticker)) +
  geom_line(linewidth = 0.4) +
  labs(x = NULL, y = "|price - 1.0|", colour = NULL, title = "Absolute deviation from $1.00")
```
:::

## Correlation Matrix

Rolling 90-day correlation between major crypto assets.

```{r}
wide <- dbGetQuery(con, sprintf("
  SELECT ticker, date, close
  FROM read_parquet('%s')
  WHERE ticker IN ('BTC','ETH','SOL','BNB','ADA','XRP')
    AND date >= '2023-01-01'
  ORDER BY ticker, date", hf("crypto_daily.parquet"))) |>
  as_tibble() |>
  group_by(ticker) |>
  mutate(ret = log(close / lag(close))) |>
  filter(!is.na(ret)) |>
  ungroup() |>
  select(date, ticker, ret) |>
  pivot_wider(names_from = ticker, values_from = ret) |>
  filter(if_all(everything(), ~ !is.na(.)))

cor_mat <- cor(wide |> select(-date), use = "complete.obs")
```

::: {.callout-note collapse="true" title="Output: Correlation matrix"}
```{r}
cor_long <- cor_mat |>
  as.data.frame() |>
  mutate(row = rownames(cor_mat)) |>
  pivot_longer(-row, names_to = "col", values_to = "cor")

ggplot(cor_long, aes(row, col, fill = cor)) +
  geom_tile() +
  geom_text(aes(label = round(cor, 2)), colour = "white", size = 3.5) +
  scale_fill_gradient2(low = "#3498db", mid = "grey20", high = "#e74c3c", midpoint = 0) +
  labs(x = NULL, y = NULL, fill = "Corr", title = "Log-return correlations (2023+)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
:::

## Coverage Table

```{r}
cr_cov <- dbGetQuery(con, sprintf("
  SELECT ticker, COUNT(*) as days, MIN(date) as first_date, MAX(date) as last_date
  FROM read_parquet('%s')
  GROUP BY ticker ORDER BY days DESC", hf("crypto_daily.parquet"))) |>
  as_tibble()

knitr::kable(cr_cov, col.names = c("Token", "Days", "From", "To"))
```


# Macro {.tabset}

19 FRED series: yields, spreads, CPI, GDP, VIX, oil, USD index. Mixed frequencies.

## Interest Rates

Treasury yields across maturities + Fed Funds rate.

```{r}
rates <- dbGetQuery(con, sprintf("
  SELECT series_id, date, value
  FROM read_parquet('%s')
  WHERE series_id IN ('DGS2','DGS10','DGS30','DFF')
    AND date >= '2020-01-01' AND value IS NOT NULL
  ORDER BY series_id, date", hf("macro_daily.parquet"))) |>
  as_tibble()
```

::: {.callout-note collapse="true" title="Output: Interest rates"}
```{r}
ggplot(rates, aes(date, value, colour = series_id)) +
  geom_line(linewidth = 0.4) +
  labs(x = NULL, y = "Yield (%)", colour = NULL,
       title = "US Treasury yields + Fed Funds rate")
```
:::

## Yield Curve

10Y-2Y spread with inversion shading.

```{r}
yc <- dbGetQuery(con, sprintf("
  SELECT date, value
  FROM read_parquet('%s')
  WHERE series_id = 'T10Y2Y' AND date >= '2018-01-01' AND value IS NOT NULL
  ORDER BY date", hf("macro_daily.parquet"))) |>
  as_tibble()
```

::: {.callout-note collapse="true" title="Output: Yield curve spread"}
```{r}
inv_start <- min(yc$date[yc$value < 0], na.rm = TRUE)
inv_end <- max(yc$date[yc$value < 0], na.rm = TRUE)

ggplot(yc, aes(date, value)) +
  geom_line(linewidth = 0.4, colour = "#e74c3c") +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  annotate("rect", xmin = inv_start, xmax = inv_end,
           ymin = -Inf, ymax = 0, fill = "#e74c3c", alpha = 0.15) +
  labs(x = NULL, y = "10Y - 2Y spread (%)",
       title = "Yield curve: red shading = inversion")
```
:::

## Credit Spreads

ICE BofA High Yield and BBB corporate option-adjusted spreads.

```{r}
spreads <- dbGetQuery(con, sprintf("
  SELECT series_id, date, value
  FROM read_parquet('%s')
  WHERE series_id IN ('BAMLH0A0HYM2','BAMLC0A4CBBB')
    AND date >= '2020-01-01' AND value IS NOT NULL
  ORDER BY series_id, date", hf("macro_daily.parquet"))) |>
  as_tibble() |>
  mutate(series_id = recode(series_id,
    BAMLH0A0HYM2 = "HY Spread", BAMLC0A4CBBB = "BBB Spread"))
```

::: {.callout-note collapse="true" title="Output: Credit spreads"}
```{r}
ggplot(spreads, aes(date, value, colour = series_id)) +
  geom_line(linewidth = 0.4) +
  labs(x = NULL, y = "OAS (bps)", colour = NULL,
       title = "ICE BofA credit spreads")
```
:::

## VIX vs S&P 500

Overlay VIX (fear index) with S&P 500.

```{r}
vix_sp <- dbGetQuery(con, sprintf("
  SELECT series_id, date, value
  FROM read_parquet('%s')
  WHERE series_id IN ('VIXCLS','SP500')
    AND date >= '2020-01-01' AND value IS NOT NULL
  ORDER BY series_id, date", hf("macro_daily.parquet"))) |>
  as_tibble()
```

::: {.callout-note collapse="true" title="Output: VIX vs S&P 500"}
```{r}
ggplot(vix_sp, aes(date, value)) +
  geom_line(linewidth = 0.4, colour = "#3498db") +
  facet_wrap(~series_id, ncol = 1, scales = "free_y") +
  labs(x = NULL, y = NULL, title = "VIX and S&P 500")
```
:::

## All Series

```{r}
ma_cov <- dbGetQuery(con, sprintf("
  SELECT series_id, COUNT(*) as n,
         MIN(date) as first_date, MAX(date) as last_date
  FROM read_parquet('%s')
  GROUP BY series_id ORDER BY series_id", hf("macro_daily.parquet"))) |>
  as_tibble()

knitr::kable(ma_cov, col.names = c("Series", "Obs", "From", "To"))
```


# Factors {.tabset}

Fama-French 3 and 5 factors plus Momentum. Daily and monthly, back to 1926.

## FF3 Factor Returns

Market, SMB, HML, and risk-free rate.

```{r}
ff3 <- dbGetQuery(con, sprintf("
  SELECT factor_name, date, value
  FROM read_parquet('%s')
  WHERE dataset = 'FF3' AND frequency = 'daily' AND date >= '2020-01-01'
  ORDER BY factor_name, date", hf("factors.parquet"))) |>
  as_tibble()
```

::: {.callout-note collapse="true" title="Output: FF3 daily returns"}
```{r}
#| fig-height: 7
ggplot(ff3, aes(date, value, colour = factor_name)) +
  geom_line(alpha = 0.6, linewidth = 0.3) +
  geom_hline(yintercept = 0, colour = "grey50", linewidth = 0.2) +
  facet_wrap(~factor_name, ncol = 1, scales = "free_y") +
  labs(x = NULL, y = "Return (%)", title = "Fama-French 3 factors (daily, 2020+)") +
  theme(legend.position = "none")
```
:::

## FF5 Cumulative

Long-run cumulative returns for all 5 factors since 2000.

```{r}
ff5 <- dbGetQuery(con, sprintf("
  SELECT factor_name, date, value
  FROM read_parquet('%s')
  WHERE dataset = 'FF5' AND frequency = 'daily'
    AND factor_name != 'RF' AND date >= '2000-01-01'
  ORDER BY factor_name, date", hf("factors.parquet"))) |>
  as_tibble() |>
  group_by(factor_name) |>
  mutate(cum_ret = cumprod(1 + value / 100) - 1) |>
  ungroup()
```

::: {.callout-note collapse="true" title="Output: FF5 cumulative returns"}
```{r}
ggplot(ff5, aes(date, cum_ret, colour = factor_name)) +
  geom_line(linewidth = 0.5) +
  scale_y_continuous(labels = percent) +
  labs(x = NULL, y = "Cumulative return", colour = NULL,
       title = "FF5 factor cumulative returns (2000+)")
```
:::

## Momentum

Cumulative return of the momentum factor — including the 2009 crash and recovery.

```{r}
mom <- dbGetQuery(con, sprintf("
  SELECT date, value
  FROM read_parquet('%s')
  WHERE dataset = 'Mom' AND frequency = 'daily' AND date >= '2000-01-01'
  ORDER BY date", hf("factors.parquet"))) |>
  as_tibble() |>
  mutate(cum_ret = cumprod(1 + value / 100) - 1)
```

::: {.callout-note collapse="true" title="Output: Momentum cumulative return"}
```{r}
ggplot(mom, aes(date, cum_ret)) +
  geom_line(linewidth = 0.5, colour = "#2ecc71") +
  geom_hline(yintercept = 0, colour = "grey50", linetype = "dashed") +
  scale_y_continuous(labels = percent) +
  labs(x = NULL, y = "Cumulative return",
       title = "Momentum factor (2000+) — note 2009 crash")
```
:::

## Factor Distributions

Distribution of daily returns by factor.

```{r}
ff5_dist <- dbGetQuery(con, sprintf("
  SELECT factor_name, value
  FROM read_parquet('%s')
  WHERE dataset = 'FF5' AND frequency = 'daily'
    AND factor_name != 'RF' AND date >= '2000-01-01'",
  hf("factors.parquet"))) |>
  as_tibble()
```

::: {.callout-note collapse="true" title="Output: Return distributions"}
```{r}
ggplot(ff5_dist, aes(value, fill = factor_name)) +
  geom_histogram(bins = 100, alpha = 0.7, show.legend = FALSE) +
  facet_wrap(~factor_name, scales = "free_y") +
  geom_vline(xintercept = 0, colour = "grey50", linetype = "dashed") +
  labs(x = "Daily return (%)", y = "Count",
       title = "FF5 daily return distributions (2000+)")
```
:::

## Summary Table

```{r}
fa_cov <- dbGetQuery(con, sprintf("
  SELECT dataset, frequency, factor_name, COUNT(*) as n,
         MIN(date) as first_date, MAX(date) as last_date
  FROM read_parquet('%s')
  GROUP BY dataset, frequency, factor_name
  ORDER BY dataset, frequency, factor_name", hf("factors.parquet"))) |>
  as_tibble()

knitr::kable(fa_cov,
  col.names = c("Dataset", "Freq", "Factor", "Obs", "From", "To"))
```


```{r}
#| label: cleanup
#| include: false
dbDisconnect(con, shutdown = TRUE)
```
