---
title: "historicaldata"
format:
  dashboard:
    orientation: rows
    theme: cosmo
    embed-resources: true
    include-in-header:
      - text: |
          <style>
          body { font-size: 1.15em; }
          /* Remove ALL internal scrollbars — MUST override Bootstrap/bslib */
          .tab-pane, .tab-content, .card, .card-body, .card-header,
          .cell-output, .cell-output-display, .cell-output-stdout,
          .datatables, .dataTables_wrapper, .dataTables_scroll,
          .bslib-card, .bslib-page-fill, .html-fill-container,
          .html-fill-item { overflow: visible !important; max-height: none !important; height: auto !important; }
          /* Plots: ensure visible size */
          .cell-output-display img { min-height: 500px; width: 100%; }
          /* Tables: no wrapping so columns stay readable */
          table.dataTable { white-space: nowrap; }
          /* Click-to-zoom: fill entire screen */
          .fullscreen-overlay {
            position: fixed !important;
            top: 0 !important; left: 0 !important;
            width: 100vw !important; height: 100vh !important;
            z-index: 9999 !important;
            background: black !important;
            padding: 10px !important;
            overflow: auto !important;
            cursor: zoom-out !important;
          }
          .fullscreen-overlay img {
            width: 100% !important; height: 100% !important;
            object-fit: contain !important;
          }
          .cell-output-display:has(img) { cursor: zoom-in; }
          /* Captions below plots */
          .fig-caption {
            font-size: 0.85em; color: #888; margin-top: 0.4em;
            text-align: left; line-height: 1.5;
          }
          details { margin-bottom: 0.5em; }
          details summary { cursor: pointer; color: #6c9bd2; font-size: 0.9em; }
          </style>
          <script>
          document.addEventListener('click', function(e) {
            // Close fullscreen
            var overlay = e.target.closest('.fullscreen-overlay');
            if (overlay) {
              overlay.classList.remove('fullscreen-overlay');
              overlay.style.cssText = '';
              return;
            }
            // Only zoom the img itself, not the parent (which includes caption)
            var img = e.target.closest('img');
            if (!img) return;
            var cell = img.closest('.cell-output-display');
            if (!cell) return;
            // Clone just the img into a temporary overlay div
            var div = document.createElement('div');
            div.className = 'fullscreen-overlay';
            div.innerHTML = '<img src="' + img.src + '">';
            div.addEventListener('click', function() { div.remove(); });
            document.body.appendChild(div);
          });
          </script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)
```

# Equity

## Row

### {.tabset}

#### AAPL Moving Averages

AAPL daily close with 50d and 200d simple moving averages. The [golden cross](https://www.investopedia.com/terms/g/goldencross.asp) (50d above 200d) and death cross (50d below) are trend-following signals.

```{r}
#| results: asis
show_code("vig_eq_aapl")
```

```{r}
#| fig-height: 8
safe_tar_read("vig_eq_aapl")
```

::: {.fig-caption}
**AAPL close + 50d/200d SMA.** Y: USD. White = close, blue = 50d SMA, red = 200d SMA. Source: Yahoo Finance via `hd_ohlcv()`. See Volatility tab. [Golden cross](https://www.investopedia.com/terms/g/goldencross.asp). Click to zoom.
:::

```{r}
df <- safe_tar_read("vig_meta_eq_aapl")
if (!is.null(df)) hd_dt(df, "Ticker metadata for the plot above.")
```

#### FAANG Returns

FAANG = Meta, Apple, Amazon, Netflix, Alphabet. Cumulative returns from 2024-01-01.

```{r}
#| results: asis
show_code("vig_eq_faang")
```

```{r}
#| fig-height: 8
safe_tar_read("vig_eq_faang")
```

::: {.fig-caption}
**FAANG cumulative returns from 2024-01-01.** Y: cum return (%). Colours: ticker. Split-adjusted close. Source: `hd_ohlcv()`. See AAPL tab.
:::

```{r}
df <- safe_tar_read("vig_meta_eq_faang")
if (!is.null(df)) hd_dt(df, "FAANG ticker metadata.")
```

#### Realised Volatility

21-day rolling annualised vol. SPY = S&P 500 benchmark.

```{r}
#| results: asis
show_code("vig_eq_vol")
```

```{r}
#| fig-height: 8
safe_tar_read("vig_eq_vol")
```

::: {.fig-caption}
**21d realised volatility (annualised).** Y: vol (%). Vol = SD(log ret) x sqrt(252). Source: `hd_ohlcv()`. [Realised vol](https://www.investopedia.com/terms/r/realizedvolatility.asp).
:::

```{r}
df <- safe_tar_read("vig_meta_eq_vol")
if (!is.null(df)) hd_dt(df, "Volatility ticker metadata.")
```

#### Coverage

```{r}
#| results: asis
show_code("vig_eq_coverage")
```

```{r}
df <- safe_tar_read("vig_eq_coverage")
if (!is.null(df)) {
  top <- df[1, ]
  hd_dt(df, paste0(
    nrow(df), " tickers. Longest: ", top$Ticker,
    " (", top$`Trading Days`, " days). Source: hd_ohlcv()."
  ))
}
```


# Crypto

## Row

### {.tabset}

#### Major Coins

BTC, ETH, SOL, BNB daily close on log scale since 2022.

```{r}
#| results: asis
show_code("vig_cr_major")
```

```{r}
#| fig-height: 8
safe_tar_read("vig_cr_major")
```

::: {.fig-caption}
**BTC, ETH, SOL, BNB log-scale close.** Y: USD (log10). Source: `hd_ohlcv()`. See Stablecoins tab.
:::

```{r}
df <- safe_tar_read("vig_meta_cr_major")
if (!is.null(df)) hd_dt(df, "Major crypto metadata.")
```

#### Stablecoin Peg

USDC/USDT vs $1.00 peg. [Depeg](https://www.investopedia.com/terms/d/depeg.asp) = liquidity crisis.

```{r}
#| results: asis
show_code("vig_cr_stable")
```

```{r}
#| fig-height: 8
safe_tar_read("vig_cr_stable")
```

::: {.fig-caption}
**USDC/USDT vs $1.00 peg.** Y: USD, zoomed $0.97–$1.03. Dashed = peg. Source: `hd_ohlcv()`.
:::

```{r}
df <- safe_tar_read("vig_meta_cr_stable")
if (!is.null(df)) hd_dt(df, "Stablecoin metadata.")
```

#### Correlation

Pearson correlation of daily log returns, 6 tokens, 2023+.

```{r}
#| results: asis
show_code("vig_cr_corr")
```

```{r}
#| fig-height: 8
safe_tar_read("vig_cr_corr")
```

::: {.fig-caption}
**Log-return correlation, 6 tokens, 2023+.** Fill: 0–1. >0.7 = limited diversification. Source: `hd_ohlcv()`.
:::

```{r}
df <- safe_tar_read("vig_meta_cr_corr")
if (!is.null(df)) hd_dt(df, "Correlation ticker metadata.")
```

#### Coverage

```{r}
#| results: asis
show_code("vig_cr_coverage")
```

```{r}
df <- safe_tar_read("vig_cr_coverage")
if (!is.null(df)) {
  hd_dt(df, paste0(nrow(df), " tokens. Source: hd_ohlcv()."))
}
```


# Macro

## Row

### {.tabset}

#### Interest Rates

Treasury yields (2Y, 10Y, 30Y) + Fed Funds rate since 2020.

```{r}
#| results: asis
show_code("vig_ma_rates")
```

```{r}
#| fig-height: 8
safe_tar_read("vig_ma_rates")
```

::: {.fig-caption}
**Treasury yields + Fed Funds rate.** Y: yield (%). DGS = constant-maturity. DFF = Fed Funds. Source: FRED via `hd_macro()`. See Yield Curve tab.
:::

#### Yield Curve

10Y–2Y spread. Negative = [inverted](https://www.investopedia.com/terms/i/invertedyieldcurve.asp), precedes recessions.

```{r}
#| results: asis
show_code("vig_ma_yc")
```

```{r}
#| fig-height: 8
safe_tar_read("vig_ma_yc")
```

::: {.fig-caption}
**10Y–2Y spread with inversion shading.** Y: spread (pp). Red = inverted. Source: FRED T10Y2Y via `hd_macro()`.
:::

#### Credit Spreads

ICE BofA OAS: HY (junk) + BBB (investment grade).

```{r}
#| results: asis
show_code("vig_ma_spreads")
```

```{r}
#| fig-height: 8
safe_tar_read("vig_ma_spreads")
```

::: {.fig-caption}
**ICE BofA OAS over Treasuries.** Y: spread (pp). Wider = more default risk. Source: FRED via `hd_macro()`.
:::

#### All Series

```{r}
#| results: asis
show_code("vig_ma_coverage")
```

```{r}
df <- safe_tar_read("vig_ma_coverage")
if (!is.null(df)) {
  hd_dt(df, paste0(nrow(df), " FRED series. Source: hd_macro()."))
}
```


# Factors

## Row

### {.tabset}

#### FF3 Daily

[Fama-French 3 factors](https://www.investopedia.com/terms/f/famaandfrenchthreefactormodel.asp): Mkt-RF, SMB, HML, RF. Daily since 2020.

```{r}
#| results: asis
show_code("vig_fa_ff3")
```

```{r}
#| fig-height: 7
safe_tar_read("vig_fa_ff3")
```

::: {.fig-caption}
**FF3 daily returns.** Y: return (%). Mkt-RF = market excess. SMB = size. HML = value. RF = T-bill. Source: Ken French via `hd_factors()`.
:::

#### FF5 Cumulative

5-factor cumulative returns since 2000. RMW = profitability, CMA = investment.

```{r}
#| results: asis
show_code("vig_fa_ff5")
```

```{r}
#| fig-height: 8
safe_tar_read("vig_fa_ff5")
```

::: {.fig-caption}
**FF5 cumulative returns (2000–2026).** Y: cum return (%). Mkt-RF dominates. HML weak post-2010. Source: Ken French via `hd_factors()`.
:::

#### Momentum

[Momentum](https://www.investopedia.com/terms/m/momentum.asp) (WML). March 2009 crash, post-2020 recovery.

```{r}
#| results: asis
show_code("vig_fa_mom")
```

```{r}
#| fig-height: 8
safe_tar_read("vig_fa_mom")
```

::: {.fig-caption}
**Momentum (WML) cumulative return (2000–2026).** 2009 = reversal crash. Source: Ken French via `hd_factors()`.
:::

#### Summary

```{r}
#| results: asis
show_code("vig_fa_coverage")
```

```{r}
df <- safe_tar_read("vig_fa_coverage")
if (!is.null(df)) {
  hd_dt(df, paste0(
    dplyr::n_distinct(df$Factor), " factors, ",
    format(sum(df$Obs), big.mark = ","), " obs. Source: hd_factors()."
  ))
}
```


# LSE ETFs

## Row

### {.tabset}

#### Most Liquid

Top 5 LSE ETFs by average daily volume, selected dynamically from metadata via `hd_search(".*[.]L$") |> slice_max(volume_avg, n = 5)`.

```{r}
#| results: asis
show_code("vig_lse_liquid")
```

```{r}
#| fig-height: 8
safe_tar_read("vig_lse_liquid")
```

::: {.fig-caption}
**Top 5 LSE ETFs by average daily volume.** Y: close (log scale, local currency). Tickers selected dynamically from metadata — not hardcoded. Source: Yahoo Finance via `hd_ohlcv()`. See Coverage tab for full list.
:::

```{r}
df <- safe_tar_read("vig_meta_lse_liquid")
if (!is.null(df)) hd_dt(df, "Most liquid LSE ETF metadata.")
```

#### FTSE vs Global

Comparison of FTSE 100 tracking ETFs vs Global equity ETFs, both listed on LSE. Uses curated groups: `hd_group("FTSE 100 ETFs")` and `hd_group("Global Equity ETFs (LSE)")`.

```{r}
#| results: asis
show_code("vig_lse_ftse_vs_global")
```

```{r}
#| fig-height: 8
safe_tar_read("vig_lse_ftse_vs_global")
```

::: {.fig-caption}
**FTSE 100 vs Global Equity ETFs.** Y: cumulative return (%). Solid/dashed = group. Source: `hd_ohlcv()` via `hd_group()`. See [FTSE 100](https://www.investopedia.com/terms/f/ftse.asp).
:::

#### GBP vs USD

Most liquid GBP-denominated vs USD-denominated LSE ETFs, selected by querying metadata currency field: `hd_search(".*[.]L$") |> filter(currency == "GBP")`.

```{r}
#| results: asis
show_code("vig_lse_currency")
```

```{r}
#| fig-height: 8
safe_tar_read("vig_lse_currency")
```

::: {.fig-caption}
**GBP vs USD denominated LSE ETFs.** Y: cumulative return (%). Faceted by currency. Top 3 each by volume. Source: `hd_ohlcv()` + metadata currency filter.
:::

```{r}
df <- safe_tar_read("vig_meta_lse_currency")
if (!is.null(df)) hd_dt(df, "GBP vs USD ticker metadata.")
```

#### Fund Families

Distribution of LSE ETFs by fund family (iShares/BlackRock, Vanguard, Xtrackers, etc.) from metadata `fund_family` field.

```{r}
#| results: asis
show_code("vig_lse_families")
```

```{r}
safe_tar_read("vig_lse_families")
```

::: {.fig-caption}
**LSE ETF count by fund family (top 15).** X: number of ETFs. Source: metadata via `hd_search()`. [ETF](https://www.investopedia.com/terms/e/etf.asp) = Exchange Traded Fund.
:::

#### Highest Yield

Top 10 LSE ETFs by dividend yield, from metadata `yield_pct` field: `hd_search(".*[.]L$") |> slice_max(yield_pct, n = 10)`.

```{r}
#| results: asis
show_code("vig_lse_yield")
```

```{r}
#| fig-height: 8
safe_tar_read("vig_lse_yield")
```

::: {.fig-caption}
**Top 10 LSE ETFs by dividend yield.** X: annualised yield (%). Source: metadata via `hd_search()`. Yield = most recent trailing distribution. See [Dividend yield](https://www.investopedia.com/terms/d/dividendyield.asp).
:::

```{r}
df <- safe_tar_read("vig_meta_lse_yield")
if (!is.null(df)) hd_dt(df, "Highest yield ETF metadata.")
```

#### Coverage

Full sortable table of all ~930 LSE ETFs with metadata. Click column headers to sort, type in filter boxes to search.

```{r}
#| results: asis
show_code("vig_lse_coverage")
```

```{r}
df <- safe_tar_read("vig_lse_coverage")
if (!is.null(df)) {
  hd_dt(df, paste0(
    nrow(df), " LSE ETFs. Columns: ticker, name, currency, fund family, ",
    "volume, yield, beta, date range. Source: metadata via hd_search()."
  ))
}
```


# Installation

## Row

### {.tabset}

#### Setup (run first)

After installing via any method below, run this once:

```{r}
#| results: asis
show_code("vig_setup")
```

This defines `hd_theme()` used by all plots. All subsequent code assumes these libraries are loaded.

#### R (pak)

```r
# Install from GitHub
pak::pak("JohnGavin/historical", subdir = "packages/historicaldata")

# Then run the Setup tab code above
```

#### Nix

```bash
# Enter reproducible Nix shell (all R + Python packages included)
nix develop github:JohnGavin/historical
```

Then start R and run the Setup tab code above. All dependencies (dplyr, ggplot2, scales, purrr, DT, arrow, duckdb) are pre-installed in the Nix shell.

#### Python (DuckDB)

No R needed — query the same data directly from Python using DuckDB's native `hf://` protocol:

```python
import duckdb

con = duckdb.connect()

# Query AAPL from HF-hosted Parquet — no download needed
df = con.sql("""
    SELECT date, close, volume
    FROM 'hf://datasets/dsfefvx/finance-historical-data/equity_daily.parquet'
    WHERE ticker = 'AAPL' AND date >= '2024-01-01'
    ORDER BY date
""").df()

# All datasets available:
# hf://datasets/dsfefvx/finance-historical-data/equity_daily.parquet   (958 tickers)
# hf://datasets/dsfefvx/finance-historical-data/crypto_daily.parquet   (14 tokens)
# hf://datasets/dsfefvx/finance-historical-data/macro_daily.parquet    (19 FRED series)
# hf://datasets/dsfefvx/finance-historical-data/factors.parquet        (7 FF factors)
# hf://datasets/dsfefvx/finance-historical-data/metadata.parquet       (994 ticker metadata)
```

#### T language

```bash
# Clone, enter Nix shell, fetch data, run pipeline
git clone https://github.com/JohnGavin/historical.git && cd historical
nix develop
python scripts/fetch_equity.py     # 51 equity tickers
Rscript scripts/fetch_crypto.R     # 14 crypto tokens
Rscript scripts/fetch_macro.R      # 19 FRED series
Rscript scripts/fetch_factors.R    # Fama-French factors
t run src/pipeline.t               # T pipeline
```

Then start R and run the Setup tab code above.
