---
title: "Factor MAX Rotation"
format:
  dashboard:
    orientation: rows
    theme: cosmo
    embed-resources: false
    include-in-header:
      - text: |
          <style>
          body { font-size: 1.15em; }
          .bslib-page-fill, .html-fill-container, .html-fill-item,
          .bslib-card, .card, .tab-pane, .tab-content, .card-body,
          .bslib-grid, .bslib-grid-item {
            height: auto !important; min-height: 0 !important;
            max-height: none !important; overflow: visible !important;
            flex: none !important;
          }
          .cell-output-display { display: block !important; position: relative !important;
            overflow: visible !important; margin-bottom: 1em; }
          .cell-output-display img { display: block; width: 100%; height: auto; min-height: 400px; }
          .datatables, .dataTables_wrapper { display: block !important;
            position: relative !important; overflow-x: auto !important; }
          table.dataTable { white-space: nowrap; }
          table caption { caption-side: top !important; }
          .fig-caption { display: block; font-size: 0.85em; color: #888;
            margin-top: 0.4em; text-align: left; line-height: 1.5; }
          </style>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)
library(dplyr)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

safe_tar_read <- function(name) {
  tryCatch(
    targets::tar_read_raw(name),
    error = function(e) {
      rds_dirs <- c("../inst/extdata/vignettes", "inst/extdata/vignettes")
      for (d in rds_dirs) {
        rds <- file.path(d, paste0(name, ".rds"))
        if (file.exists(rds)) return(readRDS(rds))
      }
      NULL
    }
  )
}

hd_dt <- function(df, caption_text) {
  if (is.null(df)) return(invisible(NULL))
  DT::datatable(df, caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold; color: #ddd;",
    caption_text), rownames = FALSE,
    options = list(pageLength = 15, scrollX = TRUE, dom = "frtip"))
}
```

# Strategy

## Row

::: {.callout-note}
## Factor MAX Rotation

Based on [Alpha Architect research](https://alphaarchitect.com/factor-max/) (Dec 2025): the **maximum daily return within a month** predicts next-month factor returns. This signal captures investor underreaction to factor-level news and is distinct from factor momentum.

**Implementation:** Each month, rank Fama-French factors by their MAX signal (highest single-day return that month). Go long the top 2 factors next month with equal weights. Rebalance monthly.
:::

## Row

### {.tabset}

#### Performance

```{r}
#| fig-height: 6
safe_tar_read("fm_cumret_plot")
```

::: {.fig-caption}
**Factor MAX vs Market.** Y: growth of $1 (log scale). Blue = Factor MAX (top-2 factors by MAX signal, equal weight). Red = Market factor (Mkt-RF). Dashed line = OOS start (2020). Source: Fama-French via `hd_factors()`. Log scale shows compounding differences.
:::

#### Metrics

```{r}
metrics <- safe_tar_read("fm_metrics")
if (!is.null(metrics)) {
  display <- metrics |>
    mutate(across(c(cagr, vol, max_dd, hit_rate, bench_cagr, bench_vol),
                  ~ paste0(round(. * 100, 1), "%"))) |>
    mutate(across(c(sharpe, bench_sharpe), ~ round(., 2)))
  hd_dt(display, "Factor MAX strategy metrics: in-sample (1963-2019), out-of-sample (2020+), and full period.")
}
```

::: {.fig-caption}
**Performance metrics.** CAGR, volatility, Sharpe ratio, max drawdown, and hit rate (fraction of months beating the market). Benchmark is the market factor (Mkt-RF). In-sample: 1963-2019. Out-of-sample: 2020+.
:::

#### Factor Selection

```{r}
sf <- safe_tar_read("fm_selection_freq")
if (!is.null(sf)) {
  sf <- sf |> mutate(pct = paste0(round(pct * 100, 1), "%"))
  hd_dt(sf, "How often each factor is selected by the MAX signal. Momentum dominates.")
}
```

::: {.fig-caption}
**Factor selection frequency.** Shows how many months (out of total) each factor was ranked in the top 2 by MAX signal. Momentum (Mom) is selected most frequently, consistent with the Alpha Architect finding that MAX captures momentum-adjacent but distinct information.
:::

# Signal

## Row

### {.tabset}

#### Heatmap

```{r}
#| fig-height: 8
safe_tar_read("fm_heatmap")
```

::: {.fig-caption}
**MAX signal rank heatmap.** Green = rank 1 (highest MAX, strongest buy signal). Dark = rank 5 (lowest). Rows: Fama-French factors (HML, SMB, RMW, CMA, Mom). Columns: calendar months. Faceted by year. Source: `hd_factors(dataset = "FF5")` + `hd_factors(dataset = "Mom")`.
:::

#### How It Works

::: {.callout-tip}
## The MAX Signal

For each factor, each month:

1. **Compute MAX:** the single highest daily return in that month
2. **Rank factors** by MAX (1 = highest)
3. **Next month:** go long the top-2 ranked factors with equal weights

**Intuition:** A factor with a high MAX day signals that the factor's underlying theme (value, momentum, etc.) is receiving positive attention. Investors underreact to this signal, creating predictable returns.

**Key distinction from momentum:** Factor momentum uses average or cumulative returns. MAX uses only the single best day. The Alpha Architect study shows these capture different information — MAX remains predictive even after controlling for momentum.
:::

# Details

## Row

### {.tabset}

#### Universe

The strategy rotates among 5 Fama-French factors (daily data from 1963):

| Factor | Description | Source |
|--------|-------------|--------|
| **HML** | High Minus Low (value) | FF5 |
| **SMB** | Small Minus Big (size) | FF5 |
| **RMW** | Robust Minus Weak (profitability) | FF5 |
| **CMA** | Conservative Minus Aggressive (investment) | FF5 |
| **Mom** | Momentum (winners minus losers) | Ken French |

Benchmark: **Mkt-RF** (market excess return).

#### Parameters

```{r}
params <- safe_tar_read("fm_params")
if (!is.null(params)) {
  param_df <- tibble(
    Parameter = c("Factors", "Top N", "Start Date", "IS End", "OOS Start", "Benchmark"),
    Value = c(paste(params$factors, collapse = ", "),
              as.character(params$top_n),
              as.character(params$start_date),
              as.character(params$is_end),
              as.character(params$oos_start),
              params$benchmark_factor)
  )
  hd_dt(param_df, "Strategy parameters. Change fm_params target and tar_make() to update all.")
}
```

#### References

1. **Research:** [Factor MAX: A New Signal for Predicting Factor Returns](https://alphaarchitect.com/factor-max/) — Alpha Architect (Dec 2025)
2. **Data:** Fama-French factors via [Ken French Data Library](https://mba.tuck.dartmouth.edu/pages/faculty/ken.french/data_library.html), accessed through `hd_factors()`
3. **Methodology:** Expanding window backtest with OOS holdout, same pattern as [Macro Defense Rotation](macro-defense-rotation.html)

#### Build Info

```{r}
#| label: build-info
#| echo: false
#| results: asis
gh_url <- tryCatch({
  remote <- system("git remote get-url origin 2>/dev/null", intern = TRUE)
  sub("\\.git$", "", sub("^git@github\\.com:", "https://github.com/", remote))
}, error = function(e) NULL)

git_sha_short <- tryCatch(
  system("git rev-parse --short HEAD 2>/dev/null", intern = TRUE),
  error = function(e) "N/A"
)
git_sha_full <- tryCatch(
  system("git rev-parse HEAD 2>/dev/null", intern = TRUE),
  error = function(e) git_sha_short
)

pkg_ver <- tryCatch(
  as.character(packageVersion("historicaldata")),
  error = function(e) "0.1.0"
)

r_ver <- as.character(getRversion())
build_time <- format(Sys.time(), "%Y-%m-%d %H:%M:%S")

ver_link <- if (!is.null(gh_url)) {
  sprintf("[%s](%s/releases/tag/v%s)", pkg_ver, gh_url, pkg_ver)
} else pkg_ver

sha_link <- if (!is.null(gh_url) && git_sha_short != "N/A") {
  sprintf("[`%s`](%s/commit/%s)", git_sha_short, gh_url, git_sha_full)
} else sprintf("`%s`", git_sha_short)

r_link <- sprintf("[%s](https://cran.r-project.org/doc/manuals/r-release/NEWS.html)", r_ver)

cat(sprintf(
  "**historicaldata** %s | **Git** %s | **R** %s | **Built** %s",
  ver_link, sha_link, r_link, build_time
))
```
