---
title: "Strategy Falsification"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    embed-resources: false
    css: falsification-theme.css
    include-in-header:
      - text: |
          <link rel="stylesheet" href="vignette-shared.css">
          <script src="vignette-shared.js"></script>
          <script type="module" src="causal-diagrams.js"></script>
          <style>/* All styles in falsification-theme.css */</style>
          <script>
          if (!localStorage.getItem('quarto-color-scheme')) {
            document.documentElement.setAttribute('data-bs-theme', 'dark');
          }
          document.addEventListener('click', function(e) {
            var overlay = e.target.closest('.fullscreen-overlay');
            if (overlay) { overlay.remove(); return; }
            var img = e.target.closest('.cell-output-display img');
            if (!img) return;
            var div = document.createElement('div');
            div.className = 'fullscreen-overlay';
            var clone = document.createElement('img');
            clone.src = img.src;
            div.appendChild(clone);
            document.body.appendChild(div);
          });
          document.addEventListener('shown.bs.tab', function(e) {
            var pane = document.querySelector(e.target.getAttribute('data-bs-target') || e.target.getAttribute('href'));
            if (pane) pane.scrollIntoView({ behavior: 'smooth', block: 'start' });
          });
          // ── UI Controls: font size + dark/light toggle ──
          document.addEventListener('DOMContentLoaded', function() {
            var toolbar = document.createElement('div');
            toolbar.id = 'ui-controls';
            toolbar.innerHTML = '<button id="font-down" title="Decrease font">A-</button>' +
              '<button id="font-up" title="Increase font">A+</button>' +
              '<button id="theme-toggle" title="Toggle dark/light"></button>';
            document.body.appendChild(toolbar);

            // Font size
            var baseSize = parseFloat(localStorage.getItem('font-size-px') || '24');
            document.documentElement.style.setProperty('--base-font-size', baseSize + 'px');
            document.getElementById('font-up').addEventListener('click', function() {
              baseSize = Math.min(baseSize + 1, 32);
              document.documentElement.style.setProperty('--base-font-size', baseSize + 'px');
              localStorage.setItem('font-size-px', baseSize);
            });
            document.getElementById('font-down').addEventListener('click', function() {
              baseSize = Math.max(baseSize - 1, 12);
              document.documentElement.style.setProperty('--base-font-size', baseSize + 'px');
              localStorage.setItem('font-size-px', baseSize);
            });

            // Dark/light toggle
            var themeBtn = document.getElementById('theme-toggle');
            var isDark = document.documentElement.getAttribute('data-bs-theme') === 'dark';
            themeBtn.textContent = isDark ? '\u2600\uFE0F' : '\uD83C\uDF19';
            themeBtn.addEventListener('click', function() {
              isDark = !isDark;
              document.documentElement.setAttribute('data-bs-theme', isDark ? 'dark' : 'light');
              themeBtn.textContent = isDark ? '\u2600\uFE0F' : '\uD83C\uDF19';
              localStorage.setItem('quarto-color-scheme', isDark ? 'dark' : 'light');
            });
          });
          </script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)
library(dplyr)

pkgload::load_all(here::here("packages/historicaldata"), quiet = TRUE)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)

# Column definitions for linked headers (open in new tab)
col_defs <- list(
  "Verdict" = list(
    title = "Classification: Genuine alpha, Borderline, or No alpha (beta)",
    href = "#assumptions"
  ),
  "Strategy" = list(title = "Strategy long name", href = NULL),
  "HAC t" = list(
    title = "Newey-West HAC-corrected t-statistic on mean excess returns",
    href = "https://en.wikipedia.org/wiki/Newey%E2%80%93West_estimator"
  ),
  "Naive Sharpe" = list(
    title = "Annualised Sharpe ratio assuming iid returns (no serial correlation correction)",
    href = "https://en.wikipedia.org/wiki/Sharpe_ratio"
  ),
  "Alpha (%)" = list(
    title = "Annualised Fama-French 5-Factor + Momentum intercept, in percent",
    href = "https://en.wikipedia.org/wiki/Fama%E2%80%93French_three-factor_model"
  ),
  "Alpha t" = list(
    title = "HAC t-statistic on the FF5+Mom alpha intercept",
    href = "#ff"
  ),
  "R\u00b2 (%)" = list(
    title = "Variance explained by the 6 Fama-French + Momentum factors, in percent",
    href = "#ff"
  )
)

# DT helper with linked column headers (open in new tab)
fals_dt <- function(df, caption_text) {
  if (is.null(df)) return(invisible(NULL))
  # Build header with title attributes and optional links
  header_cells <- lapply(names(df), function(nm) {
    cd <- col_defs[[nm]]
    tip <- if (!is.null(cd)) cd$title else nm
    href <- if (!is.null(cd)) cd$href else NULL
    label <- if (!is.null(href)) {
      as.character(htmltools::tags$a(
        href = href, target = "_blank",
        style = "color: #6c9bd2; text-decoration: underline dotted;",
        title = tip, nm
      ))
    } else {
      as.character(htmltools::tags$span(title = tip, nm))
    }
    label
  })
  sketch <- htmltools::withTags(table(
    class = "display",
    thead(tr(lapply(header_cells, function(h) th(htmltools::HTML(h))))
  )))
  DT::datatable(
    df,
    container = sketch,
    escape = FALSE,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold; color: #ddd;",
      htmltools::HTML(caption_text)
    ),
    rownames = FALSE,
    filter = "top",
    options = list(
      pageLength = 15, scrollX = TRUE, autoWidth = FALSE,
      dom = "frtip",
      search = list(regex = TRUE, caseInsensitive = TRUE),
      initComplete = DT::JS(
        "function(settings, json) {",
        "  $(this.api().table().container()).css({'color': '#ddd', 'background-color': '#1a1a1a'});",
        "  $(this.api().table().header()).css({'color': '#ddd', 'background-color': '#222'});",
        "  $('input', this.api().table().container()).css({'color': '#ddd', 'background-color': '#333', 'border-color': '#555'});",
        "}"
      )
    )
  ) |>
    DT::formatStyle(columns = names(df), color = "#ddd", backgroundColor = "#1a1a1a")
}
```

::: {.callout-note}
**Falsification framework.** Every strategy must survive multiple statistical tests before being considered genuine. A strategy that looks profitable may simply be disguised factor exposure (beta) rather than true alpha. This dashboard applies Harvey, Liu & Zhu (2016) and Lopez de Prado (2018) methodology.
:::

# Scorecard

## Row

### {.tabset}

#### Strategy Verdicts

```{r}
#| label: scorecard
scorecard <- safe_tar_read("fals_vig_scorecard")
caption <- safe_tar_read("fals_vig_scorecard_caption")
if (!is.null(scorecard)) fals_dt(scorecard, caption)
```

#### Alpha vs Factor Exposure

```{r}
#| label: ff-alpha-plot
#| fig-height: 6
#| fig-width: 10
safe_tar_read("fals_vig_ff_alpha_plot")
```

`r { cap <- safe_tar_read("fals_vig_ff_caption"); if (!is.null(cap)) paste0("*", cap, "*") else "" }`

#### How to Read This

**Verdict criteria:**

- **Genuine alpha**: Alpha t > 2.0 AND R² (%) < 15. The strategy generates returns that known factors cannot explain, and factor exposure is low.
- **Borderline**: Alpha t near 2.0. May be genuine but evidence is weak. Requires out-of-sample confirmation.
- **No alpha (beta)**: Alpha t < 2.0 OR R² (%) > 20. The strategy's returns are explained by known factor exposure (market beta, momentum, value, etc.).

**Key insight**: A strategy can have positive returns and still have zero alpha. If those returns come from market exposure (high R²), you can get the same exposure more cheaply via index funds.


# Null Environments {#nulls}

## Row

### {.tabset}

#### Rejection Heatmap

```{r}
#| label: null-heatmap
#| fig-height: 6
#| fig-width: 10
safe_tar_read("fals_vig_null_heatmap")
```

`r { cap <- safe_tar_read("fals_vig_null_heatmap_caption"); if (!is.null(cap)) paste0("*", cap, "*") else "" }`

#### Rejection Table

```{r}
#| label: null-table
null_tbl <- safe_tar_read("fals_vig_null_table")
cap <- safe_tar_read("fals_vig_null_heatmap_caption")
if (!is.null(null_tbl)) fals_dt(null_tbl, cap)
```

#### Methodology

**What are null environments?**

A null environment generates synthetic return series with no genuine signal, but with realistic statistical properties. We test: "could a strategy generate the observed t-statistic just by chance in this environment?"

For each null environment, we generate M = 100 synthetic series, apply the strategy's logic, and compute the [<abbr title="Newey-West heteroskedasticity-autocorrelation consistent estimator">HAC</abbr> t-statistic](https://en.wikipedia.org/wiki/Newey%E2%80%93West_estimator). The rejection rate is the fraction that exceeds the observed t-statistic.

**The 6 null environments** (implemented in [`falsification.R`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/falsification.R)):

- **White Noise**: iid N(0, sigma). Baseline — if a strategy "works" here, it's pure luck.
- **Regime Vol**: Two-state volatility (high/low). Tests whether signal is just vol timing.
- **MA(1)**: Autocorrelated returns (theta = 0.1). Tests whether signal exploits serial correlation friction.
- **Factor Null**: Random factor loadings matching observed factor covariance structure. Tests whether signal is just factor exposure.
- **<abbr title="Generalized Autoregressive Conditional Heteroskedasticity">GARCH</abbr>(1,1)**: omega = 1e-6, alpha = 0.09, beta = 0.90. Tests whether signal exploits volatility clustering.
- **<abbr title="Glosten-Jagannathan-Runkle GARCH — adds asymmetric leverage effect">GJR-GARCH</abbr>**: Adds asymmetric leverage (gamma = 0.05). Tests whether signal exploits the leverage effect.

**Interpretation:**

- Rejection rate < 8%: Strategy is easily replicated under the null — **no genuine signal**
- Rejection rate 8-20%: Borderline — may be noise
- Rejection rate > 20%: The null cannot explain the strategy's performance — **genuine signal**


# HAC Correction {#hac}

## Row

### {.tabset}

#### Naive vs HAC t-Statistics

```{r}
#| label: hac-plot
#| fig-height: 6
#| fig-width: 10
safe_tar_read("fals_vig_hac_plot")
```

`r { cap <- safe_tar_read("fals_vig_hac_caption"); if (!is.null(cap)) paste0("*", cap, "*") else "" }`

#### HAC Details

```{r}
#| label: hac-table
hac_comp <- safe_tar_read("fals_vig_hac_comparison")
cap <- safe_tar_read("fals_vig_hac_caption")
if (!is.null(hac_comp)) fals_dt(hac_comp, cap)
```

#### Methodology

**Why HAC correction matters:**

Financial returns exhibit volatility clustering ([GARCH](https://en.wikipedia.org/wiki/Autoregressive_conditional_heteroskedasticity) persistence). Standard Sharpe ratios assume iid returns, which inflates the Sharpe by 20-30% on average.

The [Newey-West correction](https://en.wikipedia.org/wiki/Newey%E2%80%93West_estimator) (implemented in [`hd_hac_sharpe()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/falsification.R#L80)):

- Uses a Bartlett kernel to downweight distant autocovariances
- Bandwidth = floor(4 * (T/100)^(2/9)) — automatic, data-driven
- Produces an HAC-consistent variance estimate for the mean return
- The HAC t-statistic = mean(r) / sqrt(HAC_var / T)

**Interpretation:**

- HAC t > 2.0: Robust risk-adjusted return, survives serial correlation correction
- HAC t < 2.0: Naive Sharpe is inflated by autocorrelation; true risk-adjusted performance is weaker


# Factor Regression {#ff}

## Row

### {.tabset}

#### Fama-French 5-Factor + Momentum Betas

```{r}
#| label: ff-table
ff_tbl <- safe_tar_read("fals_vig_ff_table")
cap <- safe_tar_read("fals_vig_ff_caption")
if (!is.null(ff_tbl)) fals_dt(ff_tbl, cap)
```

#### Methodology

**[<abbr title="Fama-French 5-Factor model (MktRF, SMB, HML, RMW, CMA)">FF5</abbr>](https://en.wikipedia.org/wiki/Fama%E2%80%93French_three-factor_model) + [Momentum](https://en.wikipedia.org/wiki/Momentum_(finance)) regression:**

$$r_{excess} = \alpha + \beta_1 \cdot MktRF + \beta_2 \cdot SMB + \beta_3 \cdot HML + \beta_4 \cdot RMW + \beta_5 \cdot CMA + \beta_6 \cdot Mom + \epsilon$$

Where:

- **Mkt-RF**: Market excess return (market - risk-free)
- **<abbr title="Small Minus Big — size factor (small-cap minus large-cap returns)">SMB</abbr>**: Small Minus Big (size factor)
- **<abbr title="High Minus Low — value factor (cheap stocks minus expensive stocks)">HML</abbr>**: High Minus Low (value factor)
- **<abbr title="Robust Minus Weak — profitability factor">RMW</abbr>**: Robust Minus Weak (profitability factor)
- **<abbr title="Conservative Minus Aggressive — investment factor">CMA</abbr>**: Conservative Minus Aggressive (investment factor)
- **<abbr title="Momentum — past 12-month winners minus losers">Mom</abbr>**: Momentum (winners - losers, 12-2 month)

**Alpha (%)** is the intercept — the annualised return not explained by any factor. **R² (%)** measures how much of the strategy's variance is explained by known factors.

**Decision rules** (matching Verdict in the Scorecard):

- Alpha t > 2.0, R² (%) < 15: **Genuine alpha** — returns come from something beyond known factors
- Alpha t > 2.0, R² (%) > 15: Alpha exists but significant factor exposure — investigate
- Alpha t < 2.0, R² (%) > 20: **Disguised beta** — just buy the factors directly
- Alpha t < 2.0, R² (%) < 10: Strategy is noise

**Data source:** [Kenneth French Data Library](https://mba.tuck.dartmouth.edu/pages/faculty/ken.french/data_library.html), daily frequency. Implemented in [`hd_factor_null_test()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/falsification.R).


# Multiplicity Correction {#multiplicity}

## Row

### {.tabset}

#### K_eff and Delta-Z

```{r}
#| label: multiplicity-table
mult_tbl <- safe_tar_read("fals_vig_multiplicity")
cap <- safe_tar_read("fals_vig_multiplicity_caption")
if (!is.null(mult_tbl)) fals_dt(mult_tbl, cap)
```

#### Methodology

**The multiple testing problem:**

Testing 5 strategies inflates Type I error. The chance that *at least one* strategy exceeds t = 2.0 by pure luck is much higher than 5%.

**<abbr title="Effective number of independent strategies (spectral participation ratio)">K_eff</abbr> (effective multiplicity):**

K_eff = K² / ||Σ||²_F — the spectral participation ratio of the pairwise correlation matrix. It measures how many *truly independent* strategies exist:

- K_eff = K: All strategies are uncorrelated (worst case for multiple testing)
- K_eff = 1: All strategies are perfectly correlated (just one independent bet)
- K_eff between 1 and K: Some independent structure

**Harvey threshold:** sqrt(2 * log(K_eff)) — the t-statistic a strategy must exceed to survive multiplicity correction at the 5% level. Based on Harvey, Liu & Zhu (2016), "...and the Cross-Section of Expected Returns."

**Delta-Z (IS-OOS gap):**

Delta-Z = max(t_IS) - max(t_OOS) — the gap between the best in-sample and best out-of-sample t-statistics. A large gap indicates overfitting:

- Delta-Z < 1: Modest overfitting, results likely robust
- Delta-Z 1-2: Some overfitting, treat with caution
- Delta-Z > 2: Severe overfitting, in-sample performance is misleading

In this framework, "in-sample" = HAC t-stat (strategy returns), "out-of-sample" = FF alpha t-stat (after removing known factor exposure). The OOS test is not a temporal holdout but a *structural* holdout — does the signal survive factor decomposition?


# Assumptions {#assumptions}

## Row

### {.tabset}

#### Framework

**What this framework tests (4 of 5):**

1. **Null environment testing**: Generates 100 synthetic series per environment (6 environments) to test whether a strategy's t-statistic could arise by chance
2. **HAC correction**: Adjusts for serial correlation in returns (GARCH persistence inflates naive Sharpe by ~20-30%)
3. **Factor regression**: Decomposes returns into known factor exposure (beta) and residual alpha
4. **Multiplicity correction**: Adjusts for testing multiple strategies simultaneously (K_eff, delta-Z)

**Not yet implemented (1 of 5):**

5. **<abbr title="Deflated Sharpe Ratio — adjusts for skewness, kurtosis, and number of trials">DSR</abbr>** (Deflated Sharpe Ratio): Lopez de Prado's adjustment for skewness, kurtosis, and number of trials. Corrects the Sharpe ratio for non-normality — a strategy with negative skew (fat left tail) has an inflated Sharpe. See [#53 comment](https://github.com/JohnGavin/historical/issues/53) for implementation plan.

**What this framework does NOT do:**

- Does not test for [look-ahead bias](https://github.com/JohnGavin/historical/blob/main/R/plan_falsification.R) (handled by the [`look-ahead-bias-prevention`](https://github.com/JohnGavin/historical/blob/main/../../.claude/rules/look-ahead-bias-prevention.md) rule)
- Does not test for transaction cost sensitivity (handled by [`execution-delay-sensitivity`](https://github.com/JohnGavin/historical/blob/main/R/plan_alpha_decay.R) — alpha decay analysis)
- Does not test for parameter sensitivity (handled by [`backtest-robustness`](https://github.com/JohnGavin/historical/blob/main/R/plan_bootstrap_ci.R) — bootstrap CI)
- Does not test for temporal stability (handled by [`backtest-partitions`](https://github.com/JohnGavin/historical/blob/main/R/plan_partitions.R) — train/test/validation split)
- Does not test for cross-geography pervasiveness (requires international data — see [#58](https://github.com/JohnGavin/historical/issues/58))

**References:**

- Harvey, C., Liu, Y., & Zhu, H. (2016). "...and the Cross-Section of Expected Returns." *Review of Financial Studies*, 29(1), 5-68.
- Lopez de Prado, M. (2018). "The Deflated Sharpe Ratio: Correcting for Selection Bias, Backtest Overfitting, and Non-Normality." *Journal of Portfolio Management*, 40(5), 94-107.
- Newey, W. K., & West, K. D. (1987). "A Simple, Positive Semi-definite, Heteroskedasticity and Autocorrelation Consistent Covariance Matrix." *Econometrica*, 55(3), 703-708.
- Fama, E., & French, K. (2015). "A Five-Factor Asset Pricing Model." *Journal of Financial Economics*, 116(1), 1-22.

#### Build Info

`r build_info()`

# Causal Graph {#causal}

## Row

### {.tabset}

#### Overview

`r { cg <- safe_tar_read("cg_dag"); if (!is.null(cg)) paste0("Simplified view: strategies as single nodes. **", cg$n_nodes, " nodes, ", cg$n_edges, " edges** in the full DAG. Click any strategy node to view its source code. Click diagram to zoom fullscreen.") else "" }`

```{r}
#| label: causal-overview-mount
#| results: asis
#| echo: false
cat('<div id="dag-overview-mount"></div>')
```

#### Full DAG

```{r}
#| label: causal-full-mount
#| results: asis
#| echo: false
cat('<div id="dag-full-mount"></div>')
```

Node colours: <abbr title="Fama-French factors: HML (value), SMB (size), Mom (momentum), RMW (profitability), CMA (investment)">Factors</abbr> (blue) | <abbr title="VIX, VVIX, Fed rate, Inflation — observable macro indicators">Macro</abbr> (gold) | <abbr title="Classified market state: Vol Regime (VIX-based), Rate Regime (Fed-based)">Regime</abbr> (orange) | <abbr title="Strategy-specific trading signals derived from factors and regimes">Signals</abbr> (purple) | <abbr title="Realised strategy returns after applying signals">Returns</abbr> (green) | <abbr title="Portfolio-level metrics: Sharpe ratio, max drawdown">Outcomes</abbr> (red) | <abbr title="Structural factors: factor crowding, premium decay, rebalancing costs">Structural</abbr> (grey). Yellow dashed = <abbr title="Conditional independence test failed (|partial r| >= 0.15) — DAG is missing an edge">violated implication</abbr>. Green = tested, holds. Source: [plan_causal_graph.R](https://github.com/JohnGavin/historical/blob/main/R/plan_causal_graph.R).

`r paste0('<p style="color:#888;font-size:0.8em;">Full DAG tab rendered: ', format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z"), ' | commit: ', system("git rev-parse --short HEAD", intern=TRUE), '</p>')`

#### DRIF

```{r}
#| label: causal-drif-mount
#| results: asis
#| echo: false
cat('<div id="dag-drif-mount"></div>')
```

<abbr title="Dynamic Rotation Into Factors — each month selects the FF factor with highest trailing return">DRIF</abbr> selects the best-performing FF factor each month. Loads on <abbr title="High Minus Low — value factor: cheap stocks minus expensive stocks">HML</abbr>, <abbr title="Small Minus Big — size factor: small-cap minus large-cap returns">SMB</abbr>, <abbr title="Momentum — past 12-month winners minus losers">Mom</abbr>, <abbr title="Robust Minus Weak — profitability factor">RMW</abbr>. Yellow dashed edge: VIX affects HML in crises (<abbr title="Partial correlation between HML and VIX = -0.17, indicating they co-move in crises">r=-0.17</abbr>). Source: [plan_drif.R](https://github.com/JohnGavin/historical/blob/main/R/plan_drif.R).

`r paste0('<p style="color:#888;font-size:0.8em;">DRIF tab rendered: ', format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z"), ' | commit: ', system("git rev-parse --short HEAD", intern=TRUE), '</p>')`

#### Factor MAX

```{r}
#| label: causal-fmax-mount
#| results: asis
#| echo: false
cat('<div id="dag-fmax-mount"></div>')
```

<abbr title="Factor Momentum MAX — selects the single FF factor with highest recent momentum">Factor MAX</abbr> selects the top-momentum factor. Premium has <abbr title="Strategy Sharpe ratio in early period vs late period. >50% decay means the edge has halved.">decayed >50%</abbr> (#73). Source: [plan_factormax.R](https://github.com/JohnGavin/historical/blob/main/R/plan_factormax.R).

#### LTR

```{r}
#| label: causal-ltr-mount
#| results: asis
#| echo: false
cat('<div id="dag-ltr-mount"></div>')
```

<abbr title="LambdaMART — XGBoost-based cross-sectional momentum strategy. Ranks stocks by predicted return using gradient-boosted trees.">LTR</abbr> uses XGBoost for cross-sectional stock ranking. Loads on <abbr title="Momentum — past 12-month winners minus losers">Mom</abbr> + <abbr title="Small Minus Big — size factor">SMB</abbr>. <abbr title="Fed rate above/below median classifies as high/low rate regime">Rate regime</abbr> feeds into signal. Source: [plan_ltr_momentum.R](https://github.com/JohnGavin/historical/blob/main/R/plan_ltr_momentum.R).

#### Implication Tests

```{r}
#| label: causal-implications
#| echo: false
cg_impl <- safe_tar_read("cg_test_implications")
if (!is.null(cg_impl) && is.data.frame(cg_impl$results)) {
  DT::datatable(
    cg_impl$results |>
      dplyr::select(Implication = label, `Partial r` = partial_cor,
                    Status = status, Holds = holds),
    rownames = FALSE, filter = "none",
    options = list(pageLength = 10, dom = "t"),
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left;",
      htmltools::HTML(paste0("Conditional independence tests via partial correlation. ",
             cg_impl$n_holds, " of ", cg_impl$n_implications,
             " implications hold (|r| &lt; 0.15). ",
             "Violated implications suggest missing edges in the DAG. ",
             "Source: <a href='https://github.com/JohnGavin/historical/blob/main/R/plan_causal_graph.R' target='_blank'>plan_causal_graph.R</a>, tested on monthly data."))
    )
  ) |>
    DT::formatRound("Partial r", digits = 3)
} else {
  cat("*Implication tests not yet built. Run `tar_make(names = 'cg_test_implications')`.*")
}
```

#### Interpretation

**What the DAG encodes:** Every arrow is a causal assumption from our strategy pipeline — "X causes Y" or "X determines Y". If the assumption is wrong, the arrow should be removed or redirected.

**How to read the tests:** Each conditional independence claim says "if our DAG is correct, then X and Y should be unrelated after controlling for Z." A partial correlation near zero confirms the claim; a large partial correlation violates it.

**Violated: HML ⊥ VIX (r = -0.17)**

The value factor (HML) and volatility index (VIX) are NOT independent. This means:

| Impact area | What it means | Action |
|-------------|---------------|--------|
| **DRIF strategy** | DRIF loads on HML. During VIX spikes, DRIF underperforms MORE than the Vol_regime node alone predicts. | Consider reducing DRIF weight in crisis regimes |
| **Multi-strategy portfolio (#52)** | Diversification between VIX overlay and value-heavy strategies is LESS than K_eff suggests — they correlate in crises, exactly when diversification matters most | Recompute weights with crisis-conditional correlations |
| **Tail independence (#55)** | K_eff_crisis < K_eff_calm is expected — this finding explains why | Run `hd_tail_keff()` to quantify |
| **DAG revision** | Need either: VIX → HML edge (direct), or latent "Crisis_mode" → {VIX, HML} | Added VIX → HML edge (dashed yellow in diagram) |

**Economic explanation:** In crises, investors sell risky assets indiscriminately. Value stocks (high book-to-market) are often distressed firms — they get sold first. Simultaneously, VIX spikes as options demand surges. The correlation is not VIX *causing* value to fall, but both responding to the same flight-to-safety mechanism.

**Temporal split finding:** The HML-VIX link was stronger pre-2010 (r=-0.26) and has weakened post-2010 (r=-0.14). This is the opposite of the crowding hypothesis — the crisis-value correlation peaked during 2000-2010 (dot-com crash + GFC), not after factor investing became popular. The VIX-HML edge in the DAG is most relevant for crisis periods, less so in recent calm markets.

```{r}
#| label: causal-split-table
#| echo: false
split <- safe_tar_read("cg_test_split")
if (!is.null(split) && !is.null(split$comparison)) {
  DT::datatable(
    split$comparison,
    rownames = FALSE, filter = "none",
    options = list(pageLength = 5, dom = "t"),
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left;",
      paste0("HML-VIX correlation by period. Pre-2010 includes dot-com crash + GFC. ",
             "Source: ", "[plan_causal_graph.R]",
             "(https://github.com/JohnGavin/historical/blob/main/R/plan_causal_graph.R).")
    )
  ) |> DT::formatRound("hml_vix_r", digits = 3)
} else {
  cat("*Split test not yet built.*")
}
```

`r { cap <- safe_tar_read("cg_caption"); if (!is.null(cap)) cap else "" }`
