---
title: "Strategy Falsification"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    embed-resources: false
    css: falsification-theme.css
    include-in-header:
      - text: |
          <link rel="stylesheet" href="vignette-shared.css">
          <script src="vignette-shared.js"></script>
          <script type="module" src="causal-diagrams.js"></script>
          <style>/* All styles in falsification-theme.css */</style>
          <script>
          if (!localStorage.getItem('quarto-color-scheme')) {
            document.documentElement.setAttribute('data-bs-theme', 'dark');
          }
          document.addEventListener('click', function(e) {
            var overlay = e.target.closest('.fullscreen-overlay');
            if (overlay) { overlay.remove(); return; }
            var img = e.target.closest('.cell-output-display img');
            if (!img) return;
            var div = document.createElement('div');
            div.className = 'fullscreen-overlay';
            var clone = document.createElement('img');
            clone.src = img.src;
            div.appendChild(clone);
            document.body.appendChild(div);
          });
          // ── UI Controls: font size + dark/light toggle ──
          document.addEventListener('DOMContentLoaded', function() {
            var toolbar = document.createElement('div');
            toolbar.id = 'ui-controls';
            toolbar.innerHTML = '<button id="font-down" title="Decrease font">A-</button>' +
              '<button id="font-up" title="Increase font">A+</button>' +
              '<button id="theme-toggle" title="Toggle dark/light"></button>';
            document.body.appendChild(toolbar);

            // Font size
            var baseSize = parseFloat(localStorage.getItem('font-size-px') || '24');
            document.documentElement.style.setProperty('--base-font-size', baseSize + 'px');
            document.getElementById('font-up').addEventListener('click', function() {
              baseSize = Math.min(baseSize + 1, 32);
              document.documentElement.style.setProperty('--base-font-size', baseSize + 'px');
              localStorage.setItem('font-size-px', baseSize);
            });
            document.getElementById('font-down').addEventListener('click', function() {
              baseSize = Math.max(baseSize - 1, 12);
              document.documentElement.style.setProperty('--base-font-size', baseSize + 'px');
              localStorage.setItem('font-size-px', baseSize);
            });

            // Dark/light toggle
            var themeBtn = document.getElementById('theme-toggle');
            var isDark = document.documentElement.getAttribute('data-bs-theme') === 'dark';
            themeBtn.textContent = isDark ? '\u2600\uFE0F' : '\uD83C\uDF19';
            themeBtn.addEventListener('click', function() {
              isDark = !isDark;
              document.documentElement.setAttribute('data-bs-theme', isDark ? 'dark' : 'light');
              themeBtn.textContent = isDark ? '\u2600\uFE0F' : '\uD83C\uDF19';
              localStorage.setItem('quarto-color-scheme', isDark ? 'dark' : 'light');
            });
          });
          </script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)
library(dplyr)

# Path-checked store + utils (cwd is docs/ under quarto render, project
# root under tar_make). Same pattern as docs/leaderboard.qmd:29-39 and
# the #379 fix for momentum-prepeak.qmd.
store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)

# Column definitions for linked headers (open in new tab)
col_defs <- list(
  "Verdict" = list(
    title = "Classification: Genuine alpha, Borderline, or No alpha (beta)",
    href = "#assumptions"
  ),
  "Strategy" = list(title = "Strategy long name", href = NULL),
  "HAC t" = list(
    title = "Newey-West HAC-corrected t-statistic on mean excess returns",
    href = "https://en.wikipedia.org/wiki/Newey%E2%80%93West_estimator"
  ),
  "Naive Sharpe" = list(
    title = "Annualised Sharpe ratio assuming iid returns (no serial correlation correction)",
    href = "https://en.wikipedia.org/wiki/Sharpe_ratio"
  ),
  "Alpha (%)" = list(
    title = "Annualised Fama-French 5-Factor + Momentum intercept, in percent",
    href = "https://en.wikipedia.org/wiki/Fama%E2%80%93French_three-factor_model"
  ),
  "Alpha t" = list(
    title = "HAC t-statistic on the FF5+Mom alpha intercept",
    href = "#ff"
  ),
  "R\u00b2 (%)" = list(
    title = "Variance explained by the 6 Fama-French + Momentum factors, in percent",
    href = "#ff"
  )
)

# DT helper with linked column headers (open in new tab)
fals_dt <- function(df, caption_text) {
  if (is.null(df)) return(invisible(NULL))
  # Build header with title attributes and optional links
  header_cells <- lapply(names(df), function(nm) {
    cd <- col_defs[[nm]]
    tip <- if (!is.null(cd)) cd$title else nm
    href <- if (!is.null(cd)) cd$href else NULL
    label <- if (!is.null(href)) {
      as.character(htmltools::tags$a(
        href = href, target = "_blank",
        style = "color: #6c9bd2; text-decoration: underline dotted;",
        title = tip, nm
      ))
    } else {
      as.character(htmltools::tags$span(title = tip, nm))
    }
    label
  })
  sketch <- htmltools::withTags(table(
    class = "display",
    thead(tr(lapply(header_cells, function(h) th(htmltools::HTML(h))))
  )))
  DT::datatable(
    df,
    container = sketch,
    escape = FALSE,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold; color: #ddd;",
      htmltools::HTML(caption_text)
    ),
    rownames = FALSE,
    filter = "top",
    options = list(
      pageLength = 15, scrollX = TRUE, autoWidth = FALSE,
      dom = "frtip",
      search = list(regex = TRUE, caseInsensitive = TRUE),
      initComplete = DT::JS(
        "function(settings, json) {",
        "  $(this.api().table().container()).css({'color': '#ddd', 'background-color': '#1a1a1a'});",
        "  $(this.api().table().header()).css({'color': '#ddd', 'background-color': '#222'});",
        "  $('input', this.api().table().container()).css({'color': '#ddd', 'background-color': '#333', 'border-color': '#555'});",
        "}"
      )
    )
  ) |>
    DT::formatStyle(columns = names(df), color = "#ddd", backgroundColor = "#1a1a1a")
}
```

::: {.callout-note}
**Falsification framework.** Every strategy must survive multiple statistical tests before being considered genuine. A strategy that looks profitable may simply be disguised factor exposure (beta) rather than true alpha. This dashboard applies Harvey, Liu & Zhu (2016) and Lopez de Prado (2018) methodology.
:::

# Scorecard

## Row

### {.tabset}

#### Strategy Verdicts

```{r}
#| label: scorecard
scorecard <- safe_tar_read("fals_vig_scorecard")
caption <- safe_tar_read("fals_vig_scorecard_caption")
if (!is.null(scorecard)) fals_dt(scorecard, caption)
```

#### Alpha vs Factor Exposure

```{r}
#| label: ff-alpha-plot
#| fig-height: 6
#| fig-width: 10
safe_tar_read("fals_vig_ff_alpha_plot")
```

`r { cap <- safe_tar_read("fals_vig_ff_caption"); if (!is.null(cap)) paste0("*", cap, "*") else "" }`

#### How to Read This

**Verdict criteria:**

- **Genuine alpha**: Alpha t > 2.0 AND R² (%) < 15. The strategy generates returns that known factors cannot explain, and factor exposure is low.
- **Borderline**: Alpha t near 2.0. May be genuine but evidence is weak. Requires out-of-sample confirmation.
- **No alpha (beta)**: Alpha t < 2.0 OR R² (%) > 20. The strategy's returns are explained by known factor exposure (market beta, momentum, value, etc.).

**Scorecard column definitions:**

- **HAC t**: Newey-West HAC-corrected t-statistic on mean excess returns. Adjusts for serial correlation in returns. Values > 2.0 are statistically significant at the 5% level.
- **Naive Sharpe**: Annualised Sharpe ratio assuming iid (independent and identically distributed) returns. Typically overstates risk-adjusted performance by 20-30% due to volatility clustering.
- **Alpha (%)**: Annualised Fama-French 5-Factor + Momentum alpha (intercept from regression). Represents returns not explained by these 6 factors. Positive = outperformance, negative = underperformance.
- **Alpha t**: HAC t-statistic on the FF5+Momentum alpha intercept. Must exceed 2.0 to be statistically significant. Higher values indicate stronger evidence against the null (returns = factor-driven beta).
- **R² (%)**: Variance explained by the 6 Fama-French + Momentum factors. Low R² (< 15%) suggests the strategy's returns come from genuine alpha. High R² (> 20%) suggests the strategy is just replicating known factor exposure.

**Key insight**: A strategy can have positive returns and still have zero alpha. If those returns come from market exposure (high R²), you can get the same exposure more cheaply via index funds.


# Null Environments {#nulls}

## Row

### {.tabset}

#### Rejection Heatmap

```{r}
#| label: null-heatmap
#| fig-height: 6
#| fig-width: 10
safe_tar_read("fals_vig_null_heatmap")
```

`r { cap <- safe_tar_read("fals_vig_null_heatmap_caption"); if (!is.null(cap)) paste0("*", cap, "*") else "" }`

#### Rejection Table

```{r}
#| label: null-table
null_tbl <- safe_tar_read("fals_vig_null_table")
cap <- safe_tar_read("fals_vig_null_heatmap_caption")
if (!is.null(null_tbl)) fals_dt(null_tbl, cap)
```

#### Methodology

**What are null environments?**

A null environment generates synthetic return series with no genuine signal, but with realistic statistical properties. We test: "could a strategy generate the observed t-statistic just by chance in this environment?"

For each null environment, we generate M = 100 synthetic series, apply the strategy's logic, and compute the [<abbr title="Newey-West heteroskedasticity-autocorrelation consistent estimator">HAC</abbr> t-statistic](https://en.wikipedia.org/wiki/Newey%E2%80%93West_estimator){target="_blank" rel="noopener noreferrer"}. The rejection rate is the fraction that exceeds the observed t-statistic.

**The 6 null environments** (implemented in [`falsification.R`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/falsification.R){target="_blank" rel="noopener noreferrer"}):

- **White Noise**: iid N(0, sigma). Baseline — if a strategy "works" here, it's pure luck.
- **Regime Vol**: Two-state volatility (high/low). Tests whether signal is just vol timing.
- **MA(1)**: Autocorrelated returns (theta = 0.1). Tests whether signal exploits serial correlation friction.
- **Factor Null**: Random factor loadings matching observed factor covariance structure. Tests whether signal is just factor exposure.
- **<abbr title="Generalized Autoregressive Conditional Heteroskedasticity">GARCH</abbr>(1,1)**: omega = 1e-6, alpha = 0.09, beta = 0.90. Tests whether signal exploits volatility clustering.
- **<abbr title="Glosten-Jagannathan-Runkle GARCH — adds asymmetric leverage effect">GJR-GARCH</abbr>**: Adds asymmetric leverage (gamma = 0.05). Tests whether signal exploits the leverage effect.

**Interpretation:**

- Rejection rate < 8%: Strategy is easily replicated under the null — **no genuine signal**
- Rejection rate 8-20%: Borderline — may be noise
- Rejection rate > 20%: The null cannot explain the strategy's performance — **genuine signal**


# HAC Correction {#hac}

## Row

### {.tabset}

#### Naive vs HAC t-Statistics

```{r}
#| label: hac-plot
#| fig-height: 6
#| fig-width: 10
safe_tar_read("fals_vig_hac_plot")
```

`r { cap <- safe_tar_read("fals_vig_hac_caption"); if (!is.null(cap)) paste0("*", cap, "*") else "" }`

#### HAC Details

```{r}
#| label: hac-table
hac_comp <- safe_tar_read("fals_vig_hac_comparison")
cap <- safe_tar_read("fals_vig_hac_caption")
if (!is.null(hac_comp)) fals_dt(hac_comp, cap)
```

#### Methodology

**Why HAC correction matters:**

Financial returns exhibit volatility clustering ([GARCH](https://en.wikipedia.org/wiki/Autoregressive_conditional_heteroskedasticity) persistence). Standard Sharpe ratios assume iid returns, which inflates the Sharpe by 20-30% on average.

The [Newey-West correction](https://en.wikipedia.org/wiki/Newey%E2%80%93West_estimator) (implemented in [`hd_hac_sharpe()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/falsification.R#L80)):

- Uses a Bartlett kernel to downweight distant autocovariances
- Bandwidth = floor(4 * (T/100)^(2/9)) — automatic, data-driven
- Produces an HAC-consistent variance estimate for the mean return
- The HAC t-statistic = mean(r) / sqrt(HAC_var / T)

**Interpretation:**

- HAC t > 2.0: Robust risk-adjusted return, survives serial correlation correction
- HAC t < 2.0: Naive Sharpe is inflated by autocorrelation; true risk-adjusted performance is weaker


# Factor Regression {#ff}

## Row

### {.tabset}

#### Fama-French 5-Factor + Momentum Betas

```{r}
#| label: ff-table
ff_tbl <- safe_tar_read("fals_vig_ff_table")
cap <- safe_tar_read("fals_vig_ff_caption")
if (!is.null(ff_tbl)) fals_dt(ff_tbl, cap)
```

#### Methodology

**[<abbr title="Fama-French 5-Factor model (MktRF, SMB, HML, RMW, CMA)">FF5</abbr>](https://en.wikipedia.org/wiki/Fama%E2%80%93French_three-factor_model) + [Momentum](https://en.wikipedia.org/wiki/Momentum_(finance)) regression:**

$$r_{excess} = \alpha + \beta_1 \cdot MktRF + \beta_2 \cdot SMB + \beta_3 \cdot HML + \beta_4 \cdot RMW + \beta_5 \cdot CMA + \beta_6 \cdot Mom + \epsilon$$

Where:

- **Mkt-RF**: Market excess return (market - risk-free)
- **<abbr title="Small Minus Big — size factor (small-cap minus large-cap returns)">SMB</abbr>**: Small Minus Big (size factor)
- **<abbr title="High Minus Low — value factor (cheap stocks minus expensive stocks)">HML</abbr>**: High Minus Low (value factor)
- **<abbr title="Robust Minus Weak — profitability factor">RMW</abbr>**: Robust Minus Weak (profitability factor)
- **<abbr title="Conservative Minus Aggressive — investment factor">CMA</abbr>**: Conservative Minus Aggressive (investment factor)
- **<abbr title="Momentum — past 12-month winners minus losers">Mom</abbr>**: Momentum (winners - losers, 12-2 month)

**Alpha (%)** is the intercept — the annualised return not explained by any factor. **R² (%)** measures how much of the strategy's variance is explained by known factors.

**Decision rules** (matching Verdict in the Scorecard):

- Alpha t > 2.0, R² (%) < 15: **Genuine alpha** — returns come from something beyond known factors
- Alpha t > 2.0, R² (%) > 15: Alpha exists but significant factor exposure — investigate
- Alpha t < 2.0, R² (%) > 20: **Disguised beta** — just buy the factors directly
- Alpha t < 2.0, R² (%) < 10: Strategy is noise

**Data source:** [Kenneth French Data Library](https://mba.tuck.dartmouth.edu/pages/faculty/ken.french/data_library.html){target="_blank" rel="noopener noreferrer"}, daily frequency. Implemented in [`hd_factor_null_test()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/falsification.R){target="_blank" rel="noopener noreferrer"}.


# Multiplicity Correction {#multiplicity}

## Row

### {.tabset}

#### K_eff and Delta-Z

```{r}
#| label: multiplicity-table
mult_tbl <- safe_tar_read("fals_vig_multiplicity")
cap <- safe_tar_read("fals_vig_multiplicity_caption")
if (!is.null(mult_tbl)) fals_dt(mult_tbl, cap)
```

#### Methodology

**The multiple testing problem:**

Testing 6 strategies inflates Type I error. The chance that *at least one* strategy exceeds t = 2.0 by pure luck is much higher than 5%.

**<abbr title="Effective number of independent strategies (spectral participation ratio)">K_eff</abbr> (effective multiplicity):**

K_eff = K² / ||Σ||²_F — the spectral participation ratio of the pairwise correlation matrix. It measures how many *truly independent* strategies exist:

- K_eff = K: All strategies are uncorrelated (worst case for multiple testing)
- K_eff = 1: All strategies are perfectly correlated (just one independent bet)
- K_eff between 1 and K: Some independent structure

**Harvey threshold:** sqrt(2 * log(K_eff)) — the t-statistic a strategy must exceed to survive multiplicity correction at the 5% level. Based on Harvey, Liu & Zhu (2016), "...and the Cross-Section of Expected Returns."

**Delta-Z (IS-OOS gap):**

Delta-Z = max(t_IS) - max(t_OOS) — the gap between the best in-sample and best out-of-sample t-statistics. A large gap indicates overfitting:

- Delta-Z < 1: Modest overfitting, results likely robust
- Delta-Z 1-2: Some overfitting, treat with caution
- Delta-Z > 2: Severe overfitting, in-sample performance is misleading

In this framework, "in-sample" = HAC t-stat (strategy returns), "out-of-sample" = FF alpha t-stat (after removing known factor exposure). The OOS test is not a temporal holdout but a *structural* holdout — does the signal survive factor decomposition?


# Covariance Regularisation {#covreg}

Does regularising the covariance estimator — via Ledoit-Wolf (LW) shrinkage or Random Matrix Theory (RMT) denoising — improve out-of-sample (OOS) global minimum-variance (GMV) portfolio performance relative to the plain sample estimator? This section presents a walk-forward diagnostic across two universes per Raviv (2026).

## Row

### {.tabset}

#### OOS Diagnostic

```{r}
#| label: covreg-table
covtab <- safe_tar_read("cov_diag_vig_table")
covcap <- safe_tar_read("cov_diag_vig_caption")
if (!is.null(covtab)) fals_dt(covtab, covcap)
```

#### Conditioning

```{r}
#| label: covreg-cond-plot
#| fig-height: 6
#| fig-width: 10
#| fig-alt: "Cleveland dot plot of mean covariance condition number (κ) on a log10 x-axis for three estimators (Sample, Ledoit-Wolf, RMT-denoise) and two universes. Lower κ means better conditioning. Regularised estimators cluster leftward (lower κ) relative to the Sample estimator, demonstrating conditioning improvement in both universes."
p <- safe_tar_read("cov_diag_vig_cond_plot")
if (!is.null(p)) p
```

`r { cap <- safe_tar_read("cov_diag_vig_caption"); if (!is.null(cap)) paste0("*", cap, "*") else "" }`

#### OOS Sharpe

```{r}
#| label: covreg-sharpe-plot
#| fig-height: 6
#| fig-width: 10
#| fig-alt: "Cleveland dot plot of annualised out-of-sample minimum-variance Sharpe ratio by estimator and universe. The 4-asset universe (circles) shows a clean regularisation benefit: LW and RMT exceed the Sample Sharpe. The wide universe result (triangles, marked survivorship-biased) shows Sample outperforming, but this is confounded by the absence of delisted tickers."
p <- safe_tar_read("cov_diag_vig_sharpe_plot")
if (!is.null(p)) p
```

`r { cap <- safe_tar_read("cov_diag_vig_caption"); if (!is.null(cap)) paste0("*", cap, "*") else "" }`

#### How to Read This

**What is tested:**

A 60-month rolling walk-forward backtest estimates global minimum-variance (GMV) portfolio weights from each covariance estimator using only the training window. The next-period (t+1) return is computed from the resulting weights — strictly no look-ahead (per the `alpha-decay-min-t+1` rule: t+0 execution is impossible). The three estimators compared are:

- **Sample** — classical sample covariance: S = (1/(n−1)) X'X.
- **Ledoit-Wolf (LW)** — linear shrinkage toward a constant-correlation target, with analytically optimal intensity δ* (no cross-validation). Guarantees positive-definiteness for any p and n. Implemented in [`hd_cov_estimate(method="ledoit_wolf")`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/cov_estimate.R#L100){target="_blank" rel="noopener noreferrer"}.
- **RMT-denoise** — Random Matrix Theory (RMT) Marchenko-Pastur clipping: eigenvalues below the noise threshold λ+ = (1+√(p/n))² are replaced by their common mean, removing noise components while preserving trace. Implemented in [`hd_cov_estimate(method="rmt_denoise")`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/cov_estimate.R#L100){target="_blank" rel="noopener noreferrer"}.

GMV weights are computed by [`hd_min_var_weights()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/min_var_weights.R#L65){target="_blank" rel="noopener noreferrer"} from the estimated covariance via unconstrained matrix inversion; see its documentation for the `normalize = TRUE` default (weights sum to 1). The walk-forward engine is [`hd_cov_oos_diagnostic()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/cov_diagnostic.R#L87){target="_blank" rel="noopener noreferrer"}.

**Condition number κ:** max eigenvalue / min eigenvalue. A well-conditioned matrix has κ = O(10–100); κ > 10,000 indicates near-singularity and numerical instability in inversion. The Conditioning tab uses a log₁₀ x-axis because condition numbers span multiple orders of magnitude across methods.

**Honest interpretation:**

The conditioning benefit of regularisation is unambiguous in both universes — Ledoit-Wolf and RMT-denoise produce substantially lower condition numbers than Sample regardless of the p/n ratio.

The out-of-sample (OOS) Sharpe comparison is clean only on the **4-asset universe** (SPY/TLT/GLD/DBC, p/n ≈ 0.07): regularised estimators outperform Sample on Sharpe, confirming the expected benefit in a well-studied, non-survivorship-biased universe.

The **wide universe** result (~30 large-cap US equities, p/n ≈ 0.50) shows Sample OOS Sharpe exceeding LW and RMT. This is **NOT** evidence against regularisation. The wide panel contains only currently-listed tickers — firms that survived to the present day — with no delisted firms included. Survivorship bias inflates all methods' OOS returns, but it especially advantages whichever estimator happens to concentrate more weight in the best-surviving stocks. This is a dataset artefact, not a property of the estimators.

**Why the default COV_METHOD stays "sample":**

The deployed strategy universes (4–6 assets) operate at p/n ≪ 0.1, where the sample covariance is already well-conditioned. The regularisation benefit is most important in wide regimes (p/n ≥ 0.1). COV_METHOD is the documented knob to flip when deploying a wide or point-in-time universe — set `COV_METHOD <- "ledoit_wolf"` in [`R/cov_config.R`](https://github.com/JohnGavin/historical/blob/main/R/cov_config.R){target="_blank" rel="noopener noreferrer"} and re-run `tar_make()`.

**References:** Raviv (2026), *WIREs Computational Statistics*, [doi:10.1002/wics.70068](https://doi.org/10.1002/wics.70068). Full digest: [`knowledge/wiki/covariance-estimation-wide-data.md`](https://github.com/JohnGavin/historical/blob/main/knowledge/wiki/covariance-estimation-wide-data.md){target="_blank" rel="noopener noreferrer"}. Issue [#498](https://github.com/JohnGavin/historical/issues/498){target="_blank" rel="noopener noreferrer"}.


# Walk-Forward Correlation {#wfc}

## Row

### {.tabset}

#### WFC 2×2 Matrix

```{r}
#| label: wfc-table
#| results: asis
wfc_tbl <- safe_tar_read("wfc_all_summary")
if (!is.null(wfc_tbl) && nrow(wfc_tbl) > 0) {
  library(dplyr)
  display_tbl <- wfc_tbl |>
    transmute(
      Strategy      = strategy,
      `Pearson ρ`   = round(wfc_pearson,  3),
      `Spearman ρ`  = round(wfc_spearman, 3),
      `N grid pts`  = wfc_n_points,
      `WFC category` = wfc_category,
      `Median OOS`  = round(median_oos,   3),
      `% OOS > 0`   = paste0(round(pct_pos_oos * 100, 0), "%"),
      Verdict       = classification
    )
  caption_text <- paste0(
    "Walk-Forward Correlation diagnostic (Tinsley 2026, SSRN 6324079). ",
    "Pearson ρ computed across IS↔OOS annualised Sharpe for every parameter ",
    "combination in each strategy’s grid. ",
    "IS window ends ", format(wfc_tbl$is_end[[1]], "%Y-%m-%d"), "; ",
    "OOS window ", format(wfc_tbl$oos_start[[1]], "%Y-%m-%d"),
    " – ", format(wfc_tbl$oos_end[[1]], "%Y-%m-%d"), ". ",
    "Computed ", format(wfc_tbl$computed_on[[1]], "%Y-%m-%d"), "."
  )
  fals_dt(display_tbl, caption_text)
} else {
  cat("*WFC summary not yet computed — run `tar_make()` to build `wfc_all_summary`.*")
}
```

#### Methodology

**What Walk-Forward Correlation measures:**

[WFC](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=6324079){target="_blank" rel="noopener noreferrer"} (Martyn Tinsley, 2026) is a single-number diagnostic for over-fitting across the *whole* optimisation surface. Traditional walk-forward analysis asks "does the best in-sample parameter survive out-of-sample?" WFC instead asks "across *every* parameter combination, does in-sample performance predict out-of-sample performance?"

**Calculation:** For each strategy with a tunable parameter vector θ, compute the in-sample Sharpe X(θ) and out-of-sample Sharpe Y(θ) for every point on the grid. WFC = Pearson ρ(X, Y) over all θ.

**Diagnostic 2×2 matrix (Tinsley p.4):**

|              | High WFC (ρ ≥ 0.70)               | Low WFC (ρ < 0.40)                    |
|--------------|-----------------------------------|---------------------------------------|
| **+OOS**     | `structural_edge` — low over-fitting | `spurious_luck` — high over-fitting  |
| **−OOS**     | `consistently_loss_making`         | `noise` — no edge                    |

**Calibration points from the paper (Figure 4):** high ≈ 0.881, moderate ≈ 0.581, low ≈ 0.234. This project uses 0.70 as the high/moderate boundary (midpoint of moderate–high range).

**Strategies included:**

- **Factor MAX** — parameter grid: `top_n ∈ {1, 2, 3, 4, 5}` (5 grid points)
- **Factor DRIF** — parameter grid: `alpha ∈ {0.00, 0.25, 0.50, 0.75, 1.00}` (5 grid points); lambda chosen by `cv.glmnet` internal CV

**Strategies not included:** OLMAR, TOM, and other non-parametric strategies have no tunable grid. XGB DRIF operates at the stock level (not a factor-rotation grid); factor-level XGB is tracked as a follow-up to [#318](https://github.com/JohnGavin/historical/issues/318){target="_blank" rel="noopener noreferrer"}.

**Reference:** Tinsley, M. (2026). "Walk Forward Correlation." Trade Like A Machine Ltd. [SSRN 6324079](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=6324079){target="_blank" rel="noopener noreferrer"}. Implemented in [`hd_wf_correlation()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/wf_correlation.R){target="_blank" rel="noopener noreferrer"}.


# Assumptions {#assumptions}

## Row

### {.tabset}

#### Framework

**What this framework tests (4 of 5):**

1. **Null environment testing**: Generates 100 synthetic series per environment (6 environments) to test whether a strategy's t-statistic could arise by chance
2. **HAC correction**: Adjusts for serial correlation in returns (GARCH persistence inflates naive Sharpe by ~20-30%)
3. **Factor regression**: Decomposes returns into known factor exposure (beta) and residual alpha
4. **Multiplicity correction**: Adjusts for testing multiple strategies simultaneously (K_eff, delta-Z)

**Not yet implemented (1 of 5):**

5. **<abbr title="Deflated Sharpe Ratio — adjusts for skewness, kurtosis, and number of trials">DSR</abbr>** (Deflated Sharpe Ratio): Lopez de Prado's adjustment for skewness, kurtosis, and number of trials. Corrects the Sharpe ratio for non-normality — a strategy with negative skew (fat left tail) has an inflated Sharpe. See [#53 comment](https://github.com/JohnGavin/historical/issues/53) for implementation plan.

**What this framework does NOT do:**

- Does not test for [look-ahead bias](https://github.com/JohnGavin/historical/blob/main/R/plan_falsification.R){target="_blank" rel="noopener noreferrer"} (handled by the [`look-ahead-bias-prevention`](https://github.com/JohnGavin/historical/blob/main/.claude/rules/look-ahead-bias-prevention.md){target="_blank" rel="noopener noreferrer"} rule)
- Does not test for transaction cost sensitivity (handled by [`execution-delay-sensitivity`](https://github.com/JohnGavin/historical/blob/main/R/plan_alpha_decay.R){target="_blank" rel="noopener noreferrer"} — alpha decay analysis)
- Does not test for parameter sensitivity (handled by [`backtest-robustness`](https://github.com/JohnGavin/historical/blob/main/R/plan_bootstrap_ci.R){target="_blank" rel="noopener noreferrer"} — bootstrap CI)
- Does not test for temporal stability (handled by [`backtest-partitions`](https://github.com/JohnGavin/historical/blob/main/R/plan_partitions.R){target="_blank" rel="noopener noreferrer"} — train/test/validation split)
- Does not test for cross-geography pervasiveness (requires international data — see [#58](https://github.com/JohnGavin/historical/issues/58){target="_blank" rel="noopener noreferrer"})

**References:**

- Harvey, C., Liu, Y., & Zhu, H. (2016). "...and the Cross-Section of Expected Returns." *Review of Financial Studies*, 29(1), 5-68.
- Lopez de Prado, M. (2018). "The Deflated Sharpe Ratio: Correcting for Selection Bias, Backtest Overfitting, and Non-Normality." *Journal of Portfolio Management*, 40(5), 94-107.
- Newey, W. K., & West, K. D. (1987). "A Simple, Positive Semi-definite, Heteroskedasticity and Autocorrelation Consistent Covariance Matrix." *Econometrica*, 55(3), 703-708.
- Fama, E., & French, K. (2015). "A Five-Factor Asset Pricing Model." *Journal of Financial Economics*, 116(1), 1-22.

#### Build Info

```{r}
#| label: build-info
#| results: asis
build_info()
```

# Causal Graph {#causal}

## Row

### {.tabset}

#### Overview

`r { cg <- safe_tar_read("cg_dag"); if (!is.null(cg)) paste0("Simplified view: strategies as single nodes. **", cg$n_nodes, " nodes, ", cg$n_edges, " edges** in the full DAG. Click any strategy node to view its source code. Click diagram to zoom fullscreen.") else "" }`

```{r}
#| label: causal-overview-mount
#| results: asis
#| echo: false
cat('<div id="dag-overview-mount"></div>')
```

#### Full DAG

```{r}
#| label: causal-full-mount
#| results: asis
#| echo: false
cat('<div id="dag-full-mount"></div>')
```

Node colours: <abbr title="Fama-French factors: HML (value), SMB (size), Mom (momentum), RMW (profitability), CMA (investment)">Factors</abbr> (blue) | <abbr title="VIX, VVIX, Fed rate, Inflation — observable macro indicators">Macro</abbr> (gold) | <abbr title="Classified market state: Vol Regime (VIX-based), Rate Regime (Fed-based)">Regime</abbr> (orange) | <abbr title="Strategy-specific trading signals derived from factors and regimes">Signals</abbr> (purple) | <abbr title="Realised strategy returns after applying signals">Returns</abbr> (green) | <abbr title="Portfolio-level metrics: Sharpe ratio, max drawdown">Outcomes</abbr> (red) | <abbr title="Structural factors: factor crowding, premium decay, rebalancing costs">Structural</abbr> (grey). Yellow dashed = <abbr title="Conditional independence test failed (|partial r| >= 0.15) — DAG is missing an edge">violated implication</abbr>. Green = tested, holds. Source: [plan_causal_graph.R](https://github.com/JohnGavin/historical/blob/main/R/plan_causal_graph.R).

`r paste0('<p style="color:#888;font-size:0.8em;">Full DAG tab rendered: ', format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z"), ' | commit: ', system("git rev-parse --short HEAD", intern=TRUE), '</p>')`

#### DRIF

```{r}
#| label: causal-drif-mount
#| results: asis
#| echo: false
cat('<div id="dag-drif-mount"></div>')
```

<abbr title="Dynamic Rotation Into Factors — each month selects the FF factor with highest trailing return">DRIF</abbr> selects the best-performing FF factor each month. Loads on <abbr title="High Minus Low — value factor: cheap stocks minus expensive stocks">HML</abbr>, <abbr title="Small Minus Big — size factor: small-cap minus large-cap returns">SMB</abbr>, <abbr title="Momentum — past 12-month winners minus losers">Mom</abbr>, <abbr title="Robust Minus Weak — profitability factor">RMW</abbr>. Yellow dashed edge: VIX affects HML in crises (<abbr title="Partial correlation between HML and VIX = -0.17, indicating they co-move in crises">r=-0.17</abbr>). Source: [plan_drif.R](https://github.com/JohnGavin/historical/blob/main/R/plan_drif.R).

`r paste0('<p style="color:#888;font-size:0.8em;">DRIF tab rendered: ', format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z"), ' | commit: ', system("git rev-parse --short HEAD", intern=TRUE), '</p>')`

#### Factor MAX

```{r}
#| label: causal-fmax-mount
#| results: asis
#| echo: false
cat('<div id="dag-fmax-mount"></div>')
```

<abbr title="Factor Momentum MAX — selects the single FF factor with highest recent momentum">Factor MAX</abbr> selects the top-momentum factor. Premium has <abbr title="Strategy Sharpe ratio in early period vs late period. >50% decay means the edge has halved.">decayed >50%</abbr> (#73). Source: [plan_factormax.R](https://github.com/JohnGavin/historical/blob/main/R/plan_factormax.R).

#### LTR

```{r}
#| label: causal-ltr-mount
#| results: asis
#| echo: false
cat('<div id="dag-ltr-mount"></div>')
```

<abbr title="LambdaMART — XGBoost-based cross-sectional momentum strategy. Ranks stocks by predicted return using gradient-boosted trees.">LTR</abbr> uses XGBoost for cross-sectional stock ranking. Loads on <abbr title="Momentum — past 12-month winners minus losers">Mom</abbr> + <abbr title="Small Minus Big — size factor">SMB</abbr>. <abbr title="Fed rate above/below median classifies as high/low rate regime">Rate regime</abbr> feeds into signal. Source: [plan_ltr_momentum.R](https://github.com/JohnGavin/historical/blob/main/R/plan_ltr_momentum.R).

#### Strategy & Dashboard Relationship Map

This diagram is the first version (Phase-0 prototype) of the master navigation
layer described in [issue #481](https://github.com/JohnGavin/historical/issues/481)
and built as part of [issue #514](https://github.com/JohnGavin/historical/issues/514).
It shows the **MVO / portfolio-construction subsystem** shipped in issues #498
(covariance estimator), #512 (returns shrinkage), and #513 (Black-Litterman),
and the two dashboards that surface its outputs.

Each node is clickable and links to the exact function definition in source,
pinned to commit `b346d80`. Hover any edge for a tooltip explaining the
data-flow. Future PRs extend this seed diagram to the full strategy graph
(factor layer, DRIF / FMAX / LTR / RSC nodes, and cross-dashboard navigation).

```{r}
#| label: causal-portcons-mount
#| results: asis
#| echo: false
cat('<div id="dag-portcons-mount" style="min-height:580px;"></div>')
```

Node colours follow the [diagram colour palette](https://github.com/JohnGavin/llm/blob/main/.claude/rules/mermaid-dashboard-pattern.md):
<abbr title="Covariance config + estimator — data inputs to the weight engines">Covariance Estimator</abbr> (subgraph) |
<abbr title="Expected-returns shrinkage applied before Black-Litterman">Expected Returns</abbr> (subgraph) |
<abbr title="GMV, BL, HRP, TRP, PSO — portfolio weight engines">Weight Engines</abbr> (subgraph) |
<abbr title="hd_cov_oos_diagnostic() — walk-forward OOS conditioning and Sharpe comparison">OOS Diagnostic</abbr> (single node) |
<abbr title="falsification.qmd #covreg and leaderboard.qmd — the pages that render the outputs">Dashboards</abbr> (subgraph).

#### Implication Tests

```{r}
#| label: causal-implications
#| echo: false
cg_impl <- safe_tar_read("cg_test_implications")
if (!is.null(cg_impl) && is.data.frame(cg_impl$results)) {
  # Add source code links to implication labels.
  # Node → file + line is the single source of truth in R/diagram_node_links.R.
  diag_links_path <- if (file.exists("../R/diagram_node_links.R")) "../R/diagram_node_links.R" else
    file.path(here::here(), "R/diagram_node_links.R")
  source(diag_links_path)
  .dnl <- diagram_node_links()
  # Build a named vector: node → full GitHub URL with #L anchor
  node_links <- stats::setNames(
    vapply(.dnl$node, gh_url, character(1L)),
    .dnl$node
  )
  link_node <- function(name) {
    for (nm in names(node_links)) {
      if (grepl(nm, name, fixed = TRUE)) {
        url <- node_links[[nm]]
        name <- gsub(nm, paste0("<a href='", url, "' target='_blank'>", nm, "</a>"), name, fixed = TRUE)
      }
    }
    name
  }
  impl_display <- cg_impl$results |>
    dplyr::mutate(label = sapply(label, link_node)) |>
    dplyr::select(Implication = label, `Partial r` = partial_cor,
                  Status = status, Holds = holds)
  DT::datatable(
    impl_display,
    escape = FALSE,
    rownames = FALSE, filter = "none",
    options = list(pageLength = 10, dom = "t"),
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left;",
      htmltools::HTML(paste0("Conditional independence tests via partial correlation. ",
             cg_impl$n_holds, " of ", cg_impl$n_implications,
             " implications hold (|r| &lt; 0.15). ",
             "Violated implications suggest missing edges in the DAG. ",
             "Source: <a href='https://github.com/JohnGavin/historical/blob/main/R/plan_causal_graph.R' target='_blank'>plan_causal_graph.R</a>, tested on monthly data."))
    )
  ) |>
    DT::formatRound("Partial r", digits = 3)
} else {
  cat("*Implication tests not yet built. Run `tar_make(names = 'cg_test_implications')`.*")
}
```

#### Interpretation

**What the DAG encodes:** Every arrow is a causal assumption from our strategy pipeline — "X causes Y" or "X determines Y". If the assumption is wrong, the arrow should be removed or redirected.

**How to read the tests:** Each conditional independence claim says "if our DAG is correct, then X and Y should be unrelated after controlling for Z." A partial correlation near zero confirms the claim; a large partial correlation violates it.

**Violated: HML ⊥ VIX (r = -0.17)**

The value factor (HML) and volatility index (VIX) are NOT independent. This means:

| Impact area | What it means | Action |
|-------------|---------------|--------|
| **[DRIF strategy](https://github.com/JohnGavin/historical/blob/main/R/plan_drif.R)** | DRIF loads on HML. During VIX spikes, DRIF underperforms MORE than the Vol_regime node alone predicts. | Consider reducing DRIF weight in crisis regimes |
| **[Multi-strategy portfolio](https://github.com/JohnGavin/historical/blob/main/R/plan_multi_strategy.R) ([#52](https://github.com/JohnGavin/historical/issues/52))** | Diversification between VIX overlay and value-heavy strategies is LESS than K_eff suggests — they correlate in crises, exactly when diversification matters most | Recompute weights with crisis-conditional correlations |
| **[Tail independence](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/falsification.R) ([#55](https://github.com/JohnGavin/historical/issues/55))** | K_eff_crisis < K_eff_calm is expected — this finding explains why | Run [`hd_tail_keff()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/falsification.R) to quantify |
| **[DAG revision](https://github.com/JohnGavin/historical/blob/main/R/plan_causal_graph.R)** | Need either: VIX → HML edge (direct), or latent "Crisis_mode" → {VIX, HML} | Added VIX → HML edge (dashed yellow in diagram) |

**Economic explanation:** In crises, investors sell risky assets indiscriminately. Value stocks (high book-to-market) are often distressed firms — they get sold first. Simultaneously, VIX spikes as options demand surges. The correlation is not VIX *causing* value to fall, but both responding to the same flight-to-safety mechanism.

**Temporal split finding:** The HML-VIX link was stronger pre-2010 (r=-0.26) and has weakened post-2010 (r=-0.14). This is the opposite of the crowding hypothesis — the crisis-value correlation peaked during 2000-2010 (dot-com crash + GFC), not after factor investing became popular. The VIX-HML edge in the DAG is most relevant for crisis periods, less so in recent calm markets.

```{r}
#| label: causal-split-table
#| echo: false
split <- safe_tar_read("cg_test_split")
if (!is.null(split) && !is.null(split$comparison)) {
  DT::datatable(
    split$comparison,
    rownames = FALSE, filter = "none",
    options = list(pageLength = 5, dom = "t"),
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left;",
      paste0("HML-VIX correlation by period. Pre-2010 includes dot-com crash + GFC. ",
             "Source: ", "[plan_causal_graph.R]",
             "(https://github.com/JohnGavin/historical/blob/main/R/plan_causal_graph.R).")
    )
  ) |> DT::formatRound("hml_vix_r", digits = 3)
} else {
  cat("*Split test not yet built.*")
}
```

`r { cap <- safe_tar_read("cg_caption"); if (!is.null(cap)) cap else "" }`

# Risk Architecture {#risk-arch}

## Row

### {.tabset}

#### Drawdown Duration + Loss Clustering

The Tinsley framework (Pillar 8) distinguishes two types of drawdown:
**survivable steady drawdowns** (losses spread across many periods, allowing
recovery) versus **unsurvivable clustered drawdowns** (losses arrive in bursts,
exhausting capital before recovery). Two metrics quantify this distinction:

- **Avg / Max DD duration** — how long drawdowns last on average and at worst.
  Longer durations imply the strategy requires patience to survive; very long
  max-duration periods indicate structural fragility.
- **Loss clustering** (`clustered = TRUE`) — detected when the Wald-Wolfowitz
  runs test rejects independence of the sign-of-return sequence (p < 0.05)
  *and* the lag-1 autocorrelation exceeds 0.2. Both signals must agree: the
  runs test catches non-random alternation patterns; the ACF catches momentum /
  persistence. A clustered strategy has correlated losses — a drawdown predicts
  more drawdown.

```{r}
#| label: risk-arch-table
#| echo: false
fals_db <- safe_tar_read("fals_results_db")
if (!is.null(fals_db) && nrow(fals_db) > 0 &&
    all(c("avg_dd_duration_days", "max_dd_duration_days",
          "loss_runs_p", "loss_acf_lag1", "loss_clustered",
          "max_consecutive_losses") %in% names(fals_db))) {

  risk_tbl <- fals_db |>
    dplyr::select(
      Strategy         = strategy_id,
      `Avg DD days`    = avg_dd_duration_days,
      `Max DD days`    = max_dd_duration_days,
      `Runs p-val`     = loss_runs_p,
      `ACF lag-1`      = loss_acf_lag1,
      `Clustered`      = loss_clustered,
      `Max Cons. Losses` = max_consecutive_losses
    ) |>
    dplyr::mutate(
      Strategy  = dplyr::case_match(
        Strategy,
        "avoid_worst" ~ "Avoid Worst",
        "drif"        ~ "DRIF",
        "fac_max"     ~ "Factor MAX",
        "rsc"         ~ "RSC",
        "ltr"         ~ "LTR",
        "tom"         ~ "TOM",
        .default      = Strategy
      ),
      Clustered = dplyr::case_when(
        is.na(Clustered)  ~ "N/A",
        isTRUE(Clustered) ~ "YES",
        TRUE              ~ "no"
      )
    )

  # Dynamic caption values
  worst_max_dd_strat <- risk_tbl |>
    dplyr::filter(!is.na(`Max DD days`)) |>
    dplyr::slice_max(`Max DD days`, n = 1L) |>
    dplyr::pull(Strategy)
  worst_max_dd_days <- risk_tbl |>
    dplyr::filter(!is.na(`Max DD days`)) |>
    dplyr::slice_max(`Max DD days`, n = 1L) |>
    dplyr::pull(`Max DD days`)
  clustered_strats <- risk_tbl |>
    dplyr::filter(Clustered == "YES") |>
    dplyr::pull(Strategy)
  clustered_label <- if (length(clustered_strats) == 0L) {
    "none"
  } else {
    paste(clustered_strats, collapse = ", ")
  }
  worst_cons_losses_strat <- risk_tbl |>
    dplyr::filter(!is.na(`Max Cons. Losses`)) |>
    dplyr::slice_max(`Max Cons. Losses`, n = 1L) |>
    dplyr::pull(Strategy)
  worst_cons_losses_n <- risk_tbl |>
    dplyr::filter(!is.na(`Max Cons. Losses`)) |>
    dplyr::slice_max(`Max Cons. Losses`, n = 1L) |>
    dplyr::pull(`Max Cons. Losses`)

  DT::datatable(
    risk_tbl,
    rownames = FALSE, filter = "none",
    options = list(pageLength = 10, dom = "t"),
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left;",
      htmltools::HTML(paste0(
        "Tinsley Pillar 8 risk-architecture metrics. ",
        "Worst max drawdown duration: <b>",
        worst_max_dd_strat, "</b> (",
        worst_max_dd_days, " days). ",
        "Strategies with clustered losses (runs p &lt; 0.05 AND ACF lag-1 &gt; 0.2): <b>",
        clustered_label, "</b>. ",
        "Longest consecutive losing run: <b>",
        worst_cons_losses_strat, "</b> (",
        worst_cons_losses_n, " trades). ",
        "Clustering indicates losses arrive in bursts rather than uniformly, ",
        "reducing practical survivability. ",
        "Source: <a href='https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/risk_metrics.R' target='_blank'>",
        "hd_dd_duration() + hd_loss_clustering()</a> + ",
        "<a href='https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/trades.R' target='_blank'>",
        "hd_trade_metrics()</a> in packages/historicaldata, ",
        "wired via <a href='https://github.com/JohnGavin/historical/blob/main/R/plan_falsification.R' target='_blank'>",
        "plan_falsification.R</a>."
      ))
    )
  ) |>
    DT::formatRound(c("Runs p-val", "ACF lag-1"), digits = 3) |>
    DT::formatStyle(
      "Clustered",
      backgroundColor = DT::styleEqual(
        c("YES", "no", "N/A"),
        c("#c0392b", "transparent", "transparent")
      ),
      color = DT::styleEqual(
        c("YES", "no", "N/A"),
        c("#ffffff", "inherit", "inherit")
      )
    )
} else {
  cat("*Risk architecture metrics not yet built. Run `tar_make(names = 'fals_results_db')`.*")
}
```

#### Methodology

**Drawdown duration** uses the high-water mark (HWM) definition: a drawdown
begins when cumulative returns fall below the prior peak and ends when they
recover to within 0.1% of that peak. Durations are converted from strategy
observations to calendar days using each strategy's annualisation factor (252
for daily, 12 for monthly).

**Loss clustering test** combines two signals:

1. [Wald-Wolfowitz runs test](https://en.wikipedia.org/wiki/Wald%E2%80%93Wolfowitz_runs_test)
   on the sign of returns. Under independence, the number of sign-runs follows
   an approximately normal distribution. Too few runs = clustered
   losses/wins; too many = alternating.
2. Lag-1 autocorrelation of raw returns. Positive ACF(1) > 0.2 indicates
   momentum (losses predict more losses).

The `clustered` flag requires **both** signals to agree to reduce false
positives. The runs test has power against any non-random pattern; ACF
specifically targets persistence.

**Source code:**
[`hd_dd_duration()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/risk_metrics.R)
and
[`hd_loss_clustering()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/risk_metrics.R)
in `packages/historicaldata/R/risk_metrics.R`. Tests in
[`test-risk-metrics.R`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/tests/testthat/test-risk-metrics.R).
Issue [#269](https://github.com/JohnGavin/historical/issues/269).

#### Related Vignettes

- **[Strategy Leaderboard](leaderboard.html)** — displays the falsification scorecard verdicts alongside performance rankings; leaderboard surfaces what this page proves
- **[Evidence & Context](evidence.html)** — catalogues strategies that failed the falsification gauntlet; companion document to this page's verdicts
- **[Momentum Pre-Peak](momentum-prepeak.html)** — the most recent gauntlet run; applies HAC, FF5, CPCV PBO, and Walk-Forward Correlation from this framework

