---
title: "historicaldata"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    nav-buttons: [github]
    embed-resources: false
    include-in-header:
      - text: |
          <style>
          body { font-size: 1.2em; }
          /* Kill dashboard fill layout */
          .bslib-page-fill, .html-fill-container, .html-fill-item,
          .bslib-card, .card, .tab-pane, .tab-content, .card-body,
          .bslib-grid, .bslib-grid-item {
            height: auto !important; min-height: 0 !important;
            max-height: none !important; overflow: visible !important;
            flex: none !important;
          }
          .card-grid { display: grid !important; grid-template-columns: 1fr 1fr !important;
                       gap: 1.2em !important; width: 100% !important; }
          .info-card { background: var(--bs-card-bg, #1a1a1a); border: 1px solid var(--bs-border-color, #333);
                       border-radius: 8px; padding: 1.2em; text-decoration: none;
                       color: inherit; transition: border-color 0.2s; display: block; }
          .info-card:hover { border-color: #6B8DAE; text-decoration: none; color: inherit; }
          .info-card h4 { margin: 0 0 0.4em; }
          .info-card p { opacity: 0.75; font-size: 0.9em; margin: 0; }
          .stats-row { display: flex !important; flex-direction: row !important;
                       flex-wrap: nowrap !important; gap: 2.5em !important;
                       justify-content: center !important; align-items: center !important;
                       margin: 1em 0 !important; width: 100% !important; }
          .stat { text-align: center; flex-shrink: 0; }
          .stat .num { font-size: 1.8em; font-weight: bold; }
          .stat .label { font-size: 0.8em; opacity: 0.6; }
          .badge { display: inline-block; padding: 0.2em 0.6em; border-radius: 4px;
                   font-size: 0.8em; margin-right: 0.5em; }
          .badge-blue { background: #1a3a5c; color: #6B8DAE; }
          .badge-green { background: #1a3c1a; color: #7A9E7E; }
          .badge-gold { background: #3c3520; color: #C4A35A; }
          /* Smaller page titles */
          .navbar .navbar-title, h1.title { font-size: 1.4em !important; }
          .nav-link { font-size: 0.95em !important; }
          @media (max-width: 600px) {
            .card-grid { grid-template-columns: 1fr !important; }
            .stats-row { flex-wrap: wrap !important; gap: 1em !important; }
          }
          </style>
          <script>
          // Default to dark mode
          if (!localStorage.getItem('quarto-color-scheme')) {
            document.documentElement.setAttribute('data-bs-theme', 'dark');
          }
          </script>
execute:
  echo: false
---

```{r}
#| label: setup
#| include: false
```

# Datasets

## Row

Historical finance database for backtesting: equities, crypto, macro, and factors. Code-only R package — data hosted on [Hugging Face](https://huggingface.co/datasets/JohnGavin/finance-data), queried via DuckDB `hf://` protocol with predicate pushdown.

<span class="badge badge-blue">R package</span>
<span class="badge badge-green">DuckDB + Parquet</span>
<span class="badge badge-gold">Hugging Face</span>

<div class="stats-row">
<div class="stat"><div class="num">1,022</div><div class="label">tickers</div></div>
<div class="stat"><div class="num">2.3M</div><div class="label">rows</div></div>
<div class="stat"><div class="num">6</div><div class="label">datasets</div></div>
<div class="stat"><div class="num">100yr</div><div class="label">history</div></div>
<div class="stat"><div class="num">30</div><div class="label">functions</div></div>
</div>

## Row

<div class="card-grid">
<a href="examples.html#equity" class="info-card">
<h4>Equity</h4>
<p>1,022 tickers (51 US + 949 LSE ETFs + 13 macro ETFs + 9 factor ETFs). Daily OHLCV from 1980. Includes Vanguard, iShares, SPDR UCITS ETFs. 44 MB Parquet.</p>
</a>
<a href="examples.html#crypto" class="info-card">
<h4>Crypto</h4>
<p>14 tokens (BTC, ETH, SOL, BNB, stablecoins, Solana DeFi). Daily from 2014. 1.1 MB.</p>
</a>
<a href="examples.html#macro" class="info-card">
<h4>Macro</h4>
<p>19 FRED series: SP500, <abbr title="CBOE Volatility Index — implied volatility of S&P 500 options">VIX</abbr>, Treasury yields, credit spreads, GDP, CPI, oil, USD. From 1947.</p>
</a>
<a href="examples.html#factors" class="info-card">
<h4>Factors</h4>
<p>Fama-French FF3, <abbr title="Fama-French 5-Factor model (MktRF, SMB, HML, RMW, CMA)">FF5</abbr>, Momentum. Daily + monthly. Back to 1926. Ken French Data Library.</p>
</a>
</div>

# Articles

## Row

<div class="card-grid">
<a href="leaderboard.html" class="info-card">
<h4>Strategy Leaderboard</h4>
<p>Rankings, equity curves, and feature importance across all 6 strategies. Transposed metrics table, PSO optimal portfolio weights, XGBoost feature analysis.</p>
</a>
<a href="examples.html" class="info-card">
<h4>Usage Examples</h4>
<p>Interactive dashboard with code examples for all 6 datasets: equity, crypto, macro, factors, LSE ETFs, and metadata queries.</p>
</a>
<a href="macro-defense-rotation.html" class="info-card">
<h4>Macro Defense Rotation</h4>
<p>Backtesting vignette: rotating among TLT/GLD/DBC/UUP to hedge macro risk. Expanding window methodology, in-sample vs out-of-sample holdout.</p>
</a>
<a href="factor-max.html" class="info-card">
<h4>Factor MAX Rotation</h4>
<p>Rotating among Fama-French factors using the MAX signal — maximum daily return predicts next-month factor returns. Based on Alpha Architect research (Dec 2025).</p>
</a>
<a href="drif.html" class="info-card">
<h4><abbr title="Dynamic Rotation Into Factors — selects best-performing FF factor each month">DRIF</abbr>: Daily Return Information</h4>
<p>Elastic net on the full distribution of past 21 daily factor returns (chronological + rank) predicts next-month returns. Subsumes MAX and momentum signals.</p>
</a>
<a href="stock-backtest.html" class="info-card">
<h4>Stock-Level Backtests</h4>
<p>MAX and DRIF signals applied to 660 individual stocks (S&P 500 + STOXX 600). Decile long-short portfolios. 4-strategy comparison: stock vs factor, MAX vs DRIF.</p>
</a>
<a href="avoid-worst-days.html" class="info-card">
<h4>Avoid Worst Days</h4>
<p>Thought experiment: removing the market's N worst days improves returns more than missing the N best days hurts. Demonstrates return asymmetry and why downside protection matters.</p>
</a>
<a href="falsification.html" class="info-card">
<h4>Falsification Tests</h4>
<p>Causal integrity tests: Fama-French alpha, null-signal heatmaps, HAC-robust Sharpe ratios, implication chains. Evidence our strategies aren't luck or factor exposure.</p>
</a>
<a href="european-overlay.html" class="info-card">
<h4>European Overlay</h4>
<p>CISS vs VIX regime comparison: ECB's Composite Indicator of Systemic Stress as European-native vol proxy. 5 European ETFs, OOS performance, attribution.</p>
</a>
<a href="jst-dashboard.html" class="info-card">
<h4><abbr title="Jordà-Schularick-Taylor Macrohistory Database">JST</abbr> Macrohistory: Long-Run Equity Returns</h4>
<p>155 years of equity premium across 18 countries (1870-2020). Cross-geography pervasiveness test, crisis timeline, decade heatmaps. Validates Swedroe evidence-based framework.</p>
</a>
<a href="pkg/index.html" class="info-card">
<h4>API Reference</h4>
<p>Full function documentation for all 30 exported functions, grouped by: Data Access, Discovery, Curated Groups, Quality &amp; Audit, Infrastructure, Visualisation.</p>
</a>
</div>

# Links

## Row

- [GitHub](https://github.com/JohnGavin/historical) — source code, issues, CI
- [Hugging Face Dataset](https://huggingface.co/datasets/JohnGavin/finance-data) — Parquet files, dataset card
- [Strategy Leaderboard](leaderboard.html) — rankings, equity curves, cost-adjusted metrics
- [API Reference](pkg/index.html) — pkgdown function docs
- [Examples](examples.html) — interactive dashboard with all datasets
- [Macro Defense Rotation](macro-defense-rotation.html) — backtesting vignette
- [Factor MAX](factor-max.html) — factor rotation strategy
- [DRIF](drif.html) — daily return information factor
- [Stock-Level Backtests](stock-backtest.html) — 660 stocks, 4-strategy comparison
- [Avoid Worst Days](avoid-worst-days.html) — return asymmetry thought experiment
- [Falsification Tests](falsification.html) — causal integrity, FF alpha, HAC-robust Sharpe
- [European Overlay](european-overlay.html) — CISS vs VIX regime comparison
- [JST Macrohistory](jst-dashboard.html) — 155 years equity premium, 18 countries, pervasiveness test

## Row

Built with the [T language](https://github.com/b-rodrigues/tlang) pipeline, [targets](https://docs.ropensci.org/targets/), [DuckDB](https://duckdb.org/), and [Quarto](https://quarto.org/). Data from Yahoo Finance, FRED, and Ken French Data Library.

# Installation

## Row

### {.tabset}

#### R (pak)

```r
# Install from GitHub
pak::pak("JohnGavin/historical", subdir = "packages/historicaldata")

# Then:
library(historicaldata)
hd_ohlcv("AAPL", from = "2024-01-01")
hd_ohlcv(hd_group("FAANG"), from = "2024-01-01")  # batch query
hd_macro(c("SP500", "VIXCLS"), from = "2024-01-01")
hd_search("vanguard")  # search metadata
```

#### Python (DuckDB)

No R needed — query the same data directly from Python using DuckDB's native `hf://` protocol:

```python
import duckdb

con = duckdb.connect()
df = con.sql("""
    SELECT date, close, volume
    FROM 'hf://datasets/JohnGavin/finance-data/equity_daily.parquet'
    WHERE ticker = 'AAPL' AND date >= '2024-01-01'
    ORDER BY date
""").df()
```

#### Nix

```bash
# Enter reproducible Nix shell (all R + Python packages included)
nix develop github:JohnGavin/historical
```

Then start R and load the package. All dependencies (dplyr, ggplot2, scales, DT, arrow, duckdb) are pre-installed in the Nix shell.

#### T language

```bash
# Clone, enter Nix shell, fetch data, run pipeline
git clone https://github.com/JohnGavin/historical.git && cd historical
nix develop
python scripts/fetch_equity.py     # 51 equity tickers
Rscript scripts/fetch_crypto.R     # 14 crypto tokens
Rscript scripts/fetch_macro.R      # 19 FRED series
Rscript scripts/fetch_factors.R    # Fama-French factors
t run src/pipeline.t               # T pipeline
```

# Build Info

## Row

```{r}
#| label: build-info
#| echo: false
#| results: asis
utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)
build_info()
```
