---
title: "historicaldata"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    nav-buttons: [github]
    embed-resources: false
    include-in-header:
      - text: |
          <style>
          body { font-size: 1.2em; }
          /* Kill dashboard fill layout */
          .bslib-page-fill, .html-fill-container, .html-fill-item,
          .bslib-card, .card, .tab-pane, .tab-content, .card-body,
          .bslib-grid, .bslib-grid-item {
            height: auto !important; min-height: 0 !important;
            max-height: none !important; overflow: visible !important;
            flex: none !important;
          }
          .card-grid { display: grid !important; grid-template-columns: 1fr 1fr !important;
                       gap: 1.2em !important; width: 100% !important; }
          .info-card { background: var(--bs-card-bg, #1a1a1a); border: 1px solid var(--bs-border-color, #333);
                       border-radius: 8px; padding: 1.2em; text-decoration: none;
                       color: inherit; transition: border-color 0.2s; display: block; }
          .info-card:hover { border-color: #6B8DAE; text-decoration: none; color: inherit; }
          .info-card h4 { margin: 0 0 0.4em; }
          .info-card p { opacity: 0.75; font-size: 0.9em; margin: 0; }
          .stats-row { display: flex !important; flex-direction: row !important;
                       flex-wrap: nowrap !important; gap: 2.5em !important;
                       justify-content: center !important; align-items: center !important;
                       margin: 1em 0 !important; width: 100% !important; }
          .stat { text-align: center; flex-shrink: 0; display: inline-block; }
          .stat .num { font-size: 1.8em; font-weight: bold; display: block; }
          .stat .label { font-size: 0.8em; opacity: 0.6; display: block; }
          .badge { display: inline-block; padding: 0.2em 0.6em; border-radius: 4px;
                   font-size: 0.8em; margin-right: 0.5em; }
          .badge-blue { background: #1a3a5c; color: #6B8DAE; }
          .badge-green { background: #1a3c1a; color: #7A9E7E; }
          .badge-gold { background: #3c3520; color: #C4A35A; }
          /* Smaller page titles */
          .navbar .navbar-title, h1.title { font-size: 1.4em !important; }
          .nav-link { font-size: 0.95em !important; }
          @media (max-width: 600px) {
            .card-grid { grid-template-columns: 1fr !important; }
            .stats-row { flex-wrap: wrap !important; gap: 1em !important; }
          }
          </style>
          <script>
          // Default to dark mode
          if (!localStorage.getItem('quarto-color-scheme')) {
            document.documentElement.setAttribute('data-bs-theme', 'dark');
          }
          </script>
execute:
  echo: false
---

```{r}
#| label: setup
#| include: false
# Compute landing-page stats dynamically at render time.
# Each remote query is wrapped in tryCatch; fallback is "—" so render
# succeeds even when HuggingFace is offline.

local({
  # Path-checked pkgload (cwd-trap fix: here::here() returns docs/ under quarto render) — #387
  pkg_path <- if (dir.exists("packages/historicaldata")) "packages/historicaldata" else
    file.path(dirname(here::here()), "packages/historicaldata")
  suppressMessages(pkgload::load_all(pkg_path, quiet = TRUE))

  n_tickers <- tryCatch({
    tickers_equity <- hd_tickers("equity_daily")
    tickers_crypto <- hd_tickers("crypto_daily")
    length(unique(c(tickers_equity, tickers_crypto)))
  }, error = function(e) NA_integer_)

  n_datasets <- tryCatch(length(hd_datasets()), error = function(e) NA_integer_)

  n_rows <- tryCatch({
    ds_names <- names(hd_datasets())
    row_counts <- sapply(ds_names, function(nm) {
      tryCatch({
        ds <- hd_datasets()[[nm]]
        duckplyr::read_parquet_duckdb(ds$url) |>
          dplyr::count() |>
          dplyr::collect() |>
          dplyr::pull(n)
      }, error = function(e) 0L)
    })
    sum(row_counts)
  }, error = function(e) NA_integer_)

  yr_history <- tryCatch({
    # Collect per-dataset min/max as character first (avoids POSIXct/numeric
    # conversion hazard when unlist() strips Date class → epoch seconds).
    date_datasets <- c("equity_daily", "crypto_daily", "macro_daily", "factors")
    date_pairs <- lapply(date_datasets, function(nm) {
      tryCatch({
        ds <- hd_datasets()[[nm]]
        row <- duckplyr::read_parquet_duckdb(ds$url) |>
          dplyr::summarise(
            min_date = min(date, na.rm = TRUE),
            max_date = max(date, na.rm = TRUE)
          ) |>
          dplyr::collect()
        # Coerce each value to Date via character to avoid epoch-seconds hazard
        list(
          min = as.Date(as.character(row$min_date[[1]])),
          max = as.Date(as.character(row$max_date[[1]]))
        )
      }, error = function(e) NULL)
    })
    date_pairs <- Filter(Negate(is.null), date_pairs)
    if (length(date_pairs) == 0L) return(NA_integer_)
    all_min <- as.Date(sapply(date_pairs, function(p) as.character(p$min)),
                       origin = "1970-01-01")
    all_max <- as.Date(sapply(date_pairs, function(p) as.character(p$max)),
                       origin = "1970-01-01")
    overall_min <- min(all_min, na.rm = TRUE)
    overall_max <- max(all_max, na.rm = TRUE)
    as.integer(round(as.numeric(difftime(overall_max, overall_min,
                                         units = "days")) / 365.25))
  }, error = function(e) NA_integer_)

  # Count exported functions from pkgctx-generated API doc (dynamic-prose-values rule).
  # Pattern "^kind: function" matches one line per exported function.
  # Source: docs/api-historicaldata.md (generated by scripts/regen_api_context.sh).
  n_functions <- tryCatch({
    api_path <- if (file.exists("api-historicaldata.md")) "api-historicaldata.md" else
      file.path(here::here(), "api-historicaldata.md")
    lines <- readLines(api_path, warn = FALSE)
    sum(grepl("^kind: function", lines))
  }, error = function(e) NA_integer_)

  # Format helpers
  fmt_tickers   <- if (is.na(n_tickers))   "—" else format(n_tickers, big.mark = ",")
  fmt_rows      <- if (is.na(n_rows))      "—" else paste0(round(n_rows / 1e6, 1), "M")
  fmt_datasets  <- if (is.na(n_datasets))  "—" else as.character(n_datasets)
  fmt_yr        <- if (is.na(yr_history))  "—" else paste0(yr_history, "yr")
  fmt_functions <- if (is.na(n_functions)) "—" else as.character(n_functions)

  # Export to parent env so inline R blocks can reference them
  assign("stat_tickers",   fmt_tickers,   envir = parent.env(environment()))
  assign("stat_rows",      fmt_rows,      envir = parent.env(environment()))
  assign("stat_datasets",  fmt_datasets,  envir = parent.env(environment()))
  assign("stat_yr",        fmt_yr,        envir = parent.env(environment()))
  assign("stat_functions", fmt_functions, envir = parent.env(environment()))
})
```

# Datasets

## Row

Historical finance database for backtesting: equities, crypto, macro, and factors. Code-only R package — data hosted on [Hugging Face](https://huggingface.co/datasets/JohnGavin/finance-data){target="_blank" rel="noopener noreferrer"}, queried via DuckDB `hf://` protocol with predicate pushdown.

<a href="https://github.com/JohnGavin/historical/tree/main/packages/historicaldata" target="_blank" rel="noopener noreferrer"><span class="badge badge-blue">R package</span></a>
<a href="https://duckdb.org/docs/data/parquet/overview.html" target="_blank" rel="noopener noreferrer"><span class="badge badge-green">DuckDB + Parquet</span></a>
<a href="https://huggingface.co/datasets/JohnGavin/finance-data" target="_blank" rel="noopener noreferrer"><span class="badge badge-gold">Hugging Face</span></a>

<div class="stats-row">
<a href="https://huggingface.co/datasets/JohnGavin/finance-data" target="_blank" rel="noopener noreferrer" class="stat-link"><span class="stat"><span class="num">`r stat_tickers`</span><span class="label">tickers</span></span></a>
<a href="https://huggingface.co/datasets/JohnGavin/finance-data" target="_blank" rel="noopener noreferrer" class="stat-link"><span class="stat"><span class="num">`r stat_rows`</span><span class="label">rows</span></span></a>
<a href="https://huggingface.co/datasets/JohnGavin/finance-data" target="_blank" rel="noopener noreferrer" class="stat-link"><span class="stat"><span class="num">`r stat_datasets`</span><span class="label">datasets</span></span></a>
<a href="jst-dashboard.html" class="stat-link"><span class="stat"><span class="num">`r stat_yr`</span><span class="label">history</span></span></a>
<a href="https://github.com/JohnGavin/historical/blob/main/docs/api-historicaldata.md" target="_blank" rel="noopener noreferrer" class="stat-link"><span class="stat"><span class="num">`r stat_functions`</span><span class="label">functions</span></span></a>
</div>

## Row

<details open>
<summary><strong>Equity</strong> — 1,022 tickers · daily OHLCV from 1980 · 44 MB Parquet</summary>
<ul>
<li><strong>US equities</strong> (51 tickers) — large-cap S&P 500 sample from Yahoo Finance · <a href="https://huggingface.co/datasets/JohnGavin/finance-data" target="_blank" rel="noopener noreferrer">HF dataset</a> · <a href="https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/registry.R#L31" target="_blank" rel="noopener noreferrer"><code>hd_datasets()[["equity_daily"]]</code></a></li>
<li><strong>LSE ETFs</strong> (949 tickers) — UK-listed UCITS ETFs from iShares, Vanguard, SPDR · <a href="https://huggingface.co/datasets/JohnGavin/finance-data" target="_blank" rel="noopener noreferrer">HF dataset</a> · <a href="https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/registry.R#L31" target="_blank" rel="noopener noreferrer"><code>equity_daily</code></a></li>
<li><strong>Macro ETFs</strong> (13 tickers) — TLT, GLD, DBC, UUP and extended macro hedge universe · <a href="https://huggingface.co/datasets/JohnGavin/finance-data" target="_blank" rel="noopener noreferrer">HF dataset</a> · <a href="https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/groups.R#L55" target="_blank" rel="noopener noreferrer"><code>hd_group("Macro Hedge ETFs")</code></a></li>
<li><strong>Factor ETFs</strong> (9 tickers) — US index ETFs (SPY, QQQ, IWM, DIA) plus sector ETFs · <a href="https://huggingface.co/datasets/JohnGavin/finance-data" target="_blank" rel="noopener noreferrer">HF dataset</a> · <a href="https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/groups.R#L38" target="_blank" rel="noopener noreferrer"><code>hd_group("US Index ETFs")</code></a></li>
</ul>
</details>

<details>
<summary><strong>Crypto</strong> — 14 tokens · daily OHLCV from 2014 · 1.1 MB Parquet</summary>
<ul>
<li><strong>Major crypto</strong> (4 tokens) — BTC, ETH, SOL, BNB · <a href="https://huggingface.co/datasets/JohnGavin/finance-data" target="_blank" rel="noopener noreferrer">HF dataset</a> · <a href="https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/registry.R#L47" target="_blank" rel="noopener noreferrer"><code>hd_datasets()[["crypto_daily"]]</code></a></li>
<li><strong>Stablecoins</strong> (2 tokens) — USDC, USDT · <a href="https://huggingface.co/datasets/JohnGavin/finance-data" target="_blank" rel="noopener noreferrer">HF dataset</a> · <a href="https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/groups.R#L45" target="_blank" rel="noopener noreferrer"><code>hd_group("Stablecoins")</code></a></li>
<li><strong>Solana DeFi</strong> (5 tokens) — SOL, RAY, HNT, BONK, PYTH · <a href="https://huggingface.co/datasets/JohnGavin/finance-data" target="_blank" rel="noopener noreferrer">HF dataset</a> · <a href="https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/groups.R#L48" target="_blank" rel="noopener noreferrer"><code>hd_group("Solana DeFi")</code></a></li>
<li><strong>DeFi altcoins</strong> (3 tokens) — XRP, ADA, DOGE, DOT · <a href="https://huggingface.co/datasets/JohnGavin/finance-data" target="_blank" rel="noopener noreferrer">HF dataset</a> · <a href="https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/groups.R#L50" target="_blank" rel="noopener noreferrer"><code>hd_group("DeFi Altcoins")</code></a></li>
</ul>
</details>

<details>
<summary><strong>Macro</strong> — 19 FRED series · mixed frequency · from 1947</summary>
<ul>
<li><strong>Equity &amp; volatility</strong> — SP500, <abbr title="CBOE Volatility Index — implied volatility of S&amp;P 500 options">VIX</abbr> · <a href="https://huggingface.co/datasets/JohnGavin/finance-data" target="_blank" rel="noopener noreferrer">HF dataset</a> · <a href="https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/registry.R#L54" target="_blank" rel="noopener noreferrer"><code>hd_datasets()[["macro_daily"]]</code></a></li>
<li><strong>Interest rates</strong> — 2Y, 10Y, 30Y Treasury yields, Fed Funds rate · <a href="https://huggingface.co/datasets/JohnGavin/finance-data" target="_blank" rel="noopener noreferrer">HF dataset</a> · <a href="https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/query.R#L269" target="_blank" rel="noopener noreferrer"><code>hd_macro(c("DGS2","DGS10","DGS30"))</code></a></li>
<li><strong>Credit spreads</strong> — ICE BofA High Yield OAS, BBB corporate spread · <a href="https://huggingface.co/datasets/JohnGavin/finance-data" target="_blank" rel="noopener noreferrer">HF dataset</a></li>
<li><strong>Macro fundamentals</strong> — GDP (quarterly), CPI, PCE (monthly), unemployment · <a href="https://huggingface.co/datasets/JohnGavin/finance-data" target="_blank" rel="noopener noreferrer">HF dataset</a></li>
<li><strong>Commodities &amp; currency</strong> — WTI crude oil spot, trade-weighted USD index · <a href="https://huggingface.co/datasets/JohnGavin/finance-data" target="_blank" rel="noopener noreferrer">HF dataset</a></li>
</ul>
</details>

<details>
<summary><strong>Factors</strong> — Fama-French FF3, FF5, Momentum · daily + monthly · from 1926</summary>
<ul>
<li><strong>FF3</strong> — MktRF, SMB, HML plus risk-free rate · <a href="https://mba.tuck.dartmouth.edu/pages/faculty/ken.french/data_library.html" target="_blank" rel="noopener noreferrer">Ken French Data Library</a> · <a href="https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/registry.R#L60" target="_blank" rel="noopener noreferrer"><code>hd_datasets()[["factors"]]</code></a></li>
<li><strong>FF5</strong> — adds <abbr title="Fama-French 5-Factor model adds Profitability (RMW) and Investment (CMA) to FF3">RMW, CMA</abbr> · <a href="https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/query.R#L328" target="_blank" rel="noopener noreferrer"><code>hd_factors("FF5")</code></a></li>
<li><strong>Momentum</strong> — prior-12-month return factor · <a href="https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/query.R#L328" target="_blank" rel="noopener noreferrer"><code>hd_factors("Mom")</code></a></li>
<li><strong>JST Macrohistory</strong> — 155 years, 18 countries (1870–2020) · <a href="https://www.macrohistory.net/database/" target="_blank" rel="noopener noreferrer">macrohistory.net</a> · <a href="https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/registry.R#L92" target="_blank" rel="noopener noreferrer"><code>hd_datasets()[["jst_macrohistory"]]</code></a></li>
</ul>
</details>

# Articles

## Row

<div class="card-grid">
<a href="leaderboard.html" class="info-card">
<h4>Strategy Leaderboard</h4>
<p>Rankings, equity curves, and feature importance across all 6 strategies. Transposed metrics table, PSO optimal portfolio weights, XGBoost feature analysis.</p>
</a>
<a href="macro-defense-rotation.html" class="info-card">
<h4>Macro Defense Rotation</h4>
<p>Backtesting vignette: rotating among TLT/GLD/DBC/UUP to hedge macro risk. Expanding window methodology, in-sample vs out-of-sample holdout.</p>
</a>
<a href="factor-max.html" class="info-card">
<h4>Factor MAX Rotation</h4>
<p>Rotating among Fama-French factors using the MAX signal — maximum daily return predicts next-month factor returns. Based on Alpha Architect research (Dec 2025).</p>
</a>
<a href="drif.html" class="info-card">
<h4><abbr title="Dynamic Rotation Into Factors — selects best-performing FF factor each month">DRIF</abbr>: Daily Return Information</h4>
<p>Elastic net on the full distribution of past 21 daily factor returns (chronological + rank) predicts next-month returns. Subsumes MAX and momentum signals.</p>
</a>
<a href="stock-backtest.html" class="info-card">
<h4>Stock-Level Backtests</h4>
<p>MAX and DRIF signals applied to 660 individual stocks (S&P 500 + STOXX 600). Decile long-short portfolios. 4-strategy comparison: stock vs factor, MAX vs DRIF.</p>
</a>
<a href="avoid-worst-days.html" class="info-card">
<h4>Avoid Worst Days</h4>
<p>Thought experiment: removing the market's N worst days improves returns more than missing the N best days hurts. Demonstrates return asymmetry and why downside protection matters.</p>
</a>
<a href="falsification.html" class="info-card">
<h4>Falsification Tests</h4>
<p>Causal integrity tests: Fama-French alpha, null-signal heatmaps, HAC-robust Sharpe ratios, implication chains. Evidence our strategies aren't luck or factor exposure.</p>
</a>
<a href="european-overlay.html" class="info-card">
<h4>European Overlay</h4>
<p>CISS vs VIX regime comparison: ECB's Composite Indicator of Systemic Stress as European-native vol proxy. 5 European ETFs, OOS performance, attribution.</p>
</a>
<a href="jst-dashboard.html" class="info-card">
<h4><abbr title="Jordà-Schularick-Taylor Macrohistory Database">JST</abbr> Macrohistory: Long-Run Equity Returns</h4>
<p>155 years of equity premium across 18 countries (1870-2020). Cross-geography pervasiveness test, crisis timeline, decade heatmaps. Validates Swedroe evidence-based framework.</p>
</a>
<a href="https://github.com/JohnGavin/historical/tree/main/packages/historicaldata/R" class="info-card">
<h4>Source Code</h4>
<p>Browse R source files on GitHub: Data Access, Discovery, Performance Metrics, Kelly Sizing, Null Hypothesis Testing, and more.</p>
</a>
</div>

# Links

## Row

- [GitHub](https://github.com/JohnGavin/historical) — source code, issues, CI
- [Hugging Face Dataset](https://huggingface.co/datasets/JohnGavin/finance-data) — Parquet files, dataset card
- [Strategy Leaderboard](leaderboard.html) — rankings, equity curves, cost-adjusted metrics
- [API Reference](https://github.com/JohnGavin/historical/tree/main/packages/historicaldata/R) — function source code
- [Macro Defense Rotation](macro-defense-rotation.html) — backtesting vignette
- [Factor MAX](factor-max.html) — factor rotation strategy
- [DRIF](drif.html) — daily return information factor
- [Stock-Level Backtests](stock-backtest.html) — 660 stocks, 4-strategy comparison
- [Avoid Worst Days](avoid-worst-days.html) — return asymmetry thought experiment
- [Falsification Tests](falsification.html) — causal integrity, FF alpha, HAC-robust Sharpe
- [European Overlay](european-overlay.html) — CISS vs VIX regime comparison
- [JST Macrohistory](jst-dashboard.html) — 155 years equity premium, 18 countries, pervasiveness test

## Row

Built with the [T language](https://github.com/b-rodrigues/tlang){target="_blank" rel="noopener noreferrer"} pipeline, [targets](https://docs.ropensci.org/targets/){target="_blank" rel="noopener noreferrer"}, [DuckDB](https://duckdb.org/){target="_blank" rel="noopener noreferrer"}, and [Quarto](https://quarto.org/){target="_blank" rel="noopener noreferrer"}. Data from Yahoo Finance, FRED, and Ken French Data Library.

# Installation

## Row

### {.tabset}

#### R (pak)

```r
# Install from GitHub
pak::pak("JohnGavin/historical", subdir = "packages/historicaldata")

# Then:
library(historicaldata)
hd_ohlcv("AAPL", from = "2024-01-01")
hd_ohlcv(hd_group("FAANG"), from = "2024-01-01")  # batch query
hd_macro(c("SP500", "VIXCLS"), from = "2024-01-01")
hd_search("vanguard")  # search metadata
```

```{r}
#| label: install-check
#| echo: false
#| comment: ""
#| caption: "Live install check — re-runs on every page render"
cat("# Live install check — re-runs on every page render\n")
res <- tryCatch({
  pkg_installed <- requireNamespace("historicaldata", quietly = TRUE)
  list(
    pkg_installed  = pkg_installed,
    pkg_version    = if (pkg_installed)
                       as.character(utils::packageVersion("historicaldata"))
                     else "not installed",
    n_exports      = if (pkg_installed)
                       length(getNamespaceExports("historicaldata"))
                     else NA_integer_,
    r_version      = as.character(getRversion()),
    duckdb_version = tryCatch(
                       as.character(utils::packageVersion("duckdb")),
                       error = function(e) "not installed"
                     )
  )
}, error = function(e) list(error = conditionMessage(e)))
str(res)
```

#### Python (DuckDB)

No R needed — query the same data directly from Python using DuckDB's native `hf://` protocol:

```python
import duckdb

con = duckdb.connect()
df = con.sql("""
    SELECT date, close, volume
    FROM 'hf://datasets/JohnGavin/finance-data/equity_daily.parquet'
    WHERE ticker = 'AAPL' AND date >= '2024-01-01'
    ORDER BY date
""").df()
```

#### Nix

```bash
# Enter reproducible Nix shell (all R + Python packages included)
nix develop github:JohnGavin/historical
```

Then start R and load the package. All dependencies (dplyr, ggplot2, scales, DT, arrow, duckdb) are pre-installed in the Nix shell.

#### T language

```bash
# Clone, enter Nix shell, fetch data, run pipeline
git clone https://github.com/JohnGavin/historical.git && cd historical
nix develop
python scripts/fetch_equity.py     # 51 equity tickers
Rscript scripts/fetch_crypto.R     # 14 crypto tokens
Rscript scripts/fetch_macro.R      # 19 FRED series
Rscript scripts/fetch_factors.R    # Fama-French factors
t run src/pipeline.t               # T pipeline
```

# Build Info

## Row

```{r}
#| label: build-info
#| echo: false
#| results: asis
utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)
build_info()
```
