---
title: "Strategy Leaderboard"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    embed-resources: false
    include-in-header:
      - text: |
          <link rel="stylesheet" href="vignette-shared.css">
          <script src="vignette-shared.js"></script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)
library(dplyr)
library(tidyr)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)
```

# Rankings

## Row

### {.tabset}

#### Full Period

```{r}
lb <- safe_tar_read("leaderboard")
if (!is.null(lb)) {
  full <- lb |>
    filter(period == "Full Period") |>
    select(strategy, level, signal, cagr, vol, sharpe, max_dd) |>
    arrange(desc(sharpe))

  # Transpose: metrics as rows, strategies as columns
  metrics_long <- full |>
    mutate(
      CAGR = paste0(round(cagr * 100, 1), "%"),
      Vol = paste0(round(vol * 100, 1), "%"),
      Sharpe = ifelse(is.na(sharpe), "—", as.character(round(sharpe, 2))),
      `Max DD` = paste0(round(max_dd * 100, 1), "%"),
      Level = level,
      Signal = signal
    ) |>
    select(strategy, Level, Signal, CAGR, Vol, Sharpe, `Max DD`) |>
    pivot_longer(-strategy, names_to = "Metric", values_to = "value") |>
    pivot_wider(names_from = strategy, values_from = value)

  hd_dt(metrics_long, "Full-period rankings sorted by Sharpe. Each column is a strategy — click the header to sort. Links: Factor MAX, Factor DRIF, Stock MAX, Stock DRIF, XGB DRIF.")
}
```

#### By Partition

```{r}
if (!is.null(lb)) {
  by_part <- lb |>
    mutate(
      CAGR = paste0(round(cagr * 100, 1), "%"),
      Vol = paste0(round(vol * 100, 1), "%"),
      Sharpe = ifelse(is.na(sharpe), "—", as.character(round(sharpe, 2))),
      `Max DD` = paste0(round(max_dd * 100, 1), "%")
    ) |>
    select(Strategy = strategy, Period = period, CAGR, Vol, Sharpe, `Max DD`)

  hd_dt(by_part, "All strategies × all partitions (Training / Testing / Validation / Full). Sort by any column to compare strategies within a partition.")
}
```

#### Strategy Details

Each strategy links to its full definition, implementation, and backtest:

| Strategy | Level | Signal | Definition |
|----------|-------|--------|------------|
| [Factor MAX](factor-max.html) | Factor (5 FF) | Max daily return in prior month | Ranks factors by single-day extreme |
| [Factor DRIF](drif.html) | Factor (5 FF) | Elastic net on 42 daily return features | Full daily return distribution |
| [Stock MAX](stock-backtest.html#stock-max) | Stock (660) | Max daily return, D1-D10 decile spread | Cross-sectional stock sorting |
| [Stock DRIF](stock-backtest.html#stock-drif) | Stock (660) | Elastic net, D1-D10 decile spread | Pooled cross-sectional elastic net |
| [XGB DRIF](stock-backtest.html#stock-drif) | Stock (660) | XGBoost monotonic on 42 features | Non-linear signal with monotonic constraints |
| PSO Optimal | Combined | Weighted portfolio of above | Grid search on training Sharpe |

```{r}
# Show PSO weights
w <- safe_tar_read("port_optimal_weights")
if (!is.null(w)) {
  w_df <- tibble(
    Strategy = names(w),
    Weight = paste0(round(w * 100), "%")
  ) |> filter(Weight != "0%")
  hd_dt(w_df, "PSO optimal weights (trained on Training partition Sharpe). Strategies with 0% weight omitted.")
}
```

# Performance

## Row

### {.tabset}

#### Equity Curves

```{r}
#| fig-height: 6
safe_tar_read("stk_all_comparison_plot")
```

::: {.fig-caption}
**Equity curves (4 factor/stock strategies).** Growth of $1, log scale. These curves should be interpreted with caution: stock-level strategies have 20-44% annual vol. Extreme returns (150x over 20 years) require ~30% sustained CAGR which is unstable. Dashed line = test partition start (2020).
:::

#### Portfolio

```{r}
#| fig-height: 6
safe_tar_read("port_comparison_plot")
```

::: {.fig-caption}
**PSO optimal vs equal-weight portfolio.** Combines all 4 strategies with weights optimised on training partition Sharpe. The PSO optimal tilts toward strategies with better risk-adjusted returns (Factor DRIF + Stock MAX).
:::

#### XGBoost vs Elastic Net

```{r}
#| fig-height: 6
safe_tar_read("xgb_vs_enet_plot")
```

::: {.fig-caption}
**XGBoost (monotonic constraints) vs Elastic Net** on same 42 DRIF features, stock-level D1-D10. XGBoost underperforms in-sample but captures non-linear relationships.
:::

# Features

## Row

### {.tabset}

#### XGBoost Feature Importance

```{r}
fi <- safe_tar_read("xgb_feature_importance")
if (!is.null(fi)) {
  fi_display <- fi |>
    head(15) |>
    mutate(
      Gain = paste0(round(Gain * 100, 1), "%"),
      Type = type,
      Day = day
    ) |>
    select(Feature, Gain, Type, Day)
  hd_dt(fi_display, "Top 15 features by XGBoost gain. Chronological features (c*) capture WHEN returns occurred. Rank features (r*) capture HOW EXTREME. The paper predicts chronological > rank — confirmed here.")
}
```

#### Interpretation

::: {.callout-tip}
## What Feature Importance Tells Us

The XGBoost feature importance confirms the Alpha Architect paper's key finding:

1. **Chronological features dominate** — c13 (day 13) alone accounts for ~48% of gain
2. **Mid-month timing matters most** — days 10-17 are the most predictive window
3. **Rank extremes add signal** — r5 (5th lowest) and r21 (highest) contribute
4. **Recent days are NOT the most important** — c20, c21 have low importance, contradicting simple momentum

This suggests the signal is driven by **institutional rebalancing patterns** (mid-month flows) rather than simple recency or momentum.
:::

# Details

## Row

### {.tabset}

#### Partitions

```{r}
params <- safe_tar_read("stk_params")
if (!is.null(params)) {
  param_df <- tibble(
    Partition = c("Training", "Testing", "Validation"),
    Period = c(
      paste(params$start_date, "to", params$is_end),
      paste(params$test_start, "to", params$test_end),
      paste(params$val_start, "to", params$val_end)
    ),
    Purpose = c("Model fitting, expanding window signal estimation",
                "Calibration, hyperparameter tuning, strategy comparison",
                "Final one-shot evaluation (sealed envelope)")
  )
  hd_dt(param_df, "Shared partitions across all strategies. Validation is not used until ready for production.")
}
```

#### Build Info

```{r}
#| label: build-info
#| echo: false
#| results: asis
build_info()
```
