---
title: "Strategy Leaderboard"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    embed-resources: false
    include-in-header:
      - _includes/color-scheme-meta.html
      - text: |
          <link rel="stylesheet" href="vignette-shared.css">
          <script src="vignette-shared.js"></script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)
library(dplyr)
library(tidyr)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)

disc_path <- if (file.exists("../R/disclosures.R")) "../R/disclosures.R" else
  file.path(here::here(), "R/disclosures.R")
source(disc_path)

# DEFLATED_SHARPE_EXEMPTIONS is the single source of truth for which
# strategies are documented exclusions from deflated-Sharpe/K_eff coverage
# (S21 gate, #728 item 4) -- sourced from R/plan_qa_gates.R rather than
# re-declared here, so the Rigour column below can never drift from the QA
# gate that actually enforces coverage.
qa_gates_path <- if (file.exists("../R/plan_qa_gates.R")) "../R/plan_qa_gates.R" else
  file.path(here::here(), "R/plan_qa_gates.R")
source(qa_gates_path)
```

# Rankings

::: {.callout-warning collapse="false" title="Survivorship bias"}
`r disclosure_survivorship()`
:::

## Row

### {.tabset}

#### Full Period

```{r}
#| label: detection-rigour-summary
#| include: false
lb_dp <- safe_tar_read("leaderboard")
if (!is.null(lb_dp)) {
  full_dp <- lb_dp |> dplyr::filter(period == "Full Period")

  n_pos          <- sum(full_dp$sharpe > 0, na.rm = TRUE)
  n_under_single <- sum(full_dp$detection_underpowered, na.rm = TRUE)
  n_under_mt     <- sum(full_dp$detection_underpowered_mt, na.rm = TRUE)
  n_det_missing  <- sum(full_dp$sharpe > 0 & is.na(full_dp$detection_underpowered), na.rm = TRUE)
  n_det_both     <- sum(
    !is.na(full_dp$detection_underpowered) & !full_dp$detection_underpowered &
      !is.na(full_dp$detection_underpowered_mt) & !full_dp$detection_underpowered_mt,
    na.rm = TRUE
  )
  keff_nonNA <- full_dp$k_eff_leaderboard[!is.na(full_dp$k_eff_leaderboard)]
  kraw_nonNA <- full_dp$k_raw_leaderboard[!is.na(full_dp$k_raw_leaderboard)]
  keff <- if (length(keff_nonNA) > 0) round(keff_nonNA[1], 1) else NA_real_
  kraw <- if (length(kraw_nonNA) > 0) kraw_nonNA[1] else NA_integer_

  rig_cols      <- intersect(c("deflated_sharpe", "dsr_pvalue", "k_eff_leaderboard"), names(full_dp))
  rig_present   <- if (length(rig_cols) == 3L) rowSums(!is.na(full_dp[, rig_cols])) else rep(0L, nrow(full_dp))
  exempt_strats <- DEFLATED_SHARPE_EXEMPTIONS$strategy
  n_rig_total   <- nrow(full_dp)
  n_rig_full    <- sum(rig_present == 3L)
  n_rig_exempt  <- sum(full_dp$strategy %in% exempt_strats)
  n_rig_partial <- sum(rig_present > 0L & rig_present < 3L & !(full_dp$strategy %in% exempt_strats))
  n_rig_none    <- sum(rig_present == 0L & !(full_dp$strategy %in% exempt_strats))

  detection_summary_text <- paste0(
    "Of **", n_pos, "** strategies with a positive Full-Period Sharpe, **", n_under_mt,
    "** are underpowered once corrected for testing ", kraw, " strategies (K_eff = ", keff, ") — ",
    if (n_under_single != n_under_mt) paste0(n_under_single, " already fail the uncorrected, single-test bar; ") else "",
    "**", n_det_both, "** clear", if (n_det_both == 1L) "s" else "", " both bars",
    if (n_det_missing > 0) paste0(" — **", n_det_missing, " have NO verdict at all, which is a defect**") else "",
    ". See [#726](https://github.com/JohnGavin/historical/issues/726). ",
    "Rigour-metric coverage (deflated Sharpe / DSR p-value / K_eff), across all ", n_rig_total,
    " leaderboard strategies: **", n_rig_full, "** full, ", n_rig_partial, " partial, ",
    n_rig_exempt, " documented exclusion", if (n_rig_exempt != 1L) "s" else "",
    if (n_rig_none > 0) paste0(", **", n_rig_none, " with zero coverage and no exemption — a defect**") else ", 0 undetected gaps",
    ". See [#728](https://github.com/JohnGavin/historical/issues/728)."
  )
} else {
  detection_summary_text <- "*`leaderboard` target not available.*"
  keff <- NA_real_
  kraw <- NA_integer_
}
```

::: {.callout-warning collapse="false" title="Statistical detection power and rigour, this build"}
`r detection_summary_text`
:::

```{r}
lb <- safe_tar_read("leaderboard")
if (!is.null(lb)) {
  full <- lb |>
    filter(period == "Full Period") |>
    arrange(desc(sharpe))

  # Deflated-Sharpe haircut % (dsr_haircut_pct) is not merged onto
  # `leaderboard` itself -- only onto strat_deflated_sharpe (#728 item 2).
  # Joined here so the Rigour badge tooltip below can show the full trio the
  # now-removed "Deflated Sharpe" tab used to show, with zero information
  # loss (Step 5 consolidation).
  strat_ds <- safe_tar_read("strat_deflated_sharpe")
  if (!is.null(strat_ds) && "dsr_haircut_pct" %in% names(strat_ds)) {
    full <- full |> left_join(strat_ds |> select(strategy, dsr_haircut_pct), by = "strategy")
  }
  if (!"dsr_haircut_pct" %in% names(full)) full$dsr_haircut_pct <- NA_real_

  detection_badge <- function(sr, under, dmin, under_mt, dmin_mt) {
    # 5-state verdict, symbol AND text on every state (never colour alone,
    # `accessibility` rule). States 1/2 both come from a single-test
    # "detectable" verdict but are split by the multiple-testing-corrected
    # verdict so a strategy that flips under correction (OLMAR-1, #726/#728)
    # is never rendered identically to one that clears both bars.
    if (is.na(sr) || sr <= 0) {
      return(list(
        html = "<span class='det-badge napp' title='Sharpe &le;0: the one-sided detection test has no positive effect to test for. Not applicable -- not the same as unproven.'>&ndash; n/a</span>",
        rank = 4L
      ))
    }
    if (is.na(under)) {
      return(list(
        html = "<span class='det-badge notcomputed' title='Positive Sharpe with NO detection verdict. This should never render live -- QA gate S20 (#726 item 3) is meant to block the build until it is fixed.'>&#10007; NOT COMPUTED</span>",
        rank = 5L
      ))
    }
    if (isTRUE(under)) {
      mt_note <- if (!is.na(under_mt)) " Multiple-testing correction is academic here -- the single test already fails." else ""
      return(list(
        html = sprintf(
          "<span class='det-badge underpowered' title='Needs %.0f years of data at alpha=0.05, one-sided.%s See #726.'>&#9888; Underpowered (needs %.0fyr)</span>",
          dmin, mt_note, dmin
        ),
        rank = 3L
      ))
    }
    # Single-test verdict is "detectable" from here on.
    if (is.na(under_mt)) {
      return(list(
        html = sprintf(
          "<span class='det-badge detectable' title='Needs %.1f years at alpha=0.05, one-sided. Multiple-testing-corrected verdict not computed (K_eff unavailable for this strategy). See #726.'>&#10003; Detectable (needs %.1fyr)</span>",
          dmin, dmin
        ),
        rank = 1L
      ))
    }
    if (isTRUE(under_mt)) {
      return(list(
        html = sprintf(
          "<span class='det-badge mtfail' title='Passes the single test (needs %.1fyr) but FAILS once corrected for testing multiple strategies (needs %.0fyr corrected). See #726.'>&#9686; Detectable, fails MT-correction (needs %.0fyr corrected)</span>",
          dmin, dmin_mt, dmin_mt
        ),
        rank = 2L
      ))
    }
    list(
      html = sprintf(
        "<span class='det-badge detectable' title='Needs %.1f years at alpha=0.05, one-sided (%.1fyr corrected for multiple testing). See #726.'>&#10003; Detectable (needs %.1fyr; %.1fyr corrected)</span>",
        dmin, dmin_mt, dmin, dmin_mt
      ),
      rank = 1L
    )
  }

  rigour_badge <- function(strategy, deflated, dsr_p, keff, kraw, sr, haircut) {
    # Rank: 1 = full coverage (best), 2 = partial (defect-in-progress),
    # 3 = documented exclusion (legitimate, not a defect), 4 = zero
    # coverage with no exemption (defect). Mirrors the .credible_rank /
    # .redundant_rank hidden-sort-column pattern already used in this table
    # so DT's default alphabetical string sort never scrambles verdict order.
    exempt_reason <- DEFLATED_SHARPE_EXEMPTIONS$reason[DEFLATED_SHARPE_EXEMPTIONS$strategy == strategy]
    if (length(exempt_reason) > 0) {
      return(list(
        html = sprintf(
          "<span class='rig-badge excluded' title='%s'>&#9675; excluded</span>",
          gsub("'", "&#39;", exempt_reason[[1]], fixed = TRUE)
        ),
        rank = 3L
      ))
    }
    present <- sum(!is.na(c(deflated, dsr_p, keff)))
    if (present == 3L) {
      haircut_txt <- if (is.na(haircut)) "" else sprintf(" (%.1f%% haircut)", haircut)
      return(list(
        html = sprintf(
          "<span class='rig-badge computed' title='Deflated Sharpe: %.2f (naive %.2f%s). DSR p-value: %.3f. K_eff = %.1f of %s raw.'>&#10003; 3/3</span>",
          deflated, sr, haircut_txt, dsr_p, keff, kraw
        ),
        rank = 1L
      ))
    }
    if (present > 0L) {
      return(list(
        html = sprintf(
          "<span class='rig-badge partial' title='Missing %d of 3 rigour metrics (deflated Sharpe / DSR p-value / K_eff). Tracked by #728, QA gate S21.'>&#9686; %d/3</span>",
          3L - present, present
        ),
        rank = 2L
      ))
    }
    list(
      html = "<span class='rig-badge notcomputed' title='ZERO rigour-metric coverage and no declared exemption. This should never render live -- QA gate S21 (#728 item 4) is meant to block the build until it is fixed.'>&#10007; 0/3 NOT COMPUTED</span>",
      rank = 4L
    )
  }

  det <- mapply(detection_badge, full$sharpe, full$detection_underpowered,
                full$detection_min_n_years, full$detection_underpowered_mt,
                full$detection_min_n_years_mt, SIMPLIFY = FALSE)
  rig <- mapply(rigour_badge, full$strategy, full$deflated_sharpe, full$dsr_pvalue,
                full$k_eff_leaderboard, full$k_raw_leaderboard, full$sharpe,
                full$dsr_haircut_pct, SIMPLIFY = FALSE)

  metrics_table <- full |>
    mutate(
      Sharpe     = ifelse(is.na(sharpe),       "—", as.character(round(sharpe, 2))),
      Detection  = vapply(det, function(x) x$html, character(1)),
      .det_rank  = vapply(det, function(x) x$rank, integer(1)),
      Rigour     = vapply(rig, function(x) x$html, character(1)),
      .rig_rank  = vapply(rig, function(x) x$rank, integer(1)),
      `Net CAGR` = ifelse(is.na(net_cagr),     "—", paste0(round(net_cagr * 100, 1), "%")),
      `Max DD`   = paste0(round(max_dd * 100, 1), "%"),
      `CVaR 95%` = ifelse(is.na(cvar_95),      "—", paste0(round(cvar_95 * 100, 1), "%")),
      Vol        = paste0(round(vol * 100, 1), "%"),
      Gross      = dplyr::case_when(
        !("gross_convention" %in% names(full)) | is.na(gross_convention) ~ "—",
        is_cap ~ paste0(format(gross_convention, nsmall = 1), "x*"),
        TRUE   ~ paste0(format(gross_convention, nsmall = 1), "x")
      ),
      `Cost (bps)` = dplyr::case_when(
        !("cost_per_trade_bps" %in% names(full)) | is.na(cost_per_trade_bps) ~ "—",
        cost_per_trade_bps == 0 ~ "0*",
        TRUE ~ as.character(round(cost_per_trade_bps))
      ),
      SSR        = ifelse(is.na(ssr),           "—", as.character(round(ssr, 2))),
      `Top5% Share` = ifelse(is.na(top5pct_share), "—", paste0(round(top5pct_share * 100, 1), "%")),
      # NA credible means cost metrics were never computed for this strategy
      # (cost_rows in R/plan_leaderboard.R only covers 5 core strategies +
      # PSO Optimal) -- that is "unknown", not "explicitly failed". Conflating
      # the two previously fed literal NA strategy names into the WARNING
      # note below (#637 follow-up).
      Credible   = dplyr::case_when(
        is.na(credible) ~ "not computed",
        !credible        ~ "NO",
        TRUE              ~ "yes"
      ),
      # Explicit sort rank for Credible: the display string ("NO" / "not
      # computed" / "yes") sorts alphabetically wrong under DT's default
      # string comparator (ascending gives NO, not computed, yes -- neither
      # ascending nor descending puts assessed-and-passed first, then
      # assessed-and-failed, then never-assessed). This hidden column
      # supplies the intended rank instead (#637 follow-up).
      .credible_rank = dplyr::case_when(
        is.na(credible) ~ 3L,
        !credible        ~ 2L,
        TRUE              ~ 1L
      ),
      # NA redundant means this strategy is outside the 5-strategy correlation
      # universe computed by R/plan_strategy_correlation.R (Stock MAX, Stock
      # DRIF, Factor MAX, Factor DRIF, LTR only) -- that is "unassessed", not
      # "assessed and found not redundant". Previously both NA and a computed
      # FALSE fell through to the same ambiguous "—" (#643, same conflation
      # class as Credible, #637 follow-up).
      Redundant  = if ("redundant" %in% names(full)) {
        dplyr::case_when(
          is.na(redundant) ~ "not computed",
          redundant         ~ "YES",
          TRUE              ~ "no"
        )
      } else "not computed",
      # Explicit sort rank for Redundant: same rationale as .credible_rank —
      # DT's default string comparator sorts "YES" / "no" / "not computed"
      # alphabetically, which does not put assessed-good first, then
      # assessed-bad, then never-assessed (#643).
      .redundant_rank = if ("redundant" %in% names(full)) {
        dplyr::case_when(
          is.na(redundant) ~ 3L,
          redundant         ~ 2L,
          TRUE              ~ 1L
        )
      } else 3L,
      `Incremental Sharpe` = if ("incremental_sharpe" %in% names(full)) {
        ifelse(is.na(incremental_sharpe), "—", as.character(round(incremental_sharpe, 3)))
      } else "—"
    ) |>
    # Detection sits immediately right of Sharpe per requirement 2 of the
    # `detection-power-required` rule ("beside its Sharpe, not in a distant
    # column"). Rigour sits next to SSR -- both are "how much should you
    # trust this Sharpe" columns, unlike Credible/Redundant which assess
    # cost-plausibility and correlation-redundancy respectively.
    select(Strategy = strategy, Credible, .credible_rank, Sharpe, Detection, .det_rank,
           `Net CAGR`, `Max DD`, `CVaR 95%`, Vol, Gross, `Cost (bps)`, SSR, Rigour, .rig_rank,
           `Top5% Share`, Redundant, .redundant_rank, `Incremental Sharpe`)

  # NA credible (not is.na(credible) | !credible) is deliberate here: `!NA`
  # is NA, and indexing a character vector with a logical vector containing
  # NA inserts a literal NA element into the result -- that is the root
  # cause of the "WARNING: NA, NA, NA, ..." defect (#637 follow-up). Only
  # strategies with an explicit, computed FALSE are flagged.
  n_flagged <- sum(!is.na(full$credible) & !full$credible)
  flag_note <- if (n_flagged > 0) {
    flagged <- full$strategy[!is.na(full$credible) & !full$credible]
    paste0(" WARNING: ", paste(flagged, collapse = ", "),
           " flagged as non-credible (>20% CAGR or >50x cum P&L — likely unrealistic).")
  } else ""

  n_biased <- if ("survivorship_biased" %in% names(full)) sum(full$survivorship_biased, na.rm = TRUE) else 0L
  bias_note <- if (n_biased > 0) {
    biased <- full$strategy[!is.na(full$survivorship_biased) & full$survivorship_biased]
    paste0(" Survivorship-biased (`survivorship_biased = TRUE`): ",
           paste(biased, collapse = ", "),
           " — draw on `stk_universe` (see issue #150).")
  } else ""

  # ── Gross-exposure disclosure note (#626) — dynamically computed, never hardcoded ──
  gh_url <- tryCatch({
    remote <- system("git remote get-url origin 2>/dev/null", intern = TRUE)
    sub("\\.git$", "", sub("^git@github\\.com:", "https://github.com/", remote))
  }, error = function(e) NULL)
  git_sha <- tryCatch(
    system("git rev-parse HEAD 2>/dev/null", intern = TRUE),
    error = function(e) "main"
  )
  git_sha <- if (length(git_sha) == 0 || !nzchar(git_sha)) "main" else git_sha
  gh_base <- if (!is.null(gh_url) && length(gh_url) > 0) gh_url else
    "https://github.com/JohnGavin/historical"
  hd_exposure_link <- hd_link(
    htmltools::tags$code("hd_exposure_metrics()"),
    sprintf("%s/blob/%s/packages/historicaldata/R/exposure.R#L67", gh_base, git_sha)
  )
  gross_registry_link <- hd_link(
    htmltools::tags$code("strategy_gross_convention"),
    sprintf("%s/blob/%s/R/plan_exposure.R#L27", gh_base, git_sha)
  )

  n_2x <- if ("gross_convention" %in% names(full)) {
    sum(full$gross_convention >= 2, na.rm = TRUE)
  } else 0L
  n_capped <- if ("is_cap" %in% names(full)) sum(full$is_cap, na.rm = TRUE) else 0L
  n_na_gross <- if ("gross_convention" %in% names(full)) sum(is.na(full$gross_convention)) else nrow(full)

  # tagList, not paste0: hd_exposure_link/gross_registry_link are real
  # htmltools <a> tags (see hd_link() in vignette_utils.R), not markdown
  # strings -- DT hands the caption straight to the browser with no
  # markdown pass, so `[text](url)` syntax previously rendered as literal
  # brackets and a bare URL instead of a clickable link.
  gross_note <- htmltools::tagList(
    " Gross exposure (", htmltools::tags$code("sum(abs(w))"), ", computed by ", hd_exposure_link,
    " and recorded per strategy in ", gross_registry_link,
    ") is not uniform across this table: ", n_2x,
    " strategies are built dollar-neutral long-short at 2.0x gross while others run long-only at 1.0x, ",
    "so Vol, Net CAGR, and Max DD are not directly comparable across rows even when Sharpe is similar ",
    "(Sharpe is leverage-invariant under proportional scaling; the other three columns are not). ",
    "A trailing `*` on the Gross column (", n_capped,
    " strategies) marks a construction ceiling the strategy can reach but does not realise every period ",
    "(e.g. Managed Futures targets volatility and only hits its 3.0x cap in low-vol regimes) — treat those as upper bounds, not realised averages. ",
    if (n_na_gross > 0) paste0(n_na_gross, " strategy row(s) show — for Gross: either the return is assembled from a pre-computed factor series with no explicit weight vector to measure, or the row is a meta-portfolio blend of other rows.") else ""
  )

  # ── Transaction-cost disclosure note (#624) — dynamically computed, never hardcoded ──
  cost_registry_link <- hd_link(
    htmltools::tags$code("strategy_cost_convention"),
    sprintf("%s/blob/%s/R/plan_cost_convention.R#L44", gh_base, git_sha)
  )
  n_zero_cost <- if ("cost_per_trade_bps" %in% names(full)) {
    sum(full$cost_per_trade_bps == 0, na.rm = TRUE)
  } else 0L
  zero_cost_strategies <- if (n_zero_cost > 0) {
    full$strategy[!is.na(full$cost_per_trade_bps) & full$cost_per_trade_bps == 0]
  } else character(0)
  n_na_cost <- if ("cost_per_trade_bps" %in% names(full)) sum(is.na(full$cost_per_trade_bps)) else nrow(full)
  cost_bps_range <- if ("cost_per_trade_bps" %in% names(full) &&
                          any(!is.na(full$cost_per_trade_bps))) {
    rng <- range(full$cost_per_trade_bps, na.rm = TRUE)
    paste0(rng[1], "–", rng[2], " bps")
  } else "not available"

  cost_note <- htmltools::tagList(
    " Per-strategy transaction-cost assumptions (recorded in ", cost_registry_link,
    ") are not uniform across this table, in both magnitude (", cost_bps_range,
    ") and unit: most rows charge ",
    "a per-trade bps figure on realised turnover, but Value (HML) and Managed Futures charge a flat ",
    "per-calendar-month bps figure regardless of turnover, and TOM/Risk State charge only on discrete ",
    "switch days — see cost_convention in the registry for each row's exact timing. ",
    if (n_zero_cost > 0) {
      paste0("A trailing `*` on the Cost (bps) column (", paste(zero_cost_strategies, collapse = ", "),
             ") marks a verified zero — no transaction-cost deduction exists in that strategy's code, ",
             "not a plausible default; it is not net-of-cost comparable to the cost-charging rows above and below it. ")
    } else "",
    if (n_na_cost > 0) paste0(n_na_cost, " strategy row(s) show — for Cost (bps): a meta-portfolio blend inherits its constituents' already-net returns with no separate cost applied at the blend level. ") else "",
    "This Cost (bps) column is the strategy's own internal assumption baked into its returns before they reach this table; ",
    "for the ", sum(!is.na(full$net_cagr)),
    " strategies with a Net CAGR value above, it is additional to — not the same as — the flat cost applied there."
  )

  # ── Diversification-value disclosure note (Pillar 7, #268; surfaced #624) ──
  corr_registry_link <- hd_link(
    htmltools::tags$code("plan_strategy_correlation.R"),
    sprintf("%s/blob/%s/R/plan_strategy_correlation.R#L31", gh_base, git_sha)
  )
  n_redundant <- if ("redundant" %in% names(full)) sum(full$redundant, na.rm = TRUE) else 0L
  n_redundant_na <- if ("redundant" %in% names(full)) sum(is.na(full$redundant)) else nrow(full)
  diversification_note <- htmltools::tagList(
    " Redundant = \"YES\" when a strategy's Full-Period monthly-return correlation with another, ",
    "higher-Sharpe strategy exceeds the threshold fixed in ", corr_registry_link,
    " (", n_redundant, " of ", nrow(full),
    " strategies currently flagged); \"not computed\" (", n_redundant_na,
    " strategies) marks a strategy outside the 5-strategy correlation universe ",
    "(Stock MAX, Stock DRIF, Factor MAX, Factor DRIF, LTR) that plan_strategy_correlation.R ",
    "currently covers — that is unassessed, not assessed-and-passed. Incremental Sharpe = the ",
    "change in equal-weight-portfolio Sharpe from including this strategy vs excluding it — ",
    "positive means it adds diversification value, negative means the portfolio is better off ",
    "without it. Both feed the new-strategy value gate (#496)."
  )

  # ── Detection-power / rigour disclosure note (#726, #728) — dynamically computed ──
  detection_note <- paste0(
    " Detection = 3-state verdict from `hd_detection_power()` (`detection-power-required` rule): ",
    "whether this row's own sample has enough data to distinguish its Sharpe from zero at ",
    "alpha=0.05 one-sided, 80% power. A strategy that passes the single test but fails once ",
    "corrected for testing ", kraw, " strategies (K_eff = ", keff,
    ") is marked distinctly (◆ “fails MT-correction”), not rendered identically to one ",
    "that clears both bars — see #726. Rigour = coverage of deflated Sharpe / DSR p-value / K_eff ",
    "(hover for the three numeric values); ○ excluded marks a documented exclusion from ",
    "`DEFLATED_SHARPE_EXEMPTIONS` (R/plan_qa_gates.R), not a gap — see #728."
  )

  # Short caption (#1 in the caption-length audit, historical#0b412a9c): the
  # unconditional text states only what the table ranks and the single most
  # important comparability warning. Everything else (column definitions,
  # per-strategy caveats) is real but was drowning that one sentence -- it
  # moves into a collapsed <details> block via hd_caption() (vignette_utils.R)
  # instead of being deleted.
  short_caption <- paste0(
    "Strategies ranked by Sharpe (highest first); Net CAGR is net of a 0.20%/month round-trip ",
    "cost where available. Values are not directly comparable across rows — gross exposure, ",
    "cost basis, and the Credible/Redundant/Detection/Rigour flags vary by strategy; expand ",
    "below for full column definitions."
  )

  full_notes <- htmltools::tagList(
    "SSR = Sharpe Stability Ratio (HAC t-stat of rolling 36-month Sharpes; ≥2 = consistent). ",
    "Top5% Share = fraction of total return from best 5% of months. ",
    "Credible = plausible CAGR (<20%) and cum P&L (<50x): \"NO\" marks a strategy that was ",
    "assessed and failed that check, \"not computed\" marks a strategy for which cost-adjusted ",
    "metrics were never assessed at all (not the same as failing the check).",
    flag_note, bias_note, gross_note, cost_note, diversification_note, detection_note
  )

  full_caption <- hd_caption(
    full_notes,
    short = short_caption,
    summary_label = "Column definitions and comparability caveats"
  )

  # Bespoke DT call (not hd_dt()): the Credible, Redundant, Detection, and
  # Rigour columns each need an explicit sort rank (.credible_rank /
  # .redundant_rank / .det_rank / .rig_rank, added above) instead of DT's
  # default alphabetical string comparator -- ascending on the display
  # string does not put assessed-good first, then assessed-bad, then
  # never-assessed. orderData points each column's sort at its hidden rank
  # column instead (#637 follow-up; Redundant added #643; Detection/Rigour
  # added #726/#728).
  credible_col  <- which(names(metrics_table) == "Credible") - 1L
  cred_rank_col <- which(names(metrics_table) == ".credible_rank") - 1L
  redundant_col <- which(names(metrics_table) == "Redundant") - 1L
  red_rank_col  <- which(names(metrics_table) == ".redundant_rank") - 1L
  det_col       <- which(names(metrics_table) == "Detection") - 1L
  det_rank_col  <- which(names(metrics_table) == ".det_rank") - 1L
  rig_col       <- which(names(metrics_table) == "Rigour") - 1L
  rig_rank_col  <- which(names(metrics_table) == ".rig_rank") - 1L
  numeric_cols <- which(sapply(metrics_table, function(x) is.numeric(x) ||
                                  grepl("^[0-9.%$,-]+$", as.character(x[1]))))

  DT::datatable(
    metrics_table,
    # full_caption is already a complete htmltools::tags$caption() object
    # (built by hd_caption() above) -- do NOT re-wrap it in another
    # tags$caption() here. <caption> has no descendant table elements
    # allowed in its content model (caption is itself one), so nesting two
    # would have produced malformed markup and an unpredictable DOM.
    caption = full_caption,
    escape = FALSE,
    rownames = FALSE,
    filter = "top",
    options = list(
      pageLength = 15,
      scrollX = TRUE,
      autoWidth = FALSE,
      dom = "frtip",
      search = list(regex = TRUE, caseInsensitive = TRUE),
      columnDefs = list(
        list(className = 'dt-right', targets = numeric_cols - 1),
        list(visible = FALSE, targets = c(cred_rank_col, red_rank_col, det_rank_col, rig_rank_col)),
        list(orderData = cred_rank_col, targets = credible_col),
        list(orderData = red_rank_col, targets = redundant_col),
        list(orderData = det_rank_col, targets = det_col),
        list(orderData = rig_rank_col, targets = rig_col)
      ),
      initComplete = DT::JS(
        "function(settings, json) {",
        "  var isDark = document.documentElement.getAttribute('data-bs-theme') === 'dark'",
        "    || document.body.classList.contains('dark-mode');",
        "  if (isDark) {",
        "    $(this.api().table().container()).css({'color': '#ddd', 'background-color': '#1a1a1a'});",
        "    $(this.api().table().header()).css({'color': '#ddd', 'background-color': '#222'});",
        "    $('input', this.api().table().container()).css({'color': '#ddd', 'background-color': '#333', 'border-color': '#555'});",
        "  }",
        "}"
      )
    )
  )
}
```

```{=html}
<style>
.det-badge, .rig-badge { font-weight: 600; padding: 1px 6px; border-radius: 3px; white-space: nowrap; }
.det-badge.detectable   { color: #1a7f37; }
.det-badge.mtfail       { color: #6f42c1; }
.det-badge.underpowered { color: #9a5b00; }
.det-badge.napp         { color: #6c757d; font-style: italic; font-weight: 400; }
.det-badge.notcomputed  { color: #fff; background: #c0392b; }
.rig-badge.computed     { color: #1a7f37; }
.rig-badge.partial      { color: #9a5b00; }
.rig-badge.excluded     { color: #6c757d; font-style: italic; font-weight: 400; }
.rig-badge.notcomputed  { color: #fff; background: #c0392b; }

[data-bs-theme="dark"] .det-badge.detectable,   body.dark-mode .det-badge.detectable   { color: #4ade80; }
[data-bs-theme="dark"] .det-badge.mtfail,       body.dark-mode .det-badge.mtfail       { color: #c084fc; }
[data-bs-theme="dark"] .det-badge.underpowered, body.dark-mode .det-badge.underpowered { color: #fbbf24; }
[data-bs-theme="dark"] .det-badge.napp,         body.dark-mode .det-badge.napp         { color: #9ca3af; }
[data-bs-theme="dark"] .rig-badge.computed,     body.dark-mode .rig-badge.computed     { color: #4ade80; }
[data-bs-theme="dark"] .rig-badge.partial,      body.dark-mode .rig-badge.partial      { color: #fbbf24; }
[data-bs-theme="dark"] .rig-badge.excluded,     body.dark-mode .rig-badge.excluded     { color: #9ca3af; }
@media (prefers-color-scheme: dark) {
  .det-badge.detectable   { color: #4ade80; }
  .det-badge.mtfail       { color: #c084fc; }
  .det-badge.underpowered { color: #fbbf24; }
  .rig-badge.computed     { color: #4ade80; }
  .rig-badge.partial      { color: #fbbf24; }
}
</style>
```

#### By Partition

```{r}
if (!is.null(lb)) {
  # NEVER show Validation data — sealed envelope (backtest-partitions rule).
  # As of #648, the `leaderboard` target itself no longer emits Validation
  # rows (R/plan_leaderboard.R strips them at assembly), so the
  # `filter(period != "Validation")` below is now belt-and-braces rather
  # than the sole control -- kept so this display never regresses even if
  # a future upstream change reintroduces a Validation row.
  params <- safe_tar_read("stk_params")

  # Create years lookup
  #
  # "Holdout" (#660, 2024-01-01..2026-04-30) is included here for
  # forward-compatibility with `period_order` below, but as of #660 the
  # assembled `leaderboard` target does not yet emit any row with this
  # label -- the source metrics targets (stk_max_metrics, stk_drif_metrics,
  # xgb_drif_metrics, fm_metrics, drif_metrics, port_metrics) only compute
  # Training/Testing/(Validation, filtered)/Full Period, so the Holdout cost
  # rows built in R/plan_leaderboard.R's `slice_portfolio()` are dropped by
  # its LEFT join before reaching `lb`. This row is therefore currently
  # inert (matches zero rows) -- see the #660 PR report for the follow-up
  # needed to surface it.
  years_map <- tibble(
    period = c("Training", "Testing", "Holdout", "Full Period"),
    Years = if (!is.null(params)) {
      c(
        paste0(format(as.Date(params$start_date), "%Y"), "-", format(as.Date(params$is_end), "%Y")),
        paste0(format(as.Date(params$test_start), "%Y"), "-", format(as.Date(params$test_end), "%Y")),
        paste0(format(as.Date(params$holdout_start), "%Y"), "-", format(as.Date(params$holdout_end), "%Y")),
        paste0(format(as.Date(params$start_date), "%Y"), "-", format(as.Date(params$test_end), "%Y"))
      )
    } else {
      c("—", "—", "—", "—")
    }
  )

  # Explicit period display order (NOT alphabetical, per rule): Testing,
  # Holdout, Full Period, Training. Testing and Holdout are adjacent because
  # Holdout is the more recent, more-informative-but-unsealed evaluation
  # window that directly follows Testing chronologically; Full Period and
  # Training close out the table the same way they always have (#660).
  # Implemented via factor levels so dplyr::arrange() sorts by this order
  # rather than alphabetically ("Full Period" < "Holdout" < "Testing" <
  # "Training").
  period_order <- c("Testing", "Holdout", "Full Period", "Training")

  by_part <- lb |>
    filter(period != "Validation") |>
    left_join(years_map, by = "period") |>
    mutate(period = factor(period, levels = period_order)) |>
    # Sort on the NUMERIC sharpe column before it is formatted to character
    # below -- sorting the formatted string ("9.5" vs "10.2") would sort
    # lexicographically and put "9.5" above "10.2".
    arrange(period, desc(sharpe)) |>
    mutate(
      CAGR = paste0(round(cagr * 100, 1), "%"),
      Vol = paste0(round(vol * 100, 1), "%"),
      Sharpe = ifelse(is.na(sharpe), "—", as.character(round(sharpe, 2))),
      `Max DD` = paste0(round(max_dd * 100, 1), "%"),
      `CVaR 95%` = ifelse(is.na(cvar_95), "—", paste0(round(cvar_95 * 100, 1), "%")),
      Period = as.character(period)
    ) |>
    select(Strategy = strategy, Period, Sharpe, Years, CAGR, Vol, `Max DD`, `CVaR 95%`)

  # Bespoke DT call (not hd_dt()): rows above are already sorted by Period
  # (Testing / Holdout / Full Period / Training) then Sharpe descending. hd_dt()'s
  # fixed options list has no `order` entry, so DT would apply its default
  # initial sort (column 0 ascending = Strategy) and silently undo the sort
  # above -- `order = list()` (empty array) disables DT's initial sort and
  # preserves the incoming row order.
  numeric_cols <- which(sapply(by_part, function(x) is.numeric(x) ||
                                  grepl("^[0-9.%$,-]+$", as.character(x[1]))))

  DT::datatable(
    by_part,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold;",
      class = "dt-caption",
      "All strategies × Training, Testing, and Holdout partitions, ordered Testing / Holdout / Full Period / Training then by Sharpe (descending). Years column shows date ranges. Holdout (2024-2026) is observed, not sealed -- treat it with a discount relative to a genuinely untouched sample (#660). Validation partition is sealed and genuinely untouched (2026-05-01 onwards) — not shown until production deployment."
    ),
    rownames = FALSE,
    filter = "top",
    options = list(
      pageLength = 15,
      scrollX = TRUE,
      autoWidth = FALSE,
      dom = "frtip",
      order = list(),
      search = list(regex = TRUE, caseInsensitive = TRUE),
      columnDefs = list(
        list(className = 'dt-right', targets = numeric_cols - 1)
      ),
      initComplete = DT::JS(
        "function(settings, json) {",
        "  var isDark = document.documentElement.getAttribute('data-bs-theme') === 'dark'",
        "    || document.body.classList.contains('dark-mode');",
        "  if (isDark) {",
        "    $(this.api().table().container()).css({'color': '#ddd', 'background-color': '#1a1a1a'});",
        "    $(this.api().table().header()).css({'color': '#ddd', 'background-color': '#222'});",
        "    $('input', this.api().table().container()).css({'color': '#ddd', 'background-color': '#333', 'border-color': '#555'});",
        "  }",
        "}"
      )
    )
  )
}
```

#### Financing Sensitivity

The leaderboard above charges each strategy a single, unsourced borrow rate — and, as of [#665](https://github.com/JohnGavin/historical/issues/665), assumes zero securities-lending income everywhere (a documented materiality judgement, not a measurement — see `lending_status` in `strategy_cost_convention`, R/plan_cost_convention.R). This tab does not correct that number; it measures how much of each strategy's edge survives if the single borrow-rate assumption is wrong. Two things to read correctly here: the 0% column is a no-financing counterfactual, not a claim that financing is free — do not treat it as "the" answer; and this sweep's Sharpe omits the risk-free rate (it reports *relative* deltas across borrow rates, where rf cancels), so its levels are **not** directly comparable to the leaderboard's own rf-adjusted Sharpe above — only the delta column is.

```{r}
#| label: financing-sensitivity
bss <- safe_tar_read("borrow_sensitivity_sweep")
scc <- safe_tar_read("strategy_cost_convention")
if (!is.null(bss)) {
  gh_url <- tryCatch({
    remote <- system("git remote get-url origin 2>/dev/null", intern = TRUE)
    sub("\\.git$", "", sub("^git@github\\.com:", "https://github.com/", remote))
  }, error = function(e) NULL)
  git_sha <- tryCatch(
    system("git rev-parse HEAD 2>/dev/null", intern = TRUE),
    error = function(e) "main"
  )
  git_sha <- if (length(git_sha) == 0 || !nzchar(git_sha)) "main" else git_sha
  gh_base <- if (!is.null(gh_url) && length(gh_url) > 0) gh_url else
    "https://github.com/JohnGavin/historical"
  sweep_link <- hd_link(
    htmltools::tags$code("borrow_sensitivity_sweep"),
    sprintf("%s/blob/%s/R/plan_cost_convention.R#L601", gh_base, git_sha)
  )
  fn_link <- hd_link(
    htmltools::tags$code("compute_borrow_sensitivity()"),
    sprintf("%s/blob/%s/R/plan_cost_convention.R#L322", gh_base, git_sha)
  )

  rates_sorted <- sort(unique(bss$borrow_rate_annual))
  max_rate <- max(rates_sorted)
  n_swept <- dplyr::n_distinct(bss$strategy)

  status_lookup <- if (!is.null(scc)) {
    dplyr::select(scc, strategy, borrow_status)
  } else {
    tibble(strategy = character(0), borrow_status = character(0))
  }

  # Wide pivot: one row per strategy, one Sharpe column per swept borrow
  # rate. rate_label's rounding formula is reused verbatim below to build
  # `rate_cols` so the two never drift apart.
  sharpe_wide <- bss |>
    mutate(rate_label = paste0("Sharpe @", round(borrow_rate_annual * 100), "%")) |>
    select(strategy, rate_label, sharpe) |>
    tidyr::pivot_wider(names_from = rate_label, values_from = sharpe)

  rate_cols <- paste0("Sharpe @", round(rates_sorted * 100), "%")
  delta_col <- paste0("Delta Sharpe (0%->", round(max_rate * 100), "%)")

  # sharpe_delta at the HIGHEST swept rate -- the largest, not smallest,
  # degradation is what a reader needs to see first (#665: "the single most
  # informative number in the issue" for the previously-unmodelled trio).
  delta_tbl <- bss |>
    filter(borrow_rate_annual == max_rate) |>
    select(strategy, sharpe_delta)
  names(delta_tbl)[names(delta_tbl) == "sharpe_delta"] <- delta_col

  sens_table <- sharpe_wide |>
    left_join(delta_tbl, by = "strategy") |>
    left_join(status_lookup, by = "strategy") |>
    select(strategy, dplyr::all_of(rate_cols), dplyr::all_of(delta_col), borrow_status) |>
    arrange(.data[[delta_col]]) |>
    mutate(dplyr::across(dplyr::where(is.numeric), ~round(., 2))) |>
    rename(Strategy = strategy, `Borrow Status` = borrow_status)

  most_affected <- head(sens_table$Strategy, 3)

  cmr_note <- if ("CMR" %in% sens_table$Strategy) {
    cmr_status <- sens_table$`Borrow Status`[sens_table$Strategy == "CMR"]
    if (!is.na(cmr_status) && cmr_status == "not_applicable") {
      paste0(
        " CMR is included for illustration only: its borrow_status is \"not_applicable\" -- ",
        "its universe is commodity futures/ETFs/price indices, where financing sits in the ",
        "futures curve, not an equity-style short-borrow rate, so this row is a sensitivity ",
        "illustration, not a cost CMR would actually pay."
      )
    } else ""
  } else ""

  sens_short <- paste0(
    "Sharpe sensitivity to the assumed borrow rate for the ", n_swept,
    " strategies with a real short leg — ordered by ", delta_col,
    " (most sensitive first); read that delta column, not raw Sharpe levels ",
    "(see below for why). Expand for the full methodology note."
  )

  sens_notes <- htmltools::tagList(
    "Sharpe at borrow rate ", paste0(round(rates_sorted * 100), "%", collapse = " / "),
    " (annualised), holding trade cost, signal, and weights fixed, for the ", n_swept,
    " strategies with a real short leg, computed by ", fn_link, " and assembled in ", sweep_link,
    " (reporting only -- feeds no published return, #665). Ordered by ", delta_col,
    " (most sensitive first): ", paste(most_affected, collapse = ", "),
    " lose the most Sharpe as the borrow assumption rises, so read this table from the top, ",
    "not from the leaderboard's own ranking. The 0% column is a no-financing counterfactual, ",
    "not a claim that financing is free. This sweep's Sharpe omits the risk-free rate (it ",
    "reports RELATIVE deltas across borrow rates, where rf cancels), so its levels are NOT ",
    "comparable to the leaderboard's own rf-adjusted Sharpe above -- ", delta_col,
    " is the column that means something here.",
    cmr_note
  )

  sens_caption <- hd_caption(
    sens_notes,
    short = sens_short,
    summary_label = "Full methodology note"
  )

  numeric_cols <- which(sapply(sens_table, function(x) is.numeric(x) ||
                                  grepl("^[0-9.%$,-]+$", as.character(x[1]))))

  DT::datatable(
    sens_table,
    # sens_caption is already a complete tags$caption() object (hd_caption()) --
    # do not re-wrap it (see the Rankings table above for why that matters).
    caption = sens_caption,
    rownames = FALSE,
    filter = "top",
    options = list(
      pageLength = 15,
      scrollX = TRUE,
      autoWidth = FALSE,
      dom = "frtip",
      order = list(),
      search = list(regex = TRUE, caseInsensitive = TRUE),
      columnDefs = list(
        list(className = 'dt-right', targets = numeric_cols - 1)
      ),
      initComplete = DT::JS(
        "function(settings, json) {",
        "  var isDark = document.documentElement.getAttribute('data-bs-theme') === 'dark'",
        "    || document.body.classList.contains('dark-mode');",
        "  if (isDark) {",
        "    $(this.api().table().container()).css({'color': '#ddd', 'background-color': '#1a1a1a'});",
        "    $(this.api().table().header()).css({'color': '#ddd', 'background-color': '#222'});",
        "    $('input', this.api().table().container()).css({'color': '#ddd', 'background-color': '#333', 'border-color': '#555'});",
        "  }",
        "}"
      )
    )
  )
}
```

#### Strategy Details

Each strategy links to its full definition, implementation, and backtest:

| Strategy | Level | Signal | Definition |
|----------|-------|--------|------------|
| [Factor MAX](stock-backtest.html#factor-level-deep-dive) | Factor (5 FF) | Max daily return in prior month | Ranks factors by single-day extreme |
| [Factor <abbr title="Dynamic Rotation Into Factors — selects best-performing FF factor each month">DRIF</abbr>](stock-backtest.html#factor-level-deep-dive) | Factor (5 FF) | Elastic net on 42 daily return features | Full daily return distribution |
| [Stock MAX](stock-backtest.html#stock-max) | Stock (660) | Max daily return, D1-D10 decile spread | Cross-sectional stock sorting |
| [Stock DRIF](stock-backtest.html#stock-drif) | Stock (660) | Elastic net, D1-D10 decile spread | Pooled cross-sectional elastic net |
| [XGB DRIF](stock-backtest.html#stock-drif) | Stock (660) | XGBoost monotonic on 42 features | Non-linear signal with monotonic constraints |
| [ETF DRIF A](stock-backtest.html#etf-replication) | ETF (9) | Elastic net on ETF daily returns | Direct ETF signal, long-only 13.4% CAGR |
| [ETF DRIF B](stock-backtest.html#etf-replication) | ETF (9) | Academic factor signal → ETF execution | Mapped via ETF_FACTOR_MAP, long-only 12.6% |
| [PSO Optimal](stock-backtest.html#comparison) | Combined | Weighted portfolio of above | [Grid search](stock-backtest.html#comparison) on training Sharpe |

```{r}
# Show PSO weights
w <- safe_tar_read("port_optimal_weights")
if (!is.null(w)) {
  w_df <- tibble(
    Strategy = names(w),
    Weight = paste0(round(w * 100), "%")
  ) |> filter(Weight != "0%")
  hd_dt(w_df, "PSO optimal weights (trained on Training partition Sharpe). Strategies with 0% weight omitted.")
}
```

# Falsification

## Row

### {.tabset}

#### Scorecard

The falsification framework tests 5 strategies against multiple statistical hurdles: <abbr title="Newey-West heteroskedasticity-autocorrelation consistent estimator">HAC</abbr>-corrected t-statistics, 6 null environment rejection rates, and <abbr title="Fama-French 5-Factor model (MktRF, SMB, HML, RMW, CMA)">FF5</abbr> + Momentum alpha regression. Strategies that pass all tests show genuine alpha; those that fail are explained by known factor exposure (beta). Note: this covers 5 strategies from the falsification pipeline — partially overlapping with the leaderboard strategies above.

```{r}
#| label: fals-scorecard
fals_sc <- safe_tar_read("fals_vig_scorecard")
fals_sc_cap <- safe_tar_read("fals_vig_scorecard_caption")
if (!is.null(fals_sc)) {
  cap_text <- if (!is.null(fals_sc_cap)) fals_sc_cap else "Strategy falsification scorecard"
  hd_dt(fals_sc, cap_text)
}
```

#### Null Rejection

Rejection rates across 6 null environments (M = 100 simulations each). A low rate means the strategy's performance is easily replicated under the null — no genuine signal. A high rate means the null cannot explain the observed returns.

```{r}
#| label: fals-heatmap
#| fig-height: 5
#| fig-width: 10
safe_tar_read("fals_vig_null_heatmap")
```

```{r}
#| label: fals-heatmap-caption
#| results: asis
cap <- safe_tar_read("fals_vig_null_heatmap_caption")
if (!is.null(cap)) cat(paste0("\n*", cap, "*\n"))
```

#### Alpha vs Factor Exposure

Scatter plot of annualised alpha (%) against R² (%) from the Fama-French 5-Factor + Momentum regression. Genuine alpha strategies cluster in the top-left (low factor exposure, positive alpha). Beta strategies cluster on the right (high R²).

```{r}
#| label: fals-alpha-plot
#| fig-height: 5
#| fig-width: 10
safe_tar_read("fals_vig_ff_alpha_plot")
```

```{r}
#| label: fals-alpha-caption
#| results: asis
cap <- safe_tar_read("fals_vig_ff_caption")
if (!is.null(cap)) cat(paste0("\n*", cap, "*\n"))
```

#### See Also

For detailed methodology, HAC analysis, factor regression betas, multiplicity correction (<abbr title="Effective number of independent strategies (spectral participation ratio)">K_eff</abbr>, delta-Z), and the full null environment heatmap, see the [Strategy Falsification dashboard](falsification.html).

# Performance

## Row

### {.tabset}

#### Equity Curves

```{r}
#| fig-height: 6
safe_tar_read("stk_all_comparison_plot")
```

::: {.fig-caption}
```{r}
#| echo: false
#| results: asis
cap <- safe_tar_read("stk_all_caption")
if (!is.null(cap)) cat(cap) else cat("*Caption not available — run tar_make() to generate.*")
```
:::

#### Portfolio

```{r}
#| fig-height: 6
safe_tar_read("port_comparison_plot")
```

::: {.fig-caption}
**PSO optimal vs equal-weight portfolio.** Combines all 4 strategies with weights optimised on training partition Sharpe. The PSO optimal tilts toward strategies with better risk-adjusted returns (Factor DRIF + Stock MAX).
:::

#### Monthly Returns

```{r}
monthly <- safe_tar_read("port_monthly_returns")
if (!is.null(monthly)) {
  # Most recent year first, columns Jan-Dec + Annual
  col_order <- c("Year", "Jan", "Feb", "Mar", "Apr", "May", "Jun",
                 "Jul", "Aug", "Sep", "Oct", "Nov", "Dec", "Annual")
  monthly <- monthly |>
    select(any_of(col_order)) |>
    arrange(desc(Year))

  # Add marginal row (mean across years)
  num_cols <- setdiff(col_order, "Year")
  monthly <- monthly |> mutate(Year = as.character(Year))
  marginal <- monthly |>
    summarise(across(all_of(num_cols), ~ format(round(mean(as.numeric(.), na.rm = TRUE), 1), nsmall = 1))) |>
    mutate(Year = "Mean")
  monthly <- bind_rows(monthly, marginal)

  DT::datatable(
    monthly,
    rownames = FALSE, filter = "none",
    options = list(
      pageLength = 60, scrollX = TRUE, dom = "t",
      order = list(list(0, "desc")),
      columnDefs = list(list(className = "dt-right", targets = seq_len(length(col_order) - 1)))
    ),
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left;",
      "Monthly returns (PSO optimal portfolio) by year, most recent first. Annual = compounded. Mean row = average across years. Cell borders: green = positive, red = negative."
    )
  ) |>
    DT::formatStyle(
      columns = num_cols,
      color = DT::styleInterval(0, c("#e74c3c", "#69d4a0")),
      border = DT::styleInterval(0, c("2px solid #e74c3c", "2px solid #69d4a0"))
    )
}
```

#### XGBoost vs Elastic Net

```{r}
#| fig-height: 6
safe_tar_read("xgb_vs_enet_plot")
```

::: {.fig-caption}
**XGBoost (monotonic constraints) vs Elastic Net** on same 42 DRIF features, stock-level D1-D10. XGBoost underperforms elastic net (-6.3% vs +1.3% full-period CAGR). Monotonic constraints may be too restrictive for cross-sectional returns. See [Stock-Level Backtests](stock-backtest.html#stock-drif) for full analysis.
:::

# Robustness

## Row

### {.tabset}

#### Bootstrap CI

```{r}
#| fig-height: 6
safe_tar_read("boot_sharpe_plot")
```

```{r}
boot_ci <- safe_tar_read("boot_ci_summary")
if (!is.null(boot_ci)) {
  display <- boot_ci |>
    transmute(
      Strategy = strategy_label,
      `Sharpe [5%, 95%]` = paste0(round(sharpe_mean, 2), " [", round(sharpe_lo, 2), ", ", round(sharpe_hi, 2), "]"),
      `CAGR [5%, 95%]` = paste0(round(cagr_mean * 100, 1), "% [", round(cagr_lo * 100, 1), "%, ", round(cagr_hi * 100, 1), "%]"),
      `Max DD [5%, 95%]` = paste0(round(dd_mean * 100, 1), "% [", round(dd_lo * 100, 1), "%, ", round(dd_hi * 100, 1), "%]"),
      `CI Crosses Zero` = ifelse(ci_crosses_zero, "YES — may be noise", "No")
    )
  hd_dt(display, "Block bootstrap (1000 draws, block=3m). If Sharpe CI crosses zero, strategy may be indistinguishable from noise.")
}
```

#### Alpha Decay

```{r}
#| fig-height: 6
# #356: Alpha decay needs a positive base Sharpe (decay is the fraction
# of base Sharpe that survives delay; if base is negative, the metric
# is meaningless). Surface the actual reason in the callout.
dp <- safe_tar_read("decay_plot")
decay_metrics <- safe_tar_read("decay_metrics")
neg_base <- !is.null(decay_metrics) && nrow(decay_metrics) > 0 &&
            all(decay_metrics$sharpe[decay_metrics$delay == min(decay_metrics$delay)] <= 0, na.rm = TRUE)
if (!is.null(dp) && !is.null(decay_metrics) && nrow(decay_metrics) > 0 && !neg_base) {
  print(dp)
} else if (neg_base) {
  cat("::: callout-warning\n",
      "**Alpha decay not meaningful — base Sharpe is non-positive for every strategy on the page.** ",
      "Decay measures the *fraction of base Sharpe that survives execution delay*; if base Sharpe ≤ 0 there is no alpha to decay. ",
      "Investigate the strategies' base Sharpe before re-running this tab. See <a href=\"https://github.com/JohnGavin/historical/issues/356\" target=\"_blank\" rel=\"noopener noreferrer\">#356</a>.\n",
      ":::\n", sep = "")
} else {
  cat("::: callout-note\n",
      "**Alpha decay analysis not yet computed.** ",
      "Run `tar_make(c(\"decay_metrics\", \"decay_plot\", \"decay_half_life\"))` ",
      "to populate this tab. See <a href=\"https://github.com/JohnGavin/historical/issues/356\" target=\"_blank\" rel=\"noopener noreferrer\">#356</a>.\n",
      ":::\n", sep = "")
}
```

```{r}
decay_hl <- safe_tar_read("decay_half_life")
if (!is.null(decay_hl) && nrow(decay_hl) > 0) {  # #356: don't render an empty DT
  display <- decay_hl |>
    transmute(
      Strategy = dplyr::recode(strategy, stk_max = "Stock MAX", stk_drif = "Stock DRIF"),
      `Base Sharpe (t+1)` = round(base_sharpe, 2),
      `Half-Life (days)` = ifelse(is.na(half_life_days), ">10 (robust)", paste0(half_life_days))
    )
  hd_dt(display, "Alpha decay: days until Sharpe drops to 50% of t+1 baseline. >10 = signal persists; <3 = fragile, execution-dependent.")
}
```

#### Regime

```{r}
#| fig-height: 6
safe_tar_read("regime_plot")
```

```{r}
regime_m <- safe_tar_read("regime_metrics")
if (!is.null(regime_m)) {
  display <- regime_m |>
    filter(period == "Full Period") |>
    transmute(
      Strategy = strategy,
      CAGR = paste0(round(cagr * 100, 1), "%"),
      Vol = paste0(round(vol * 100, 1), "%"),
      Sharpe = round(sharpe, 2),
      `Max DD` = paste0(round(max_dd * 100, 1), "%")
    )
  hd_dt(display, "Regime-adjusted vs base portfolio (full period). Regime scaling: low=1.0, med=0.7, high=0.4. High-risk months shaded red in plot.")
}
```

#### Kelly Sizing

```{r}
#| fig-height: 6
safe_tar_read("kelly_comparison_plot")
```

```{r}
kelly_t <- safe_tar_read("kelly_sizing_table")
if (!is.null(kelly_t)) {
  display <- kelly_t |>
    transmute(
      Strategy = strategy,
      `Mean Monthly Edge` = paste0(round(edge * 100, 2), "%"),
      `Monthly Variance` = round(variance, 4),
      `Full Kelly` = paste0(round(kelly_full * 100, 1), "%"),
      `Quarter Kelly` = paste0(round(kelly_frac * 100, 2), "%")
    )
  hd_dt(display, "Kelly criterion edge estimates (training data only). Quarter Kelly = 25% of full Kelly, capped at 5%. Negative edge = do not trade.")
}
```

# Structural Breaks

## Row {height=100%}

### {.tabset}

#### Interpretation

A **structural break** at date *T* means the returns process appears to have changed before and after that date — the distribution of vol-normalised returns differs significantly on each side. This is **evidence worth investigating**, not an automatic allocation trigger.

The `resulting-prohibition` rule applies directly here: a detected break does not justify revising strategy weights without identifying new *structural* evidence for why the change occurred. Past underperformance alone is not such evidence — Swedroe documents 12–39-year underperformance episodes for well-established risk premia (`underperformance-prior`). Conflating a drawdown with a structural change risks abandoning a strategy at exactly the wrong time.

Rob Carver's empirical finding (2026) is the calibrating prior: across instrument/forecast pairs, only ~13% show breaks at the 1% level, and whole-history estimation often beats break-adjusted estimation out-of-sample. **Break-splitting is a guard against over-splitting, not an always-on re-estimator.**

For the full framework context see the [market-behavior gap analysis](https://github.com/JohnGavin/historical/blob/main/knowledge/wiki/market-behavior-gap-analysis.md) (§ Structural Stability), which this section closes (gap now ✅).

#### Summary Table

```{r}
#| label: sb-table
sb_sum <- safe_tar_read("structural_breaks_summary")
sb_cap <- safe_tar_read("structural_breaks_caption")
snames  <- safe_tar_read("strategy_names")

if (!is.null(sb_sum) && !is.null(snames)) {
  # strategy_names join: structural_breaks_summary uses short_name labels
  sb_disp <- sb_sum |>
    left_join(
      snames |> select(short_name, long_name),
      by = c("strategy" = "short_name")
    ) |>
    mutate(Strategy = dplyr::coalesce(long_name, strategy)) |>
    transmute(
      Strategy,
      `Break Date(s)`     = break_dates,
      `N Breaks`          = ifelse(is.na(n_breaks), "—", as.character(n_breaks)),
      `Whole Sharpe`      = ifelse(
        is.na(whole_sharpe), "—",
        format(round(whole_sharpe,      3L), nsmall = 3L)
      ),
      `Post-Break Sharpe` = ifelse(
        is.na(post_break_sharpe), "—",
        format(round(post_break_sharpe, 3L), nsmall = 3L)
      ),
      `Divergence %`      = ifelse(
        is.na(sharpe_divergence_pct), "—",
        paste0(sharpe_divergence_pct, "%")
      ),
      # `material_divergence` is NA for two structurally different reasons
      # that were previously conflated into the same "—" (#643, same
      # conflation class as Credible, #637 follow-up):
      #   1. The break test itself errored or was skipped for this strategy
      #      (n_breaks is NA_integer_, see the `entry$error` branch in
      #      R/plan_structural_breaks.R) -- genuinely "not computed".
      #   2. The break test ran and found NO break at all (n_breaks == 0) --
      #      there is nothing to diverge from, so "no material divergence"
      #      is a real, assessed answer, not an unknown one. A break WAS
      #      found (n_breaks > 0) but the divergence percentage could not be
      #      reliably computed (near-zero whole_sharpe denominator or too
      #      few post-break observations) is its own indeterminate case and
      #      stays "not computed" -- a break with an unmeasurable divergence
      #      is not the same as "no material divergence".
      Material            = dplyr::case_when(
        is.na(n_breaks)                            ~ "not computed",
        n_breaks == 0L                             ~ "no",
        n_breaks > 0L & is.na(material_divergence) ~ "not computed",
        material_divergence                        ~ "YES",
        TRUE                                       ~ "no"
      ),
      # Explicit sort rank for Material -- same rationale as .credible_rank /
      # .redundant_rank: DT's default string comparator sorts "YES" / "no" /
      # "not computed" alphabetically, not assessed-good, assessed-bad,
      # never-assessed (#643).
      .material_rank      = dplyr::case_when(
        is.na(n_breaks)                            ~ 3L,
        n_breaks == 0L                             ~ 1L,
        n_breaks > 0L & is.na(material_divergence) ~ 3L,
        material_divergence                        ~ 2L,
        TRUE                                       ~ 1L
      ),
      Note                = ifelse(is.na(note), "", note)
    )

  cap_text <- if (!is.null(sb_cap)) sb_cap else
    "Structural break analysis — run tar_make() to populate."

  # Explicit sort rank for Material -- same rationale as .credible_rank /
  # .redundant_rank above: DT's default string comparator sorts the display
  # string alphabetically, not assessed-good, assessed-bad, never-assessed.
  material_col      <- which(names(sb_disp) == "Material") - 1L
  material_rank_col <- which(names(sb_disp) == ".material_rank") - 1L

  DT::datatable(
    sb_disp,
    caption  = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold;",
      class = "dt-caption",
      cap_text
    ),
    rownames = FALSE,
    filter   = "top",
    options  = list(
      pageLength   = 20L,
      scrollX      = TRUE,
      dom          = "frtip",
      columnDefs   = list(
        list(visible = FALSE, targets = material_rank_col),
        list(orderData = material_rank_col, targets = material_col)
      ),
      initComplete = DT::JS(
        "function(settings, json) {",
        "  var isDark = document.documentElement.getAttribute('data-bs-theme') === 'dark'",
        "    || document.body.classList.contains('dark-mode');",
        "  if (isDark) {",
        "    $(this.api().table().container()).css({'color': '#ddd', 'background-color': '#1a1a1a'});",
        "    $(this.api().table().header()).css({'color': '#ddd', 'background-color': '#222'});",
        "    $('input', this.api().table().container()).css({'color': '#ddd', 'background-color': '#333', 'border-color': '#555'});",
        "  }",
        "}"
      )
    )
  ) |>
    DT::formatStyle(
      "Material",
      target          = "row",
      backgroundColor = DT::styleEqual("YES", "rgba(255, 99, 71, 0.18)")
    )
}
```

#### Sharpe Comparison

```{r}
#| label: sb-dotplot
#| fig-height: 7
#| fig-width: 9
# Path-checked pkgload (cwd-trap fix: here::here() returns docs/ under quarto render)
pkg_path <- if (dir.exists("packages/historicaldata")) "packages/historicaldata" else
  file.path(dirname(here::here()), "packages/historicaldata")
pkgload::load_all(pkg_path, quiet = TRUE)

sb_sum <- safe_tar_read("structural_breaks_summary")
snames  <- safe_tar_read("strategy_names")

if (!is.null(sb_sum) && !is.null(snames)) {
  plot_df <- sb_sum |>
    filter(!is.na(whole_sharpe)) |>
    left_join(
      snames |> select(short_name, long_name),
      by = c("strategy" = "short_name")
    ) |>
    mutate(
      label     = dplyr::coalesce(long_name, strategy),
      has_break = factor(
        ifelse(!is.na(n_breaks) & n_breaks > 0L, "Break at 1%", "No break"),
        levels = c("No break", "Break at 1%")
      )
    )

  strat_order <- plot_df |>
    arrange(whole_sharpe) |>
    pull(label)

  long_df <- plot_df |>
    select(label, has_break,
           `Whole History` = whole_sharpe,
           `Post-Break`    = post_break_sharpe) |>
    tidyr::pivot_longer(
      c(`Whole History`, `Post-Break`),
      names_to  = "Period",
      values_to = "Sharpe"
    ) |>
    filter(!is.na(Sharpe)) |>
    mutate(
      label  = factor(label,  levels = strat_order),
      Period = factor(Period, levels = c("Whole History", "Post-Break"))
    )

  ggplot(long_df, aes(x = Sharpe, y = label, colour = Period, shape = has_break)) +
    geom_vline(xintercept = 0, linetype = "dotted",
               colour = "grey50", linewidth = 0.4) +
    geom_point(size = 3.5, alpha = 0.9) +
    scale_colour_manual(
      values = hd_palette(2),
      name   = "Estimate period"
    ) +
    scale_shape_manual(
      values = c("No break" = 16L, "Break at 1%" = 17L),
      name   = "Break status"
    ) +
    labs(
      x     = "Annualised Sharpe Ratio",
      y     = NULL,
      title = "Whole-history vs post-break Sharpe"
    ) +
    hd_theme()
}
```

```{r}
#| label: sb-dotplot-caption
#| results: asis
sb_cap <- safe_tar_read("structural_breaks_caption")
fig_cap <- paste0(
  "**Figure — Cleveland dot plot:** whole-history Sharpe (circle) vs ",
  "post-break Sharpe (where a break was detected) for all leaderboard strategies. ",
  "**Data:** return series pre-computed in the docs targets pipeline; see ",
  "[plan_structural_breaks.R](https://github.com/JohnGavin/historical/blob/main/R/plan_structural_breaks.R). ",
  "**Variables:** annualised Sharpe ratio; Whole History = full return series, ",
  "Post-Break = segment after the first detected break (triangle shape). ",
  "**Labels:** `long_name` from the `strategy_names` target ",
  "([plan_strategy_names.R](https://github.com/JohnGavin/historical/blob/main/R/plan_strategy_names.R)), ",
  "ordered by whole-history Sharpe ascending. ",
  "**Methodology:** iterative forward-split t-test at 1% level with 5-year minimum segment; see ",
  "[hd_structural_breaks()](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/structural_breaks.R). ",
  "**Caveats:** scanning all candidate split dates inflates false-break rate above the nominal 1% level; ",
  "a detected break is evidence to investigate, not an allocation trigger (`resulting-prohibition`); ",
  "whole-history estimation often beats break-adjusted OOS (Carver 2026). ",
  "**Chart type:** Cleveland dot plot per visualization-standards (dotchart first choice). ",
  if (!is.null(sb_cap)) paste0("Dynamic summary: ", sb_cap) else ""
)
cat(paste0("\n*", fig_cap, "*\n"))
```

# Details

## Row

### {.tabset}

#### Correlations

```{r}
corr <- safe_tar_read("strategy_correlation")
if (!is.null(corr)) {
  corr_display <- corr |>
    mutate(across(-strategy, ~ round(as.numeric(.), 2)))
  knitr::kable(corr_display, caption = "Pairwise correlation (Pearson, monthly returns).")
}
```

#### Assumptions

Per `backtesting-assumptions` rule, all strategies use:

| Assumption | Value |
|------------|-------|
| Execution | Market close (4 PM ET) |
| Transaction cost | 0.50% per trade (1.00% round-trip) |
| Estimated turnover | 80% per month (decile sort) |
| Short borrow cost | 3% per annum (0.25%/month) |
| Return winsorisation | ±20% per month |
| Cash return | 3-month US Treasury rate (RF from FF3) |
| Dividends | Reinvested (total return basis) |
| Taxes | Excluded (pre-tax) |

**Total monthly cost for L/S:** turnover × cost × 2 legs × 2 (buy+sell) + borrow = 0.80 × 0.005 × 4 + 0.0025 = **~1.85%/month**. This is why most stock-level L/S strategies are unprofitable after costs.

#### Partitions

```{r}
params <- safe_tar_read("stk_params")
if (!is.null(params)) {
  # Four-tier re-cut (#660): Holdout is the observed-but-unsealed tier that
  # absorbs the 2024-2026 span burned by #660 (docs/stock-backtest.qmd
  # published Validation figures and drew a strategy conclusion from them).
  # It is computed automatically, unlike Validation -- see
  # .claude/rules/backtest-partitions.md.
  param_df <- tibble(
    Partition = c("Training", "Testing", "Holdout", "Validation"),
    Period = c(
      paste(params$start_date, "to", params$is_end),
      paste(params$test_start, "to", params$test_end),
      paste(params$holdout_start, "to", params$holdout_end),
      paste(params$val_start, "to", "present")
    ),
    Purpose = c(
      "Model fitting, expanding window signal estimation",
      "Calibration, hyperparameter tuning, strategy comparison",
      "OBSERVED, NOT SEALED — usable for evaluation with a stated discount; never a one-shot result (#660)",
      "SEALED — not computed or shown until production deployment"
    )
  )
  hd_dt(param_df, "Shared partitions across all strategies. Holdout is observed but not sealed (#660). Validation is sealed: metrics are NOT computed until a one-shot production decision.")
}
```

#### Related Vignettes

- **[Strategy Falsification](falsification.html)** — provides the statistical survival verdict (HAC t-stat, null rejection rate, FF5 alpha) underlying the leaderboard scorecard
- **[Stock Backtest](stock-backtest.html)** — per-asset equity-level view of Factor MAX and DRIF strategies ranked here
- **[Factor MAX](stock-backtest.html#factor-level-deep-dive)** — detailed breakdown of the top-ranked factor rotation strategy

#### Build Info

```{r}
#| label: build-info
#| echo: false
#| results: asis
build_info()
```
