---
title: "Strategy Leaderboard"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    embed-resources: false
    include-in-header:
      - text: |
          <link rel="stylesheet" href="vignette-shared.css">
          <script src="vignette-shared.js"></script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)
library(dplyr)
library(tidyr)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)
```

# Rankings

## Row

### {.tabset}

#### Full Period

```{r}
lb <- safe_tar_read("leaderboard")
if (!is.null(lb)) {
  full <- lb |>
    filter(period == "Full Period") |>
    arrange(max_dd)

  # Transpose: risk first, then return (net only — no gross)
  metrics_long <- full |>
    mutate(
      `Max DD` = paste0(round(max_dd * 100, 1), "%"),
      `CVaR 95%` = ifelse(is.na(cvar_95), "—", paste0(round(cvar_95 * 100, 1), "%")),
      Vol = paste0(round(vol * 100, 1), "%"),
      `Net CAGR` = ifelse(is.na(net_cagr), "—", paste0(round(net_cagr * 100, 1), "%")),
      Sharpe = ifelse(is.na(sharpe), "—", as.character(round(sharpe, 2))),
      Credible = ifelse(is.na(credible) | !credible, "NO", "yes")
    ) |>
    select(strategy, `Max DD`, `CVaR 95%`, Vol, `Net CAGR`, Sharpe, Credible) |>
    pivot_longer(-strategy, names_to = "Metric", values_to = "value") |>
    pivot_wider(names_from = strategy, values_from = value)

  # Flag non-credible strategies
  n_flagged <- sum(!full$credible, na.rm = TRUE)
  flag_note <- if (n_flagged > 0) {
    flagged <- full$strategy[!full$credible]
    paste0(" WARNING: ", paste(flagged, collapse=", "),
           " flagged as non-credible (>20% CAGR or >50x cum P&L — likely unrealistic).")
  } else ""

  hd_dt(metrics_long, paste0("Risk-first rankings (sorted by max drawdown). Risk: DD, CVaR, Vol. Return: Net CAGR (after 0.20% round-trip/month), Cum P&L (net). Sharpe secondary.", flag_note))
}
```

#### By Partition

```{r}
if (!is.null(lb)) {
  # NEVER show Validation data — sealed envelope (backtest-partitions rule)
  by_part <- lb |>
    filter(period != "Validation") |>
    mutate(
      CAGR = paste0(round(cagr * 100, 1), "%"),
      Vol = paste0(round(vol * 100, 1), "%"),
      Sharpe = ifelse(is.na(sharpe), "—", as.character(round(sharpe, 2))),
      `Max DD` = paste0(round(max_dd * 100, 1), "%")
    ) |>
    select(Strategy = strategy, Period = period, CAGR, Vol, Sharpe, `Max DD`)

  hd_dt(by_part, "All strategies × Training and Testing partitions. Validation partition is sealed — not shown until production deployment.")
}
```

#### Strategy Details

Each strategy links to its full definition, implementation, and backtest:

| Strategy | Level | Signal | Definition |
|----------|-------|--------|------------|
| [Factor MAX](factor-max.html) | Factor (5 FF) | Max daily return in prior month | Ranks factors by single-day extreme |
| [Factor DRIF](drif.html) | Factor (5 FF) | Elastic net on 42 daily return features | Full daily return distribution |
| [Stock MAX](stock-backtest.html#stock-max) | Stock (660) | Max daily return, D1-D10 decile spread | Cross-sectional stock sorting |
| [Stock DRIF](stock-backtest.html#stock-drif) | Stock (660) | Elastic net, D1-D10 decile spread | Pooled cross-sectional elastic net |
| [XGB DRIF](stock-backtest.html#stock-drif) | Stock (660) | XGBoost monotonic on 42 features | Non-linear signal with monotonic constraints |
| [ETF DRIF A](stock-backtest.html#etf-replication) | ETF (9) | Elastic net on ETF daily returns | Direct ETF signal, long-only 13.4% CAGR |
| [ETF DRIF B](stock-backtest.html#etf-replication) | ETF (9) | Academic factor signal → ETF execution | Mapped via ETF_FACTOR_MAP, long-only 12.6% |
| [PSO Optimal](stock-backtest.html#comparison) | Combined | Weighted portfolio of above | [Grid search](stock-backtest.html#comparison) on training Sharpe |

```{r}
# Show PSO weights
w <- safe_tar_read("port_optimal_weights")
if (!is.null(w)) {
  w_df <- tibble(
    Strategy = names(w),
    Weight = paste0(round(w * 100), "%")
  ) |> filter(Weight != "0%")
  hd_dt(w_df, "PSO optimal weights (trained on Training partition Sharpe). Strategies with 0% weight omitted.")
}
```

# Performance

## Row

### {.tabset}

#### Equity Curves

```{r}
#| fig-height: 6
safe_tar_read("stk_all_comparison_plot")
```

::: {.fig-caption}
**Equity curves (4 factor/stock strategies).** Growth of $1, log scale. Stock-level strategies (MAX, DRIF) lose money net of costs despite the same signal being profitable at factor level. The difference is execution cost, not diversification: stock-level decile sorts produce ~80% monthly turnover across 130 positions at 0.50%/trade = ~1.85%/month total cost (~22%/yr). Factor-level trades 2-4 positions with ~40% turnover at 0.10%/trade = ~0.16%/month (~2%/yr). A 7% gross spread survives 2% cost drag (factor) but not 22% (stock). Dashed line = test partition start (2020).
:::

#### Portfolio

```{r}
#| fig-height: 6
safe_tar_read("port_comparison_plot")
```

::: {.fig-caption}
**PSO optimal vs equal-weight portfolio.** Combines all 4 strategies with weights optimised on training partition Sharpe. The PSO optimal tilts toward strategies with better risk-adjusted returns (Factor DRIF + Stock MAX).
:::

#### Monthly Returns

```{r}
monthly <- safe_tar_read("port_monthly_returns")
if (!is.null(monthly)) {
  # Most recent year first, columns Jan-Dec + Annual
  col_order <- c("Year", "Jan", "Feb", "Mar", "Apr", "May", "Jun",
                 "Jul", "Aug", "Sep", "Oct", "Nov", "Dec", "Annual")
  monthly <- monthly |>
    select(any_of(col_order)) |>
    arrange(desc(Year))

  # Add marginal row (mean across years)
  num_cols <- setdiff(col_order, "Year")
  marginal <- monthly |>
    summarise(across(all_of(num_cols), ~ format(round(mean(as.numeric(.), na.rm = TRUE), 1), nsmall = 1))) |>
    mutate(Year = "Mean")
  monthly <- bind_rows(monthly, marginal)

  DT::datatable(
    monthly,
    rownames = FALSE, filter = "none",
    options = list(
      pageLength = 25, scrollX = TRUE, dom = "t",
      columnDefs = list(list(className = "dt-right", targets = seq_len(length(col_order) - 1)))
    ),
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left;",
      "Monthly returns (PSO optimal portfolio) by year, most recent first. Annual = compounded. Mean row = average across years. Cell borders: green = positive, red = negative."
    )
  ) |>
    DT::formatStyle(
      columns = num_cols,
      color = DT::styleInterval(0, c("#e74c3c", "#2ecc71")),
      border = DT::styleInterval(0, c("2px solid #e74c3c", "2px solid #2ecc71"))
    )
}
```

#### XGBoost vs Elastic Net

```{r}
#| fig-height: 6
safe_tar_read("xgb_vs_enet_plot")
```

::: {.fig-caption}
**XGBoost (monotonic constraints) vs Elastic Net** on same 42 DRIF features, stock-level D1-D10. XGBoost underperforms elastic net (-6.3% vs +1.3% full-period CAGR). Monotonic constraints may be too restrictive for cross-sectional returns. See [Stock-Level Backtests](stock-backtest.html#stock-drif) for full analysis.
:::

# Robustness

## Row

### {.tabset}

#### Bootstrap CI

```{r}
#| fig-height: 6
safe_tar_read("boot_sharpe_plot")
```

```{r}
boot_ci <- safe_tar_read("boot_ci_summary")
if (!is.null(boot_ci)) {
  display <- boot_ci |>
    transmute(
      Strategy = strategy_label,
      `Sharpe [5%, 95%]` = paste0(round(sharpe_mean, 2), " [", round(sharpe_lo, 2), ", ", round(sharpe_hi, 2), "]"),
      `CAGR [5%, 95%]` = paste0(round(cagr_mean * 100, 1), "% [", round(cagr_lo * 100, 1), "%, ", round(cagr_hi * 100, 1), "%]"),
      `Max DD [5%, 95%]` = paste0(round(dd_mean * 100, 1), "% [", round(dd_lo * 100, 1), "%, ", round(dd_hi * 100, 1), "%]"),
      `CI Crosses Zero` = ifelse(ci_crosses_zero, "YES — may be noise", "No")
    )
  hd_dt(display, "Block bootstrap (1000 draws, block=3m). If Sharpe CI crosses zero, strategy may be indistinguishable from noise.")
}
```

#### Alpha Decay

```{r}
#| fig-height: 6
safe_tar_read("decay_plot")
```

```{r}
decay_hl <- safe_tar_read("decay_half_life")
if (!is.null(decay_hl)) {
  display <- decay_hl |>
    transmute(
      Strategy = dplyr::recode(strategy, stk_max = "Stock MAX", stk_drif = "Stock DRIF"),
      `Base Sharpe (t+1)` = round(base_sharpe, 2),
      `Half-Life (days)` = ifelse(is.na(half_life_days), ">10 (robust)", paste0(half_life_days))
    )
  hd_dt(display, "Alpha decay: days until Sharpe drops to 50% of t+1 baseline. >10 = signal persists; <3 = fragile, execution-dependent.")
}
```

#### Regime

```{r}
#| fig-height: 6
safe_tar_read("regime_plot")
```

```{r}
regime_m <- safe_tar_read("regime_metrics")
if (!is.null(regime_m)) {
  display <- regime_m |>
    filter(period == "Full Period") |>
    transmute(
      Strategy = strategy,
      CAGR = paste0(round(cagr * 100, 1), "%"),
      Vol = paste0(round(vol * 100, 1), "%"),
      Sharpe = round(sharpe, 2),
      `Max DD` = paste0(round(max_dd * 100, 1), "%")
    )
  hd_dt(display, "Regime-adjusted vs base portfolio (full period). Regime scaling: low=1.0, med=0.7, high=0.4. High-risk months shaded red in plot.")
}
```

#### Kelly Sizing

```{r}
#| fig-height: 6
safe_tar_read("kelly_comparison_plot")
```

```{r}
kelly_t <- safe_tar_read("kelly_sizing_table")
if (!is.null(kelly_t)) {
  display <- kelly_t |>
    transmute(
      Strategy = strategy,
      `Mean Monthly Edge` = paste0(round(edge * 100, 2), "%"),
      `Monthly Variance` = round(variance, 4),
      `Full Kelly` = paste0(round(kelly_full * 100, 1), "%"),
      `Quarter Kelly` = paste0(round(kelly_frac * 100, 2), "%")
    )
  hd_dt(display, "Kelly criterion edge estimates (training data only). Quarter Kelly = 25% of full Kelly, capped at 5%. Negative edge = do not trade.")
}
```

# Details

## Row

### {.tabset}

#### Correlations

```{r}
corr <- safe_tar_read("strategy_correlation")
if (!is.null(corr)) {
  corr_display <- corr |>
    mutate(across(-strategy, ~ round(as.numeric(.), 2)))
  knitr::kable(corr_display, caption = "Pairwise correlation (Pearson, monthly returns).")
}
```

#### Assumptions

Per `backtesting-assumptions` rule, all strategies use:

| Assumption | Value |
|------------|-------|
| Execution | Market close (4 PM ET) |
| Transaction cost | 0.50% per trade (1.00% round-trip) |
| Estimated turnover | 80% per month (decile sort) |
| Short borrow cost | 3% per annum (0.25%/month) |
| Return winsorisation | ±20% per month |
| Cash return | 3-month US Treasury rate (RF from FF3) |
| Dividends | Reinvested (total return basis) |
| Taxes | Excluded (pre-tax) |

**Total monthly cost for L/S:** turnover × cost × 2 legs × 2 (buy+sell) + borrow = 0.80 × 0.005 × 4 + 0.0025 = **~1.85%/month**. This is why most stock-level L/S strategies are unprofitable after costs.

#### Partitions

```{r}
params <- safe_tar_read("stk_params")
if (!is.null(params)) {
  param_df <- tibble(
    Partition = c("Training", "Testing", "Validation"),
    Period = c(
      paste(params$start_date, "to", params$is_end),
      paste(params$test_start, "to", params$test_end),
      paste(params$val_start, "to", "present")
    ),
    Purpose = c("Model fitting, expanding window signal estimation",
                "Calibration, hyperparameter tuning, strategy comparison",
                "SEALED — not computed or shown until production deployment")
  )
  hd_dt(param_df, "Shared partitions across all strategies. Validation is sealed: metrics are NOT computed until a one-shot production decision.")
}
```

#### Build Info

```{r}
#| label: build-info
#| echo: false
#| results: asis
build_info()
```
