---
title: "Strategy Leaderboard"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    embed-resources: false
    include-in-header:
      - text: |
          <link rel="stylesheet" href="vignette-shared.css">
          <script src="vignette-shared.js"></script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)
library(dplyr)
library(tidyr)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)

disc_path <- if (file.exists("../R/disclosures.R")) "../R/disclosures.R" else
  file.path(here::here(), "R/disclosures.R")
source(disc_path)
```

# Rankings

::: {.callout-warning collapse="false" title="Survivorship bias"}
`r disclosure_survivorship()`
:::

## Row

### {.tabset}

#### Full Period

```{r}
lb <- safe_tar_read("leaderboard")
if (!is.null(lb)) {
  full <- lb |>
    filter(period == "Full Period") |>
    arrange(desc(sharpe))

  metrics_table <- full |>
    mutate(
      Sharpe     = ifelse(is.na(sharpe),       "—", as.character(round(sharpe, 2))),
      `Net CAGR` = ifelse(is.na(net_cagr),     "—", paste0(round(net_cagr * 100, 1), "%")),
      `Max DD`   = paste0(round(max_dd * 100, 1), "%"),
      `CVaR 95%` = ifelse(is.na(cvar_95),      "—", paste0(round(cvar_95 * 100, 1), "%")),
      Vol        = paste0(round(vol * 100, 1), "%"),
      Gross      = dplyr::case_when(
        !("gross_convention" %in% names(full)) | is.na(gross_convention) ~ "—",
        is_cap ~ paste0(format(gross_convention, nsmall = 1), "x*"),
        TRUE   ~ paste0(format(gross_convention, nsmall = 1), "x")
      ),
      `Cost (bps)` = dplyr::case_when(
        !("cost_per_trade_bps" %in% names(full)) | is.na(cost_per_trade_bps) ~ "—",
        cost_per_trade_bps == 0 ~ "0*",
        TRUE ~ as.character(round(cost_per_trade_bps))
      ),
      SSR        = ifelse(is.na(ssr),           "—", as.character(round(ssr, 2))),
      `Top5% Share` = ifelse(is.na(top5pct_share), "—", paste0(round(top5pct_share * 100, 1), "%")),
      # NA credible means cost metrics were never computed for this strategy
      # (cost_rows in R/plan_leaderboard.R only covers 5 core strategies +
      # PSO Optimal) -- that is "unknown", not "explicitly failed". Conflating
      # the two previously fed literal NA strategy names into the WARNING
      # note below (#637 follow-up).
      Credible   = dplyr::case_when(
        is.na(credible) ~ "—",
        !credible        ~ "NO",
        TRUE              ~ "yes"
      ),
      Redundant  = if ("redundant" %in% names(full)) {
        ifelse(is.na(redundant), "—", ifelse(redundant, "YES", "no"))
      } else "—",
      `Incremental Sharpe` = if ("incremental_sharpe" %in% names(full)) {
        ifelse(is.na(incremental_sharpe), "—", as.character(round(incremental_sharpe, 3)))
      } else "—"
    ) |>
    select(Strategy = strategy, Credible, Sharpe, `Net CAGR`, `Max DD`, `CVaR 95%`, Vol, Gross,
           `Cost (bps)`, SSR, `Top5% Share`, Redundant, `Incremental Sharpe`)

  # NA credible (not is.na(credible) | !credible) is deliberate here: `!NA`
  # is NA, and indexing a character vector with a logical vector containing
  # NA inserts a literal NA element into the result -- that is the root
  # cause of the "WARNING: NA, NA, NA, ..." defect (#637 follow-up). Only
  # strategies with an explicit, computed FALSE are flagged.
  n_flagged <- sum(!is.na(full$credible) & !full$credible)
  flag_note <- if (n_flagged > 0) {
    flagged <- full$strategy[!is.na(full$credible) & !full$credible]
    paste0(" WARNING: ", paste(flagged, collapse = ", "),
           " flagged as non-credible (>20% CAGR or >50x cum P&L — likely unrealistic).")
  } else ""

  n_biased <- if ("survivorship_biased" %in% names(full)) sum(full$survivorship_biased, na.rm = TRUE) else 0L
  bias_note <- if (n_biased > 0) {
    biased <- full$strategy[!is.na(full$survivorship_biased) & full$survivorship_biased]
    paste0(" Survivorship-biased (`survivorship_biased = TRUE`): ",
           paste(biased, collapse = ", "),
           " — draw on `stk_universe` (see issue #150).")
  } else ""

  # ── Gross-exposure disclosure note (#626) — dynamically computed, never hardcoded ──
  gh_url <- tryCatch({
    remote <- system("git remote get-url origin 2>/dev/null", intern = TRUE)
    sub("\\.git$", "", sub("^git@github\\.com:", "https://github.com/", remote))
  }, error = function(e) NULL)
  git_sha <- tryCatch(
    system("git rev-parse HEAD 2>/dev/null", intern = TRUE),
    error = function(e) "main"
  )
  git_sha <- if (length(git_sha) == 0 || !nzchar(git_sha)) "main" else git_sha
  gh_base <- if (!is.null(gh_url) && length(gh_url) > 0) gh_url else
    "https://github.com/JohnGavin/historical"
  hd_exposure_link <- sprintf(
    "[`hd_exposure_metrics()`](%s/blob/%s/packages/historicaldata/R/exposure.R#L67)",
    gh_base, git_sha
  )
  gross_registry_link <- sprintf(
    "[`strategy_gross_convention`](%s/blob/%s/R/plan_exposure.R#L27)",
    gh_base, git_sha
  )

  n_2x <- if ("gross_convention" %in% names(full)) {
    sum(full$gross_convention >= 2, na.rm = TRUE)
  } else 0L
  n_capped <- if ("is_cap" %in% names(full)) sum(full$is_cap, na.rm = TRUE) else 0L
  n_na_gross <- if ("gross_convention" %in% names(full)) sum(is.na(full$gross_convention)) else nrow(full)

  gross_note <- paste0(
    " Gross exposure (`sum(abs(w))`, computed by ", hd_exposure_link,
    " and recorded per strategy in ", gross_registry_link,
    ") is not uniform across this table: ", n_2x,
    " strategies are built dollar-neutral long-short at 2.0x gross while others run long-only at 1.0x, ",
    "so Vol, Net CAGR, and Max DD are not directly comparable across rows even when Sharpe is similar ",
    "(Sharpe is leverage-invariant under proportional scaling; the other three columns are not). ",
    "A trailing `*` on the Gross column (", n_capped,
    " strategies) marks a construction **ceiling** the strategy can reach but does not realise every period ",
    "(e.g. Managed Futures targets volatility and only hits its 3.0x cap in low-vol regimes) — treat those as upper bounds, not realised averages. ",
    if (n_na_gross > 0) paste0(n_na_gross, " strategy row(s) show — for Gross: either the return is assembled from a pre-computed factor series with no explicit weight vector to measure, or the row is a meta-portfolio blend of other rows.") else ""
  )

  # ── Transaction-cost disclosure note (#624) — dynamically computed, never hardcoded ──
  cost_registry_link <- sprintf(
    "[`strategy_cost_convention`](%s/blob/%s/R/plan_cost_convention.R#L44)",
    gh_base, git_sha
  )
  n_zero_cost <- if ("cost_per_trade_bps" %in% names(full)) {
    sum(full$cost_per_trade_bps == 0, na.rm = TRUE)
  } else 0L
  zero_cost_strategies <- if (n_zero_cost > 0) {
    full$strategy[!is.na(full$cost_per_trade_bps) & full$cost_per_trade_bps == 0]
  } else character(0)
  n_na_cost <- if ("cost_per_trade_bps" %in% names(full)) sum(is.na(full$cost_per_trade_bps)) else nrow(full)
  cost_bps_range <- if ("cost_per_trade_bps" %in% names(full) &&
                          any(!is.na(full$cost_per_trade_bps))) {
    rng <- range(full$cost_per_trade_bps, na.rm = TRUE)
    paste0(rng[1], "–", rng[2], " bps")
  } else "not available"

  cost_note <- paste0(
    " Per-strategy transaction-cost assumptions (recorded in ", cost_registry_link,
    ") are **not uniform** across this table, in both magnitude (", cost_bps_range,
    ") and unit: most rows charge ",
    "a per-trade bps figure on realised turnover, but Value (HML) and Managed Futures charge a flat ",
    "per-calendar-month bps figure regardless of turnover, and TOM/Risk State charge only on discrete ",
    "switch days — see `cost_convention` in the registry for each row's exact timing. ",
    if (n_zero_cost > 0) {
      paste0("A trailing `*` on the Cost (bps) column (", paste(zero_cost_strategies, collapse = ", "),
             ") marks a **verified zero** — no transaction-cost deduction exists in that strategy's code, ",
             "not a plausible default; it is not net-of-cost comparable to the cost-charging rows above and below it. ")
    } else "",
    if (n_na_cost > 0) paste0(n_na_cost, " strategy row(s) show — for Cost (bps): a meta-portfolio blend inherits its constituents' already-net returns with no separate cost applied at the blend level. ") else "",
    "This Cost (bps) column is the strategy's own internal assumption baked into its returns before they reach this table; ",
    "for the ", sum(!is.na(full$net_cagr)),
    " strategies with a Net CAGR value above, it is additional to — not the same as — the flat cost applied there."
  )

  # ── Diversification-value disclosure note (Pillar 7, #268; surfaced #624) ──
  corr_registry_link <- sprintf(
    "[`plan_strategy_correlation.R`](%s/blob/%s/R/plan_strategy_correlation.R#L31)",
    gh_base, git_sha
  )
  n_redundant <- if ("redundant" %in% names(full)) sum(full$redundant, na.rm = TRUE) else 0L
  diversification_note <- paste0(
    " Redundant = TRUE when a strategy's Full-Period monthly-return correlation with another, ",
    "higher-Sharpe strategy exceeds the threshold fixed in ", corr_registry_link,
    " (", n_redundant, " of ", nrow(full),
    " strategies currently flagged); Incremental Sharpe = the change in equal-weight-portfolio Sharpe from ",
    "including this strategy vs excluding it — positive means it adds diversification value, negative means ",
    "the portfolio is better off without it. Both feed the new-strategy value gate (#496)."
  )

  hd_dt(metrics_table, paste0(
    "Strategies ranked by Sharpe (highest first). ",
    "Net CAGR after 0.20% round-trip/month cost where available. ",
    "SSR = Sharpe Stability Ratio (HAC t-stat of rolling 36-month Sharpes; ≥2 = consistent). ",
    "Top5% Share = fraction of total return from best 5% of months. ",
    "Credible = plausible CAGR (<20%) and cum P&L (<50x); — where cost-adjusted metrics are not computed for this strategy.",
    flag_note, bias_note, gross_note, cost_note, diversification_note
  ))
}
```

#### By Partition

```{r}
if (!is.null(lb)) {
  # NEVER show Validation data — sealed envelope (backtest-partitions rule)
  params <- safe_tar_read("stk_params")

  # Create years lookup
  years_map <- tibble(
    period = c("Training", "Testing", "Full Period"),
    Years = if (!is.null(params)) {
      c(
        paste0(format(as.Date(params$start_date), "%Y"), "-", format(as.Date(params$is_end), "%Y")),
        paste0(format(as.Date(params$test_start), "%Y"), "-", format(as.Date(params$test_end), "%Y")),
        paste0(format(as.Date(params$start_date), "%Y"), "-", format(as.Date(params$test_end), "%Y"))
      )
    } else {
      c("—", "—", "—")
    }
  )

  # Explicit period display order (NOT alphabetical, per rule): Testing,
  # Full Period, Training. Implemented via factor levels so dplyr::arrange()
  # sorts by this order rather than alphabetically ("Full Period" < "Testing"
  # < "Training").
  period_order <- c("Testing", "Full Period", "Training")

  by_part <- lb |>
    filter(period != "Validation") |>
    left_join(years_map, by = "period") |>
    mutate(period = factor(period, levels = period_order)) |>
    # Sort on the NUMERIC sharpe column before it is formatted to character
    # below -- sorting the formatted string ("9.5" vs "10.2") would sort
    # lexicographically and put "9.5" above "10.2".
    arrange(period, desc(sharpe)) |>
    mutate(
      CAGR = paste0(round(cagr * 100, 1), "%"),
      Vol = paste0(round(vol * 100, 1), "%"),
      Sharpe = ifelse(is.na(sharpe), "—", as.character(round(sharpe, 2))),
      `Max DD` = paste0(round(max_dd * 100, 1), "%"),
      `CVaR 95%` = ifelse(is.na(cvar_95), "—", paste0(round(cvar_95 * 100, 1), "%")),
      Period = as.character(period)
    ) |>
    select(Strategy = strategy, Period, Sharpe, Years, CAGR, Vol, `Max DD`, `CVaR 95%`)

  # Bespoke DT call (not hd_dt()): rows above are already sorted by Period
  # (Testing / Full Period / Training) then Sharpe descending. hd_dt()'s
  # fixed options list has no `order` entry, so DT would apply its default
  # initial sort (column 0 ascending = Strategy) and silently undo the sort
  # above -- `order = list()` (empty array) disables DT's initial sort and
  # preserves the incoming row order.
  numeric_cols <- which(sapply(by_part, function(x) is.numeric(x) ||
                                  grepl("^[0-9.%$,-]+$", as.character(x[1]))))

  DT::datatable(
    by_part,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold;",
      class = "dt-caption",
      "All strategies × Training and Testing partitions, ordered Testing / Full Period / Training then by Sharpe (descending). Years column shows date ranges. Validation partition is sealed — not shown until production deployment."
    ),
    rownames = FALSE,
    filter = "top",
    options = list(
      pageLength = 15,
      scrollX = TRUE,
      autoWidth = FALSE,
      dom = "frtip",
      order = list(),
      search = list(regex = TRUE, caseInsensitive = TRUE),
      columnDefs = list(
        list(className = 'dt-right', targets = numeric_cols - 1)
      ),
      initComplete = DT::JS(
        "function(settings, json) {",
        "  var isDark = document.documentElement.getAttribute('data-bs-theme') === 'dark'",
        "    || document.body.classList.contains('dark-mode');",
        "  if (isDark) {",
        "    $(this.api().table().container()).css({'color': '#ddd', 'background-color': '#1a1a1a'});",
        "    $(this.api().table().header()).css({'color': '#ddd', 'background-color': '#222'});",
        "    $('input', this.api().table().container()).css({'color': '#ddd', 'background-color': '#333', 'border-color': '#555'});",
        "  }",
        "}"
      )
    )
  )
}
```

#### Strategy Details

Each strategy links to its full definition, implementation, and backtest:

| Strategy | Level | Signal | Definition |
|----------|-------|--------|------------|
| [Factor MAX](stock-backtest.html#factor-level-deep-dive) | Factor (5 FF) | Max daily return in prior month | Ranks factors by single-day extreme |
| [Factor <abbr title="Dynamic Rotation Into Factors — selects best-performing FF factor each month">DRIF</abbr>](stock-backtest.html#factor-level-deep-dive) | Factor (5 FF) | Elastic net on 42 daily return features | Full daily return distribution |
| [Stock MAX](stock-backtest.html#stock-max) | Stock (660) | Max daily return, D1-D10 decile spread | Cross-sectional stock sorting |
| [Stock DRIF](stock-backtest.html#stock-drif) | Stock (660) | Elastic net, D1-D10 decile spread | Pooled cross-sectional elastic net |
| [XGB DRIF](stock-backtest.html#stock-drif) | Stock (660) | XGBoost monotonic on 42 features | Non-linear signal with monotonic constraints |
| [ETF DRIF A](stock-backtest.html#etf-replication) | ETF (9) | Elastic net on ETF daily returns | Direct ETF signal, long-only 13.4% CAGR |
| [ETF DRIF B](stock-backtest.html#etf-replication) | ETF (9) | Academic factor signal → ETF execution | Mapped via ETF_FACTOR_MAP, long-only 12.6% |
| [PSO Optimal](stock-backtest.html#comparison) | Combined | Weighted portfolio of above | [Grid search](stock-backtest.html#comparison) on training Sharpe |

```{r}
# Show PSO weights
w <- safe_tar_read("port_optimal_weights")
if (!is.null(w)) {
  w_df <- tibble(
    Strategy = names(w),
    Weight = paste0(round(w * 100), "%")
  ) |> filter(Weight != "0%")
  hd_dt(w_df, "PSO optimal weights (trained on Training partition Sharpe). Strategies with 0% weight omitted.")
}
```

# Falsification

## Row

### {.tabset}

#### Scorecard

The falsification framework tests 5 strategies against multiple statistical hurdles: <abbr title="Newey-West heteroskedasticity-autocorrelation consistent estimator">HAC</abbr>-corrected t-statistics, 6 null environment rejection rates, and <abbr title="Fama-French 5-Factor model (MktRF, SMB, HML, RMW, CMA)">FF5</abbr> + Momentum alpha regression. Strategies that pass all tests show genuine alpha; those that fail are explained by known factor exposure (beta). Note: this covers 5 strategies from the falsification pipeline — partially overlapping with the leaderboard strategies above.

```{r}
#| label: fals-scorecard
fals_sc <- safe_tar_read("fals_vig_scorecard")
fals_sc_cap <- safe_tar_read("fals_vig_scorecard_caption")
if (!is.null(fals_sc)) {
  cap_text <- if (!is.null(fals_sc_cap)) fals_sc_cap else "Strategy falsification scorecard"
  hd_dt(fals_sc, cap_text)
}
```

#### Null Rejection

Rejection rates across 6 null environments (M = 100 simulations each). A low rate means the strategy's performance is easily replicated under the null — no genuine signal. A high rate means the null cannot explain the observed returns.

```{r}
#| label: fals-heatmap
#| fig-height: 5
#| fig-width: 10
safe_tar_read("fals_vig_null_heatmap")
```

```{r}
#| label: fals-heatmap-caption
#| results: asis
cap <- safe_tar_read("fals_vig_null_heatmap_caption")
if (!is.null(cap)) cat(paste0("\n*", cap, "*\n"))
```

#### Alpha vs Factor Exposure

Scatter plot of annualised alpha (%) against R² (%) from the Fama-French 5-Factor + Momentum regression. Genuine alpha strategies cluster in the top-left (low factor exposure, positive alpha). Beta strategies cluster on the right (high R²).

```{r}
#| label: fals-alpha-plot
#| fig-height: 5
#| fig-width: 10
safe_tar_read("fals_vig_ff_alpha_plot")
```

```{r}
#| label: fals-alpha-caption
#| results: asis
cap <- safe_tar_read("fals_vig_ff_caption")
if (!is.null(cap)) cat(paste0("\n*", cap, "*\n"))
```

#### See Also

For detailed methodology, HAC analysis, factor regression betas, multiplicity correction (<abbr title="Effective number of independent strategies (spectral participation ratio)">K_eff</abbr>, delta-Z), and the full null environment heatmap, see the [Strategy Falsification dashboard](falsification.html).

# Performance

## Row

### {.tabset}

#### Equity Curves

```{r}
#| fig-height: 6
safe_tar_read("stk_all_comparison_plot")
```

::: {.fig-caption}
```{r}
#| echo: false
#| results: asis
cap <- safe_tar_read("stk_all_caption")
if (!is.null(cap)) cat(cap) else cat("*Caption not available — run tar_make() to generate.*")
```
:::

#### Portfolio

```{r}
#| fig-height: 6
safe_tar_read("port_comparison_plot")
```

::: {.fig-caption}
**PSO optimal vs equal-weight portfolio.** Combines all 4 strategies with weights optimised on training partition Sharpe. The PSO optimal tilts toward strategies with better risk-adjusted returns (Factor DRIF + Stock MAX).
:::

#### Monthly Returns

```{r}
monthly <- safe_tar_read("port_monthly_returns")
if (!is.null(monthly)) {
  # Most recent year first, columns Jan-Dec + Annual
  col_order <- c("Year", "Jan", "Feb", "Mar", "Apr", "May", "Jun",
                 "Jul", "Aug", "Sep", "Oct", "Nov", "Dec", "Annual")
  monthly <- monthly |>
    select(any_of(col_order)) |>
    arrange(desc(Year))

  # Add marginal row (mean across years)
  num_cols <- setdiff(col_order, "Year")
  monthly <- monthly |> mutate(Year = as.character(Year))
  marginal <- monthly |>
    summarise(across(all_of(num_cols), ~ format(round(mean(as.numeric(.), na.rm = TRUE), 1), nsmall = 1))) |>
    mutate(Year = "Mean")
  monthly <- bind_rows(monthly, marginal)

  DT::datatable(
    monthly,
    rownames = FALSE, filter = "none",
    options = list(
      pageLength = 60, scrollX = TRUE, dom = "t",
      order = list(list(0, "desc")),
      columnDefs = list(list(className = "dt-right", targets = seq_len(length(col_order) - 1)))
    ),
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left;",
      "Monthly returns (PSO optimal portfolio) by year, most recent first. Annual = compounded. Mean row = average across years. Cell borders: green = positive, red = negative."
    )
  ) |>
    DT::formatStyle(
      columns = num_cols,
      color = DT::styleInterval(0, c("#e74c3c", "#69d4a0")),
      border = DT::styleInterval(0, c("2px solid #e74c3c", "2px solid #69d4a0"))
    )
}
```

#### XGBoost vs Elastic Net

```{r}
#| fig-height: 6
safe_tar_read("xgb_vs_enet_plot")
```

::: {.fig-caption}
**XGBoost (monotonic constraints) vs Elastic Net** on same 42 DRIF features, stock-level D1-D10. XGBoost underperforms elastic net (-6.3% vs +1.3% full-period CAGR). Monotonic constraints may be too restrictive for cross-sectional returns. See [Stock-Level Backtests](stock-backtest.html#stock-drif) for full analysis.
:::

# Robustness

## Row

### {.tabset}

#### Bootstrap CI

```{r}
#| fig-height: 6
safe_tar_read("boot_sharpe_plot")
```

```{r}
boot_ci <- safe_tar_read("boot_ci_summary")
if (!is.null(boot_ci)) {
  display <- boot_ci |>
    transmute(
      Strategy = strategy_label,
      `Sharpe [5%, 95%]` = paste0(round(sharpe_mean, 2), " [", round(sharpe_lo, 2), ", ", round(sharpe_hi, 2), "]"),
      `CAGR [5%, 95%]` = paste0(round(cagr_mean * 100, 1), "% [", round(cagr_lo * 100, 1), "%, ", round(cagr_hi * 100, 1), "%]"),
      `Max DD [5%, 95%]` = paste0(round(dd_mean * 100, 1), "% [", round(dd_lo * 100, 1), "%, ", round(dd_hi * 100, 1), "%]"),
      `CI Crosses Zero` = ifelse(ci_crosses_zero, "YES — may be noise", "No")
    )
  hd_dt(display, "Block bootstrap (1000 draws, block=3m). If Sharpe CI crosses zero, strategy may be indistinguishable from noise.")
}
```

#### Deflated Sharpe

The Deflated Sharpe Ratio (<abbr title="Deflated Sharpe Ratio: Sharpe adjusted for non-normality (skewness, kurtosis) and multiple testing across K trials — Lopez de Prado 2018">DSR</abbr>) adjusts the naive Sharpe for both non-normality of returns (skewness and kurtosis) and multiple testing, answering the question: given that we tested K strategies, is this Sharpe ratio statistically significant? The multiple-testing penalty uses the Vertox correlation-aware effective count K_eff_strat (not the raw count M = 4), because the 4 strategies are correlated — using the raw count would over-penalise and reject strategies that carry genuine independent signal. See [Lopez de Prado (2018)](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=2460551) for the DSR derivation and [Vertox — The Effective Number of Tested Strategies](https://www.vertoxquant.com/p/the-effective-number-of-tested-strategies) for the K_eff_strat methodology.

```{r}
#| label: deflated-sharpe
x <- safe_tar_read("strat_deflated_sharpe")
if (!is.null(x)) {
  keff <- round(x$k_eff_strat[1], 2)
  kraw <- x$k_raw[1]
  cap <- paste0(
    "Deflated Sharpe (Lopez de Prado 2018). ",
    "Multiple-testing penalty uses Vertox K_eff_strat = ", keff,
    " effective strategies vs ", kraw,
    " raw — the 4 strategies are correlated, so the raw count over-penalises. ",
    "DSR adjusts for skewness, kurtosis, and the effective number of trials."
  )
  display <- x |>
    transmute(
      Strategy        = strategy,
      `Naive Sharpe`  = round(naive_sharpe, 2),
      `Deflated Sharpe` = round(deflated_sharpe, 2),
      `Haircut %`     = paste0(round(dsr_haircut_pct, 1), "%"),
      `DSR p-value`   = round(dsr_pvalue, 3)
    )
  hd_dt(display, cap)
}
```

#### Alpha Decay

```{r}
#| fig-height: 6
# #356: Alpha decay needs a positive base Sharpe (decay is the fraction
# of base Sharpe that survives delay; if base is negative, the metric
# is meaningless). Surface the actual reason in the callout.
dp <- safe_tar_read("decay_plot")
decay_metrics <- safe_tar_read("decay_metrics")
neg_base <- !is.null(decay_metrics) && nrow(decay_metrics) > 0 &&
            all(decay_metrics$sharpe[decay_metrics$delay == min(decay_metrics$delay)] <= 0, na.rm = TRUE)
if (!is.null(dp) && !is.null(decay_metrics) && nrow(decay_metrics) > 0 && !neg_base) {
  print(dp)
} else if (neg_base) {
  cat("::: callout-warning\n",
      "**Alpha decay not meaningful — base Sharpe is non-positive for every strategy on the page.** ",
      "Decay measures the *fraction of base Sharpe that survives execution delay*; if base Sharpe ≤ 0 there is no alpha to decay. ",
      "Investigate the strategies' base Sharpe before re-running this tab. See <a href=\"https://github.com/JohnGavin/historical/issues/356\" target=\"_blank\" rel=\"noopener noreferrer\">#356</a>.\n",
      ":::\n", sep = "")
} else {
  cat("::: callout-note\n",
      "**Alpha decay analysis not yet computed.** ",
      "Run `tar_make(c(\"decay_metrics\", \"decay_plot\", \"decay_half_life\"))` ",
      "to populate this tab. See <a href=\"https://github.com/JohnGavin/historical/issues/356\" target=\"_blank\" rel=\"noopener noreferrer\">#356</a>.\n",
      ":::\n", sep = "")
}
```

```{r}
decay_hl <- safe_tar_read("decay_half_life")
if (!is.null(decay_hl) && nrow(decay_hl) > 0) {  # #356: don't render an empty DT
  display <- decay_hl |>
    transmute(
      Strategy = dplyr::recode(strategy, stk_max = "Stock MAX", stk_drif = "Stock DRIF"),
      `Base Sharpe (t+1)` = round(base_sharpe, 2),
      `Half-Life (days)` = ifelse(is.na(half_life_days), ">10 (robust)", paste0(half_life_days))
    )
  hd_dt(display, "Alpha decay: days until Sharpe drops to 50% of t+1 baseline. >10 = signal persists; <3 = fragile, execution-dependent.")
}
```

#### Regime

```{r}
#| fig-height: 6
safe_tar_read("regime_plot")
```

```{r}
regime_m <- safe_tar_read("regime_metrics")
if (!is.null(regime_m)) {
  display <- regime_m |>
    filter(period == "Full Period") |>
    transmute(
      Strategy = strategy,
      CAGR = paste0(round(cagr * 100, 1), "%"),
      Vol = paste0(round(vol * 100, 1), "%"),
      Sharpe = round(sharpe, 2),
      `Max DD` = paste0(round(max_dd * 100, 1), "%")
    )
  hd_dt(display, "Regime-adjusted vs base portfolio (full period). Regime scaling: low=1.0, med=0.7, high=0.4. High-risk months shaded red in plot.")
}
```

#### Kelly Sizing

```{r}
#| fig-height: 6
safe_tar_read("kelly_comparison_plot")
```

```{r}
kelly_t <- safe_tar_read("kelly_sizing_table")
if (!is.null(kelly_t)) {
  display <- kelly_t |>
    transmute(
      Strategy = strategy,
      `Mean Monthly Edge` = paste0(round(edge * 100, 2), "%"),
      `Monthly Variance` = round(variance, 4),
      `Full Kelly` = paste0(round(kelly_full * 100, 1), "%"),
      `Quarter Kelly` = paste0(round(kelly_frac * 100, 2), "%")
    )
  hd_dt(display, "Kelly criterion edge estimates (training data only). Quarter Kelly = 25% of full Kelly, capped at 5%. Negative edge = do not trade.")
}
```

# Structural Breaks

## Row {height=100%}

### {.tabset}

#### Interpretation

A **structural break** at date *T* means the returns process appears to have changed before and after that date — the distribution of vol-normalised returns differs significantly on each side. This is **evidence worth investigating**, not an automatic allocation trigger.

The `resulting-prohibition` rule applies directly here: a detected break does not justify revising strategy weights without identifying new *structural* evidence for why the change occurred. Past underperformance alone is not such evidence — Swedroe documents 12–39-year underperformance episodes for well-established risk premia (`underperformance-prior`). Conflating a drawdown with a structural change risks abandoning a strategy at exactly the wrong time.

Rob Carver's empirical finding (2026) is the calibrating prior: across instrument/forecast pairs, only ~13% show breaks at the 1% level, and whole-history estimation often beats break-adjusted estimation out-of-sample. **Break-splitting is a guard against over-splitting, not an always-on re-estimator.**

For the full framework context see the [market-behavior gap analysis](https://github.com/JohnGavin/historical/blob/main/knowledge/wiki/market-behavior-gap-analysis.md) (§ Structural Stability), which this section closes (gap now ✅).

#### Summary Table

```{r}
#| label: sb-table
sb_sum <- safe_tar_read("structural_breaks_summary")
sb_cap <- safe_tar_read("structural_breaks_caption")
snames  <- safe_tar_read("strategy_names")

if (!is.null(sb_sum) && !is.null(snames)) {
  # strategy_names join: structural_breaks_summary uses short_name labels
  sb_disp <- sb_sum |>
    left_join(
      snames |> select(short_name, long_name),
      by = c("strategy" = "short_name")
    ) |>
    mutate(Strategy = dplyr::coalesce(long_name, strategy)) |>
    transmute(
      Strategy,
      `Break Date(s)`     = break_dates,
      `N Breaks`          = ifelse(is.na(n_breaks), "—", as.character(n_breaks)),
      `Whole Sharpe`      = ifelse(
        is.na(whole_sharpe), "—",
        format(round(whole_sharpe,      3L), nsmall = 3L)
      ),
      `Post-Break Sharpe` = ifelse(
        is.na(post_break_sharpe), "—",
        format(round(post_break_sharpe, 3L), nsmall = 3L)
      ),
      `Divergence %`      = ifelse(
        is.na(sharpe_divergence_pct), "—",
        paste0(sharpe_divergence_pct, "%")
      ),
      Material            = ifelse(
        is.na(material_divergence), "—",
        ifelse(material_divergence, "YES", "no")
      ),
      Note                = ifelse(is.na(note), "", note)
    )

  cap_text <- if (!is.null(sb_cap)) sb_cap else
    "Structural break analysis — run tar_make() to populate."

  DT::datatable(
    sb_disp,
    caption  = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold;",
      class = "dt-caption",
      cap_text
    ),
    rownames = FALSE,
    filter   = "top",
    options  = list(
      pageLength   = 20L,
      scrollX      = TRUE,
      dom          = "frtip",
      initComplete = DT::JS(
        "function(settings, json) {",
        "  var isDark = document.documentElement.getAttribute('data-bs-theme') === 'dark'",
        "    || document.body.classList.contains('dark-mode');",
        "  if (isDark) {",
        "    $(this.api().table().container()).css({'color': '#ddd', 'background-color': '#1a1a1a'});",
        "    $(this.api().table().header()).css({'color': '#ddd', 'background-color': '#222'});",
        "    $('input', this.api().table().container()).css({'color': '#ddd', 'background-color': '#333', 'border-color': '#555'});",
        "  }",
        "}"
      )
    )
  ) |>
    DT::formatStyle(
      "Material",
      target          = "row",
      backgroundColor = DT::styleEqual("YES", "rgba(255, 99, 71, 0.18)")
    )
}
```

#### Sharpe Comparison

```{r}
#| label: sb-dotplot
#| fig-height: 7
#| fig-width: 9
# Path-checked pkgload (cwd-trap fix: here::here() returns docs/ under quarto render)
pkg_path <- if (dir.exists("packages/historicaldata")) "packages/historicaldata" else
  file.path(dirname(here::here()), "packages/historicaldata")
pkgload::load_all(pkg_path, quiet = TRUE)

sb_sum <- safe_tar_read("structural_breaks_summary")
snames  <- safe_tar_read("strategy_names")

if (!is.null(sb_sum) && !is.null(snames)) {
  plot_df <- sb_sum |>
    filter(!is.na(whole_sharpe)) |>
    left_join(
      snames |> select(short_name, long_name),
      by = c("strategy" = "short_name")
    ) |>
    mutate(
      label     = dplyr::coalesce(long_name, strategy),
      has_break = factor(
        ifelse(!is.na(n_breaks) & n_breaks > 0L, "Break at 1%", "No break"),
        levels = c("No break", "Break at 1%")
      )
    )

  strat_order <- plot_df |>
    arrange(whole_sharpe) |>
    pull(label)

  long_df <- plot_df |>
    select(label, has_break,
           `Whole History` = whole_sharpe,
           `Post-Break`    = post_break_sharpe) |>
    tidyr::pivot_longer(
      c(`Whole History`, `Post-Break`),
      names_to  = "Period",
      values_to = "Sharpe"
    ) |>
    filter(!is.na(Sharpe)) |>
    mutate(
      label  = factor(label,  levels = strat_order),
      Period = factor(Period, levels = c("Whole History", "Post-Break"))
    )

  ggplot(long_df, aes(x = Sharpe, y = label, colour = Period, shape = has_break)) +
    geom_vline(xintercept = 0, linetype = "dotted",
               colour = "grey50", linewidth = 0.4) +
    geom_point(size = 3.5, alpha = 0.9) +
    scale_colour_manual(
      values = hd_palette(2),
      name   = "Estimate period"
    ) +
    scale_shape_manual(
      values = c("No break" = 16L, "Break at 1%" = 17L),
      name   = "Break status"
    ) +
    labs(
      x     = "Annualised Sharpe Ratio",
      y     = NULL,
      title = "Whole-history vs post-break Sharpe"
    ) +
    hd_theme()
}
```

```{r}
#| label: sb-dotplot-caption
#| results: asis
sb_cap <- safe_tar_read("structural_breaks_caption")
fig_cap <- paste0(
  "**Figure — Cleveland dot plot:** whole-history Sharpe (circle) vs ",
  "post-break Sharpe (where a break was detected) for all leaderboard strategies. ",
  "**Data:** return series pre-computed in the docs targets pipeline; see ",
  "[plan_structural_breaks.R](https://github.com/JohnGavin/historical/blob/main/R/plan_structural_breaks.R). ",
  "**Variables:** annualised Sharpe ratio; Whole History = full return series, ",
  "Post-Break = segment after the first detected break (triangle shape). ",
  "**Labels:** `long_name` from the `strategy_names` target ",
  "([plan_strategy_names.R](https://github.com/JohnGavin/historical/blob/main/R/plan_strategy_names.R)), ",
  "ordered by whole-history Sharpe ascending. ",
  "**Methodology:** iterative forward-split t-test at 1% level with 5-year minimum segment; see ",
  "[hd_structural_breaks()](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/structural_breaks.R). ",
  "**Caveats:** scanning all candidate split dates inflates false-break rate above the nominal 1% level; ",
  "a detected break is evidence to investigate, not an allocation trigger (`resulting-prohibition`); ",
  "whole-history estimation often beats break-adjusted OOS (Carver 2026). ",
  "**Chart type:** Cleveland dot plot per visualization-standards (dotchart first choice). ",
  if (!is.null(sb_cap)) paste0("Dynamic summary: ", sb_cap) else ""
)
cat(paste0("\n*", fig_cap, "*\n"))
```

# Details

## Row

### {.tabset}

#### Correlations

```{r}
corr <- safe_tar_read("strategy_correlation")
if (!is.null(corr)) {
  corr_display <- corr |>
    mutate(across(-strategy, ~ round(as.numeric(.), 2)))
  knitr::kable(corr_display, caption = "Pairwise correlation (Pearson, monthly returns).")
}
```

#### Assumptions

Per `backtesting-assumptions` rule, all strategies use:

| Assumption | Value |
|------------|-------|
| Execution | Market close (4 PM ET) |
| Transaction cost | 0.50% per trade (1.00% round-trip) |
| Estimated turnover | 80% per month (decile sort) |
| Short borrow cost | 3% per annum (0.25%/month) |
| Return winsorisation | ±20% per month |
| Cash return | 3-month US Treasury rate (RF from FF3) |
| Dividends | Reinvested (total return basis) |
| Taxes | Excluded (pre-tax) |

**Total monthly cost for L/S:** turnover × cost × 2 legs × 2 (buy+sell) + borrow = 0.80 × 0.005 × 4 + 0.0025 = **~1.85%/month**. This is why most stock-level L/S strategies are unprofitable after costs.

#### Partitions

```{r}
params <- safe_tar_read("stk_params")
if (!is.null(params)) {
  param_df <- tibble(
    Partition = c("Training", "Testing", "Validation"),
    Period = c(
      paste(params$start_date, "to", params$is_end),
      paste(params$test_start, "to", params$test_end),
      paste(params$val_start, "to", "present")
    ),
    Purpose = c("Model fitting, expanding window signal estimation",
                "Calibration, hyperparameter tuning, strategy comparison",
                "SEALED — not computed or shown until production deployment")
  )
  hd_dt(param_df, "Shared partitions across all strategies. Validation is sealed: metrics are NOT computed until a one-shot production decision.")
}
```

#### Related Vignettes

- **[Strategy Falsification](falsification.html)** — provides the statistical survival verdict (HAC t-stat, null rejection rate, FF5 alpha) underlying the leaderboard scorecard
- **[Stock Backtest](stock-backtest.html)** — per-asset equity-level view of Factor MAX and DRIF strategies ranked here
- **[Factor MAX](stock-backtest.html#factor-level-deep-dive)** — detailed breakdown of the top-ranked factor rotation strategy

#### Build Info

```{r}
#| label: build-info
#| echo: false
#| results: asis
build_info()
```
