---
title: "Macro Defense Rotation"
format:
  dashboard:
    orientation: rows
    theme: cosmo
    embed-resources: true
    include-in-header:
      - text: |
          <style>
          body { font-size: 1.15em; }
          /* Black navbar/header strip */
          .navbar, .quarto-dashboard-header, .navbar-dark,
          .bslib-page-dashboard > .navbar {
            background-color: black !important;
            color: white !important;
          }
          .navbar .navbar-brand, .navbar .nav-link, .navbar-text,
          .navbar .navbar-nav .nav-link {
            color: white !important;
          }
          .navbar .nav-link.active, .navbar .nav-link:hover {
            color: white !important;
            background-color: black !important;
            border: 1px solid white !important;
            border-radius: 4px;
          }
          .navbar .nav-link { color: #aaa !important; }
          /* Allow navbar tabs to wrap to second line */
          .navbar-nav { flex-wrap: wrap !important; }
          .navbar .nav-link { font-size: 0.9em; padding: 0.3em 0.6em !important; }
          .bslib-page-fill, .html-fill-container, .html-fill-item,
          .bslib-card, .card, .tab-pane, .tab-content, .card-body,
          .bslib-grid, .bslib-grid-item {
            height: auto !important; min-height: 0 !important;
            max-height: none !important; overflow: visible !important;
            flex: none !important;
          }
          .cell-output-display { display: block !important; position: relative !important;
            overflow: visible !important; margin-bottom: 1em; }
          .cell-output-display img { display: block; width: 100%; height: auto; min-height: 400px; }
          .fig-caption { display: block; position: relative; clear: both;
            margin-top: 0.5em; margin-bottom: 1em; font-size: 0.85em; color: #888; }
          .datatables, .dataTables_wrapper { display: block !important;
            position: relative !important; overflow-x: auto !important; margin-bottom: 1em; }
          table.dataTable { white-space: nowrap; }
          table caption { caption-side: top !important; }
          details { margin-bottom: 0.5em; }
          details summary { cursor: pointer; color: #6c9bd2; font-size: 0.9em; }
          .fullscreen-overlay {
            position: fixed !important; top: 0 !important; left: 0 !important;
            width: 100vw !important; height: 100vh !important;
            z-index: 9999 !important; background: black !important;
            padding: 10px !important; cursor: zoom-out !important;
          }
          .fullscreen-overlay img { width: 100% !important; height: 100% !important;
            object-fit: contain !important; }
          .cell-output-display:has(img) { cursor: zoom-in; }
          </style>
          <script>
          document.addEventListener('click', function(e) {
            var overlay = e.target.closest('.fullscreen-overlay');
            if (overlay) { overlay.classList.remove('fullscreen-overlay'); overlay.style.cssText = ''; return; }
            var img = e.target.closest('img');
            if (!img) return;
            var div = document.createElement('div');
            div.className = 'fullscreen-overlay';
            div.innerHTML = '<img src="' + img.src + '">';
            div.addEventListener('click', function() { div.remove(); });
            document.body.appendChild(div);
          });
          </script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)
library(dplyr)

pkgload::load_all(here::here("packages/historicaldata"), quiet = TRUE)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

hd_dt <- function(df, caption_text) {
  if (is.null(df)) return(invisible(NULL))
  # Format numbers
  for (col in names(df)) {
    if (is.numeric(df[[col]]) && col %in% c("cagr", "vol", "max_dd", "hit_rate"))
      df[[col]] <- paste0(round(df[[col]] * 100, 1), "%")
    if (is.numeric(df[[col]]) && col %in% c("sharpe", "sortino", "calmar"))
      df[[col]] <- round(df[[col]], 2)
  }
  DT::datatable(df, caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold; color: #ddd;",
    caption_text), rownames = FALSE, options = list(pageLength = 15, scrollX = TRUE, dom = "frtip"))
}
```

::: {.callout-warning}
**EXPLORATORY BACKTEST — IN-SAMPLE.** Tabs 2-6 show in-sample results (2008-2022). Tab 7 shows out-of-sample (2023+). This is not investment advice — it is an educational exercise in quantitative backtesting methodology using the `historicaldata` R package.
:::

# Universe

## Row

### {.tabset}

#### Price History

Normalised price history of the 4 defensive ETFs + SPY benchmark since 2007, rebased to 100.

```{r}
#| fig-height: 8
prices <- tar_read(bt_prices) |>
  filter(ticker != "BIL") |>  # exclude cash proxy from price chart
  group_by(ticker) |>
  mutate(normalised = adjusted / first(adjusted) * 100) |>
  ungroup()

n_tickers <- length(unique(prices$ticker))
ggplot(prices, aes(as.Date(date), normalised, colour = ticker)) +
  geom_line(linewidth = 0.5) +
  scale_y_log10() +
  scale_colour_manual(values = hd_palette(n_tickers)) +
  labs(x = NULL, y = "Normalised price (log, base=100)", colour = NULL,
       title = "Defense First universe: TLT/GLD/DBC/UUP + SPY (2007+)") +
  hd_theme()
```

::: {.fig-caption}
**TLT (Treasuries), GLD (Gold), DBC (Commodities), UUP (USD Index), SPY (S&P 500).** Y: normalised price (log scale, rebased to 100). Each covers a different macro risk regime. Source: `hd_ohlcv()` via `hd_group("Defense First")`.
:::

#### Correlation

```{r}
#| fig-height: 6
rets <- tar_read(bt_returns) |>
  filter(as.Date(date) <= as.Date("2022-12-31")) |>
  select(ticker, date, ret) |>
  tidyr::pivot_wider(names_from = ticker, values_from = ret) |>
  filter(if_all(-date, ~ !is.na(.)))

cor_mat <- cor(rets |> select(-date))
cor_long <- cor_mat |> as.data.frame() |>
  mutate(row = rownames(cor_mat)) |>
  tidyr::pivot_longer(-row, names_to = "col", values_to = "cor")

ggplot(cor_long, aes(row, col, fill = cor)) +
  geom_tile(colour = "grey30") +
  geom_text(aes(label = round(cor, 2)), colour = "white", size = 4) +
  scale_fill_gradient2(low = "#00BFFF", mid = "grey30", high = "#FF6347",
                       midpoint = 0, limits = c(-1, 1)) +
  labs(x = NULL, y = NULL, fill = "Corr",
       title = "Monthly return correlation (in-sample, 2007-2022)") +
  hd_theme() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

::: {.fig-caption}
**Monthly return correlations for the rotation universe (in-sample).** Low/negative correlations between hedges confirm diversification benefit. GLD-UUP strongly negative (-0.41). SPY-TLT negative (-0.31). Source: `hd_ohlcv()`.
:::

#### Metadata

```{r}
pkgload::load_all(here::here("packages/historicaldata"), quiet = TRUE)
meta <- hd_ticker_meta(c("TLT", "GLD", "DBC", "UUP", "SPY"))
hd_dt(meta, "Defense First universe: 4 hedges + SPY benchmark.")
```

#### Regime Map

Which hedge ranked #1 each month based on the expanding-window composite momentum signal.

```{r}
#| fig-height: 2
regime <- tar_read(bt_regime_map) |>
  mutate(date = as.Date(date),
         rank_1 = factor(rank_1, levels = c("TLT", "GLD", "DBC", "UUP")))

ggplot(regime, aes(date, 1, fill = rank_1)) +
  geom_tile(height = 1) +
  scale_fill_manual(values = c(TLT = "#6B8DAE", GLD = "#C4A35A",
                               DBC = "#A67B5B", UUP = "#7A9E7E"),
                    na.value = "grey50") +
  labs(x = NULL, y = NULL, fill = "Top hedge",
       title = "Regime map: which hedge ranked #1 each month (expanding window)") +
  hd_theme() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
```

::: {.fig-caption}
**Monthly regime map.** Colour = top-ranked hedge by composite momentum (avg of 1/3/6/12-month returns). TLT (steel blue) = deflation/easing. GLD (muted gold) = monetary instability. DBC (brown) = inflation/commodities. UUP (sage) = USD strength. Expanding window: each month uses all prior data, no lookahead.
:::


# In-Sample (2008-2022)

## Row

### {.tabset}

#### Equity Curve

**EXPLORATORY / IN-SAMPLE.** Defense First rotation vs SPY buy-and-hold.

```{r}
#| fig-height: 8
port <- tar_read(bt_portfolio) |> mutate(date = as.Date(date))

ggplot(port, aes(date)) +
  geom_line(aes(y = cum_port, colour = "Defense First"), linewidth = 0.6) +
  geom_line(aes(y = cum_bench, colour = "SPY"), linewidth = 0.6) +
  scale_y_log10() +
  scale_colour_manual(values = c("Defense First" = "#00BFFF", "SPY" = "#FF6347")) +
  labs(x = NULL, y = "Cumulative return (log)", colour = NULL,
       title = "In-sample equity curve: Defense First vs SPY (2008-2022)") +
  hd_theme()
```

::: {.fig-caption}
**In-sample equity curves (log scale).** Defense First (blue) vs SPY (red). The strategy sacrifices upside for dramatically lower drawdowns. SPY CAGR higher but max DD -46% vs Defense First -14%. EXPLORATORY — in-sample only.
:::

#### Drawdowns

```{r}
#| fig-height: 8
dd <- tar_read(bt_drawdowns) |> mutate(date = as.Date(date))

ggplot(dd, aes(date)) +
  geom_area(aes(y = dd_bench), fill = "#FF6347", alpha = 0.3) +
  geom_line(aes(y = dd_bench, colour = "SPY"), linewidth = 0.4) +
  geom_area(aes(y = dd_port), fill = "#00BFFF", alpha = 0.3) +
  geom_line(aes(y = dd_port, colour = "Defense First"), linewidth = 0.4) +
  scale_y_continuous(labels = scales::percent) +
  scale_colour_manual(values = c("Defense First" = "#00BFFF", "SPY" = "#FF6347")) +
  labs(x = NULL, y = "Drawdown", colour = NULL,
       title = "Drawdown comparison: Defense First vs SPY (in-sample)") +
  hd_theme()
```

::: {.fig-caption}
**Drawdown chart.** SPY (red) suffered -46% in 2008-09 and -34% in 2020. Defense First (blue) max DD -14%. The strategy's primary value is loss mitigation. In-sample only.
:::

#### Metrics

```{r}
hd_dt(tar_read(bt_metrics_is),
  "In-sample performance metrics (2008-2022). EXPLORATORY — no out-of-sample peeking.")
```


# Prediction vs Actual

## Row

### {.tabset}

#### Scatter

Does the composite momentum signal predict next month's return?

```{r}
#| fig-height: 7
exp <- tar_read(bt_expanding) |> mutate(date = as.Date(date))

ggplot(exp, aes(predicted_return, actual_return)) +
  geom_point(alpha = 0.4, colour = "#00BFFF") +
  geom_smooth(method = "lm", colour = "#FF6347", se = TRUE) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  labs(x = "Predicted return (momentum signal)", y = "Actual return (next month)",
       title = "Prediction vs actual: monthly returns (in-sample)") +
  hd_theme()
```

::: {.fig-caption}
**Predicted vs actual monthly returns.** X: composite momentum signal (weighted trailing returns). Y: realised portfolio return next month. Positive slope = signal has predictive power. Wide scatter = noisy prediction (expected for monthly returns). In-sample.
:::

#### Hit Rate Over Time

```{r}
#| fig-height: 7
exp_roll <- exp |>
  mutate(hit = as.numeric(beat_benchmark),
         cs = cumsum(hit),
         rolling_hit = ifelse(row_number() >= 12,
           (cs - lag(cs, 12)) / 12, NA_real_))

ggplot(exp_roll, aes(date)) +
  geom_line(aes(y = rolling_hit), colour = "#00BFFF", linewidth = 0.5) +
  geom_hline(yintercept = 0.5, linetype = "dashed", colour = "grey50") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  labs(x = NULL, y = "12-month rolling hit rate",
       title = "Rolling hit rate: fraction of months beating SPY") +
  hd_theme()
```

::: {.fig-caption}
**12-month rolling hit rate** (fraction of months portfolio beats SPY). Above 50% dashed line = strategy outperforming on a rolling basis. The strategy is designed for defense (reducing losses) not for beating SPY in all regimes. In-sample.
:::


# Strategy Variants

## Row

### {.tabset}

#### Comparison Table

```{r}
hd_dt(tar_read(bt_comparison),
  "In-sample strategy variant comparison (2008-2022). All EXPLORATORY.")
```

#### Equity Curves

```{r}
#| fig-height: 8
exp <- tar_read(bt_expanding) |> mutate(date = as.Date(date))
ewt <- tar_read(bt_equalwt) |> mutate(date = as.Date(date))
top <- tar_read(bt_topone) |> mutate(date = as.Date(date))

variants <- bind_rows(
  exp |> transmute(date, ret = actual_return, strategy = "Momentum (40/30/20/10)"),
  exp |> transmute(date, ret = benchmark_return, strategy = "SPY buy & hold"),
  ewt |> transmute(date, ret = port_ret, strategy = "Equal weight (25%)"),
  top |> transmute(date, ret = port_ret, strategy = "Top 1 only (100%)")
) |>
  group_by(strategy) |>
  mutate(cum = cumprod(1 + ret)) |>
  ungroup()

ggplot(variants, aes(date, cum, colour = strategy)) +
  geom_line(linewidth = 0.5) +
  scale_y_log10() +
  scale_colour_manual(values = hd_palette(4)) +
  labs(x = NULL, y = "Cumulative return (log)", colour = NULL,
       title = "Strategy variants: in-sample equity curves (2008-2022)") +
  hd_theme()
```

::: {.fig-caption}
**4 strategy variants compared.** Momentum rotation (40/30/20/10), equal weight (25% each), top-1 concentrated (100%), and SPY benchmark. All in-sample. Momentum rotation balances return and risk.
:::


# Allocation

## Row

### {.tabset}

#### Weight History

```{r}
#| fig-height: 8
exp <- tar_read(bt_expanding) |>
  mutate(date = as.Date(date)) |>
  tidyr::pivot_longer(cols = starts_with("w_"), names_to = "rank", values_to = "weight") |>
  mutate(ticker = case_when(
    rank == "w_1" ~ rank_1, rank == "w_2" ~ rank_2,
    rank == "w_3" ~ rank_3, rank == "w_4" ~ rank_4
  ))

ggplot(exp, aes(date, weight, fill = ticker)) +
  geom_area(position = "stack") +
  scale_fill_manual(values = c(TLT = "#6B8DAE", GLD = "#C4A35A",
                               DBC = "#A67B5B", UUP = "#7A9E7E"),
                    na.value = "grey50") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL, y = "Portfolio weight", fill = NULL,
       title = "Monthly allocation history (expanding window)") +
  hd_theme()
```

::: {.fig-caption}
**Monthly weight allocation across TLT/GLD/DBC/UUP (stacked).** Regime shifts visible: 2008 GFC → TLT, 2022 inflation → DBC/UUP. In-sample.
:::

#### Individual Weights

Unstacked view — each ETF's weight shown separately.

```{r}
#| fig-height: 10
ggplot(exp, aes(date, weight, colour = ticker)) +
  geom_line(linewidth = 0.5) +
  facet_wrap(~ticker, ncol = 1) +
  scale_colour_manual(values = c(TLT = "#6B8DAE", GLD = "#C4A35A",
                                 DBC = "#A67B5B", UUP = "#7A9E7E"),
                      na.value = "grey50") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL, y = "Weight", colour = NULL,
       title = "Individual ETF weights over time (unstacked)") +
  hd_theme() +
  theme(legend.position = "none")
```

::: {.fig-caption}
**Individual ETF weights (unstacked, faceted).** Each panel shows one ETF's allocation over time. Weights cycle between 40% (rank 1) and 10% (rank 4) as momentum shifts.
:::


# Out-of-Sample (2023-2026)

## Row

### {.tabset}

#### IS vs OOS Metrics

**OUT-OF-SAMPLE (2023-2026).** Run ONCE — no parameter re-tuning after seeing this.

```{r}
hd_dt(tar_read(bt_oos_vs_is),
  "In-sample vs out-of-sample comparison. OOS run ONCE with no re-tuning.")
```

#### OOS Equity Curve

```{r}
#| fig-height: 8
oos <- tar_read(bt_oos) |>
  mutate(date = as.Date(date),
         cum_port = cumprod(1 + actual_return),
         cum_bench = cumprod(1 + benchmark_return))

ggplot(oos, aes(date)) +
  geom_line(aes(y = cum_port, colour = "Defense First"), linewidth = 0.6) +
  geom_line(aes(y = cum_bench, colour = "SPY"), linewidth = 0.6) +
  scale_colour_manual(values = c("Defense First" = "#00BFFF", "SPY" = "#FF6347")) +
  labs(x = NULL, y = "Cumulative return", colour = NULL,
       title = "Out-of-sample equity curve (2023+)") +
  hd_theme()
```

::: {.fig-caption}
**Out-of-sample equity curves (2023+).** This period was NOT used for strategy design or tuning. If OOS performance degrades vs in-sample, it may indicate overfitting. Compare Sharpe ratios: IS vs OOS. One-time evaluation — no re-tuning permitted.
:::

#### Full-Period Scatter (IS + OOS)

Prediction vs actual including the out-of-sample period.

```{r}
#| fig-height: 7
# Combine IS expanding + OOS
exp_is <- tar_read(bt_expanding) |> mutate(date = as.Date(date), period = "In-sample")
oos_raw <- tar_read(bt_oos) |> mutate(date = as.Date(date), period = "Out-of-sample")

# OOS doesn't have predicted_return — compute from replication data
rep <- tar_read(bt_replication) |> mutate(date = as.Date(date))
full <- bind_rows(
  exp_is |> select(date, actual_return, benchmark_return, period),
  oos_raw |> select(date, actual_return, benchmark_return, period)
)

ggplot(full, aes(benchmark_return, actual_return, colour = period)) +
  geom_point(alpha = 0.5, size = 2) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.8, alpha = 0.2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
  geom_hline(yintercept = 0, colour = "grey30") +
  geom_vline(xintercept = 0, colour = "grey30") +
  scale_colour_manual(values = c("In-sample" = "#6B8DAE", "Out-of-sample" = "#C4A35A")) +
  scale_fill_manual(values = c("In-sample" = "#6B8DAE", "Out-of-sample" = "#C4A35A")) +
  labs(x = "SPY monthly return", y = "Portfolio monthly return", colour = NULL,
       title = "Portfolio vs SPY returns: in-sample + out-of-sample") +
  hd_theme()
```

::: {.fig-caption}
**Full-period scatter: portfolio monthly return vs SPY monthly return.** Each point = one month. Steel blue = in-sample (2008-2022). Gold = out-of-sample (2023+). Solid lines = OLS regression with 95% confidence bands, fitted independently per period. Dashed diagonal = 1:1 line (points below = strategy underperformed SPY that month). The strategy clusters closer to zero on both axes — lower volatility, smaller gains AND smaller losses. Regression slopes < 1 confirm the strategy has lower beta than SPY. Source: `hd_ohlcv()` via `hd_group("Defense First")`.
:::

#### Full-Period Hit Rate

Rolling 12-month fraction of months beating SPY, including OOS.

```{r}
#| fig-height: 7
full_hit <- full |>
  mutate(hit = as.numeric(actual_return > benchmark_return),
         cs = cumsum(hit),
         rolling_hit = ifelse(row_number() >= 12, (cs - lag(cs, 12)) / 12, NA_real_))

ggplot(full_hit, aes(date, rolling_hit, colour = period)) +
  geom_line(linewidth = 0.5) +
  geom_hline(yintercept = 0.5, linetype = "dashed", colour = "grey50") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  scale_colour_manual(values = c("In-sample" = "#6B8DAE", "Out-of-sample" = "#C4A35A")) +
  labs(x = NULL, y = "12-month rolling hit rate", colour = NULL,
       title = "Rolling hit rate: in-sample + out-of-sample") +
  hd_theme()
```

::: {.fig-caption}
**Full-period rolling hit rate.** Steel blue = in-sample. Gold = out-of-sample. The strategy beats SPY roughly 35-40% of months — it's a defense strategy, not an alpha strategy. Value comes from limiting drawdowns, not from monthly outperformance.
:::


# Replication

## Row

### {.tabset}

#### Article vs Our Results

To replicate the article's exact numbers: (1) fixed lookback windows (not expanding), (2) start Jan 2009 (not 2008), (3) full period including 2023-2026 (no OOS holdout).

```{r}
hd_dt(tar_read(bt_replication_metrics),
  "Article claims vs our replication. Fixed lookback, 2009-2026, no OOS holdout.")
```

::: {.fig-caption}
**Replication comparison.** "Article claims" = numbers from the source. "Our replication" = same methodology (fixed lookback) on our data. Differences may arise from: data source (yfinance vs article's source), exact month-end dates, dividend/split adjustments.
:::

#### Replication Equity Curve

```{r}
#| fig-height: 8
rep <- tar_read(bt_replication) |>
  mutate(date = as.Date(date),
         cum_port = cumprod(1 + actual_return),
         cum_bench = cumprod(1 + benchmark_return))

ggplot(rep, aes(date)) +
  geom_line(aes(y = cum_port, colour = "Defense First (fixed)"), linewidth = 0.6) +
  geom_line(aes(y = cum_bench, colour = "SPY"), linewidth = 0.6) +
  scale_y_log10() +
  scale_colour_manual(values = c("Defense First (fixed)" = "#00BFFF", "SPY" = "#FF6347")) +
  labs(x = NULL, y = "Cumulative return (log)", colour = NULL,
       title = "Article replication: fixed lookback, 2009-2026") +
  hd_theme()
```

::: {.fig-caption}
**Replication of article methodology.** Fixed 1/3/6/12-month lookback (not expanding window). Jan 2009 start. Full period including 2023+. Compare CAGR, Sharpe, max DD with article's claims above.
:::

#### Expanding vs Fixed

```{r}
#| fig-height: 8
exp <- tar_read(bt_expanding) |> mutate(date = as.Date(date))
rep <- tar_read(bt_replication) |> mutate(date = as.Date(date))

# Align to common dates
both <- bind_rows(
  exp |> transmute(date, ret = actual_return, method = "Expanding window (our)"),
  rep |> transmute(date, ret = actual_return, method = "Fixed lookback (article)")
) |>
  group_by(method) |>
  mutate(cum = cumprod(1 + ret)) |>
  ungroup()

ggplot(both, aes(date, cum, colour = method)) +
  geom_line(linewidth = 0.5) +
  scale_y_log10() +
  scale_colour_manual(values = c("Expanding window (our)" = "#00BFFF",
                                  "Fixed lookback (article)" = "#FFD700")) +
  labs(x = NULL, y = "Cumulative return (log)", colour = NULL,
       title = "Expanding window vs fixed lookback comparison") +
  hd_theme()
```

::: {.fig-caption}
**Methodological comparison.** Expanding window (blue, our approach) vs fixed lookback (gold, article's approach). Fixed lookback is more responsive but more prone to whipsaws. Expanding window is more conservative but reduces overfitting risk.
:::


# Details

## Row

### {.tabset}

#### Strategy Summary

**Defense First** is a macro rotation strategy that rotates among 4 defensive ETFs, each covering a different risk regime:

| ETF | Asset Class | Hedges Against | Regime |
|-----|------------|----------------|--------|
| TLT | Long-duration US Treasuries | Deflation, Fed easing | Rates falling |
| GLD | Gold | Monetary instability, falling real rates | Currency debasement |
| DBC | Broad commodities | Stagflation, supply shocks | Inflation rising |
| UUP | US Dollar index | Global stress, funding crises | Flight to USD |

**Benchmark:** SPY (S&P 500 ETF). The strategy aims to protect against drawdowns, not necessarily beat SPY in bull markets.

**Source:** [Defense First: When Not Losing is Winning](https://notfinancial.substack.com/p/defense-first-when-not-losing-is)

#### Implementation

**Signal:** Composite momentum — equal-weighted average of 1, 3, 6, and 12-month simple returns for each ETF.

**Allocation by rank:**

| Rank | Weight |
|------|--------|
| 1st (strongest momentum) | 40% |
| 2nd | 30% |
| 3rd | 20% |
| 4th (weakest) | 10% |

**Fallback:** If all 4 hedge ETFs score below BIL (1-3 month T-bills = cash proxy), allocate 100% to SPY instead.

**Rebalance:** Monthly at month-end close.

**Our modification:** We use an **expanding window** (all available history) instead of the article's fixed lookback. This is more conservative — the signal adapts more slowly but reduces overfitting risk. See the Replication tab for the fixed-lookback comparison.

**Transaction costs:** Not modelled. Article estimates ~0.60% annual drag at 25bps per trade, reducing CAGR from 10.83% to 10.17%.

#### Notes

**Expanding window vs fixed lookback:**

- **Fixed lookback** (article): each month, compute 1/3/6/12-month returns using only those trailing months. More responsive to regime changes. Risk: whipsaws in volatile periods.
- **Expanding window** (our approach): each month, compute momentum using ALL data from start to current month. More conservative. Risk: slower to adapt to new regimes.

**In-sample vs out-of-sample:**

- **In-sample (2008-2022):** Used for strategy development and variant comparison. All results in tabs 2-6 are in-sample. Labelled EXPLORATORY.
- **Out-of-sample (2023-2026):** Held out during development. Run ONCE at the end. No parameter re-tuning permitted after seeing OOS results.

**Caveats:**

- This is educational, not investment advice
- No transaction costs modelled
- Survivorship bias: we only test ETFs that exist today
- Data source differences may cause replication gaps
- Monthly rebalancing assumes month-end execution at close price

#### Glossary

| Term | Definition |
|------|-----------|
| CAGR | Compound Annual Growth Rate — annualised return |
| Sharpe Ratio | (Return - Risk-free rate) / Volatility. Higher = better risk-adjusted return |
| Calmar Ratio | CAGR / Max Drawdown. Higher = better drawdown-adjusted return |
| Sortino Ratio | Like Sharpe but uses downside deviation only |
| Max Drawdown | Largest peak-to-trough decline |
| Hit Rate | Fraction of months portfolio beats benchmark |
| Expanding Window | Uses all data from start to current date (grows each month) |
| Fixed Lookback | Uses only the last N months (window size stays constant) |
| OOS | Out-of-sample — data not used during strategy development |
| Momentum | Tendency for assets that performed well recently to continue |

#### References

1. **Source article:** [Defense First: When Not Losing is Winning](https://notfinancial.substack.com/p/defense-first-when-not-losing-is) — Not Financial Advice (Substack)
2. **Data source:** [historicaldata](https://huggingface.co/datasets/dsfefvx/finance-historical-data) R package via DuckDB `hf://` protocol
3. **TLT:** [iShares 20+ Year Treasury Bond ETF](https://www.ishares.com/us/products/239454/)
4. **GLD:** [SPDR Gold Shares](https://www.spdrgoldshares.com/)
5. **DBC:** [Invesco DB Commodity Index Tracking Fund](https://www.invesco.com/us/financial-products/etfs/product-detail?audienceType=Investor&productId=ETF-DBC)
6. **UUP:** [Invesco DB US Dollar Index Bullish Fund](https://www.invesco.com/us/financial-products/etfs/product-detail?audienceType=Investor&productId=ETF-UUP)
7. **Backtesting methodology:** Expanding window to avoid lookahead bias. See [Advances in Financial Machine Learning](https://www.wiley.com/en-us/Advances+in+Financial+Machine+Learning-p-9781119482086) (de Prado, 2018) for methodology.

#### Build Info

```{r}
#| label: build-info
#| echo: false
#| results: asis
gh_url <- tryCatch({
  remote <- system("git remote get-url origin 2>/dev/null", intern = TRUE)
  sub("\\.git$", "", sub("^git@github\\.com:", "https://github.com/", remote))
}, error = function(e) NULL)

git_sha_short <- tryCatch(
  system("git rev-parse --short HEAD 2>/dev/null", intern = TRUE),
  error = function(e) "N/A"
)
git_sha_full <- tryCatch(
  system("git rev-parse HEAD 2>/dev/null", intern = TRUE),
  error = function(e) git_sha_short
)

pkg_ver <- tryCatch(
  as.character(packageVersion("historicaldata")),
  error = function(e) "0.1.0"
)

r_ver <- as.character(getRversion())
build_time <- format(Sys.time(), "%Y-%m-%d %H:%M:%S")

ver_link <- if (!is.null(gh_url)) {
  sprintf("[%s](%s/releases/tag/v%s)", pkg_ver, gh_url, pkg_ver)
} else pkg_ver

sha_link <- if (!is.null(gh_url) && git_sha_short != "N/A") {
  sprintf("[`%s`](%s/commit/%s)", git_sha_short, gh_url, git_sha_full)
} else sprintf("`%s`", git_sha_short)

r_link <- sprintf("[%s](https://cran.r-project.org/doc/manuals/r-release/NEWS.html)", r_ver)

cat(sprintf(
  "**historicaldata** %s | **Git** %s | **R** %s | **Built** %s",
  ver_link, sha_link, r_link, build_time
))
```
