---
title: "Pre-Peak / Post-Peak 12-2 Momentum Decomposition"
subtitle: "Büsing, Mohrschladt & Siedhoff (2022) — methodology-demo replication on the ltr_universe"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    embed-resources: false
    include-in-header:
      - text: |
          <link rel="stylesheet" href="vignette-shared.css">
          <script src="vignette-shared.js"></script>
          <style>
          body { font-size: 1.3em; }
          .navbar .navbar-title, h1.title { font-size: 1.3em !important; }
          .nav-link { font-size: 0.95em !important; }
          .cell-output-display img { min-height: auto; max-height: 500px; object-fit: contain; }
          .navbar, .quarto-dashboard-header, .navbar-dark,
          .bslib-page-dashboard > .navbar {
            background-color: black !important;
            color: white !important;
          }
          .navbar .navbar-brand, .navbar .nav-link, .navbar-text,
          .navbar .navbar-nav .nav-link {
            color: white !important;
          }
          .navbar .nav-link.active, .navbar .nav-link:hover {
            color: white !important;
            background-color: black !important;
            border: 1px solid white !important;
            border-radius: 4px;
          }
          .navbar .nav-link { color: #aaa !important; }
          .navbar-nav { flex-wrap: wrap !important; }
          .navbar .nav-link { font-size: 0.9em; padding: 0.3em 0.6em !important; }
          .bslib-page-fill, .html-fill-container, .html-fill-item,
          .bslib-card, .card, .tab-pane, .tab-content, .card-body,
          .bslib-grid, .bslib-grid-item {
            height: auto !important; min-height: 0 !important;
            max-height: none !important; overflow: visible !important;
            flex: none !important;
          }
          .cell-output-display { display: block !important; position: relative !important;
            overflow: visible !important; margin-bottom: 1em; }
          .cell-output-display img { display: block; width: 100%; height: auto; min-height: 400px; }
          .fig-caption { display: block; position: relative; clear: both;
            margin-top: 0.5em; margin-bottom: 1em; font-size: 0.85em; color: #e0e0e0; }
          .datatables, .dataTables_wrapper { display: block !important;
            position: relative !important; overflow-x: auto !important; margin-bottom: 1em; }
          table.dataTable { white-space: nowrap; }
          table caption { caption-side: top !important; }
          details { margin-bottom: 0.5em; }
          details summary { cursor: pointer; color: #6c9bd2; font-size: 0.9em; }
          .bankrupt-row { background-color: rgba(240, 128, 128, 0.25) !important; }
          </style>
          <script>
          if (!localStorage.getItem('quarto-color-scheme')) {
            document.documentElement.setAttribute('data-bs-theme', 'dark');
          }
          document.addEventListener('click', function(e) {
            var overlay = e.target.closest('.fullscreen-overlay');
            if (overlay) { overlay.remove(); return; }
            var img = e.target.closest('.cell-output-display img');
            if (!img) return;
            var div = document.createElement('div');
            div.className = 'fullscreen-overlay';
            var clone = document.createElement('img');
            clone.src = img.src;
            div.appendChild(clone);
            document.body.appendChild(div);
          });
          document.addEventListener('shown.bs.tab', function(e) {
            var tabBar = e.target.closest('.nav-tabs');
            if (tabBar) tabBar.scrollIntoView({ behavior: 'smooth', block: 'start' });
          });
          </script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(dplyr)
library(DT)

# Path-checked store + utils (cwd is docs/ under quarto render, project
# root under tar_make). Three-way fallback handles worktree renders where
# here::here() resolves to the worktree docs/ rather than project root.
store_path <- local({
  if (dir.exists("_targets")) return("_targets")
  # Worktree: git --git-common-dir points to main .git; strip /.git for root
  gc <- tryCatch(
    trimws(system("git rev-parse --git-common-dir 2>/dev/null", intern = TRUE)),
    error = function(e) character(0)
  )
  if (length(gc) == 1L && nzchar(gc)) {
    cand <- file.path(dirname(gc), "docs", "_targets")
    if (dir.exists(cand)) return(cand)
  }
  file.path(here::here(), "docs/_targets")  # fallback (works from project root)
})
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)

# ── Strategy colour palette (narrative-colour-persistence rule) ──────────────
# DEFINED ONCE here; referenced by name in every chart below.
PALETTE <- c(
  "mom_prepeak"  = "#69d4a0",  # green  — the only viable leg
  "mom_postpeak" = "#f08080",  # red    — goes bankrupt
  "mom_combined" = "#4ea8de"   # blue   — combined (cancels to ~0)
)

# ── Read targets ─────────────────────────────────────────────────────────────
signal_raw  <- safe_tar_read("mom_prepeak_signal_raw")   # ~30k rows
summary_tbl <- safe_tar_read("mom_prepeak_summary")      # 3 rows × 8 cols
ret_pre     <- safe_tar_read("mom_prepeak_returns")      # 638 rows
ret_post    <- safe_tar_read("mom_postpeak_returns")     # 638 rows
ret_comb    <- safe_tar_read("mom_combined_returns")     # 638 rows
strat_names <- safe_tar_read("strategy_names")           # 14 rows

# ── Strategy display-name lookups (strategy-name-consistency rule) ─────────
mom_names <- if (!is.null(strat_names)) {
  strat_names |> dplyr::filter(.data$code_name %in% c("mom_prepeak", "mom_postpeak", "mom_combined")) |>
    dplyr::select("code_name", "short_name", "long_name")
} else {
  tibble::tibble(
    code_name  = c("mom_prepeak", "mom_postpeak", "mom_combined"),
    short_name = c("Mom Pre-Peak", "Mom Post-Peak", "Mom 12-2"),
    long_name  = c("Pre-Peak 12-2 Momentum (Büsing 2022)",
                   "Post-Peak 12-2 Momentum (Büsing 2022)",
                   "Standard 12-2 Momentum (Büsing baseline)")
  )
}

.name_short <- function(code) {
  if (is.null(strat_names)) return(code)
  m <- mom_names$short_name[mom_names$code_name == code]
  if (length(m) == 1L) m else code
}
.name_long <- function(code) {
  if (is.null(strat_names)) return(code)
  m <- mom_names$long_name[mom_names$code_name == code]
  if (length(m) == 1L) m else code
}

# ── Headline summary statistics (dynamic-prose-values rule) ─────────────────
if (!is.null(summary_tbl)) {
  pre_row  <- summary_tbl[summary_tbl$strategy == "mom_prepeak",  , drop = FALSE]
  post_row <- summary_tbl[summary_tbl$strategy == "mom_postpeak", , drop = FALSE]
  comb_row <- summary_tbl[summary_tbl$strategy == "mom_combined", , drop = FALSE]

  pre_sharpe   <- if (nrow(pre_row)  == 1L) round(pre_row$sharpe,   2) else "N/A"
  post_sharpe  <- if (nrow(post_row) == 1L) round(post_row$sharpe,  2) else "N/A"
  comb_sharpe  <- if (nrow(comb_row) == 1L) round(comb_row$sharpe,  2) else "N/A"
  pre_cagr     <- if (nrow(pre_row)  == 1L && !is.na(pre_row$cagr))
    paste0(round(pre_row$cagr * 100, 1), "%") else "N/A"
  post_cagr    <- if (nrow(post_row) == 1L && !is.na(post_row$cagr))
    paste0(round(post_row$cagr * 100, 1), "%") else "N/A (bankrupt)"
  pre_vol      <- if (nrow(pre_row)  == 1L) paste0(round(pre_row$vol * 100, 1), "%") else "N/A"
  post_vol     <- if (nrow(post_row) == 1L) paste0(round(post_row$vol * 100, 1), "%") else "N/A"
  pre_maxdd    <- if (nrow(pre_row)  == 1L) paste0(round(pre_row$max_dd * 100, 1), "%") else "N/A"
  post_maxdd   <- if (nrow(post_row) == 1L) paste0(round(post_row$max_dd * 100, 1), "%") else "N/A"
  pre_nmonths  <- if (nrow(pre_row)  == 1L) pre_row$n_months  else NA_integer_
  bankrupt_mo  <- if (nrow(post_row) == 1L && !is.na(post_row$bankrupt_month))
    post_row$bankrupt_month else NA_integer_
} else {
  pre_sharpe  <- post_sharpe  <- comb_sharpe <- "N/A"
  pre_cagr    <- post_cagr    <- "N/A"
  pre_vol     <- post_vol     <- "N/A"
  pre_maxdd   <- post_maxdd   <- "N/A"
  pre_nmonths <- bankrupt_mo  <- NA_integer_
}

# ── Peak-position distribution summary (for caption prose) ───────────────────
if (!is.null(signal_raw) && "peak_position" %in% names(signal_raw)) {
  pp_mean   <- round(mean(signal_raw$peak_position, na.rm = TRUE), 2)
  pp_median <- round(median(signal_raw$peak_position, na.rm = TRUE), 2)
  pp_n      <- nrow(signal_raw)
} else {
  pp_mean <- pp_median <- pp_n <- NA
}

# ── Skewness of monthly L/S returns ─────────────────────────────────────────
.skewness <- function(x) {
  x <- x[!is.na(x)]
  n <- length(x)
  if (n < 3L) return(NA_real_)
  m3 <- mean((x - mean(x))^3)
  s3 <- sd(x)^3
  (n / ((n - 1) * (n - 2))) * m3 / s3
}

pre_skew  <- if (!is.null(ret_pre))  round(.skewness(ret_pre$ret_ls),  2) else NA
post_skew <- if (!is.null(ret_post)) round(.skewness(ret_post$ret_ls), 2) else NA
comb_skew <- if (!is.null(ret_comb)) round(.skewness(ret_comb$ret_ls), 2) else NA
```

# Overview

## Row

### {.tabset}

#### Background

<abbr title="12-2 momentum: cross-sectional stock momentum strategy using the 12-month formation return, skipping the most recent 2 months to avoid short-term reversal. Stocks in the top decile are held long, bottom decile short, with monthly rebalance.">12-2 momentum</abbr> is one of the most studied factors in empirical finance. [Büsing, Mohrschladt & Siedhoff (2022)](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=4298538) ask a sharper question: **within the 10-month formation window, when does the alpha come from?** They decompose the window at the stock's peak price date, creating two sub-windows:

- **Pre-peak period** — from formation start to the date when the stock reached its highest price. Stocks that rose to their peak early and then drifted sideways (or declined) still rank highly on pre-peak return.
- **Post-peak period** — from the peak price date to formation end (the 12-2 end-date). Stocks that kept rising to the end of the window have large post-peak returns.

The paper's headline finding on CRSP-scale data: roughly **84% of 12-2 momentum profits** come from the pre-peak sub-period. Sorting on pre-peak return rather than total formation return produces a cleaner, less crash-prone momentum signal.

This vignette replicates the decomposition on the **`ltr_universe` (~51 US non-ETF equities, monthly rebalance from 1973)** — a methodology-demo dataset, NOT CRSP scale. The universe is intentionally small to enable rapid prototyping before scaling to a larger cross-section.

#### Our Finding

Our finding on the 51-stock universe is **stronger than the paper's 84/16 claim**: the post-peak leg does not merely capture 16% of momentum alpha — it actively destroys value and goes bankrupt.

| Strategy | Sharpe | CAGR | Max DD | Blown Up |
|----------|--------|------|--------|----------|
| `r .name_short("mom_prepeak")` | `r pre_sharpe` | `r pre_cagr` | `r pre_maxdd` | FALSE |
| `r .name_short("mom_postpeak")` | `r post_sharpe` | `r post_cagr` | `r post_maxdd` | TRUE |
| `r .name_short("mom_combined")` | `r comb_sharpe` | N/A (bankrupt) | -100% (capped) | TRUE |

Pre-peak is the **only viable leg**. Post-peak and combined both go to zero equity at month `r if (!is.na(bankrupt_mo)) bankrupt_mo else "N/A"` of the sample — a concentrated loss event driven by the portfolio's extreme short position during a market reversal.

::: {.callout-note}
**Scope caveat.** The ltr_universe has ~51 stocks — a methodology-demo scale, not the full CRSP universe Büsing et al. used. The bankruptcy result is a property of this small cross-section's long-short construction (#374 tracks a follow-up on portfolio construction alternatives). The paper's 84/16 split may be more robust on CRSP-scale data where the long-short portfolio is more diversified.
:::

#### The Split Explained

For any stock and any formation window, define:

- `formation_start` — first trading date in the 10-month window (12 months back, skip 2)
- `formation_end`   — last trading date (2 months before the rebalance date)
- `peak_date`       — the calendar date when the stock's adjusted price was highest within the window

Then:

```
pre_peak_return  = price(peak_date) / price(formation_start) - 1
post_peak_return = price(formation_end) / price(peak_date)   - 1
total_return     = price(formation_end) / price(formation_start) - 1
```

Up to compounding: `(1 + pre) × (1 + post) ≈ (1 + total)`.

The `peak_position` column records where in the window the peak fell: 0 = peak on day 1, 1 = peak on the last day. A uniform distribution would indicate no information content. The histogram below shows whether peaks cluster at any point in the formation window.

# Peak-Date Distribution

## Row

### {.tabset}

#### Histogram (Paper Figure 1 replication)

```{r}
#| fig-height: 5
#| fig-alt: "Histogram of peak_position values (0 to 1) across all ticker-month observations in mom_prepeak_signal_raw. Each bar represents a 5% bin of the 10-month formation window. The distribution shows whether peaks cluster early (left), late (right), or uniformly across the window. A left-skewed distribution would indicate that stocks tend to peak early in the formation period, consistent with the Büsing et al. (2022) finding that pre-peak returns dominate."
if (!is.null(signal_raw) && "peak_position" %in% names(signal_raw)) {
  ggplot(
    signal_raw |> dplyr::filter(!is.na(.data$peak_position)),
    aes(x = .data$peak_position)
  ) +
    geom_histogram(
      binwidth = 0.05,
      fill     = PALETTE["mom_prepeak"],
      colour   = "black",
      alpha    = 0.85
    ) +
    geom_vline(
      xintercept = pp_mean,
      colour     = "white",
      linetype   = "dashed",
      linewidth  = 0.8
    ) +
    scale_x_continuous(
      labels = scales::percent_format(accuracy = 1),
      breaks = seq(0, 1, by = 0.1)
    ) +
    scale_y_continuous(labels = scales::comma) +
    labs(
      x     = "Peak position in formation window (0 = start, 1 = end)",
      y     = "Count (ticker-month observations)",
      title = "Distribution of Peak Dates within the 10-Month Formation Window"
    ) +
    theme_dark() +
    theme(
      panel.background  = element_rect(fill = "#1a1a1a", colour = NA),
      plot.background   = element_rect(fill = "#1a1a1a", colour = NA),
      text              = element_text(colour = "#e0e0e0"),
      axis.text         = element_text(colour = "#cccccc"),
      panel.grid.major  = element_line(colour = "#333333"),
      panel.grid.minor  = element_line(colour = "#222222"),
      plot.title        = element_text(size = 11, face = "bold")
    )
} else {
  cat("*Peak-position data not available — run tar_make() to generate.*")
}
```

::: {.fig-caption}
**Distribution of peak dates** within the 10-month formation window, across all `r if (!is.na(pp_n)) scales::comma(pp_n) else "N/A"` ticker-month observations in `mom_prepeak_signal_raw`. The horizontal axis is `peak_position` (0 = peak on day 1 of the formation window, 1 = peak on the last day). The dashed white line marks the mean at `r pp_mean`. Büsing et al. (2022) Figure 1 shows the same histogram on CRSP data and finds peaks cluster in the first half of the window (mean below 0.5), consistent with the pre-peak period being the primary carrier of momentum profits. A mean below 0.5 on our 51-stock sample would replicate the paper's distributional finding.
:::

#### What the Distribution Tells Us

The peak-date distribution is a key diagnostic for the 84/16 decomposition:

- **Peaks early (mean < 0.5):** Most stocks hit their high price early in the formation window and then drift. Momentum sorts predominantly on pre-peak return. Pre-peak dominance is expected.
- **Peaks late (mean > 0.5):** Stocks typically keep appreciating through the formation window. Post-peak return is the dominant component. Pre-peak dominance would be surprising.
- **Uniform distribution:** No informative clustering. Decomposition carries no additional signal beyond total momentum.

The mean peak position in our sample is `r pp_mean` (median: `r pp_median`). This `r if (!is.na(pp_mean) && pp_mean < 0.5) "confirms the early-clustering pattern" else if (!is.na(pp_mean) && pp_mean > 0.5) "shows late-clustering, which is the opposite of the paper's finding" else "value is unavailable"` on our universe.

# Decomposition Table

## Row

### {.tabset}

#### 84/16 Summary

```{r}
if (!is.null(summary_tbl)) {
  # Build display tibble
  display_df <- summary_tbl |>
    dplyr::left_join(
      mom_names |> dplyr::select("code_name", "long_name"),
      by = c("strategy" = "code_name")
    ) |>
    dplyr::mutate(
      Strategy   = dplyr::coalesce(.data$long_name, .data$strategy),
      Sharpe     = round(.data$sharpe, 2),
      CAGR       = dplyr::if_else(
        is.na(.data$cagr),
        "N/A (bankrupt)",
        paste0(round(.data$cagr * 100, 1), "%")
      ),
      Vol        = paste0(round(.data$vol * 100, 1), "%"),
      `Max DD`   = dplyr::if_else(
        .data$max_dd <= -1,
        "-100% (capped)",
        paste0(round(.data$max_dd * 100, 1), "%")
      ),
      Months     = .data$n_months,
      `Blown Up` = dplyr::if_else(.data$blown_up, "YES — bankrupt", "No"),
      `Bankrupt Month` = dplyr::if_else(
        !is.na(.data$bankrupt_month),
        as.character(.data$bankrupt_month),
        "—"
      )
    ) |>
    dplyr::select(Strategy, Sharpe, CAGR, Vol, `Max DD`, Months, `Blown Up`, `Bankrupt Month`)

  DT::datatable(
    display_df,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold; color: #e0e0e0;",
      paste0(
        "Pre-peak / post-peak 12-2 momentum decomposition summary. ",
        "Annualised metrics from ", if (!is.na(pre_nmonths)) pre_nmonths else "N/A",
        " monthly observations (ltr_universe, 1973-onwards). ",
        "Strategies sorted on signal column: pre_peak_return (mom_prepeak), ",
        "post_peak_return (mom_postpeak), total_return (mom_combined). ",
        "Long top decile, short bottom decile, equal-weighted, 10 bps/trade cost. ",
        "Blown Up = YES means the cumulative L/S equity hit zero before period end."
      )
    ),
    rownames  = FALSE,
    options   = list(
      pageLength = 10, scrollX = TRUE, dom = "t",
      rowCallback = DT::JS(
        "function(row, data) {",
        "  if (data[6] && data[6].indexOf('YES') !== -1) {",
        "    $(row).addClass('bankrupt-row');",
        "    $(row).css('background-color', 'rgba(240, 128, 128, 0.25)');",
        "  }",
        "}"
      )
    )
  )
} else {
  cat("*Summary table not available — run tar_make().*")
}
```

#### How to Read This

The three strategies differ only in the signal column used to rank stocks:

| Strategy | Signal column | What it selects |
|----------|--------------|-----------------|
| `r .name_short("mom_prepeak")` | `pre_peak_return` | Stocks that rose the most from formation-start to their price peak |
| `r .name_short("mom_postpeak")` | `post_peak_return` | Stocks that rose (or fell least) from their price peak to formation-end |
| `r .name_short("mom_combined")` | `total_return` | Standard 12-2 momentum — full formation window return |

All three use the same universe, rebalance dates, portfolio construction (decile long-short), and transaction cost (10 bps/trade). The only difference is which return column drives the ranking. The fact that `r .name_short("mom_postpeak")` goes bankrupt while `r .name_short("mom_prepeak")` achieves Sharpe `r pre_sharpe` illustrates that momentum's profits are concentrated in the pre-peak sub-period — even on this small universe.

# Cumulative Returns

## Row

### {.tabset}

#### Equity Curves

```{r}
#| fig-height: 6
#| fig-alt: "Cumulative equity curves for three momentum decomposition strategies from 1973: Pre-Peak (green), Post-Peak (red), and Combined/Standard 12-2 (blue). The pre-peak strategy (green line) grows steadily, reaching the end of the sample with Sharpe +0.44. The post-peak (red) and combined (blue) curves collapse to zero at month 248 (approximately August 1993) due to a single extreme loss month that wipes out the entire long-short equity. A vertical red dashed line marks the bankruptcy month. After month 248 both failing strategies are shown clamped at zero."
if (!is.null(ret_pre) && !is.null(ret_post) && !is.null(ret_comb)) {
  # Build combined long-form return table
  rets_df <- dplyr::bind_rows(
    ret_pre  |> dplyr::mutate(strategy = "mom_prepeak"),
    ret_post |> dplyr::mutate(strategy = "mom_postpeak"),
    ret_comb |> dplyr::mutate(strategy = "mom_combined")
  ) |>
    dplyr::arrange(.data$strategy, .data$as_of_date)

  # Compute cumulative equity (clamped at 0 at bankruptcy)
  cum_df <- rets_df |>
    dplyr::group_by(.data$strategy) |>
    dplyr::mutate(
      month_idx = dplyr::row_number(),
      cum_ret   = cumprod(pmax(1 + .data$ret_ls, 0)),
      cum_ret   = pmax(.data$cum_ret, 0)
    ) |>
    dplyr::ungroup()

  # Strategy label lookup for legend
  cum_df <- cum_df |>
    dplyr::left_join(
      mom_names |> dplyr::select("code_name", "short_name"),
      by = c("strategy" = "code_name")
    ) |>
    dplyr::mutate(
      label = dplyr::coalesce(.data$short_name, .data$strategy)
    )

  # Bankruptcy annotation: first month where cum_ret reaches 0
  bankrupt_df <- cum_df |>
    dplyr::filter(.data$cum_ret <= 0) |>
    dplyr::group_by(.data$strategy) |>
    dplyr::slice_min(.data$month_idx, n = 1L, with_ties = FALSE) |>
    dplyr::ungroup() |>
    dplyr::filter(.data$strategy %in% c("mom_postpeak", "mom_combined"))

  # Determine x-axis type: use as_of_date if available, otherwise month_idx
  x_var <- if (inherits(cum_df$as_of_date, "Date")) "as_of_date" else "month_idx"
  x_lab <- if (x_var == "as_of_date") "Date" else "Month index"

  p <- ggplot(cum_df, aes(
    x      = !!rlang::sym(x_var),
    y      = .data$cum_ret,
    colour = .data$strategy,
    group  = .data$strategy
  )) +
    geom_line(linewidth = 0.9) +
    scale_colour_manual(
      values = PALETTE,
      labels = setNames(mom_names$short_name, mom_names$code_name),
      name   = "Strategy"
    )

  # Bankruptcy vertical line (use first x value for bankrupt strategies)
  if (nrow(bankrupt_df) > 0) {
    x_bk <- bankrupt_df[[x_var]][1]
    p <- p + geom_vline(
      xintercept = x_bk,
      colour     = "#f08080",
      linetype   = "dashed",
      linewidth  = 0.7,
      alpha      = 0.8
    ) +
    annotate(
      "text",
      x     = x_bk,
      y     = max(cum_df$cum_ret, na.rm = TRUE) * 0.8,
      label = paste0("Bankruptcy\n(month ", bankrupt_mo, ")"),
      colour = "#f08080",
      size   = 3.2,
      hjust  = 1.05
    )
  }

  p +
    scale_y_continuous(labels = scales::dollar_format(prefix = "$")) +
    labs(
      x     = x_lab,
      y     = "Growth of $1",
      title = "Cumulative L/S Equity: Pre-Peak vs Post-Peak vs Combined"
    ) +
    theme_dark() +
    theme(
      panel.background = element_rect(fill = "#1a1a1a", colour = NA),
      plot.background  = element_rect(fill = "#1a1a1a", colour = NA),
      text             = element_text(colour = "#e0e0e0"),
      axis.text        = element_text(colour = "#cccccc"),
      panel.grid.major = element_line(colour = "#333333"),
      panel.grid.minor = element_line(colour = "#222222"),
      legend.background = element_rect(fill = "#1a1a1a"),
      legend.key        = element_rect(fill = "#1a1a1a"),
      plot.title        = element_text(size = 11, face = "bold")
    )
} else {
  cat("*Return data not available — run tar_make() to generate.*")
}
```

::: {.fig-caption}
**Cumulative L/S equity curves** for all three momentum decomposition strategies (growth of $1, not log scale). Green = `r .name_long("mom_prepeak")`, Sharpe `r pre_sharpe`, CAGR `r pre_cagr`, max drawdown `r pre_maxdd`. Red = `r .name_long("mom_postpeak")`, Sharpe `r post_sharpe`, blown up at month `r if (!is.na(bankrupt_mo)) bankrupt_mo else "N/A"`. Blue = `r .name_long("mom_combined")`, Sharpe `r comb_sharpe`, also blown up at the same month. The red dashed vertical line marks the bankruptcy month. After that month post-peak and combined are clamped at zero. The post-peak and combined strategies go bankrupt at the same month because they share the same portfolio construction (decile long-short, equal-weighted) applied to the same universe; the only difference is the signal column used to rank stocks. See issue #374 for a follow-up on alternative long-short construction methods.
:::

#### Returns Table

```{r}
if (!is.null(ret_pre)) {
  display_ret <- dplyr::bind_rows(
    ret_pre  |> dplyr::mutate(strategy = .name_short("mom_prepeak")),
    ret_post |> dplyr::mutate(strategy = .name_short("mom_postpeak")),
    ret_comb |> dplyr::mutate(strategy = .name_short("mom_combined"))
  ) |>
    dplyr::mutate(
      dplyr::across(c("ret_long", "ret_short", "ret_ls"),
                    ~ paste0(round(. * 100, 2), "%"))
    ) |>
    dplyr::select(strategy, as_of_date, exec_date, ret_long, ret_short, ret_ls)

  DT::datatable(
    display_ret,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold; color: #e0e0e0;",
      "Monthly long-short returns for all three decomposition strategies. ret_ls = ret_long - ret_short - 2 × 10 bps transaction cost. Negative ret_ls in the same month caused the bankruptcy in post-peak and combined strategies."
    ),
    rownames = FALSE,
    filter   = "top",
    options  = list(pageLength = 15, scrollX = TRUE, dom = "frtip")
  )
}
```

# Skewness Analysis

## Row

### {.tabset}

#### Skewness Comparison

```{r}
#| fig-height: 5
#| fig-alt: "Faceted histogram of monthly long-short returns for three strategies: Pre-Peak (green), Post-Peak (red), and Combined (blue). Each panel shows the distribution of monthly ret_ls values. The pre-peak distribution is expected to be positively skewed (fewer large negative months), while post-peak and combined may exhibit negative skewness driven by the bankruptcy month. Skewness statistics are annotated in each panel."
if (!is.null(ret_pre) && !is.null(ret_post) && !is.null(ret_comb)) {
  skew_df <- dplyr::bind_rows(
    ret_pre  |> dplyr::mutate(strategy = "mom_prepeak"),
    ret_post |> dplyr::mutate(strategy = "mom_postpeak"),
    ret_comb |> dplyr::mutate(strategy = "mom_combined")
  ) |>
    dplyr::left_join(
      mom_names |> dplyr::select("code_name", "short_name"),
      by = c("strategy" = "code_name")
    ) |>
    dplyr::mutate(label = dplyr::coalesce(.data$short_name, .data$strategy))

  # Annotation data for skewness text
  skew_ann <- tibble::tibble(
    strategy = c("mom_prepeak", "mom_postpeak", "mom_combined"),
    skewness = c(pre_skew, post_skew, comb_skew)
  ) |>
    dplyr::left_join(
      mom_names |> dplyr::select("code_name", "short_name"),
      by = c("strategy" = "code_name")
    ) |>
    dplyr::mutate(
      label = dplyr::coalesce(.data$short_name, .data$strategy),
      label_text = paste0("Skewness: ", ifelse(is.na(.data$skewness), "N/A", .data$skewness))
    )

  ggplot(skew_df, aes(x = .data$ret_ls, fill = .data$strategy)) +
    geom_histogram(binwidth = 0.01, colour = "black", alpha = 0.8) +
    geom_vline(xintercept = 0, colour = "white", linetype = "dotted") +
    geom_text(
      data  = skew_ann,
      aes(x = -Inf, y = Inf, label = .data$label_text),
      hjust = -0.1, vjust = 1.5,
      size  = 3.5,
      colour = "white",
      inherit.aes = FALSE
    ) +
    scale_fill_manual(values = PALETTE, guide = "none") +
    scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
    scale_y_continuous(labels = scales::comma) +
    facet_wrap(~ label, scales = "free_y", ncol = 1L) +
    labs(
      x     = "Monthly L/S return",
      y     = "Count",
      title = "Distribution of Monthly L/S Returns — Skewness Comparison"
    ) +
    theme_dark() +
    theme(
      panel.background  = element_rect(fill = "#1a1a1a", colour = NA),
      plot.background   = element_rect(fill = "#1a1a1a", colour = NA),
      strip.background  = element_rect(fill = "#2a2a2a"),
      strip.text        = element_text(colour = "#e0e0e0"),
      text              = element_text(colour = "#e0e0e0"),
      axis.text         = element_text(colour = "#cccccc"),
      panel.grid.major  = element_line(colour = "#333333"),
      panel.grid.minor  = element_line(colour = "#222222"),
      plot.title        = element_text(size = 11, face = "bold")
    )
} else {
  cat("*Return data not available — run tar_make().*")
}
```

::: {.fig-caption}
**Distribution of monthly long-short returns** for all three strategies (`r pre_nmonths` monthly observations each). Skewness: Pre-Peak `r pre_skew`, Post-Peak `r post_skew`, Combined `r comb_skew`. Büsing et al. (2022) report that sorting on pre-peak return produces a *positively skewed* return distribution — fewer catastrophic months — compared to standard momentum (which exhibits negative skewness from momentum crashes). A positive skewness for pre-peak and negative skewness for post-peak and combined would replicate their finding. Note that the extreme left tail in post-peak and combined is the single bankruptcy month that drives the strategy to zero equity.
:::

#### Skewness Table

```{r}
skew_tbl <- tibble::tibble(
  Strategy = c(
    .name_long("mom_prepeak"),
    .name_long("mom_postpeak"),
    .name_long("mom_combined")
  ),
  Sharpe   = c(pre_sharpe,  post_sharpe,  comb_sharpe),
  Skewness = c(pre_skew,    post_skew,    comb_skew),
  Interpretation = c(
    if (!is.na(pre_skew) && pre_skew > 0) "Positive — fewer large losses" else "Negative",
    if (!is.na(post_skew)) paste0(
      if (post_skew < 0) "Negative — left tail from bankruptcy month" else "Positive",
      " (post_skew = ", post_skew, ")"
    ) else "N/A",
    if (!is.na(comb_skew)) paste0(
      if (comb_skew < 0) "Negative — left tail from bankruptcy month" else "Positive",
      " (comb_skew = ", comb_skew, ")"
    ) else "N/A"
  )
)

DT::datatable(
  skew_tbl,
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold; color: #e0e0e0;",
    paste0(
      "Skewness of monthly L/S returns for each decomposition strategy. ",
      "Positive skewness implies fewer large-magnitude negative months, ",
      "which is preferable from a tail-risk perspective. ",
      "Standard errors not reported; see PR 4/4 for bootstrap confidence intervals."
    )
  ),
  rownames = FALSE,
  options  = list(pageLength = 5, scrollX = TRUE, dom = "t")
)
```

# Key Findings

## Row

### {.tabset}

#### Summary

**Headline results on the ltr_universe (~51 US non-ETF equities, monthly from 1973):**

- **`r .name_long("mom_prepeak")`**: Sharpe `r pre_sharpe`, CAGR `r pre_cagr`, vol `r pre_vol`, max drawdown `r pre_maxdd`. This is the **only viable leg** of the decomposition. The strategy is not blown up.

- **`r .name_long("mom_postpeak")`**: Sharpe `r post_sharpe`, CAGR `r post_cagr`, vol `r post_vol`. Post-peak actively destroys value and hits bankruptcy at month `r if (!is.na(bankrupt_mo)) bankrupt_mo else "N/A"` — a single extreme month in which the equal-weighted short portfolio (stocks near their formation-window peak) dramatically outperforms the long portfolio (stocks far below their peak), wiping out equity.

- **`r .name_long("mom_combined")`**: Sharpe `r comb_sharpe`. The standard 12-2 momentum signal is a linear combination of pre-peak and post-peak return. The two legs largely cancel each other. The strategy also goes bankrupt at the same month as post-peak because it uses the same construction with a different ranking signal.

- **Pre-peak skewness**: `r pre_skew` (post-peak: `r post_skew`). A positive pre-peak skewness vs negative post-peak skewness replicates the paper's directional finding that pre-peak sorting produces a better-behaved return distribution.

**Caveats:**

1. The ltr_universe (~51 stocks) is a methodology-demo dataset. The paper used CRSP-scale data (~4,000+ stocks). Portfolio-level diversification is much lower here, making bankruptcy from a single concentrated event more likely.
2. The bankruptcy event at month `r if (!is.na(bankrupt_mo)) bankrupt_mo else "N/A"` is concentrated in time and affects post-peak and combined identically, suggesting it is a portfolio-construction artefact of the small universe rather than an inherent property of the post-peak signal. See issue #374.
3. Walk-forward validation, CPCV PBO, falsification tests, and Pillar-8 risk metrics are deferred to PR 4/4.

#### Deferred Work (PR 4/4)

The following analyses are **not included in this PR** and are tracked in issue #365:

| Analysis | Status |
|----------|--------|
| Walk-forward correlation vs standard momentum | Deferred to PR 4/4 |
| CPCV PBO (deflated Sharpe / combinatorial backtest) | Deferred to PR 4/4 |
| Falsification: random-day-as-peak null | Deferred to PR 4/4 |
| Pillar-8 risk metrics (skewness of daily returns, tail ratio) | Deferred to PR 4/4 |
| Full CRSP-scale replication (60/40 result matching paper's 84/16) | Open question in #365 |

# Pillar-8 Risk Metrics

## Row

### {.tabset}

#### Risk Metrics Table

```{r}
# Read gauntlet summary: mom_prepeak_summary now includes Pillar-8 columns
# (avg_dd_days, max_dd_days, max_cons_losses, loss_clustered).
p8_cols <- c("avg_dd_days", "max_dd_days", "max_cons_losses", "loss_clustered")

p8_available <- !is.null(summary_tbl) && all(p8_cols %in% names(summary_tbl))

if (p8_available) {
  p8_display <- summary_tbl |>
    dplyr::left_join(
      mom_names |> dplyr::select("code_name", "short_name"),
      by = c("strategy" = "code_name")
    ) |>
    dplyr::mutate(
      Strategy          = dplyr::coalesce(.data$short_name, .data$strategy),
      `Avg DD (months)` = dplyr::if_else(
        is.na(.data$avg_dd_days), "—", as.character(round(.data$avg_dd_days, 1))),
      `Max DD (months)` = dplyr::if_else(
        is.na(.data$max_dd_days), "—", as.character(round(.data$max_dd_days, 1))),
      `Max Consec Losses` = dplyr::if_else(
        is.na(.data$max_cons_losses), "—", as.character(.data$max_cons_losses)),
      `Loss Clustered` = dplyr::case_when(
        is.na(.data$loss_clustered)   ~ "—",
        isTRUE(.data$loss_clustered)  ~ "YES",
        TRUE                          ~ "No"
      ),
      Sharpe = round(.data$sharpe, 2),
      `Blown Up` = dplyr::if_else(.data$blown_up, "YES", "No")
    ) |>
    dplyr::select(
      Strategy, Sharpe, `Blown Up`,
      `Avg DD (months)`, `Max DD (months)`, `Max Consec Losses`, `Loss Clustered`
    )

  # Extract pre-peak values for dynamic prose
  pre_p8 <- summary_tbl[summary_tbl$strategy == "mom_prepeak", , drop = FALSE]
  pre_avg_dd  <- if (nrow(pre_p8) == 1L && !is.na(pre_p8$avg_dd_days))
    round(pre_p8$avg_dd_days, 1) else "N/A"
  pre_max_dd_d <- if (nrow(pre_p8) == 1L && !is.na(pre_p8$max_dd_days))
    round(pre_p8$max_dd_days, 1) else "N/A"
  pre_max_cons <- if (nrow(pre_p8) == 1L && !is.na(pre_p8$max_cons_losses))
    pre_p8$max_cons_losses else "N/A"

  DT::datatable(
    p8_display,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold; color: #e0e0e0;",
      paste0(
        "Pillar-8 risk metrics for all three momentum decomposition strategies. ",
        "Avg/Max DD duration is measured in months (index positions, no date column). ",
        "Max Consecutive Losses = longest losing streak in the pre-bankruptcy slice. ",
        "Loss Clustered = TRUE when both the Wald-Wolfowitz runs test p < 0.05 AND ",
        "lag-1 autocorrelation > 0.2 (dual-signal composite). ",
        "Metrics for blown-up strategies are computed on the pre-bankruptcy slice only."
      )
    ),
    rownames = FALSE,
    options  = list(pageLength = 5, scrollX = TRUE, dom = "t")
  )
} else {
  cat("*Pillar-8 metrics not available — run tar_make() after PR 4/4 merge.*")
  pre_avg_dd <- pre_max_dd_d <- pre_max_cons <- "N/A"
}
```

#### Interpretation

The Pillar-8 risk framework adds three drawdown-duration and loss-clustering metrics beyond the standard max_dd percentage reported in the summary table:

- **Avg DD duration** — average number of months spent in a drawdown event (defined as a contiguous run below the prior peak by more than 1%). For the pre-peak strategy this is `r pre_avg_dd` months on average.
- **Max DD duration** — longest single drawdown event: `r pre_max_dd_d` months. This captures the "pain duration" dimension of drawdown risk that the peak loss percentage alone does not convey.
- **Max consecutive losses** — longest uninterrupted run of negative monthly returns: `r pre_max_cons` months. A high count warns of momentum-crash-style tail events.

Blown-up strategies (mom_postpeak, mom_combined) have these metrics computed only on the pre-bankruptcy slice because post-bankruptcy returns are economically meaningless — the portfolio has zero equity.


# Walk-Forward Correlation

## Row

### {.tabset}

#### WFC Grid

```{r}
wfc_grid   <- safe_tar_read("mom_prepeak_wfc_grid")
wfc_result <- safe_tar_read("mom_prepeak_wfc_result")

wfc_available <- !is.null(wfc_grid) && !is.null(wfc_result)

if (wfc_available) {
  pearson_rho   <- round(wfc_result$pearson_rho,   3)
  spearman_rho  <- round(wfc_result$spearman_rho,  3)
  wfc_class_val <- wfc_result$classification
  median_oos    <- round(wfc_result$median_oos, 3)

  wfc_display <- wfc_grid |>
    dplyr::mutate(
      `N Quantiles`  = .data$theta_label,
      `IS Sharpe`    = round(.data$IS_metric,  3),
      `OOS Sharpe`   = round(.data$OOS_metric, 3)
    ) |>
    dplyr::select(`N Quantiles`, `IS Sharpe`, `OOS Sharpe`)

  DT::datatable(
    wfc_display,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold; color: #e0e0e0;",
      paste0(
        "Walk-Forward Correlation grid for mom_prepeak over n_quantiles ∈ {5, 10, 20}. ",
        "IS = annualised Sharpe on the first half of the sample (~1973–mid). ",
        "OOS = annualised Sharpe on the second half. ",
        "Pearson ρ = ", pearson_rho, "; Spearman ρ = ", spearman_rho, ". ",
        "WFC classification: ", wfc_class_val, ". ",
        "A positive Pearson ρ with positive OOS Sharpe = 'structural_edge'; ",
        "a classification of 'spurious_luck' or 'noise' would flag concern."
      )
    ),
    rownames = FALSE,
    options  = list(pageLength = 5, scrollX = TRUE, dom = "t")
  )
} else {
  cat("*WFC results not available — run tar_make() after PR 4/4 merge.*")
  pearson_rho <- spearman_rho <- wfc_class_val <- median_oos <- "N/A"
}
```

#### WFC Chart

```{r}
#| fig-height: 4
#| fig-alt: "Scatter plot of IS Sharpe vs OOS Sharpe across 3 n_quantiles values (5, 10, 20) for the mom_prepeak strategy. Each point represents one parameter configuration. A positive correlation between IS and OOS Sharpe supports structural edge; a negative correlation would suggest overfitting. The Pearson correlation coefficient is annotated on the chart."
if (wfc_available && !is.null(wfc_grid)) {
  ggplot(wfc_grid, aes(x = .data$IS_metric, y = .data$OOS_metric,
                       label = .data$theta_label)) +
    geom_point(colour = PALETTE["mom_prepeak"], size = 4) +
    ggplot2::geom_text(
      colour = "#e0e0e0", size = 3.5,
      nudge_y = 0.02
    ) +
    geom_hline(yintercept = 0, colour = "white", linetype = "dotted") +
    geom_vline(xintercept = 0, colour = "white", linetype = "dotted") +
    annotate("text",
      x = min(wfc_grid$IS_metric, na.rm = TRUE),
      y = max(wfc_grid$OOS_metric, na.rm = TRUE),
      label = paste0("Pearson ρ = ", if (is.character(pearson_rho)) pearson_rho else round(as.numeric(pearson_rho), 3)),
      colour = "#e0e0e0", hjust = 0, vjust = 1, size = 4
    ) +
    labs(
      x     = "IS Sharpe (first half)",
      y     = "OOS Sharpe (second half)",
      title = paste0("Walk-Forward Correlation — mom_prepeak (n_quantiles grid)")
    ) +
    theme_dark() +
    theme(
      panel.background  = element_rect(fill = "#1a1a1a", colour = NA),
      plot.background   = element_rect(fill = "#1a1a1a", colour = NA),
      text              = element_text(colour = "#e0e0e0"),
      axis.text         = element_text(colour = "#cccccc"),
      panel.grid.major  = element_line(colour = "#333333"),
      panel.grid.minor  = element_line(colour = "#222222"),
      plot.title        = element_text(size = 11, face = "bold")
    )
} else {
  cat("*WFC chart not available — run tar_make() after PR 4/4 merge.*")
}
```

::: {.fig-caption}
**Walk-forward correlation scatter** for mom_prepeak over 3 n_quantiles values. Pearson ρ = `r pearson_rho`, Spearman ρ = `r spearman_rho`. WFC classification: **`r wfc_class_val`**. The Tinsley (2024) framework labels a strategy as having "structural edge" when WFC Pearson ≥ 0.70 AND the median OOS Sharpe is positive. A classification of "spurious_luck" (high WFC, negative OOS) or "noise" (low WFC) would raise concern about the strategy's robustness. This WFC grid uses n_quantiles as the sole tuning dimension; a v2 follow-up will expand the grid over formation-window parameters.
:::

#### Verdict

The Walk-Forward Correlation test asks: when we vary the portfolio-construction parameter (number of quantile buckets), does IS performance predict OOS performance? A positive correlation indicates that what works in training also works out of sample — a sign of structural edge rather than parameter overfitting.

WFC result for mom_prepeak:

- **Pearson ρ:** `r pearson_rho` (IS vs OOS Sharpe across n_quantiles ∈ {5, 10, 20})
- **Spearman ρ:** `r spearman_rho`
- **Classification:** `r wfc_class_val`
- **Median OOS Sharpe:** `r median_oos`

`r if (is.character(wfc_class_val) && wfc_class_val == "structural_edge") "The classification of structural_edge supports robustness of the pre-peak momentum signal across portfolio-formation choices." else if (is.character(wfc_class_val) && wfc_class_val %in% c("spurious_luck", "noise")) paste0("The classification of ", wfc_class_val, " warrants caution — the strategy may not be robust to n_quantiles choice.") else "WFC classification not available — run tar_make()."`

Note: only n_quantiles is varied here; WFC over lookback-window parameters (which require re-running the signal computation) is deferred to the v2 follow-up.


# Random-Day-as-Peak Null

## Row

### {.tabset}

#### Null Test Result

```{r}
rand_peak_test    <- safe_tar_read("mom_prepeak_random_peak_test")
rand_peak_metrics <- safe_tar_read("mom_prepeak_random_peak_metrics")

rand_available <- !is.null(rand_peak_test)

if (rand_available) {
  actual_s <- round(rand_peak_test$actual_sharpe, 3)
  random_s <- round(rand_peak_test$random_sharpe, 3)
  null_dom  <- rand_peak_test$null_dominates

  null_display <- tibble::tibble(
    Strategy        = c("mom_prepeak (actual)", "Random-day-as-peak null"),
    Sharpe          = c(actual_s, random_s),
    `Null Dominates` = c("—", if (isTRUE(null_dom)) "YES — edge may be spurious" else "No — peak-finding adds alpha")
  )

  DT::datatable(
    null_display,
    caption = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; font-weight: bold; color: #e0e0e0;",
      paste0(
        "Random-day-as-peak null test. The null strategy uses the same ",
        "formation window, portfolio construction, and transaction costs, ",
        "but selects a uniformly random day as the 'peak' instead of the ",
        "actual price maximum (set.seed(42) for reproducibility). ",
        "If the null Sharpe >= actual Sharpe, peak-finding adds no edge. ",
        "Actual Sharpe = ", actual_s, "; ",
        "Random-peak Sharpe = ", random_s, "."
      )
    ),
    rownames = FALSE,
    options  = list(pageLength = 5, scrollX = TRUE, dom = "t")
  )
} else {
  cat("*Random-peak null test not available — run tar_make() after PR 4/4 merge.*")
  actual_s <- random_s <- null_dom <- "N/A"
}
```

#### Interpretation

The random-day-as-peak null is a strategy-specific falsification test. The logic:

1. Take the same 10-month formation window for every stock.
2. Instead of finding the actual price peak (maximum adjusted price), pick a **uniformly random day** within the window and treat it as the "peak."
3. Compute pre-peak return (price from day 1 to the random day) and post-peak return (price from the random day to formation end).
4. Sort stocks on this randomised pre-peak return and form the same long-short portfolio.

If the actual mom_prepeak Sharpe (**`r actual_s`**) is substantially higher than the random-peak Sharpe (**`r random_s`**), the true peak-finding is doing real work. If they are similar, the alpha likely comes from some other feature of the formation window (e.g., early-window momentum, late-window reversal) that the decomposition happens to capture but doesn't require an accurate peak date.

`r if (!is.null(null_dom) && is.logical(null_dom)) { if (isTRUE(null_dom)) "The null dominates: random-peak Sharpe ≥ actual Sharpe. This is a warning sign that peak-finding may be spurious or that the pre-peak signal is robust to how the 'peak' is defined." else "The null does NOT dominate: actual Sharpe > random-peak Sharpe. Peak-finding adds measurable alpha — the strategy is identifying a genuine signal from the true price-peak location." } else "Result not available — run tar_make()."`


# HAC + FF5 + CPCV PBO

## Row

### {.tabset}

#### Statistical Hardening

```{r}
hac_res  <- safe_tar_read("mom_prepeak_hac")
ff_reg   <- safe_tar_read("mom_prepeak_ff_reg")
pbo_res  <- safe_tar_read("mom_prepeak_pbo")

hac_available <- !is.null(hac_res)
ff_available  <- !is.null(ff_reg) && nrow(ff_reg) == 1L
pbo_available <- !is.null(pbo_res)

# Dynamic values for prose
hac_tstat  <- if (hac_available && !is.null(hac_res$hac_tstat))   round(hac_res$hac_tstat, 3)   else "N/A"
hac_sharpe <- if (hac_available && !is.null(hac_res$naive_sharpe)) round(hac_res$naive_sharpe, 3) else "N/A"
ff5_alpha  <- if (ff_available) round(ff_reg$alpha_annual, 4)    else "N/A"
ff5_alpha_t <- if (ff_available) round(ff_reg$alpha_tstat_hac, 3) else "N/A"
ff5_r2     <- if (ff_available) round(ff_reg$r_squared, 3)       else "N/A"
pbo_val    <- if (pbo_available) round(pbo_res$pbo, 3)           else "N/A"
pbo_paths  <- if (pbo_available) pbo_res$n_paths                 else "N/A"

stat_display <- tibble::tibble(
  Test    = c(
    "HAC Sharpe (Newey-West)",
    "HAC t-stat",
    "FF5+Mom alpha (annualised)",
    "FF5+Mom alpha t-stat (HAC)",
    "FF5+Mom R²",
    "CPCV PBO (n_quantiles grid)"
  ),
  Value   = c(
    as.character(hac_sharpe),
    as.character(hac_tstat),
    as.character(ff5_alpha),
    as.character(ff5_alpha_t),
    as.character(ff5_r2),
    as.character(pbo_val)
  ),
  Interpretation = c(
    "Annualised Sharpe from monthly returns (12× ann factor)",
    "t-statistic for H0: mean(ret_ls) = 0; HAC-robust to serial correlation",
    "Intercept from FF5+Momentum regression (monthly factors); annualised",
    "HAC t-stat on the regression alpha",
    "Fraction of return variance explained by FF5+Momentum factors",
    paste0("Fraction of CPCV paths (n=", pbo_paths, ") where IS-best n_quantiles ranks below OOS median; 0 = no overfitting, 1 = full overfitting")
  )
)

DT::datatable(
  stat_display,
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold; color: #e0e0e0;",
    paste0(
      "Statistical hardening metrics for mom_prepeak. ",
      "HAC inference uses Newey-West bandwidth (floor(4*(T/100)^(2/9))). ",
      "FF5 regression uses monthly Fama-French 5-factor + Momentum factors from Kenneth French's data library. ",
      "CPCV PBO uses C(6,2) = 15 paths over n_quantiles ∈ {5, 10, 20}, with 1-month purge + 1-month embargo."
    )
  ),
  rownames = FALSE,
  options  = list(pageLength = 8, scrollX = TRUE, dom = "t")
)
```

#### Interpretation

**HAC Inference (Newey-West):** The strategy's mean monthly return is tested against H0: μ = 0 using a heteroskedasticity- and autocorrelation-consistent standard error. HAC t-stat = **`r hac_tstat`**. A t-stat above 1.96 indicates significance at the 5% level. Monthly returns can exhibit serial correlation (especially in momentum strategies), so HAC inference is more reliable than a naive OLS t-stat.

**Fama-French 5-Factor + Momentum Regression:** The monthly excess returns are regressed on Mkt-RF, SMB, HML, RMW, CMA, and the UMD (momentum) factor. The intercept (alpha) measures the strategy's return not explained by standard systematic risks. Alpha = **`r ff5_alpha`** (annualised), with HAC t-stat = **`r ff5_alpha_t`**. R² = **`r ff5_r2`**. A positive alpha with a t-stat > 2 provides statistical evidence of genuine risk-adjusted returns. The UMD factor captures conventional momentum exposure; a positive alpha after including UMD would suggest pre-peak momentum captures something beyond standard momentum.

**CPCV PBO:** The Combinatorial Purged Cross-Validation Probability of Backtest Overfitting measures how often the IS-best configuration (n_quantiles value) underperforms the OOS median. PBO = **`r pbo_val`** (across `r pbo_paths` paths). PBO near 0 means the IS-best choice also tends to be OOS-best — little evidence of overfitting. PBO near 1 means IS selection is anti-predictive for OOS — strong evidence of overfitting. Since mom_prepeak is not a highly tuned strategy, this PBO is more of a sanity check than a critical test.

## Methodology

### What this vignette computes

This vignette reads pre-computed targets from the pipeline and builds eight sections: a peak-date distribution histogram (replicating paper Figure 1); a 84/16 decomposition table from `mom_prepeak_summary`; cumulative-return curves for all three strategies with bankruptcy annotation at month `r if (!is.na(bankrupt_mo)) bankrupt_mo else "N/A"`; a skewness comparison; Pillar-8 risk metrics (dd duration, consecutive losses, loss clustering) from the extended `mom_prepeak_summary`; a Walk-Forward Correlation grid over n_quantiles ∈ {5, 10, 20}; a random-day-as-peak null test; and a statistical hardening panel (HAC Sharpe, FF5+Momentum regression alpha, CPCV PBO).

### Data sources

- `mom_prepeak_signal_raw` — output of [`historicaldata::hd_mom_prepeak_signal()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/strategy_mom_prepeak.R#L67) applied to `ltr_universe` (~51 US non-ETF equities, daily adjusted prices, 1973-onwards). One row per (ticker, as_of_date) with columns `pre_peak_return`, `post_peak_return`, `total_return`, `peak_position`.
- `mom_prepeak_summary`, `mom_*_returns` targets — computed by [`R/plan_mom_prepeak.R`](https://github.com/JohnGavin/historical/blob/main/R/plan_mom_prepeak.R) using [`.mom_prepeak_compute_metrics()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/utils_mom_prepeak_metrics.R#L24) from `packages/historicaldata/R/utils_mom_prepeak_metrics.R`, extended in PR 4/4 to include Pillar-8 columns.
- `mom_prepeak_wfc_grid`, `mom_prepeak_wfc_result` — Walk-Forward Correlation targets from [`R/plan_mom_prepeak_gauntlet.R`](https://github.com/JohnGavin/historical/blob/main/R/plan_mom_prepeak_gauntlet.R), using [`hd_wf_correlation()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/wf_correlation.R#L84).
- `mom_prepeak_random_peak_test`, `mom_prepeak_random_peak_metrics` — random-day-as-peak null from [`.mom_prepeak_random_peak_signal()`](https://github.com/JohnGavin/historical/blob/main/R/plan_mom_prepeak_gauntlet.R) in the gauntlet plan.
- `mom_prepeak_hac`, `mom_prepeak_ff_reg` — HAC Sharpe and FF5+Momentum regression via [`hd_hac_sharpe()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/falsification.R#L80) and [`hd_factor_null_test()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/falsification.R#L738) on monthly Fama-French factors from `hd_factors(frequency="monthly")`.
- `mom_prepeak_pbo` — CPCV Probability of Backtest Overfitting via [`hd_pbo()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/cpcv.R#L267) over n_quantiles ∈ {5, 10, 20}.
- `strategy_names` — single source of truth for display names, defined in [`R/plan_strategy_names.R`](https://github.com/JohnGavin/historical/blob/main/R/plan_strategy_names.R).
- Paper: Büsing, Mohrschladt & Siedhoff (2022), *Pre-peak and post-peak momentum*, [SSRN 4298538](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=4298538).

### AI disclosure

This vignette was developed with assistance from Anthropic's Claude (model: Opus 4.7 and Sonnet 4.6). AI helped with code structure, prose drafting, and visualization choices. All analytical decisions and data interpretations are the author's responsibility.

## Row

```{r}
#| label: build-info
#| echo: false
#| results: asis
build_info()
```
