---
title: "Negative Results"
format:
  html:
    embed-resources: false
    theme:
      dark: darkly
      light: cosmo
    toc: true
    toc-depth: 2
    code-fold: true
    include-in-header:
      - text: |
          <style>
          body { font-size: 1.3em; }
          caption, figcaption, .figure-caption { color: #fff !important; }
          .negative-card {
            background: #1a1a2e; border: 1px solid #d73027; border-radius: 8px;
            padding: 16px; margin: 16px 0;
          }
          .negative-card h3 { color: #d73027; margin-top: 0; }
          .lesson { background: #111; border-left: 4px solid #e6a817; padding: 12px 16px; margin: 12px 0; }
          </style>
          <script>
          if (!localStorage.getItem('quarto-color-scheme'))
            document.documentElement.setAttribute('data-bs-theme', 'dark');
          </script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(dplyr)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)
```

## Why Document Negative Results?

Negative results are as valuable as positive ones. A strategy that fails the falsification framework teaches us:

- Which signals are **disguised beta** (market exposure dressed up as alpha)
- Which intuitions about markets are **wrong** (or at least not tradeable)
- Where the **boundary** lies between genuine alpha and noise

Publication bias in finance means failed strategies are rarely documented. This page corrects that.

## Failed Strategies

```{r}
#| label: neg-scorecard
summary <- safe_tar_read("fals_summary")
nms <- safe_tar_read("fals_vig_names")
if (!is.null(summary) && !is.null(nms)) {
  neg <- summary |>
    filter(ff_alpha_tstat < 2.0) |>
    mutate(
      Strategy = nms$long_name[match(strategy, nms$code_name)],
      `HAC t` = round(hac_tstat, 2),
      `Alpha t` = round(ff_alpha_tstat, 2),
      `R² (%)` = round(ff_r_squared * 100, 1),
      `Alpha (%)` = round(ff_alpha_annual * 100, 0),
      Diagnosis = case_when(
        ff_r_squared > 0.50 ~ "Pure market beta",
        ff_r_squared > 0.15 ~ "Significant factor exposure",
        TRUE ~ "Noise — neither alpha nor beta"
      )
    ) |>
    select(Strategy, `HAC t`, `Alpha t`, `R² (%)`, `Alpha (%)`, Diagnosis) |>
    arrange(`Alpha t`)

  hd_dt(neg, paste0(
    nrow(neg), " of ", nrow(summary),
    " strategies failed the falsification framework (Alpha t < 2.0). ",
    "These strategies' returns are explained by known factor exposure or noise."
  ))
}
```

## Avoid Worst Days (VIX Protection) {#avoid-worst}

::: {.negative-card}
### Verdict: Pure market beta (R² = 41%)

**What it does:** Exit to cash when yesterday's VIX > 30 or yesterday's absolute return > 3%. Re-enter when VIX < 25 and 5-day cooloff expires. All signals use t-1 (previous day) values.

**Why it fails:**

- The strategy is essentially a crude vol-timing filter on SPY
- When VIX is high, you're out of the market — but so is the VIX signal embedded in the market itself
- The 41% R² means the strategy's returns are 41% explained by Mkt-RF (market factor alone)
- Alpha = -1% per year: the strategy **destroys** value relative to just holding the market
- The t+1 execution constraint is critical: the pre-fix version (using same-day VIX) showed 18.4% CAGR; honest t+1 execution drops it to 5.3%

**Key lesson:** VIX is not a leading indicator — it's a concurrent indicator. By the time VIX > 30, the worst days have already happened.
:::

::: {.lesson}
**What we learned:** Any strategy that triggers on elevated VIX is timing exits AFTER the crash, not before it. The worst days and best days cluster together (median 5 calendar days apart). You cannot avoid one without missing the other. See the [temporal clustering analysis](avoid-worst-days.html).
:::

## Risk State Classification (VIX Overlay) {#risk-state}

::: {.negative-card}
### Verdict: Pure market beta (R² = 88%)

**What it does:** Three-signal design: VVIX (vol-of-vol) as early warning, VIX term structure 5-day change, VIX term structure level. Market-specific percentile thresholds from training data only. Overlay on SPY: reduce exposure when signals are elevated.

**Why it fails:**

- 88% R² — this is almost entirely market beta
- Alpha = -1% per year (negative): the overlay HURTS performance
- Despite using VVIX (which IS an early warning — it rises before VIX spikes), the signal doesn't translate into tradeable alpha
- The 5-day change in term structure adds noise, not signal
- Percentile thresholds from training data don't generalise: crisis regimes are structurally different from calm regimes

**Key lesson:** Adding more VIX-derived signals doesn't help. VVIX, term structure, and VIX level are all highly correlated. Three correlated signals ≈ one signal with more noise.
:::

::: {.lesson}
**What we learned:** Implied volatility signals (VIX, VVIX, term structure) are priced in. The market already incorporates this information. A strategy based on these signals is, by construction, replicating what the market already does — hence the 88% R² with the market factor.
:::

## LTR Cross-Sectional Momentum (Borderline) {#ltr}

::: {.negative-card}
### Verdict: Borderline (Alpha t = 1.99, R² = 7%)

**What it does:** XGBoost LambdaMART ranks ~500 US stocks monthly using 21 momentum/volatility features. Long top decile, short bottom decile.

**Why it's borderline:**

- Alpha t = 1.99 — just below the 2.0 threshold
- R² = 7% — low factor exposure, which is good
- Alpha = 166% per year — suspiciously high
- The demo uses ~51 stocks (not the full US universe) which inflates returns
- Walk-forward from 2005 with annual retraining — but limited out-of-sample validation

**Key lesson:** The signal may be genuine but the implementation is too small-scale to confirm. Needs: full US universe (3000+ stocks), transaction costs for monthly L/S rebalancing, and multi-market replication.
:::

::: {.lesson}
**What we learned:** Cross-sectional momentum is well-documented academically. Our implementation is a proof-of-concept, not a production signal. The 166% annual alpha is an artifact of the small universe, not evidence of a genuine edge.
:::

## Pattern: What Predicts Failure?

```{r}
#| label: neg-pattern
if (!is.null(summary) && !is.null(nms)) {
  pattern <- summary |>
    mutate(
      Strategy = nms$long_name[match(strategy, nms$code_name)],
      `Asset Class` = nms$asset_class[match(strategy, nms$code_name)],
      `Frequency` = nms$frequency[match(strategy, nms$code_name)],
      `R² (%)` = round(ff_r_squared * 100, 1),
      `Alpha t` = round(ff_alpha_tstat, 2),
      Verdict = ifelse(ff_alpha_tstat > 2.0, "Alpha", "No alpha")
    ) |>
    select(Strategy, `Asset Class`, Frequency, `R² (%)`, `Alpha t`, Verdict) |>
    arrange(desc(`Alpha t`))

  hd_dt(pattern, paste0(
    "All 5 strategies sorted by Alpha t. ",
    "The two overlay strategies (daily, VIX-based) both fail. ",
    "The two factor strategies (monthly, momentum-based) both pass. ",
    "Pattern: daily VIX-based overlays = beta; monthly factor rotation = alpha."
  ))
}
```

## Implications for Portfolio Construction

1. **Do not include** Avoid Worst Days or Risk State in the multi-strategy portfolio (#52)
2. **VIX-based overlays** add cost and complexity without alpha — use passive SPY instead
3. **Factor rotation** (DRIF, Factor MAX) is where the genuine alpha lives
4. **Cross-sectional momentum** (LTR) needs scaling before inclusion — promising but unconfirmed at production scale

## References

- Full falsification methodology: [Strategy Falsification](falsification.html)
- Avoid Worst Days analysis: [avoid-worst-days.html](avoid-worst-days.html)
- Strategy leaderboard: [leaderboard.html](leaderboard.html)
- Harvey, Liu & Zhu (2016). "...and the Cross-Section of Expected Returns"

```{r}
#| label: build-info
#| results: asis
build_info()
```
