---
title: "Prototype Results: AAPL + BTC"
format: html
execute:
  echo: true
  warning: false
---

## Overview

Prototype validating the historical data pipeline with two tickers
(AAPL equity, BTC crypto) from four sources (Kaggle + Yahoo, backfill + CoinGecko).

```{r}
#| label: setup
library(arrow)
library(dplyr)
```

## Cross-Reference Results

```{r}
#| label: xref
# Read from pipeline output
# xref <- arrow::read_ipc_file("pipeline-output/analysis/xref_report.arrow")
# knitr::kable(xref)
```

## Data Quality Summary

```{r}
#| label: quality
# equity <- arrow::read_ipc_file("pipeline-output/analysis/equity.arrow")
# crypto <- arrow::read_ipc_file("pipeline-output/analysis/crypto.arrow")
#
# tibble(
#   dataset = c("equity", "crypto"),
#   n_rows = c(nrow(equity), nrow(crypto)),
#   n_tickers = c(n_distinct(equity$ticker), n_distinct(crypto$ticker)),
#   date_from = c(min(equity$date), min(crypto$date)),
#   date_to = c(max(equity$date), max(crypto$date))
# ) |> knitr::kable()
```

## AAPL Stock Split Check (2020-08-31)

```{r}
#| label: split-check
# Check that adjusted prices are continuous across the 4:1 split
# aapl <- equity |> filter(ticker == "AAPL")
# split_window <- aapl |> filter(date >= "2020-08-25", date <= "2020-09-05")
# knitr::kable(split_window)
```

## Lessons Learned

<!-- Fill in after running the prototype -->

1. **Cross-reference discrepancies:** TBD
2. **Split/dividend handling:** TBD
3. **DuckDB httpfs latency:** TBD
4. **T sandbox friction:** TBD
5. **Schema decisions:** TBD
6. **Parquet sizing:** TBD
7. **Plan changes needed:** TBD
