---
title: "Real or Simulated?"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    embed-resources: false
    include-in-header:
      - text: |
          <link rel="stylesheet" href="vignette-shared.css">
          <script src="https://cdn.plot.ly/plotly-2.27.0.min.js"></script>
          <style>
          body { font-size: 1.5em; }
          .navbar .navbar-title, h1.title { font-size: 1.3em !important; }
          .nav-link { font-size: 0.85em !important; }
          .navbar, .quarto-dashboard-header, .navbar-dark,
          .bslib-page-dashboard > .navbar {
            background-color: black !important; color: white !important;
          }
          .navbar .navbar-brand, .navbar .nav-link, .navbar-text,
          .navbar .navbar-nav .nav-link { color: white !important; }
          .navbar .nav-link.active, .navbar .nav-link:hover {
            color: white !important; background-color: black !important;
            border: 1px solid white !important; border-radius: 4px;
          }
          .navbar .nav-link { color: #aaa !important; font-size: 0.9em; padding: 0.3em 0.6em !important; }
          .navbar-nav { flex-wrap: wrap !important; }
          .bslib-page-fill, .html-fill-container, .html-fill-item,
          .bslib-card, .card, .tab-pane, .tab-content, .card-body,
          .bslib-grid, .bslib-grid-item {
            height: auto !important; min-height: 0 !important;
            max-height: none !important; overflow: visible !important;
            flex: none !important;
          }
          #quiz-container { padding: 0.5em; }
          .quiz-charts { display: flex; gap: 1em; margin-bottom: 1em; }
          .quiz-charts > div { flex: 1; min-height: 300px; }
          .quiz-btn { padding: 0.8em 2em; margin: 0.3em; font-size: 1.1em;
            border: 2px solid #666; border-radius: 8px; cursor: pointer;
            background: #222; color: #ddd; transition: all 0.2s; }
          .quiz-btn:hover { border-color: #4a90d9; color: white; }
          .quiz-btn.correct { background: #1a9850 !important; border-color: #1a9850 !important; }
          .quiz-btn.wrong { background: #d73027 !important; border-color: #d73027 !important; }
          .quiz-btn:disabled { opacity: 0.5; cursor: default; }
          .quiz-score { font-size: 1.2em; color: #6c9bd2; margin: 0.5em 0; }
          .quiz-info { color: #aaa; font-size: 0.9em; margin: 0.5em 0; }
          .quiz-reveal { background: #111; border: 1px solid #444; border-radius: 8px;
            padding: 1em; margin: 1em 0; display: none; }
          .quiz-reveal.show { display: block; }
          .quiz-length-selector { margin: 0.5em 0; }
          .quiz-length-selector label { color: #aaa; margin-right: 0.5em; }
          .quiz-length-selector select { background: #222; color: #ddd; border: 1px solid #555;
            padding: 0.3em 0.5em; border-radius: 4px; }
          .quiz-difficulty { display: inline-block; padding: 0.2em 0.6em; border-radius: 4px;
            font-size: 0.8em; font-weight: bold; }
          .quiz-difficulty.easy { background: #1a9850; color: white; }
          .quiz-difficulty.medium { background: #e6a817; color: black; }
          .quiz-difficulty.hard { background: #d73027; color: white; }
          .quiz-difficulty.pseudo { background: #7b3294; color: white; }
          </style>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)
```

# Quiz

## Row

Can you tell real financial returns from simulated ones? Each round shows two time series — one is real strategy returns, the other is generated from a statistical null model. Guess which is real. Shorter visible windows earn bonus points.

### {.tabset}

#### Play

```{r}
#| label: quiz-data
#| results: asis
quiz_json <- safe_tar_read("quiz_json")
cat('<div id="quiz-container">\n')
cat('  <div class="quiz-score">Score: <span id="score">0</span> / <span id="total">0</span> points | Round <span id="round-num">1</span> / <span id="round-total">20</span></div>\n')
cat('  <div class="quiz-info"><span id="difficulty-badge"></span> <span id="round-info"></span></div>\n')
cat('  <div class="quiz-length-selector">\n')
cat('    <label>Visible length:</label>\n')
cat('    <select id="length-select" onchange="updateLength()">\n')
cat('      <option value="100">100 points (1x)</option>\n')
cat('      <option value="50">50 points (1.5x)</option>\n')
cat('      <option value="25">25 points (2x)</option>\n')
cat('      <option value="15">15 points (3x)</option>\n')
cat('    </select>\n')
cat('  </div>\n')
cat('  <div class="quiz-charts"><div id="chart-a"></div><div id="chart-b"></div></div>\n')
cat('  <div style="text-align:center;">\n')
cat('    <button class="quiz-btn" id="btn-a" onclick="guess(\'A\')">A is Real</button>\n')
cat('    <button class="quiz-btn" id="btn-b" onclick="guess(\'B\')">B is Real</button>\n')
cat('    <button class="quiz-btn" id="btn-next" onclick="nextRound()" style="display:none;">Next Round &rarr;</button>\n')
cat('  </div>\n')
cat('  <div class="quiz-reveal" id="reveal"></div>\n')
cat('</div>\n')

# Embed quiz data as JSON
cat('<script id="quiz-data-json" type="application/json">\n')
if (!is.null(quiz_json)) cat(quiz_json)
cat('\n</script>\n')

# Quiz logic
cat('<script>\n')
cat('
var quizData = JSON.parse(document.getElementById("quiz-data-json").textContent || "[]");
var currentRound = 0;
var score = 0;
var totalPoints = 0;
var answered = false;
var lengthMultipliers = {100: 1, 50: 1.5, 25: 2, 15: 3};
var diffMultipliers = {easy: 1, medium: 2, hard: 3, pseudo: 4};

function renderRound() {
  if (currentRound >= quizData.length) { showFinal(); return; }
  var r = quizData[currentRound];
  answered = false;
  document.getElementById("round-num").textContent = currentRound + 1;
  document.getElementById("round-total").textContent = quizData.length;
  document.getElementById("btn-a").disabled = false;
  document.getElementById("btn-b").disabled = false;
  document.getElementById("btn-a").className = "quiz-btn";
  document.getElementById("btn-b").className = "quiz-btn";
  document.getElementById("btn-next").style.display = "none";
  document.getElementById("reveal").className = "quiz-reveal";
  document.getElementById("reveal").innerHTML = "";

  var badge = document.getElementById("difficulty-badge");
  badge.className = "quiz-difficulty " + r.difficulty;
  badge.textContent = r.difficulty.charAt(0).toUpperCase() + r.difficulty.slice(1);
  document.getElementById("round-info").textContent = "(" + diffMultipliers[r.difficulty] + "x difficulty)";

  updateCharts();
}

function updateCharts() {
  var r = quizData[currentRound];
  var len = parseInt(document.getElementById("length-select").value);
  var a = r.series_a.slice(-len);
  var b = r.series_b.slice(-len);
  // Convert returns to cumulative growth
  var cumA = [1], cumB = [1];
  for (var i = 0; i < a.length; i++) { cumA.push(cumA[i] * (1 + a[i])); }
  for (var i = 0; i < b.length; i++) { cumB.push(cumB[i] * (1 + b[i])); }

  var layout = {
    paper_bgcolor: "#000", plot_bgcolor: "#000",
    font: {color: "#e0e0e0"}, margin: {t: 30, b: 30, l: 40, r: 10},
    xaxis: {gridcolor: "#333", title: "Observation"},
    yaxis: {gridcolor: "#333", title: "Growth of $1"}
  };
  Plotly.react("chart-a", [{y: cumA, type: "scatter", mode: "lines",
    line: {color: "#4a90d9", width: 2}, name: "Series A"}],
    Object.assign({}, layout, {title: {text: "Series A", font: {size: 14, color: "#e0e0e0"}}}),
    {scrollZoom: true, displayModeBar: false});
  Plotly.react("chart-b", [{y: cumB, type: "scatter", mode: "lines",
    line: {color: "#e6a817", width: 2}, name: "Series B"}],
    Object.assign({}, layout, {title: {text: "Series B", font: {size: 14, color: "#e0e0e0"}}}),
    {scrollZoom: true, displayModeBar: false});
}

function updateLength() { if (!answered) updateCharts(); }

function guess(choice) {
  if (answered) return;
  answered = true;
  var r = quizData[currentRound];
  var correct = (choice === r.answer);
  var len = parseInt(document.getElementById("length-select").value);
  var pts = diffMultipliers[r.difficulty] * lengthMultipliers[len];
  if (correct) score += pts;
  totalPoints += pts;

  document.getElementById("score").textContent = score;
  document.getElementById("total").textContent = totalPoints;
  document.getElementById("btn-a").disabled = true;
  document.getElementById("btn-b").disabled = true;
  if (r.answer === "A") {
    document.getElementById("btn-a").className = "quiz-btn correct";
    if (choice === "B") document.getElementById("btn-b").className = "quiz-btn wrong";
  } else {
    document.getElementById("btn-b").className = "quiz-btn correct";
    if (choice === "A") document.getElementById("btn-a").className = "quiz-btn wrong";
  }
  document.getElementById("btn-next").style.display = "inline-block";

  // Show reveal with full series
  var reveal = document.getElementById("reveal");
  reveal.className = "quiz-reveal show";
  var verdict = correct ? "<span style=\\"color:#1a9850;font-weight:bold\\">Correct!</span>"
                        : "<span style=\\"color:#d73027;font-weight:bold\\">Wrong.</span>";
  reveal.innerHTML = verdict + " " + r.answer + " was real.<br>" +
    "<b>Real:</b> " + r.real_source + "<br>" +
    "<b>Simulated:</b> " + r.sim_source + " (" + r.difficulty + ")<br>" +
    "<b>Points:</b> " + (correct ? "+" + pts : "0") + " (difficulty " +
    diffMultipliers[r.difficulty] + "x, length " + lengthMultipliers[len] + "x)" +
    "<div id=\\"reveal-chart\\" style=\\"height:250px;margin-top:0.5em;\\"></div>";

  // Full reveal chart with both series labelled
  var cumReal = [1], cumSim = [1];
  for (var i = 0; i < r.full_real.length; i++) { cumReal.push(cumReal[i] * (1 + r.full_real[i])); }
  for (var i = 0; i < r.full_sim.length; i++) { cumSim.push(cumSim[i] * (1 + r.full_sim[i])); }
  var revLayout = {
    paper_bgcolor: "#111", plot_bgcolor: "#111",
    font: {color: "#e0e0e0"}, margin: {t: 30, b: 30, l: 40, r: 10},
    xaxis: {gridcolor: "#333"}, yaxis: {gridcolor: "#333", title: "Growth of $1"},
    legend: {orientation: "h", x: 0.5, xanchor: "center", y: -0.15}
  };
  Plotly.react("reveal-chart", [
    {y: cumReal, type: "scatter", mode: "lines", line: {color: "#2ecc71", width: 2}, name: "Real: " + r.real_source},
    {y: cumSim, type: "scatter", mode: "lines", line: {color: "#e74c3c", width: 2, dash: "dash"}, name: "Sim: " + r.sim_source}
  ], revLayout, {displayModeBar: false});
}

function nextRound() {
  currentRound++;
  renderRound();
}

function showFinal() {
  var container = document.getElementById("quiz-container");
  var pct = totalPoints > 0 ? Math.round(score / totalPoints * 100) : 0;
  container.innerHTML = "<h2 style=\\"color:#6c9bd2\\">Quiz Complete</h2>" +
    "<p style=\\"font-size:1.3em\\">Final score: <b>" + score + " / " + totalPoints + "</b> points (" + pct + "%)</p>" +
    "<p>You completed " + quizData.length + " rounds across 4 difficulty levels.</p>" +
    "<p><a href=\\"falsification.html\\">See the full falsification methodology</a></p>" +
    "<button class=\\"quiz-btn\\" onclick=\\"location.reload()\\">Play Again</button>";
}

// Arrow key navigation
document.addEventListener("keydown", function(e) {
  if (e.key === "ArrowRight" && answered) nextRound();
});

// Start
if (quizData.length > 0) renderRound();
else document.getElementById("quiz-container").innerHTML = "<p>Quiz data not available. Run tar_make() first.</p>";
')
cat('</script>\n')
```

#### About

**Can you tell real from simulated?**

This quiz tests your ability to distinguish genuine financial time series from statistically generated ones. Each round presents two cumulative return series — one real, one simulated — and asks you to identify the real one.

**Difficulty levels:**

- **Easy** (1x): Simulated from White Noise (iid Gaussian returns). No volatility clustering, no structure. Should be distinguishable by the lack of realistic patterns.
- **Medium** (2x): Simulated from [GARCH(1,1)](https://en.wikipedia.org/wiki/Autoregressive_conditional_heteroskedasticity). Has volatility clustering — much harder to distinguish from real data.
- **Hard** (3x): Simulated from [GJR-GARCH](https://en.wikipedia.org/wiki/GARCH#GJR-GARCH) with asymmetric leverage. Negative returns increase future volatility more than positive returns — closely mimics real market dynamics.
- **Pseudo** (4x): The "simulated" series is actually the real series played backwards. Tests whether you can detect the temporal asymmetry of real markets (crashes are fast, recoveries are slow).

**Length bonus:**

Seeing fewer data points makes the task harder. Select a shorter window for bonus multipliers: 50 points (1.5x), 25 points (2x), or 15 points (3x).

**Why this matters:**

If you cannot distinguish a strategy's returns from a null model, neither can a statistical test. The [falsification framework](falsification.html) formalises this intuition with 6 null environments and 100 simulations each. This quiz lets you experience the same challenge visually.

**Data source:** Real returns from 5 strategies in the [strategy leaderboard](leaderboard.html). Simulated returns from [`hd_null_env_*()`](https://github.com/JohnGavin/historical/blob/main/packages/historicaldata/R/falsification.R) null generators.

#### Build Info

```{r}
#| label: build-info
#| results: asis
build_info()
```
