---
title: "Stock-Level Backtests"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    embed-resources: false
    include-in-header:
      - text: |
          <link rel="stylesheet" href="vignette-shared.css">
          <script src="vignette-shared.js"></script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)
library(dplyr)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)
```

# Comparison

## Row

### {.tabset}

#### Summary

Two signals applied at two levels, producing four strategies:

| | Signal | Stock-Level (660 stocks, D1-D10) | Factor-Level (5 FF factors, top 2) |
|--|--------|----------------------------------|-------------------------------------|
| **MAX** | Highest single-day return in prior month | [Stock MAX](#stock-max) | [Factor MAX](factor-max.html) |
| **DRIF** | Elastic net on 42 daily return features | [Stock DRIF](#stock-drif) | [Factor DRIF](drif.html) |

**Stock-level:** Each month, rank all 660 stocks by signal. Long top decile (~38 stocks), short bottom decile. Monthly rebalance. Equal weight within each decile.

**Factor-level:** Each month, rank 5 Fama-French factors (HML, SMB, RMW, CMA, Mom) by signal. Long top 2. Monthly rebalance. Equal weight.

#### Equity Curves

```{r}
#| fig-height: 6
safe_tar_read("stk_all_comparison_plot")
```

::: {.fig-caption}
**Equity curves: 4 strategies, growth of $1 (log scale).** These curves should be interpreted with caution: (1) stock-level strategies have ~44% annual volatility, (2) a 150x return over 20 years requires ~30% CAGR which is unstable, (3) the stock MAX curve's spike is driven by a few extreme months. Dashed line = test period start (2020). Source: HuggingFace equity_daily + FF5/Momentum.
:::

#### Metrics

```{r}
stk_max <- safe_tar_read("stk_max_metrics")
stk_drif <- safe_tar_read("stk_drif_metrics")
fac_max <- safe_tar_read("fm_metrics")
fac_drif <- safe_tar_read("drif_metrics")

if (!is.null(stk_max) && !is.null(stk_drif) && !is.null(fac_max) && !is.null(fac_drif)) {
  make_row <- function(m, name) {
    m |> mutate(
      Strategy = name,
      CAGR = paste0(round(cagr * 100, 1), "%"),
      Vol = paste0(round(vol * 100, 1), "%"),
      Sharpe = round(sharpe, 2),
      `Max DD` = paste0(round(max_dd * 100, 1), "%")
    ) |> select(Strategy, Period = period, CAGR, Vol, Sharpe, `Max DD`)
  }

  all_metrics <- bind_rows(
    make_row(stk_max, "Stock MAX"),
    make_row(stk_drif, "Stock DRIF"),
    make_row(fac_max, "Factor MAX"),
    make_row(fac_drif, "Factor DRIF")
  )

  hd_dt(all_metrics, "All strategies by partition (Training / Testing / Validation / Full). Factor-level links: Factor MAX, Factor DRIF.")
}
```

#### Key Findings

**P&L**

| Finding | Evidence |
|---------|----------|
| Stock MAX has the highest raw CAGR (12.1%) | But with 44% vol — risk-adjusted Sharpe is modest (0.23) |
| Factor DRIF has the best full-period Sharpe (0.26) | Stable signal from elastic net on 6 academic factors |
| Stock DRIF full-period CAGR is only 1.3% | Elastic net regularisation is very aggressive at stock level |

**Out-of-Sample / Validation**

| Finding | Evidence |
|---------|----------|
| Stock DRIF has the best OOS Sharpe (0.79) | Elastic net generalises well — low IS Sharpe but high OOS |
| Factor MAX degrades OOS (Sharpe -0.36) | Single-signal approach overfits at factor level |
| Stock MAX OOS looks extreme (45% CAGR) | Driven by COVID recovery + meme stocks — not stable |

**Risk**

| Finding | Evidence |
|---------|----------|
| Stock-level strategies have 2-5x higher vol than factor-level | 44% vol (stock MAX) vs 9% (factor MAX) |
| Stock MAX max drawdown is -74% | Unacceptable for most allocators |
| Factor strategies have drawdowns <36% | Lower risk but lower return |

**Robustness**

| Finding | Evidence |
|---------|----------|
| DRIF > MAX at both levels (Sharpe, OOS) | Full daily return distribution captures more signal than single extreme |
| Factor-level is more stable, stock-level has more upside | Classic diversification vs concentration tradeoff |

# Stock MAX

## Row

### {.tabset}

#### Definition

::: {.callout-note}
## Strategy: Stock-Level Factor MAX

**What:** Each month, sort all stocks by the maximum daily return observed in the prior month. Go long the top decile, short the bottom decile.

**Rationale (Market Dynamics):** A stock with an unusually high single-day return signals that new information arrived — an earnings surprise, an upgrade, a sector rotation. Markets underreact to this information at the individual stock level because:

1. **Attention limitation** — investors can't monitor 660 stocks daily; extreme returns in small/mid caps go unnoticed
2. **Anchoring** — after a large daily move, investors anchor to the pre-move price, underweighting the new information
3. **Liquidity friction** — the initial move is partially driven by liquidity demand, not fully by information. The price continues to adjust over the following month.

**Step-by-step example (one month):**

1. Collect daily returns for all 660 stocks in March 2024
2. For each stock, find the single highest daily return in March (the "MAX")
3. Rank all stocks by MAX: stock with the biggest single-day gain ranks #1
4. Divide into 10 deciles (~66 stocks each). D1 = highest MAX, D10 = lowest
5. April 2024: go long D1, short D10, equal weight within each decile
6. End of April: close all positions, repeat
:::

#### Performance

```{r}
#| fig-height: 6
safe_tar_read("stk_max_plot")
```

::: {.fig-caption}
**Stock-level MAX: D1 long / D10 short, ~38 stocks per decile.** Blue = long-short spread, red = long-only (D1). Signal: maximum daily return in prior month. Source: 660 stocks from HuggingFace equity_daily.
:::

#### Metrics

```{r}
metrics <- safe_tar_read("stk_max_metrics")
if (!is.null(metrics)) {
  display <- metrics |>
    mutate(across(c(cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2),
           avg_long = round(avg_long), avg_short = round(avg_short))
  hd_dt_wide(display, "Stock MAX metrics by partition.")
}
```

#### vs Factor MAX

```{r}
#| fig-height: 6
safe_tar_read("stk_max_vs_factor_plot")
```

::: {.fig-caption}
**Stock vs Factor MAX.** Stock-level exploits cross-sectional dispersion across 660 stocks. Factor-level rotates among 5 academic factors. Same MAX signal, different universes. Factor-level definition: [Factor MAX](factor-max.html).
:::

#### Pros & Cons

| | Detail |
|--|--------|
| **Pros** | Simple signal (one number per stock-month). High raw returns. Works across markets. |
| **Cons** | Very high volatility (44%). Max drawdown -74%. Sensitive to extreme events. High turnover (~monthly full rebalance). |
| **Works well** | High-volatility regimes (VIX > 25). Small/mid-cap stocks with less analyst coverage. |
| **Works badly** | Low-volatility, trending markets where MAX signals are noise. |
| **Factor exposure** | Loads on size (small-cap tilt in D1), volatility, and short-term reversal. Partially explained by these factors — not pure alpha. |

# Stock DRIF

## Row

### {.tabset}

#### Definition

::: {.callout-note}
## Strategy: Stock-Level DRIF (Daily Return Information Factor)

**What:** Each month, use an elastic net regression on the past 21 daily returns (both in chronological order and ranked by magnitude — 42 features total) to predict next-month stock returns. Go long the top decile by predicted return, short the bottom decile.

**Rationale (Market Dynamics):** The full sequence of daily returns contains more predictive information than any single summary statistic (mean, max, volatility). Specifically:

1. **Chronological features** (c1...c21) capture when returns occurred — recent price pressure and liquidity effects carry more weight than older ones
2. **Rank features** (r1...r21) capture how extreme returns were — behavioral biases cause investors to overweight tail events
3. **Elastic net** combines these 42 features with regularisation, avoiding overfitting while capturing non-obvious interactions

**Step-by-step example (one month):**

1. For each of 660 stocks, collect the 21 daily returns in March 2024
2. Create 42 features: c1...c21 (chronological order) + r1...r21 (sorted by magnitude)
3. Train elastic net (alpha=0.5, 5-fold CV) on all prior stock-months (expanding window, 60-month minimum)
4. Predict each stock's April return from its March features
5. Rank stocks by predicted return. D1 = highest predicted, D10 = lowest
6. April: long D1, short D10, equal weight
7. End of April: close, repeat

**Key distinction from MAX:** MAX uses 1 feature (the single highest daily return). DRIF uses 42 features (the full daily return distribution). The Alpha Architect paper shows DRIF subsumes MAX — MAX is a special case of DRIF.
:::

#### Performance

```{r}
#| fig-height: 6
safe_tar_read("stk_drif_plot")
```

::: {.fig-caption}
**Stock-level DRIF: pooled elastic net, D1/D10 decile spread.** Blue = long-short, red = long-only (D1). Expanding window with 60-month minimum training. Source: 660 stocks, 42 features per stock-month.
:::

#### Metrics

```{r}
metrics <- safe_tar_read("stk_drif_metrics")
if (!is.null(metrics)) {
  display <- metrics |>
    mutate(across(c(cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2),
           avg_long = round(avg_long), avg_short = round(avg_short))
  hd_dt_wide(display, "Stock DRIF metrics by partition.")
}
```

#### Pros & Cons

| | Detail |
|--|--------|
| **Pros** | Uses full daily return information (42 features). Best OOS Sharpe (0.79). Elastic net prevents overfitting. |
| **Cons** | Weak in-sample (1.3% CAGR). Compute-intensive (10 min for 660 stocks × 600 months). High turnover. |
| **Works well** | When the cross-section is large (more stocks = better elastic net training). Post-2020 period. |
| **Works badly** | Small universes (<100 stocks). When feature matrix is rank-deficient. |
| **Factor exposure** | Lower factor loading than MAX (elastic net regularises out common factors). More likely to represent genuine alpha. |
| **OOS > IS anomaly** | Possible explanations: universe composition change post-2010, elastic net needs many years to learn, or OOS window (62 months) is too short to be conclusive. |

#### References

1. [DRIF: factor-level implementation](drif.html) — same methodology on 5 FF factors
2. [Factor MAX](factor-max.html) — simpler signal, comparison benchmark
3. [Alpha Architect: Daily Stock Returns](https://alphaarchitect.com/daily-stock-returns/) — original research
4. [Alpha Architect: Factor MAX](https://alphaarchitect.com/factor-max/) — MAX signal research

# ETF Replication

## Row

### {.tabset}

#### Definition

::: {.callout-note}
## Why ETFs?

Stock-level long-short strategies lose ~20%/yr after realistic costs (1.85%/month for 660-stock decile sort). Factor ETFs (VLUE, MTUM, QUAL, etc.) offer institutional liquidity at ~0.20%/month — 10x cheaper.

**Two approaches tested:**

| | A: DRIF on ETF Returns | B: Academic Signal → ETFs |
|--|----------------------|--------------------------|
| **Signal** | Elastic net on ETF daily prices | Factor-level DRIF (FF5+Mom) mapped to ETFs |
| **Execution** | Trade the predicted ETFs directly | Map HML→VLUE, Mom→MTUM, etc. |
| **Training data** | 2013-2026 (ETF inception) | 1963-2026 (academic factors) |
| **Key risk** | Short history (10 ETFs × 10 years) | Weak mapping (VLUE/HML cor = 0.34) |

**Result:** Long-short fails for both. **Long-only works:** 13.4% CAGR (A) and 12.6% CAGR (B). The signal identifies winners, not losers.
:::

#### Equity Curves

```{r}
#| fig-height: 6
safe_tar_read("etf_comparison_plot")
```

::: {.fig-caption}
**ETF replication: Approach A vs B (long-short, net of costs).** Both approaches struggle as long-short because all factor ETFs have positive market beta — shorting one while longing another is mostly a bet on relative factor performance. The long-only legs (not shown) return 13%+ CAGR.
:::

#### Approach A Metrics

```{r}
metrics_a <- safe_tar_read("etf_a_metrics")
if (!is.null(metrics_a)) {
  display <- metrics_a |>
    mutate(across(c(cagr, long_cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2))
  hd_dt_wide(display, "Approach A: DRIF applied directly to 9 factor ETF daily returns. Long-short (L/S) and long-only metrics. Cost: 0.10%/trade, 0.5%/yr borrow.")
}
```

#### Approach B Metrics

```{r}
metrics_b <- safe_tar_read("etf_b_metrics")
if (!is.null(metrics_b)) {
  display <- metrics_b |>
    mutate(across(c(cagr, long_cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2))
  hd_dt_wide(display, "Approach B: Academic factor DRIF signal mapped to ETFs via ETF_FACTOR_MAP. Mapping: HML→VLUE/VTV/IWD, Mom→MTUM, RMW→QUAL, etc.")
}
```

#### Key Findings

| Finding | Evidence |
|---------|----------|
| **Long-only is the practical strategy** | A long-only: 13.4% CAGR. B long-only: 12.6%. Compare S&P 500 ~10%. |
| **Short leg is unprofitable** | A long-short: -0.8%. B long-short: -1.9%. ETFs all have market beta — shorting one is mostly shorting the market. |
| **A > B** | Approach A (direct ETF signal) has higher Sharpe (-0.11 vs -0.31) and lower DD (-23.7% vs -29.0%). The VLUE/HML mapping (cor=0.34) in B adds noise. |
| **Cost advantage is real** | ETF costs ~0.20%/month vs 1.85%/month stock-level. The long-only gross returns survive after costs. |
| **Validation is strong** | Both show ~20% long-only CAGR in 2023-2026. Caveat: overlaps a bull market. |

# Details

## Row

### {.tabset}

#### Universe

660 individual stocks from HuggingFace `equity_daily`:

| Group | Tickers | History |
|-------|--------:|---------|
| S&P 500 (US) | ~500 | 2000-2026 (most), 1962+ (blue chips) |
| STOXX 600 majors (EU) | ~143 | 2000-2026 |
| Factor ETFs | 9 | 2013-2026 |
| **Total** | **~660** | |

Minimum 252 trading days (1 year) to enter universe. Monthly rebalance. ~38 stocks per decile.

#### Partitions

```{r}
params <- safe_tar_read("stk_params")
if (!is.null(params)) {
  param_df <- tibble(
    Parameter = c("Training", "Testing", "Validation",
                  "Min History", "Lookback Days", "Deciles"),
    Value = c(
      paste(params$start_date, "to", params$is_end),
      paste(params$test_start, "to", params$test_end),
      paste(params$val_start, "to", params$val_end),
      paste(params$min_history_days, "days"),
      as.character(params$lookback_days),
      as.character(params$n_deciles)
    )
  )
  hd_dt(param_df, "Shared parameters. Training: model fitting. Testing: calibration. Validation: sealed one-shot evaluation.")
}
```

#### Build Info

```{r}
#| label: build-info
#| echo: false
#| results: asis
build_info()
```
