---
title: "Stock-Level Backtests"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    embed-resources: false
    include-in-header:
      - text: |
          <link rel="stylesheet" href="vignette-shared.css">
          <script src="vignette-shared.js"></script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)
library(dplyr)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)

disc_path <- if (file.exists("../R/disclosures.R")) "../R/disclosures.R" else
  file.path(here::here(), "R/disclosures.R")
source(disc_path)
```

::: {.callout-warning collapse="false" title="Survivorship bias"}
`r disclosure_survivorship()`
:::

# Comparison

## Row

### {.tabset}

#### Summary

Two signals applied at two levels, producing four strategies:

| | Signal | Stock-Level (660 stocks, D1-D10) | Factor-Level (5 FF factors, top 2) |
|--|--------|----------------------------------|-------------------------------------|
| **MAX** | Highest single-day return in prior month | [Stock MAX](#stock-max) | [Factor MAX](factor-max.html) |
| **<abbr title="Dynamic Rotation Into Factors — selects best-performing FF factor each month">DRIF</abbr>** | Elastic net on 42 daily return features | [Stock DRIF](#stock-drif) | [Factor DRIF](drif.html) |

**Stock-level:** Each month, rank all 660 stocks by signal. Long top decile (~38 stocks), short bottom decile. Monthly rebalance. Equal weight within each decile.

**Factor-level:** Each month, rank 5 Fama-French factors (<abbr title="High Minus Low — value factor (cheap stocks minus expensive stocks)">HML</abbr>, <abbr title="Small Minus Big — size factor (small-cap minus large-cap returns)">SMB</abbr>, <abbr title="Robust Minus Weak — profitability factor">RMW</abbr>, <abbr title="Conservative Minus Aggressive — investment factor">CMA</abbr>, <abbr title="Momentum — past 12-month winners minus losers">Mom</abbr>) by signal. Long top 2. Monthly rebalance. Equal weight.

#### Equity Curves

```{r}
#| fig-height: 6
safe_tar_read("stk_all_comparison_plot")
```

::: {.fig-caption}
```{r}
#| echo: false
#| results: asis
cap <- safe_tar_read("stk_all_caption")
if (!is.null(cap)) cat(cap) else cat("*Caption not available — run tar_make() to generate.*")
```
:::

#### Metrics

```{r}
stk_max <- safe_tar_read("stk_max_metrics")
stk_drif <- safe_tar_read("stk_drif_metrics")
fac_max <- safe_tar_read("fm_metrics")
fac_drif <- safe_tar_read("drif_metrics")

if (!is.null(stk_max) && !is.null(stk_drif) && !is.null(fac_max) && !is.null(fac_drif)) {
  make_row <- function(m, name) {
    m |> mutate(
      Strategy = name,
      CAGR = paste0(round(cagr * 100, 1), "%"),
      Vol = paste0(round(vol * 100, 1), "%"),
      Sharpe = round(sharpe, 2),
      `Max DD` = paste0(round(max_dd * 100, 1), "%")
    ) |> select(Strategy, Period = period, CAGR, Vol, Sharpe, `Max DD`)
  }

  all_metrics <- bind_rows(
    make_row(stk_max, "Stock MAX"),
    make_row(stk_drif, "Stock DRIF"),
    make_row(fac_max, "Factor MAX"),
    make_row(fac_drif, "Factor DRIF")
  )

  hd_dt(all_metrics, paste0(
    "All strategies by partition (Training / Testing / Validation / Full). ",
    "Factor-level links: Factor MAX, Factor DRIF. ",
    "Stock MAX and Stock DRIF draw on `stk_universe` (survivorship-biased — see issue #150)."
  ))
}
```

#### Key Findings

**P&L**

| Finding | Evidence |
|---------|----------|
| Stock MAX full-period CAGR: `r { m <- safe_tar_read("stk_max_metrics"); v <- if(!is.null(m)) m$cagr[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,1),"%") else "N/A" }` | Vol: `r { m <- safe_tar_read("stk_max_metrics"); v <- if(!is.null(m)) m$vol[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,0),"%") else "N/A" }` — full-period Sharpe: `r { m <- safe_tar_read("stk_max_metrics"); v <- if(!is.null(m)) m$sharpe[m$period=="Full Period"] else numeric(0); if(length(v)==1L) round(v,2) else "N/A" }` |
| Factor DRIF full-period Sharpe: `r { m <- safe_tar_read("drif_metrics"); v <- if(!is.null(m)) m$sharpe[m$period=="Full Period"] else numeric(0); if(length(v)==1L) round(v,2) else "N/A" }` | Full-period covers the entire sample including Training. Validation Sharpe: `r { m <- safe_tar_read("drif_metrics"); v <- if(!is.null(m)) m$sharpe[m$period=="Validation"] else numeric(0); if(length(v)==1L) round(v,2) else "N/A" }` |
| Stock DRIF full-period CAGR is only `r { m <- safe_tar_read("stk_drif_metrics"); if(!is.null(m)) paste0(round(m$cagr[m$period=="Full Period"]*100,1),"%") else "N/A" }` | Elastic net regularisation is very aggressive at stock level |

**Out-of-Sample / Validation**

::: {.callout-warning}
**Current Validation Sharpe (sealed partition):** `r { get_val <- function(tgt, col, per) { m <- safe_tar_read(tgt); v <- if (!is.null(m)) m[[col]][m$period == per] else numeric(0); if (length(v) == 1L) round(v, 2) else "N/A" }; paste0("Stock DRIF: ", get_val("stk_drif_metrics","sharpe","Validation"), " | Stock MAX: ", get_val("stk_max_metrics","sharpe","Validation"), " | Factor MAX: ", get_val("fm_metrics","sharpe","Validation"), " | Factor DRIF: ", get_val("drif_metrics","sharpe","Validation")) }`. Validation Sharpe values are shown above for each strategy. The Testing partition OOS results (below) reflect the calibration window used during strategy development, not the sealed evaluation.
:::

| Finding | Evidence |
|---------|----------|
| Stock DRIF showed the strongest Testing-period Sharpe during elastic net development (Testing Sharpe: `r { m <- safe_tar_read("stk_drif_metrics"); if(!is.null(m)) round(m$sharpe[m$period=="Testing"],2) else "N/A" }`) | Validation Sharpe is now `r { m <- safe_tar_read("stk_drif_metrics"); v <- if(!is.null(m)) m$sharpe[m$period=="Validation"] else numeric(0); if(length(v)==1L) round(v,2) else "N/A" }` — the elastic net that appeared to generalise in Testing has not held in the sealed Validation period |
| Factor MAX Testing-period Sharpe: `r { m <- safe_tar_read("fm_metrics"); if(!is.null(m)) round(m$sharpe[m$period=="Testing"],2) else "N/A" }` | Single-signal approach at factor level; Validation Sharpe: `r { m <- safe_tar_read("fm_metrics"); v <- if(!is.null(m)) m$sharpe[m$period=="Validation"] else numeric(0); if(length(v)==1L) round(v,2) else "N/A" }` |
| Stock MAX OOS CAGR looks extreme (`r { m <- safe_tar_read("stk_max_metrics"); if(!is.null(m)) paste0(round(m$cagr[m$period=="Testing"]*100,0),"%") else "N/A" }`) | Driven by COVID recovery + meme stocks — Validation CAGR: `r { m <- safe_tar_read("stk_max_metrics"); v <- if(!is.null(m)) m$cagr[m$period=="Validation"] else numeric(0); if(length(v)==1L) paste0(round(v*100,1),"%") else "N/A" }` |

**Risk**

| Finding | Evidence |
|---------|----------|
| Stock-level vol (stock MAX: `r { m <- safe_tar_read("stk_max_metrics"); v <- if(!is.null(m)) m$vol[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,0),"%") else "N/A" }`) vs factor-level vol (factor MAX: `r { m <- safe_tar_read("fm_metrics"); v <- if(!is.null(m)) m$vol[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,0),"%") else "N/A" }`) | Compare the two values above — see metrics table for all strategies |
| Stock MAX max drawdown is `r { m <- safe_tar_read("stk_max_metrics"); v <- if(!is.null(m)) m$max_dd[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,0),"%") else "N/A" }` | Unacceptable for most allocators |
| Factor strategies have drawdowns `r { m <- safe_tar_read("fm_metrics"); v <- if(!is.null(m)) m$max_dd[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0("<",round(abs(v)*100,0),"%") else "N/A" }` | Lower risk but lower return |

**Robustness**

| Finding | Evidence |
|---------|----------|
| Full daily return distribution (42 features) carries more signal than single-day MAX on paper; whether this advantage persists in the Validation period requires checking the current metrics table | Factor-level Validation Sharpe: `r { m <- safe_tar_read("drif_metrics"); v <- if(!is.null(m)) m$sharpe[m$period=="Validation"] else numeric(0); if(length(v)==1L) round(v,2) else "N/A" }` vs Factor MAX Validation Sharpe: `r { m <- safe_tar_read("fm_metrics"); v <- if(!is.null(m)) m$sharpe[m$period=="Validation"] else numeric(0); if(length(v)==1L) round(v,2) else "N/A" }` |
| Factor-level shows lower volatility than stock-level across all periods | Classic diversification vs concentration tradeoff — confirmed by current metrics table |

# Stock MAX

## Row

### {.tabset}

#### Definition

::: {.callout-note}
## Strategy: Stock-Level Factor MAX

**What:** Each month, sort all stocks by the maximum daily return observed in the prior month. Go long the top decile, short the bottom decile.

**Rationale (Market Dynamics):** A stock with an unusually high single-day return signals that new information arrived — an earnings surprise, an upgrade, a sector rotation. Markets underreact to this information at the individual stock level because:

1. **Attention limitation** — investors can't monitor 660 stocks daily; extreme returns in small/mid caps go unnoticed
2. **Anchoring** — after a large daily move, investors anchor to the pre-move price, underweighting the new information
3. **Liquidity friction** — the initial move is partially driven by liquidity demand, not fully by information. The price continues to adjust over the following month.

**Step-by-step example (one month):**

1. Collect daily returns for all 660 stocks in March 2024
2. For each stock, find the single highest daily return in March (the "MAX")
3. Rank all stocks by MAX: stock with the biggest single-day gain ranks #1
4. Divide into 10 deciles (~66 stocks each). D1 = highest MAX, D10 = lowest
5. April 2024: go long D1, short D10, equal weight within each decile
6. End of April: close all positions, repeat
:::

#### Performance

```{r}
#| fig-height: 6
safe_tar_read("stk_max_plot")
```

::: {.fig-caption}
**Stock-level MAX: D1 long / D10 short, ~38 stocks per decile.** Blue = long-short spread, red = long-only (D1). Signal: maximum daily return in prior month. Source: 660 stocks from HuggingFace equity_daily.
:::

#### Metrics

```{r}
metrics <- safe_tar_read("stk_max_metrics")
if (!is.null(metrics)) {
  display <- metrics |>
    mutate(across(c(cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2),
           avg_long = round(avg_long), avg_short = round(avg_short))
  hd_dt_wide(display, "Stock MAX metrics by partition.")
}
```

#### vs Factor MAX

```{r}
#| fig-height: 6
safe_tar_read("stk_max_vs_factor_plot")
```

::: {.fig-caption}
**Stock vs Factor MAX.** Stock-level exploits cross-sectional dispersion across 660 stocks. Factor-level rotates among 5 academic factors. Same MAX signal, different universes. Factor-level definition: [Factor MAX](factor-max.html).
:::

#### Pros & Cons

| | Detail |
|--|--------|
| **Pros** | Simple signal (one number per stock-month). High raw returns. Works across markets. |
| **Cons** | Very high volatility (`r { m <- safe_tar_read("stk_max_metrics"); v <- if(!is.null(m)) m$vol[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,0),"%") else "N/A" }`). Max drawdown `r { m <- safe_tar_read("stk_max_metrics"); v <- if(!is.null(m)) m$max_dd[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,0),"%") else "N/A" }`. Sensitive to extreme events. High turnover (~monthly full rebalance). |
| **Works well** | High-volatility regimes (<abbr title="CBOE Volatility Index — implied volatility of S&P 500 options">VIX</abbr> > 25). Small/mid-cap stocks with less analyst coverage. |
| **Works badly** | Low-volatility, trending markets where MAX signals are noise. |
| **Factor exposure** | Loads on size (small-cap tilt in D1), volatility, and short-term reversal. Partially explained by these factors — not pure alpha. |

# Stock DRIF

## Row

### {.tabset}

#### Definition

::: {.callout-note}
## Strategy: Stock-Level DRIF (Daily Return Information Factor)

**What:** Each month, use an <a href="https://en.wikipedia.org/wiki/Elastic_net_regularization" target="_blank" rel="noopener noreferrer"><abbr title="A regularized regression method combining L1 (lasso) and L2 (ridge) penalties">elastic net regression</abbr></a> on the past 21 daily returns (both in chronological order and ranked by magnitude — 42 features total) to predict next-month stock returns. Go long the top decile by predicted return, short the bottom decile.

**Rationale (Market Dynamics):** The full sequence of daily returns contains more predictive information than any single summary statistic (mean, max, volatility). Specifically:

1. **Chronological features** (c1...c21) capture when returns occurred — recent price pressure and liquidity effects carry more weight than older ones
2. **Rank features** (r1...r21) capture how extreme returns were — behavioral biases cause investors to overweight tail events
3. **Elastic net** combines these 42 features with regularisation, avoiding overfitting while capturing non-obvious interactions

**Step-by-step example (one month):**

1. For each of 660 stocks, collect the 21 daily returns in March 2024
2. Create 42 features: c1...c21 (chronological order) + r1...r21 (sorted by magnitude)
3. Train elastic net (mixed L1/L2 regularisation, 5-fold CV) on all prior stock-months (expanding window, 60-month minimum)
4. Predict each stock's April return from its March features
5. Rank stocks by predicted return. D1 = highest predicted, D10 = lowest
6. April: long D1, short D10, equal weight
7. End of April: close, repeat

**Key distinction from MAX:** MAX uses 1 feature (the single highest daily return). DRIF uses 42 features (the full daily return distribution). The Alpha Architect paper shows DRIF subsumes MAX — MAX is a special case of DRIF.
:::

#### Performance

```{r}
#| fig-height: 6
safe_tar_read("stk_drif_plot")
```

::: {.fig-caption}
**Stock-level DRIF: pooled elastic net, D1/D10 decile spread.** Blue = long-short, red = long-only (D1). Expanding window with 60-month minimum training. Source: 660 stocks, 42 features per stock-month.
:::

#### Metrics

```{r}
metrics <- safe_tar_read("stk_drif_metrics")
if (!is.null(metrics)) {
  display <- metrics |>
    mutate(across(c(cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2),
           avg_long = round(avg_long), avg_short = round(avg_short))
  hd_dt_wide(display, "Stock DRIF metrics by partition.")
}
```

#### Pros & Cons

| | Detail |
|--|--------|
| **Pros (long-only)** | Uses full daily return information (42 features <!-- structural: 21 chronological + 21 rank daily returns — part of strategy definition, not a computed metric -->). Elastic net prevents overfitting. During strategy development (Testing period), showed Testing Sharpe of `r { m <- safe_tar_read("stk_drif_metrics"); if(!is.null(m)) round(m$sharpe[m$period=="Testing"],2) else "N/A" }`. |
| **Cons** | Weak in-sample long-short (`r { m <- safe_tar_read("stk_drif_metrics"); if(!is.null(m)) paste0(round(m$cagr[m$period=="Training"]*100,1),"%") else "N/A" }` CAGR). Compute-intensive (10 min for 660 stocks × 600 months). High turnover. Validation Sharpe `r { m <- safe_tar_read("stk_drif_metrics"); v <- if(!is.null(m)) m$sharpe[m$period=="Validation"] else numeric(0); if(length(v)==1L) round(v,2) else "N/A" }` — generalisation to sealed Validation has not held. |
| **Works well** | When the cross-section is large (more stocks = better elastic net training). Long-only implementation avoids cost drag of short leg. |
| **Works badly** | Small universes (<100 stocks). Long-short after costs (D1-D10 spread insufficient to overcome 1.85%/month cost). |
| **Factor exposure** | Lower factor loading than MAX (elastic net regularises out common factors). More likely to represent genuine alpha. |
| **OOS > IS anomaly** | Possible explanations: universe composition change post-2010, elastic net needs many years to learn, or OOS window (62 months) is too short to be conclusive. |
| **Implementation choice** | Long-only D1 is the practical strategy — avoids short costs and borrows, captures most of the signal. Long-short shown for comparison but unprofitable after realistic costs. |

#### References

1. [DRIF: factor-level implementation](drif.html) — same methodology on 5 FF factors
2. [Factor MAX](factor-max.html) — simpler signal, comparison benchmark
3. [Alpha Architect: Daily Stock Returns](https://alphaarchitect.com/daily-stock-returns/){target="_blank" rel="noopener noreferrer"} — original research
4. [Alpha Architect: Factor MAX](https://alphaarchitect.com/factor-max/){target="_blank" rel="noopener noreferrer"} — MAX signal research

# XGBoost DRIF

## Row

### {.tabset}

#### Definition

::: {.callout-note}
## Strategy: XGBoost DRIF (Monotonic Constraints)

**What:** Replace elastic net with XGBoost (gradient boosted trees) on the same 42 DRIF features. All features are constrained to be monotonically increasing — higher feature value must always predict higher return.

**Hypothesis:** XGBoost can capture non-linear relationships (interaction effects, threshold effects) that elastic net misses, while monotonic constraints preserve economic intuition.

**Result:** XGBoost full-period CAGR: `r { m <- safe_tar_read("xgb_drif_metrics"); v <- if(!is.null(m)) m$cagr[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,1),"%") else "N/A" }` vs Elastic Net: `r { m <- safe_tar_read("stk_drif_metrics"); v <- if(!is.null(m)) m$cagr[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,1),"%") else "N/A" }`. The monotonic constraints are likely too restrictive for cross-sectional stock returns where the relationship between daily return features and next-month returns is not universally monotonic.
:::

#### XGBoost vs Elastic Net

```{r}
#| fig-height: 6
safe_tar_read("xgb_vs_enet_plot")
```

::: {.fig-caption}
**XGBoost (monotonic) vs Elastic Net** on same 42 DRIF features, stock-level D1-D10. The XGBoost curve shows occasional extreme monthly spikes — these are months where monotonic constraints force all 42 features into the same direction, creating artificially concentrated decile predictions. The two large jumps likely reflect months where the constrained model's predictions became highly correlated across stocks, amplifying the D1-D10 spread beyond what the underlying signal supports.
:::

#### Why It Fails

::: {.callout-warning}
## XGBoost Feature Importance Is Unreliable Here

Feature importance from this model is **not shown** because it is misleading. With all 42 features forced monotonically increasing and depth=3 trees, XGBoost routes most splits through whichever single feature best separates training data. The result (one feature appearing to explain ~50% of variation) is an artifact of the constraints, not a real finding. Different training windows produce different "dominant" features. The elastic net — which can assign negative coefficients — is better suited for this cross-sectional task.
:::

#### Metrics

```{r}
metrics <- safe_tar_read("xgb_drif_metrics")
if (!is.null(metrics)) {
  display <- metrics |>
    filter(period != "Validation") |>
    mutate(across(c(cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2),
           avg_long = round(avg_long), avg_short = round(avg_short))
  hd_dt_wide(display, "XGBoost DRIF metrics by partition. Compare to Stock DRIF metrics in the Stock DRIF tab.")
}
```

#### Pros & Cons

| | Detail |
|--|--------|
| **Pros** | Captures non-linear interactions. Monotonic constraints preserve economic intuition. Feature importance is interpretable. |
| **Cons** | Underperforms elastic net at stock level. Monotonic constraints too restrictive for 42 features. Occasional extreme monthly predictions. Compute-intensive (7+ minutes). |
| **Works well** | Unknown — does not outperform simpler model in any tested regime. |
| **Works badly** | Cross-sectional stock prediction where features have non-monotonic relationships with returns. |
| **Verdict** | **Not recommended.** Elastic net is simpler, faster, and more accurate for this application. |

# ETF Replication

## Row

### {.tabset}

#### Definition

::: {.callout-note}
## Why ETFs?

Stock-level long-short strategies lose ~20%/yr after realistic costs (1.85%/month for 660-stock decile sort). Factor ETFs (VLUE, MTUM, QUAL, etc.) offer institutional liquidity at ~0.20%/month — 10x cheaper.

**Two approaches tested:**

| | A: DRIF on ETF Returns | B: Academic Signal → ETFs |
|--|----------------------|--------------------------|
| **Signal** | Elastic net on ETF daily prices | Factor-level DRIF (FF5+Mom) mapped to ETFs |
| **Execution** | Trade the predicted ETFs directly | Map HML→VLUE, Mom→MTUM, etc. |
| **Training data** | 2013-2026 (ETF inception) | 1963-2026 (academic factors) |
| **Key risk** | Short history (10 ETFs × 10 years) | Weak mapping (VLUE/HML cor = 0.34) |

**Result:** Long-short fails for both. **Long-only works:** `r { m <- safe_tar_read("etf_a_metrics"); if(!is.null(m)) paste0(round(m$long_cagr[m$period=="Full Period"]*100,1),"%") else "N/A" }` CAGR (A) and `r { m <- safe_tar_read("etf_b_metrics"); if(!is.null(m)) paste0(round(m$long_cagr[m$period=="Full Period"]*100,1),"%") else "N/A" }` CAGR (B). The signal identifies winners, not losers.
:::

#### Equity Curves

```{r}
#| fig-height: 6
safe_tar_read("etf_comparison_plot")
```

::: {.fig-caption}
**ETF replication: Approach A vs B (long-short, net of costs).** Both approaches struggle as long-short because all factor ETFs have positive market beta — shorting one while longing another is mostly a bet on relative factor performance. The long-only legs (not shown) return `r { m <- safe_tar_read("etf_a_metrics"); if(!is.null(m)) paste0(round(m$long_cagr[m$period=="Full Period"]*100,0),"%+") else "13%+" }` CAGR.
:::

#### Approach A Metrics

```{r}
metrics_a <- safe_tar_read("etf_a_metrics")
if (!is.null(metrics_a)) {
  display <- metrics_a |>
    mutate(across(c(cagr, long_cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2))
  hd_dt_wide(display, "Approach A: DRIF applied directly to 9 factor ETF daily returns. Long-short (L/S) and long-only metrics. Cost: 0.10%/trade, 0.5%/yr borrow.")
}
```

#### Approach B Metrics

```{r}
metrics_b <- safe_tar_read("etf_b_metrics")
if (!is.null(metrics_b)) {
  display <- metrics_b |>
    mutate(across(c(cagr, long_cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2))
  hd_dt_wide(display, "Approach B: Academic factor DRIF signal mapped to ETFs via ETF_FACTOR_MAP. Mapping: HML→VLUE/VTV/IWD, Mom→MTUM, RMW→QUAL, etc.")
}
```

#### Key Findings

| Finding | Evidence |
|---------|----------|
| **Long-only is the practical strategy** | A long-only: `r { m <- safe_tar_read("etf_a_metrics"); if(!is.null(m)) paste0(round(m$long_cagr[m$period=="Full Period"]*100,1),"%") else "N/A" }` CAGR. B long-only: `r { m <- safe_tar_read("etf_b_metrics"); if(!is.null(m)) paste0(round(m$long_cagr[m$period=="Full Period"]*100,1),"%") else "N/A" }`. Compare S&P 500 ~10%. |
| **Short leg is unprofitable** | A long-short: `r { m <- safe_tar_read("etf_a_metrics"); if(!is.null(m)) paste0(round(m$cagr[m$period=="Full Period"]*100,1),"%") else "N/A" }`. B long-short: `r { m <- safe_tar_read("etf_b_metrics"); if(!is.null(m)) paste0(round(m$cagr[m$period=="Full Period"]*100,1),"%") else "N/A" }`. ETFs all have market beta — shorting one is mostly shorting the market. |
| **A > B** | Approach A (direct ETF signal) has higher Sharpe (`r { m <- safe_tar_read("etf_a_metrics"); if(!is.null(m)) round(m$sharpe[m$period=="Full Period"],2) else "N/A" }` vs `r { m <- safe_tar_read("etf_b_metrics"); if(!is.null(m)) round(m$sharpe[m$period=="Full Period"],2) else "N/A" }`) and lower DD (`r { m <- safe_tar_read("etf_a_metrics"); if(!is.null(m)) paste0(round(m$max_dd[m$period=="Full Period"]*100,1),"%") else "N/A" }` vs `r { m <- safe_tar_read("etf_b_metrics"); if(!is.null(m)) paste0(round(m$max_dd[m$period=="Full Period"]*100,1),"%") else "N/A" }`). The VLUE/HML mapping (cor=0.34) in B adds noise. |
| **Cost advantage is real** | ETF costs ~0.20%/month vs 1.85%/month stock-level. The long-only gross returns survive after costs. |
| **Validation is strong** | Both show ~20% long-only CAGR in 2023-2026. Caveat: overlaps a bull market. |

# Details

## Row

### {.tabset}

#### Universe

660 individual stocks from HuggingFace `equity_daily`:

| Group | Tickers | History |
|-------|--------:|---------|
| S&P 500 (US) | ~500 | 2000-2026 (most), 1962+ (blue chips) |
| STOXX 600 majors (EU) | ~143 | 2000-2026 |
| Factor ETFs | 9 | 2013-2026 |
| **Total** | **~660** | |

Minimum 252 trading days (1 year) to enter universe. Monthly rebalance. ~38 stocks per decile.

#### Partitions

```{r}
params <- safe_tar_read("stk_params")
if (!is.null(params)) {
  param_df <- tibble(
    Parameter = c("Training", "Testing", "Validation",
                  "Min History", "Lookback Days", "Deciles"),
    Value = c(
      paste(params$start_date, "to", params$is_end),
      paste(params$test_start, "to", params$test_end),
      paste(params$val_start, "to", params$val_end),
      paste(params$min_history_days, "days"),
      as.character(params$lookback_days),
      as.character(params$n_deciles)
    )
  )
  hd_dt(param_df, "Shared parameters. Training: model fitting. Testing: calibration. Validation: sealed one-shot evaluation.")
}
```

#### Build Info

```{r}
#| label: build-info
#| echo: false
#| results: asis
build_info()
```
