---
title: "Stock-Level Backtests"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    embed-resources: false
    include-in-header:
      - text: |
          <link rel="stylesheet" href="vignette-shared.css">
          <script src="vignette-shared.js"></script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)
library(dplyr)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)
```

# Comparison

## Row

::: {.callout-note}
## 4 Strategies × 2 Levels

Cross-sectional backtests on 660 individual stocks (S&P 500 + STOXX 600 majors), compared with factor-level implementations on 5 Fama-French factors. Two signals: MAX (highest single-day return) and DRIF (elastic net on 42 daily return features).

**Stock-level:** Each month, rank all 660 stocks by signal, long top decile (~38 stocks), short bottom decile. Monthly rebalance.

**Factor-level:** Each month, rank 5 factors by signal, long top 2. Monthly rebalance.
:::

## Row

### {.tabset}

#### All Strategies

```{r}
#| fig-height: 6
safe_tar_read("stk_all_comparison_plot")
```

::: {.fig-caption}
**4-strategy comparison.** Y: growth of $1 (log scale). Stock-level uses D1 long / D10 short across 660 stocks. Factor-level uses top-2 of 5 FF factors. Dashed line = OOS start (2020). Source: HuggingFace equity_daily + FF5/Momentum via `hd_factors()`.
:::

#### Metrics

```{r}
stk_max <- safe_tar_read("stk_max_metrics")
stk_drif <- safe_tar_read("stk_drif_metrics")
fac_max <- safe_tar_read("fm_metrics")
fac_drif <- safe_tar_read("drif_metrics")

if (!is.null(stk_max) && !is.null(stk_drif) && !is.null(fac_max) && !is.null(fac_drif)) {
  combined <- bind_rows(
    stk_max |> filter(period == "Full Period") |> mutate(strategy = "Stock MAX"),
    stk_drif |> filter(period == "Full Period") |> mutate(strategy = "Stock DRIF"),
    fac_max |> filter(period == "Full Period") |> mutate(strategy = "Factor MAX"),
    fac_drif |> filter(period == "Full Period") |> mutate(strategy = "Factor DRIF")
  ) |>
    mutate(across(c(cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2)) |>
    select(Strategy = strategy, CAGR = cagr, Vol = vol, Sharpe = sharpe, `Max DD` = max_dd)

  # OOS
  combined_oos <- bind_rows(
    stk_max |> filter(period == "Out-of-Sample") |> mutate(strategy = "Stock MAX"),
    stk_drif |> filter(period == "Out-of-Sample") |> mutate(strategy = "Stock DRIF"),
    fac_max |> filter(period == "Out-of-Sample") |> mutate(strategy = "Factor MAX"),
    fac_drif |> filter(period == "Out-of-Sample") |> mutate(strategy = "Factor DRIF")
  ) |>
    mutate(across(c(cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2)) |>
    select(Strategy = strategy, `OOS CAGR` = cagr, `OOS Vol` = vol,
           `OOS Sharpe` = sharpe, `OOS Max DD` = max_dd)

  full <- inner_join(combined, combined_oos, by = "Strategy")
  hd_dt(full, "4-strategy comparison: full period and out-of-sample (2020+). Stock-level uses D1-D10 decile spread across 660 stocks. Factor-level uses top-2 of 5 factors.")
}
```

#### Key Findings

::: {.callout-tip}
## What the Comparison Shows

1. **Factor DRIF has the best risk-adjusted full-period Sharpe (0.26)** — elastic net on factor daily returns produces the most stable signal

2. **Stock DRIF has the best OOS Sharpe (0.79)** — the pooled elastic net generalises well to unseen data, even though in-sample returns are weak

3. **Stock MAX has the highest raw returns (12.1% CAGR)** but with extreme volatility (44%) and drawdown (-74%). The signal is real but noisy.

4. **Factor MAX is the weakest** — a single extreme daily return at the factor level contains less cross-sectional information than at the stock level

5. **The DRIF paper's insight is confirmed:** the full daily return distribution (42 features) captures more predictive information than the single maximum (1 feature), especially out-of-sample
:::

# Stock MAX

## Row

### {.tabset}

#### Performance

```{r}
#| fig-height: 6
safe_tar_read("stk_max_plot")
```

::: {.fig-caption}
**Stock-level Factor MAX.** D1 long / D10 short, ~38 stocks per decile. Signal: maximum daily return in the prior month. Dashed = OOS start. Source: 660 stocks from HuggingFace equity_daily.
:::

#### Metrics

```{r}
metrics <- safe_tar_read("stk_max_metrics")
if (!is.null(metrics)) {
  display <- metrics |>
    mutate(across(c(cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2),
           avg_long = round(avg_long), avg_short = round(avg_short))
  hd_dt_wide(display, "Stock-level MAX: decile long-short portfolio metrics.")
}
```

#### vs Factor MAX

```{r}
#| fig-height: 6
safe_tar_read("stk_max_vs_factor_plot")
```

::: {.fig-caption}
**Stock-level vs factor-level MAX.** The stock-level signal exploits cross-sectional dispersion across 660 stocks. Factor-level rotates among 5 academic factors. Both use the same MAX signal (highest single-day return).
:::

# Stock DRIF

## Row

### {.tabset}

#### Performance

```{r}
#| fig-height: 6
safe_tar_read("stk_drif_plot")
```

::: {.fig-caption}
**Stock-level DRIF.** Pooled elastic net (alpha=0.5, 5-fold CV) on 42 features per stock-month (21 chronological + 21 rank daily returns). D1 long / D10 short. Expanding window with 60-month minimum.
:::

#### Metrics

```{r}
metrics <- safe_tar_read("stk_drif_metrics")
if (!is.null(metrics)) {
  display <- metrics |>
    mutate(across(c(cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2),
           avg_long = round(avg_long), avg_short = round(avg_short))
  hd_dt_wide(display, "Stock-level DRIF: pooled elastic net, decile long-short portfolio.")
}
```

#### Why OOS > IS?

::: {.callout-warning}
## In-Sample Weakness, Out-of-Sample Strength

Stock-level DRIF shows an unusual pattern: OOS Sharpe (0.79) is much higher than IS Sharpe (-0.02). Possible explanations:

1. **Elastic net regularisation is very aggressive** with 42 features × 400 stocks — the model learns slowly, needing many years before the signal emerges

2. **Universe composition changed** — more stocks entered the dataset post-2010, improving the cross-sectional signal quality

3. **Regime shift** — the post-2020 period (COVID recovery, meme stocks, factor rotation) may coincidentally favour the DRIF signal

4. **Small OOS window** — only 62 months of OOS data; results may not be robust

This warrants further investigation: split OOS into sub-periods, test with different expanding window minimums, and compare with a random signal baseline.
:::

# Details

## Row

### {.tabset}

#### Universe

660 individual stocks from HuggingFace `equity_daily`:

| Group | Tickers | History |
|-------|--------:|---------|
| S&P 500 (US) | ~500 | 2000-2026 (most), 1962+ (blue chips) |
| STOXX 600 majors (EU) | ~143 | 2000-2026 |
| Factor ETFs | 9 | 2013-2026 |
| **Total** | **~660** | |

Minimum 252 trading days (1 year) to enter universe. Monthly rebalance. ~38 stocks per decile.

#### Parameters

```{r}
params <- safe_tar_read("stk_params")
if (!is.null(params)) {
  param_df <- tibble(
    Parameter = c("Min History (days)", "Lookback Days", "Deciles",
                  "Start Date", "IS End", "OOS Start"),
    Value = c(as.character(params$min_history_days),
              as.character(params$lookback_days),
              as.character(params$n_deciles),
              as.character(params$start_date),
              as.character(params$is_end),
              as.character(params$oos_start))
  )
  hd_dt(param_df, "Shared parameters for stock-level MAX and DRIF backtests.")
}
```

#### Build Info

```{r}
#| label: build-info
#| echo: false
#| results: asis
build_info()
```
