---
title: "Stock-Level Backtests"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    embed-resources: false
    include-in-header:
      - _includes/color-scheme-meta.html
      - text: |
          <link rel="stylesheet" href="vignette-shared.css">
          <script src="vignette-shared.js"></script>
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(targets)
library(ggplot2)
library(DT)
library(dplyr)

store_path <- if (dir.exists("_targets")) "_targets" else
  file.path(here::here(), "docs/_targets")
tar_config_set(store = store_path)

utils_path <- if (file.exists("vignette_utils.R")) "vignette_utils.R" else
  file.path(here::here(), "docs/vignette_utils.R")
source(utils_path)

disc_path <- if (file.exists("../R/disclosures.R")) "../R/disclosures.R" else
  file.path(here::here(), "R/disclosures.R")
source(disc_path)
```

# Comparison

::: {.callout-note .small collapse="false"}
⚠ Stock MAX, Stock DRIF, and XGBoost DRIF below carry known survivorship
bias in the stock universe. Full detail: [Biases and Caveats](#biases-and-caveats).
:::

## Row

### {.tabset}

#### Summary

Two signals applied at two levels, producing four strategies:

| | Signal | Stock-Level (660 stocks, D1-D10) | Factor-Level (5 FF factors, top 2) |
|--|--------|----------------------------------|-------------------------------------|
| **MAX** | Highest single-day return in prior month | [Stock MAX](#stock-max) | [Factor MAX](#factor-level-deep-dive) |
| **<abbr title="Dynamic Rotation Into Factors — selects best-performing FF factor each month">DRIF</abbr>** | Elastic net on 42 daily return features | [Stock DRIF](#stock-drif) | [Factor DRIF](#factor-level-deep-dive) |

**Stock-level:** Each month, rank all 660 stocks by signal. Long top decile (~38 stocks), short bottom decile. Monthly rebalance. Equal weight within each decile.

**Factor-level:** Each month, rank 5 Fama-French factors (<abbr title="High Minus Low — value factor (cheap stocks minus expensive stocks)">HML</abbr>, <abbr title="Small Minus Big — size factor (small-cap minus large-cap returns)">SMB</abbr>, <abbr title="Robust Minus Weak — profitability factor">RMW</abbr>, <abbr title="Conservative Minus Aggressive — investment factor">CMA</abbr>, <abbr title="Momentum — past 12-month winners minus losers">Mom</abbr>) by signal. Long top 2. Monthly rebalance. Equal weight.

#### Equity Curves

```{r}
#| fig-height: 6
{ p <- safe_tar_read("stk_all_comparison_plot"); if (!is.null(p)) p else { cat("*Chart not available — run tar_make() to generate.*\n"); invisible(NULL) } }
```

::: {.fig-caption}
```{r}
#| echo: false
#| results: asis
cap <- safe_tar_read("stk_all_caption")
if (!is.null(cap)) cat(cap) else cat("*Caption not available — run tar_make() to generate.*")
```
:::

#### Metrics

```{r}
stk_max <- safe_tar_read("stk_max_metrics")
stk_drif <- safe_tar_read("stk_drif_metrics")
fac_max <- safe_tar_read("fm_metrics")
fac_drif <- safe_tar_read("drif_metrics")

if (!is.null(stk_max) && !is.null(stk_drif) && !is.null(fac_max) && !is.null(fac_drif)) {
  make_row <- function(m, name) {
    m |> mutate(
      Strategy = name,
      CAGR = paste0(round(cagr * 100, 1), "%"),
      Vol = paste0(round(vol * 100, 1), "%"),
      Sharpe = round(sharpe, 2),
      `Max DD` = paste0(round(max_dd * 100, 1), "%")
    ) |> select(Strategy, Period = period, CAGR, Vol, Sharpe, `Max DD`)
  }

  all_metrics <- bind_rows(
    make_row(stk_max, "Stock MAX"),
    make_row(stk_drif, "Stock DRIF"),
    make_row(fac_max, "Factor MAX"),
    make_row(fac_drif, "Factor DRIF")
  ) |>
    filter(Period != "Validation")

  hd_dt(all_metrics, paste0(
    "All strategies by partition (Training / Testing / Full). ",
    "Validation is sealed — not reported until a one-shot production evaluation ",
    "(see `scripts/evaluate_validation.R`). ",
    "Factor-level links: Factor MAX, Factor DRIF. ",
    "Stock MAX and Stock DRIF draw on `stk_universe` (survivorship-biased — see issue #150)."
  ))
}
```

#### Key Findings

**P&L**

| Finding | Evidence |
|---------|----------|
| Stock MAX full-period CAGR: `r { m <- safe_tar_read("stk_max_metrics"); v <- if(!is.null(m)) m$cagr[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,1),"%") else "N/A" }` | Vol: `r { m <- safe_tar_read("stk_max_metrics"); v <- if(!is.null(m)) m$vol[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,0),"%") else "N/A" }` — full-period Sharpe: `r { m <- safe_tar_read("stk_max_metrics"); v <- if(!is.null(m)) m$sharpe[m$period=="Full Period"] else numeric(0); if(length(v)==1L) round(v,2) else "N/A" }` |
| Factor DRIF full-period Sharpe: `r { m <- safe_tar_read("drif_metrics"); v <- if(!is.null(m)) m$sharpe[m$period=="Full Period"] else numeric(0); if(length(v)==1L) round(v,2) else "N/A" }` | Full-period covers the entire sample including Training. Validation Sharpe is sealed — not reported until a one-shot production evaluation (see `scripts/evaluate_validation.R`). |
| Stock DRIF full-period CAGR is only `r { m <- safe_tar_read("stk_drif_metrics"); if(!is.null(m)) paste0(round(m$cagr[m$period=="Full Period"]*100,1),"%") else "N/A" }` | Elastic net regularisation is very aggressive at stock level |

**Out-of-Sample / Validation**

::: {.callout-warning}
**Validation (sealed partition):** Validation Sharpe for these strategies is not reported on this page — it is sealed until a one-shot production evaluation (see `scripts/evaluate_validation.R`). The Testing partition OOS results (below) reflect the calibration window used during strategy development, not the sealed evaluation.
:::

| Finding | Evidence |
|---------|----------|
| Stock DRIF Testing-period Sharpe during elastic net development: `r { m <- safe_tar_read("stk_drif_metrics"); v <- if(!is.null(m)) m$sharpe[m$period=="Testing"] else numeric(0); if(length(v)==1L) round(v,2) else "N/A" }` | Testing period reflects the calibration window used during development. Validation is sealed — not reported (see `scripts/evaluate_validation.R`). |
| Factor MAX Testing-period Sharpe: `r { m <- safe_tar_read("fm_metrics"); if(!is.null(m)) round(m$sharpe[m$period=="Testing"],2) else "N/A" }` | Single-signal approach at factor level; Validation is sealed — not reported (see `scripts/evaluate_validation.R`). |
| Stock MAX OOS CAGR looks extreme (`r { m <- safe_tar_read("stk_max_metrics"); if(!is.null(m)) paste0(round(m$cagr[m$period=="Testing"]*100,0),"%") else "N/A" }`) | Driven by COVID recovery + meme stocks. Validation CAGR is sealed — not reported (see `scripts/evaluate_validation.R`). |

**Risk**

| Finding | Evidence |
|---------|----------|
| Stock-level vol (stock MAX: `r { m <- safe_tar_read("stk_max_metrics"); v <- if(!is.null(m)) m$vol[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,0),"%") else "N/A" }`) vs factor-level vol (factor MAX: `r { m <- safe_tar_read("fm_metrics"); v <- if(!is.null(m)) m$vol[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,0),"%") else "N/A" }`) | Compare the two values above — see metrics table for all strategies |
| Stock MAX max drawdown is `r { m <- safe_tar_read("stk_max_metrics"); v <- if(!is.null(m)) m$max_dd[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,0),"%") else "N/A" }` | Unacceptable for most allocators |
| Factor strategies have drawdowns `r { m <- safe_tar_read("fm_metrics"); v <- if(!is.null(m)) m$max_dd[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0("<",round(abs(v)*100,0),"%") else "N/A" }` | Lower risk but lower return |

**Robustness**

| Finding | Evidence |
|---------|----------|
| Full daily return distribution (42 features) carries more signal than single-day MAX on paper | Whether this advantage persists cannot be assessed here — Validation is sealed and not reported (see `scripts/evaluate_validation.R`) |
| Factor-level vs stock-level volatility comparison — see current metrics table | Classic diversification vs concentration tradeoff |

# Stock MAX

## Row

### {.tabset}

#### Definition

::: {.callout-note}
## Strategy: Stock-Level Factor MAX

**What:** Each month, sort all stocks by the maximum daily return observed in the prior month. Go long the top decile, short the bottom decile.

**Rationale (Market Dynamics):** A stock with an unusually high single-day return signals that new information arrived — an earnings surprise, an upgrade, a sector rotation. Markets underreact to this information at the individual stock level because:

1. **Attention limitation** — investors can't monitor 660 stocks daily; extreme returns in small/mid caps go unnoticed
2. **Anchoring** — after a large daily move, investors anchor to the pre-move price, underweighting the new information
3. **Liquidity friction** — the initial move is partially driven by liquidity demand, not fully by information. The price continues to adjust over the following month.

**Step-by-step example (one month):**

1. Collect daily returns for all 660 stocks in March 2024
2. For each stock, find the single highest daily return in March (the "MAX")
3. Rank all stocks by MAX: stock with the biggest single-day gain ranks #1
4. Divide into 10 deciles (~66 stocks each). D1 = highest MAX, D10 = lowest
5. April 2024: go long D1, short D10, equal weight within each decile
6. End of April: close all positions, repeat
:::

#### Performance

```{r}
#| fig-height: 6
{ p <- safe_tar_read("stk_max_plot"); if (!is.null(p)) p else { cat("*Chart not available — run tar_make() to generate.*\n"); invisible(NULL) } }
```

::: {.fig-caption}
**Stock-level MAX: D1 long / D10 short, ~38 stocks per decile.** Blue = long-short spread, red = long-only (D1). Signal: maximum daily return in prior month. Source: 660 stocks from HuggingFace equity_daily.
:::

#### Metrics

```{r}
metrics <- safe_tar_read("stk_max_metrics")
if (!is.null(metrics)) {
  display <- metrics |>
    filter(period != "Validation") |>
    mutate(across(c(cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2),
           avg_long = round(avg_long), avg_short = round(avg_short))
  hd_dt_wide(display, "Stock MAX metrics by partition (Validation sealed — see scripts/evaluate_validation.R).")
}
```

#### vs Factor MAX

```{r}
#| fig-height: 6
{ p <- safe_tar_read("stk_max_vs_factor_plot"); if (!is.null(p)) p else { cat("*Chart not available — run tar_make() to generate.*\n"); invisible(NULL) } }
```

::: {.fig-caption}
**Stock vs Factor MAX.** Stock-level exploits cross-sectional dispersion across 660 stocks. Factor-level rotates among 5 academic factors. Same MAX signal, different universes. Factor-level definition: [Factor MAX](#factor-level-deep-dive).
:::

#### Pros & Cons

| | Detail |
|--|--------|
| **Pros** | Simple signal (one number per stock-month). High raw returns. Works across markets. |
| **Cons** | Very high volatility (`r { m <- safe_tar_read("stk_max_metrics"); v <- if(!is.null(m)) m$vol[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,0),"%") else "N/A" }`). Max drawdown `r { m <- safe_tar_read("stk_max_metrics"); v <- if(!is.null(m)) m$max_dd[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,0),"%") else "N/A" }`. Sensitive to extreme events. High turnover (~monthly full rebalance). |
| **Works well** | High-volatility regimes (<abbr title="CBOE Volatility Index — implied volatility of S&P 500 options">VIX</abbr> > 25). Small/mid-cap stocks with less analyst coverage. |
| **Works badly** | Low-volatility, trending markets where MAX signals are noise. |
| **Factor exposure** | Loads on size (small-cap tilt in D1), volatility, and short-term reversal. Partially explained by these factors — not pure alpha. |

# Stock DRIF

## Row

### {.tabset}

#### Definition

::: {.callout-note}
## Strategy: Stock-Level DRIF (Daily Return Information Factor)

**What:** Each month, use an <a href="https://en.wikipedia.org/wiki/Elastic_net_regularization" target="_blank" rel="noopener noreferrer"><abbr title="A regularized regression method combining L1 (lasso) and L2 (ridge) penalties">elastic net regression</abbr></a> on the past 21 daily returns (both in chronological order and ranked by magnitude — 42 features total) to predict next-month stock returns. Go long the top decile by predicted return, short the bottom decile.

**Rationale (Market Dynamics):** The full sequence of daily returns contains more predictive information than any single summary statistic (mean, max, volatility). Specifically:

1. **Chronological features** (c1...c21) capture when returns occurred — recent price pressure and liquidity effects carry more weight than older ones
2. **Rank features** (r1...r21) capture how extreme returns were — behavioral biases cause investors to overweight tail events
3. **Elastic net** combines these 42 features with regularisation, avoiding overfitting while capturing non-obvious interactions

**Step-by-step example (one month):**

1. For each of 660 stocks, collect the 21 daily returns in March 2024
2. Create 42 features: c1...c21 (chronological order) + r1...r21 (sorted by magnitude)
3. Train elastic net (mixed L1/L2 regularisation, 5-fold CV) on all prior stock-months (expanding window, 60-month minimum)
4. Predict each stock's April return from its March features
5. Rank stocks by predicted return. D1 = highest predicted, D10 = lowest
6. April: long D1, short D10, equal weight
7. End of April: close, repeat

**Key distinction from MAX:** MAX uses 1 feature (the single highest daily return). DRIF uses 42 features (the full daily return distribution). The Alpha Architect paper shows DRIF subsumes MAX — MAX is a special case of DRIF.
:::

#### Performance

```{r}
#| fig-height: 6
{ p <- safe_tar_read("stk_drif_plot"); if (!is.null(p)) p else { cat("*Chart not available — run tar_make() to generate.*\n"); invisible(NULL) } }
```

::: {.fig-caption}
**Stock-level DRIF: pooled elastic net, D1/D10 decile spread.** Blue = long-short, red = long-only (D1). Expanding window with 60-month minimum training. Source: 660 stocks, 42 features per stock-month.
:::

#### Metrics

```{r}
metrics <- safe_tar_read("stk_drif_metrics")
if (!is.null(metrics)) {
  display <- metrics |>
    filter(period != "Validation") |>
    mutate(across(c(cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2),
           avg_long = round(avg_long), avg_short = round(avg_short))
  hd_dt_wide(display, "Stock DRIF metrics by partition (Validation sealed — see scripts/evaluate_validation.R).")
}
```

#### Pros & Cons

| | Detail |
|--|--------|
| **Pros (long-only)** | Uses full daily return information (42 features <!-- structural: 21 chronological + 21 rank daily returns — part of strategy definition, not a computed metric -->). Elastic net prevents overfitting. During strategy development (Testing period), showed Testing Sharpe of `r { m <- safe_tar_read("stk_drif_metrics"); if(!is.null(m)) round(m$sharpe[m$period=="Testing"],2) else "N/A" }`. |
| **Cons** | Weak in-sample long-short (`r { m <- safe_tar_read("stk_drif_metrics"); if(!is.null(m)) paste0(round(m$cagr[m$period=="Training"]*100,1),"%") else "N/A" }` CAGR). Compute-intensive (10 min for 660 stocks × 600 months). High turnover. Generalisation beyond the Testing partition has not been assessed — Validation is sealed and is not read here (see `scripts/evaluate_validation.R`). |
| **Works well** | When the cross-section is large (more stocks = better elastic net training). Long-only implementation avoids cost drag of short leg. |
| **Works badly** | Small universes (<100 stocks). Long-short after costs (D1-D10 spread insufficient to overcome 1.85%/month cost). |
| **Factor exposure** | Lower factor loading than MAX (elastic net regularises out common factors). More likely to represent genuine alpha. |
| **OOS > IS anomaly** | Possible explanations: universe composition change post-2010, elastic net needs many years to learn, or OOS window (62 months) is too short to be conclusive. |
| **Implementation choice** | Long-only D1 is the practical strategy — avoids short costs and borrows, captures most of the signal. Long-short shown for comparison but unprofitable after realistic costs. |

#### References

1. [DRIF: factor-level implementation](#factor-level-deep-dive) — same methodology on 5 FF factors
2. [Factor MAX](#factor-level-deep-dive) — simpler signal, comparison benchmark
3. [Alpha Architect: Daily Stock Returns](https://alphaarchitect.com/daily-stock-returns/){target="_blank" rel="noopener noreferrer"} — original research
4. [Alpha Architect: Factor MAX](https://alphaarchitect.com/factor-max/){target="_blank" rel="noopener noreferrer"} — MAX signal research

# XGBoost DRIF

## Row

### {.tabset}

#### Definition

::: {.callout-note}
## Strategy: XGBoost DRIF (Monotonic Constraints)

**What:** Replace elastic net with XGBoost (gradient boosted trees) on the same 42 DRIF features. All features are constrained to be monotonically increasing — higher feature value must always predict higher return.

**Hypothesis:** XGBoost can capture non-linear relationships (interaction effects, threshold effects) that elastic net misses, while monotonic constraints preserve economic intuition.

**Result:** XGBoost full-period CAGR: `r { m <- safe_tar_read("xgb_drif_metrics"); v <- if(!is.null(m)) m$cagr[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,1),"%") else "N/A" }` vs Elastic Net: `r { m <- safe_tar_read("stk_drif_metrics"); v <- if(!is.null(m)) m$cagr[m$period=="Full Period"] else numeric(0); if(length(v)==1L) paste0(round(v*100,1),"%") else "N/A" }`. The monotonic constraints are likely too restrictive for cross-sectional stock returns where the relationship between daily return features and next-month returns is not universally monotonic.
:::

#### XGBoost vs Elastic Net

```{r}
#| fig-height: 6
{ p <- safe_tar_read("xgb_vs_enet_plot"); if (!is.null(p)) p else { cat("*Chart not available — run tar_make() to generate.*\n"); invisible(NULL) } }
```

::: {.fig-caption}
**XGBoost (monotonic) vs Elastic Net** on same 42 DRIF features, stock-level D1-D10. The XGBoost curve shows occasional extreme monthly spikes — these are months where monotonic constraints force all 42 features into the same direction, creating artificially concentrated decile predictions. The two large jumps likely reflect months where the constrained model's predictions became highly correlated across stocks, amplifying the D1-D10 spread beyond what the underlying signal supports.
:::

#### Why It Fails

::: {.callout-warning}
## XGBoost Feature Importance Is Unreliable Here

Feature importance from this model is **not shown** because it is misleading. With all 42 features forced monotonically increasing and depth=3 trees, XGBoost routes most splits through whichever single feature best separates training data. The result (one feature appearing to explain ~50% of variation) is an artifact of the constraints, not a real finding. Different training windows produce different "dominant" features. The elastic net — which can assign negative coefficients — is better suited for this cross-sectional task.
:::

#### Metrics

```{r}
metrics <- safe_tar_read("xgb_drif_metrics")
if (!is.null(metrics)) {
  display <- metrics |>
    filter(period != "Validation") |>
    mutate(across(c(cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2),
           avg_long = round(avg_long), avg_short = round(avg_short))
  hd_dt_wide(display, "XGBoost DRIF metrics by partition. Compare to Stock DRIF metrics in the Stock DRIF tab.")
}
```

#### Pros & Cons

| | Detail |
|--|--------|
| **Pros** | Captures non-linear interactions. Monotonic constraints preserve economic intuition. Feature importance is interpretable. |
| **Cons** | Underperforms elastic net at stock level. Monotonic constraints too restrictive for 42 features. Occasional extreme monthly predictions. Compute-intensive (7+ minutes). |
| **Works well** | Unknown — does not outperform simpler model in any tested regime. |
| **Works badly** | Cross-sectional stock prediction where features have non-monotonic relationships with returns. |
| **Verdict** | **Not recommended.** Elastic net is simpler, faster, and more accurate for this application. |

# ETF Replication

## Row

### {.tabset}

#### Definition

::: {.callout-note}
## Why ETFs?

Stock-level long-short strategies lose ~20%/yr after realistic costs (1.85%/month for 660-stock decile sort). Factor ETFs (VLUE, MTUM, QUAL, etc.) offer institutional liquidity at ~0.20%/month — 10x cheaper.

**Two approaches tested:**

| | A: DRIF on ETF Returns | B: Academic Signal → ETFs |
|--|----------------------|--------------------------|
| **Signal** | Elastic net on ETF daily prices | Factor-level DRIF (FF5+Mom) mapped to ETFs |
| **Execution** | Trade the predicted ETFs directly | Map HML→VLUE, Mom→MTUM, etc. |
| **Training data** | 2013-2026 (ETF inception) | 1963-2026 (academic factors) |
| **Key risk** | Short history (10 ETFs × 10 years) | Weak mapping (VLUE/HML cor = 0.34) |

**Result:** Long-short fails for both. **Long-only works:** `r { m <- safe_tar_read("etf_a_metrics"); if(!is.null(m)) paste0(round(m$long_cagr[m$period=="Full Period"]*100,1),"%") else "N/A" }` CAGR (A) and `r { m <- safe_tar_read("etf_b_metrics"); if(!is.null(m)) paste0(round(m$long_cagr[m$period=="Full Period"]*100,1),"%") else "N/A" }` CAGR (B). The signal identifies winners, not losers.
:::

#### Equity Curves

```{r}
#| fig-height: 6
{ p <- safe_tar_read("etf_comparison_plot"); if (!is.null(p)) p else { cat("*Chart not available — run tar_make() to generate.*\n"); invisible(NULL) } }
```

::: {.fig-caption}
**ETF replication: Approach A vs B (long-short, net of costs).** Both approaches struggle as long-short because all factor ETFs have positive market beta — shorting one while longing another is mostly a bet on relative factor performance. The long-only legs (not shown) return `r { m <- safe_tar_read("etf_a_metrics"); if(!is.null(m)) paste0(round(m$long_cagr[m$period=="Full Period"]*100,0),"%+") else "13%+" }` CAGR.
:::

#### Approach A Metrics

```{r}
metrics_a <- safe_tar_read("etf_a_metrics")
if (!is.null(metrics_a)) {
  display <- metrics_a |>
    filter(period != "Validation") |>
    mutate(across(c(cagr, long_cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2))
  hd_dt_wide(display, "Approach A: DRIF applied directly to 9 factor ETF daily returns. Long-short (L/S) and long-only metrics. Cost: 0.10%/trade, 0.5%/yr borrow. Validation sealed — see scripts/evaluate_validation.R.")
}
```

#### Approach B Metrics

```{r}
metrics_b <- safe_tar_read("etf_b_metrics")
if (!is.null(metrics_b)) {
  display <- metrics_b |>
    filter(period != "Validation") |>
    mutate(across(c(cagr, long_cagr, vol, max_dd), ~ paste0(round(. * 100, 1), "%")),
           sharpe = round(sharpe, 2))
  hd_dt_wide(display, "Approach B: Academic factor DRIF signal mapped to ETFs via ETF_FACTOR_MAP. Mapping: HML→VLUE/VTV/IWD, Mom→MTUM, RMW→QUAL, etc. Validation sealed — see scripts/evaluate_validation.R.")
}
```

#### Key Findings

| Finding | Evidence |
|---------|----------|
| **Long-only is the practical strategy** | A long-only: `r { m <- safe_tar_read("etf_a_metrics"); if(!is.null(m)) paste0(round(m$long_cagr[m$period=="Full Period"]*100,1),"%") else "N/A" }` CAGR. B long-only: `r { m <- safe_tar_read("etf_b_metrics"); if(!is.null(m)) paste0(round(m$long_cagr[m$period=="Full Period"]*100,1),"%") else "N/A" }`. Compare S&P 500 ~10%. |
| **Short leg is unprofitable** | A long-short: `r { m <- safe_tar_read("etf_a_metrics"); if(!is.null(m)) paste0(round(m$cagr[m$period=="Full Period"]*100,1),"%") else "N/A" }`. B long-short: `r { m <- safe_tar_read("etf_b_metrics"); if(!is.null(m)) paste0(round(m$cagr[m$period=="Full Period"]*100,1),"%") else "N/A" }`. ETFs all have market beta — shorting one is mostly shorting the market. |
| **Approach A vs B** | A full-period Sharpe: `r { m <- safe_tar_read("etf_a_metrics"); if(!is.null(m)) round(m$sharpe[m$period=="Full Period"],2) else "N/A" }`. B full-period Sharpe: `r { m <- safe_tar_read("etf_b_metrics"); if(!is.null(m)) round(m$sharpe[m$period=="Full Period"],2) else "N/A" }`. A max DD: `r { m <- safe_tar_read("etf_a_metrics"); if(!is.null(m)) paste0(round(m$max_dd[m$period=="Full Period"]*100,1),"%") else "N/A" }`. The VLUE/HML mapping (cor=0.34) in B adds noise. |
| **Cost advantage is real** | ETF costs ~0.20%/month vs 1.85%/month stock-level. The long-only gross returns survive after costs. |
| **Validation** | Sealed — not reported until a one-shot production evaluation (see `scripts/evaluate_validation.R`). |

# Factor-Level Deep Dive {#factor-level-deep-dive}

## Row

::: {.callout-note}
Factor-level strategies rotate among 5 Fama-French factors (HML, SMB, RMW, CMA, Mom) each month. Both signals select the top-2 factors with equal weights and rebalance monthly. See the [Comparison](#comparison) tab for the 2×2 stock vs factor overview. The former standalone Factor MAX and DRIF pages now redirect here (#514 consolidation 2/3).
:::

## Row

### {.tabset}

#### Factor MAX

::: {.callout-tip}
## Factor MAX Rotation

Based on [Alpha Architect research](https://alphaarchitect.com/factor-max/){target="_blank" rel="noopener noreferrer"} (Dec 2025): the **maximum daily return within a month** predicts next-month factor returns. Each month, rank the 5 Fama-French factors by their MAX signal (highest single-day return). Go long the top-2 factors with equal weights. Rebalance monthly.
:::

```{r}
#| fig-height: 6
{ p <- safe_tar_read("fm_cumret_plot"); if (!is.null(p)) p else { cat("*Chart not available — run tar_make() to generate.*\n"); invisible(NULL) } }
```

::: {.fig-caption}
**Factor MAX vs Market.** Y: growth of $1 (log scale). Blue = Factor MAX (top-2 factors by MAX signal, equal weight). Red = Market factor (Mkt-RF). Dashed line = OOS start (2020). Source: Fama-French via `hd_factors()`. Log scale shows compounding differences.
:::

```{r}
metrics <- safe_tar_read("fm_metrics")
if (!is.null(metrics)) {
  display <- metrics |>
    filter(period != "Validation") |>
    mutate(across(c(cagr, vol, max_dd, hit_rate, bench_cagr, bench_vol),
                  ~ paste0(round(. * 100, 1), "%"))) |>
    mutate(across(c(sharpe, bench_sharpe), ~ round(., 2)))
  hd_dt(display, "Factor MAX strategy metrics: in-sample (1963-2019), out-of-sample (2020+), and full period. Validation sealed — see scripts/evaluate_validation.R.")
}
```

```{r}
sf <- safe_tar_read("fm_selection_freq")
if (!is.null(sf)) {
  sf <- sf |> mutate(pct = paste0(round(pct * 100, 1), "%"))
  hd_dt(sf, "How often each factor is selected by the MAX signal. Momentum dominates.")
}
```

```{r}
#| fig-height: 8
{ p <- safe_tar_read("fm_heatmap"); if (!is.null(p)) p else { cat("*Chart not available — run tar_make() to generate.*\n"); invisible(NULL) } }
```

::: {.fig-caption}
**MAX signal rank heatmap.** Green = rank 1 (highest MAX, strongest buy signal). Dark = rank 5 (lowest). Rows: Fama-French factors (HML, SMB, RMW, CMA, Mom). Columns: calendar months. Faceted by year. Source: `hd_factors(dataset = "FF5")` + `hd_factors(dataset = "Mom")`.
:::

```{r}
#| fig-height: 6
{ p <- safe_tar_read("fm_etf_plot"); if (!is.null(p)) p else { cat("*Chart not available — run tar_make() to generate.*\n"); invisible(NULL) } }
```

::: {.fig-caption}
**Factor ETFs vs academic factors.** VLUE (iShares Value) vs HML (Fama-French value factor), MTUM (iShares Momentum) vs Mom (Fama-French momentum). ETFs include market beta; academic factors are market-neutral. Source: `hd_ohlcv()` + `hd_factors()`.
:::

```{r}
corr <- safe_tar_read("fm_etf_corr")
if (!is.null(corr)) {
  display <- corr |>
    dplyr::mutate(ETF_CAGR = paste0(ETF_CAGR, "%"), Factor_CAGR = paste0(Factor_CAGR, "%"))
  hd_dt(display, "Correlation between factor ETFs and academic factors (monthly returns, 2013-2026). Low correlations (VLUE/HML = 0.34, MTUM/Mom = -0.09) show ETFs capture different exposures than academic factors.")
}
```

```{r}
params <- safe_tar_read("fm_params")
if (!is.null(params)) {
  param_df <- tibble(
    Parameter = c("Factors", "Top N", "Start Date", "IS End", "OOS Start", "Benchmark"),
    Value = c(paste(params$factors, collapse = ", "),
              as.character(params$top_n),
              as.character(params$start_date),
              as.character(params$is_end),
              as.character(params$oos_start),
              params$benchmark_factor)
  )
  hd_dt(param_df, "Factor MAX parameters. Change fm_params target and tar_make() to update all.")
}
```

**References:** [Alpha Architect: Factor MAX](https://alphaarchitect.com/factor-max/){target="_blank" rel="noopener noreferrer"} — original research (Dec 2025). Data: [Ken French Data Library](https://mba.tuck.dartmouth.edu/pages/faculty/ken.french/data_library.html){target="_blank" rel="noopener noreferrer"} via `hd_factors()`.

#### DRIF (Factor)

::: {.callout-tip}
## DRIF: Daily Return Information Factor

Based on [Alpha Architect research](https://alphaarchitect.com/daily-stock-returns/){target="_blank" rel="noopener noreferrer"}: an elastic net regression on the **full distribution of past 21 daily returns** — chronological order and rank order (42 features total) — predicts next-month factor returns. Each month, train an expanding-window elastic net (`glmnet::cv.glmnet`, 5-fold CV) and go long the top-2 predicted factors. The DRIF signal subsumes many anomalies including MAX, short-term reversal, and volatility effects.
:::

```{r}
#| fig-height: 6
{ p <- safe_tar_read("drif_cumret_plot"); if (!is.null(p)) p else { cat("*Chart not available — run tar_make() to generate.*\n"); invisible(NULL) } }
```

::: {.fig-caption}
**DRIF vs Market.** Y: growth of $1 (log scale). Blue = DRIF (top-2 factors by elastic net prediction, equal weight). Red = Market factor (Mkt-RF). Dashed line = OOS start (2020). Source: Fama-French FF5 + Momentum via `hd_factors()`.
:::

::: {.callout-warning}
**Validation (sealed partition):** Not reported here — DRIF Validation metrics are sealed until a one-shot production evaluation (see `scripts/evaluate_validation.R`). The in-sample and Testing Sharpes above reflect the development period only.
:::

```{r}
metrics <- safe_tar_read("drif_metrics")
if (!is.null(metrics)) {
  display <- metrics |>
    dplyr::filter(period != "Validation") |>
    dplyr::mutate(across(c(cagr, vol, max_dd, hit_rate, bench_cagr, bench_vol),
                  ~ paste0(round(. * 100, 1), "%"))) |>
    dplyr::mutate(across(c(sharpe, bench_sharpe), ~ round(., 2)))
  hd_dt_wide(display, "DRIF strategy metrics: in-sample (1968-2019), out-of-sample (2020+), and full period. Uses 60-month minimum expanding window. Validation sealed — see scripts/evaluate_validation.R.")
}
```

```{r}
sf <- safe_tar_read("drif_selection_freq")
if (!is.null(sf)) {
  sf <- sf |> dplyr::mutate(pct = paste0(round(pct * 100, 1), "%"))
  hd_dt(sf, "How often each factor is selected by the DRIF signal. Momentum leads but less dominant than in Factor MAX.")
}
```

```{r}
#| fig-height: 6
{ p <- safe_tar_read("drif_vs_max_plot"); if (!is.null(p)) p else { cat("*Chart not available — run tar_make() to generate.*\n"); invisible(NULL) } }
```

::: {.fig-caption}
**DRIF vs Factor MAX.** Both strategies applied to the same FF5+Momentum factor universe. DRIF uses 42 features (21 chronological + 21 rank daily returns) via elastic net; Factor MAX uses only the single highest daily return. Full-period Sharpe — DRIF: `r { m <- safe_tar_read("drif_metrics"); v <- if(!is.null(m)) m$sharpe[m$period=="Full Period"] else numeric(0); if(length(v)==1L) round(v,2) else "N/A" }`, Factor MAX: `r { m <- safe_tar_read("fm_metrics"); v <- if(!is.null(m)) m$sharpe[m$period=="Full Period"] else numeric(0); if(length(v)==1L) round(v,2) else "N/A" }`.
:::

```{r}
params <- safe_tar_read("drif_params")
if (!is.null(params)) {
  param_df <- dplyr::tibble(
    Parameter = c("Factors", "Lookback Days", "Top N", "Elastic Net Alpha",
                  "Min Train Months", "Start Date", "IS End", "OOS Start"),
    Value = c(paste(params$factors, collapse = ", "),
              as.character(params$lookback_days),
              as.character(params$top_n),
              as.character(params$alpha),
              as.character(params$min_train_months),
              as.character(params$start_date),
              as.character(params$is_end),
              as.character(params$oos_start))
  )
  hd_dt(param_df, "DRIF parameters. Change drif_params target and tar_make() to update all.")
}
```

The Alpha Architect paper finds chronological features are ~1.7x more predictive than rank features:

| Dimension | Monthly Spread | Intuition |
|-----------|---------------|-----------|
| **Chronological** | ~1.5% | Recent price pressure, liquidity, order flow |
| **Rank** | ~0.9% | Behavioral biases around extreme returns |
| **Combined (DRIF)** | ~1.6% | Full daily return distribution |

::: {.callout-warning}
## Factor-Level vs Stock-Level Limitation

This is a **factor-level** implementation — rotating among 5-6 academic factors. The original DRIF paper applies the signal cross-sectionally to individual stocks (3000+ in CRSP). Key differences: cross-section of 5 factors vs 3000+ stocks (much less statistical power); factor daily returns are more correlated than stock returns; no shorting (long top-2 only). A proper stock-level implementation is in the [Stock DRIF](#stock-drif) section above.
:::

**References:** [Alpha Architect: DRIF](https://alphaarchitect.com/daily-stock-returns/){target="_blank" rel="noopener noreferrer"} — original research (2026). Data: [Ken French Data Library](https://mba.tuck.dartmouth.edu/pages/faculty/ken.french/data_library.html){target="_blank" rel="noopener noreferrer"} via `hd_factors()`.

# Details

## Row

### {.tabset}

#### Universe

660 individual stocks from HuggingFace `equity_daily`:

| Group | Tickers | History |
|-------|--------:|---------|
| S&P 500 (US) | ~500 | 2000-2026 (most), 1962+ (blue chips) |
| STOXX 600 majors (EU) | ~143 | 2000-2026 |
| Factor ETFs | 9 | 2013-2026 |
| **Total** | **~660** | |

Minimum 252 trading days (1 year) to enter universe. Monthly rebalance. ~38 stocks per decile.

#### Partitions

```{r}
params <- safe_tar_read("stk_params")
if (!is.null(params)) {
  param_df <- tibble(
    Parameter = c("Training", "Testing", "Validation",
                  "Min History", "Lookback Days", "Deciles"),
    Value = c(
      paste(params$start_date, "to", params$is_end),
      paste(params$test_start, "to", params$test_end),
      paste(params$val_start, "to", params$val_end),
      paste(params$min_history_days, "days"),
      as.character(params$lookback_days),
      as.character(params$n_deciles)
    )
  )
  hd_dt(param_df, "Shared parameters. Training: model fitting. Testing: calibration. Validation: sealed one-shot evaluation.")
}
```

#### Related Vignettes

- **[Factor MAX (Factor-Level Deep Dive)](#factor-level-deep-dive)** — factor-level MAX signal; this page applies that signal cross-sectionally to 660 individual equities
- **[DRIF (Factor-Level Deep Dive)](#factor-level-deep-dive)** — factor-level DRIF; this page applies the same elastic-net selection to individual stocks
- **[Strategy Leaderboard](leaderboard.html)** — overall strategy rankings across all strategies including stock-level implementations

#### Build Info

```{r}
#| label: build-info
#| echo: false
#| results: asis
build_info_tabset()
```

# Biases and Caveats

## Row

### {.tabset}

#### Survivorship Bias

::: {.callout-warning collapse="false" title="Survivorship bias"}
`r disclosure_survivorship()`
:::

Affected on this dashboard: [Stock MAX](#stock-max), [Stock DRIF](#stock-drif),
and [XGBoost DRIF](#xgboost-drif) — all draw on the same `stk_universe`. See
the [Comparison](#comparison) page for the strategy overview these results
feed into.

```{r}
#| label: caveats-survivorship-table
# A prose sentence naming the three affected strategies reads as thin on its
# own -- expand it into a table (owner feedback: a tab with one paragraph
# looks broken). Pulled from the same `leaderboard` target and the same
# `survivorship_biased` flag the Strategy Leaderboard dashboard renders
# (R/plan_leaderboard.R:552) -- never a hand-typed strategy list, which
# would silently drift the moment the flag changes
# (`reproducible-ingestion`/`fail-loud-not-null` rules).
sb_lb <- safe_tar_read("leaderboard")
if (!is.null(sb_lb) && "survivorship_biased" %in% names(sb_lb)) {
  sb_tbl <- sb_lb |>
    filter(period == "Full Period", !is.na(survivorship_biased), survivorship_biased) |>
    mutate(
      Universe = "`stk_universe`",
      Sharpe = ifelse(is.na(sharpe), "—", as.character(round(sharpe, 2)))
    ) |>
    select(Strategy = strategy, Universe, Sharpe)

  if (nrow(sb_tbl) > 0) {
    hd_dt(sb_tbl, paste0(
      "Strategies on this dashboard flagged `survivorship_biased = TRUE` in the ",
      "[Strategy Leaderboard](leaderboard.html#rankings) target -- see issue #150 for the ",
      "data-availability constraint. `stk_universe` restricts to the top-N tickers by ",
      "current market cap; delisted firms are absent from history."
    ))
  } else {
    cat("*No strategy on this dashboard is currently flagged `survivorship_biased = TRUE`.*\n")
  }
} else {
  cat("*`leaderboard` target not available -- per-strategy flag table cannot be computed.*\n")
}
```

#### About This Page

This dashboard-wide caveat used to sit at the very top of the [Comparison](#comparison)
page, ahead of any content. It moved here so a first-time reader sees the
strategy comparison before the caveat, while a compact pointer at the top of
[Comparison](#comparison) keeps it one click away rather than hidden.

**Caveats collected on this page:**

- **[Survivorship Bias](#biases-and-caveats)** — `stk_universe` (used by Stock MAX,
  Stock DRIF, XGB DRIF above) restricts to top-N tickers by current market cap;
  delisted firms are absent from history. See [issue #150](https://github.com/JohnGavin/historical/issues/150).

The **[Strategy Leaderboard](leaderboard.html#biases-and-caveats)** dashboard
collects the same caveat, plus a second one (statistical detection power and
rigour), in its own equivalent page — the leaderboard ranks the strategies
built on this same `stk_universe` alongside every other strategy in the
project. See [issue #150](https://github.com/JohnGavin/historical/issues/150)
for the underlying data-availability constraint and remediation status.
